### 2017-04-07 paper/comoving-followup.tex
sketch layout

\documentclass[modern, letterpaper]{aastex61}

% TODO:
% -

% style notes
% -----------
% - Use the Makefile; don't be typing ``pdflatex'' or some bullshit.
% - Line break between sentences to make the git diffs readable.
% - Use \, as a multiply operator.
% - Reserve () for function arguments; use [] or {} for outer shit.
% - Use \sectionname not Section, \figname not Figure, \documentname not Article

\include{gitstuff}
\include{preamble}

% Project-specific
\newcommand{\gaia}{\project{Gaia}}

\begin{document}

\title{Co-moving stars in \textsl{Gaia DR1}:
       a catalog of genuine co-moving star pairs from low-resolution spectroscopic follow-up}

\author{Adrian~M.~Price-Whelan}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}
\affiliation{To whom correspondence should be addressed:
             \texttt{adrn@astro.princeton.edu}}

\author{Semyeong~Oh}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\author{Melissa~Ness}
\affiliation{todo}

\author{David~W.~Hogg}
\affiliation{todo}

\author{David~N.~Spergel}
\affiliation{Flatiron Institute,
             Simons Foundation,
             162 Fifth Avenue,
             New York, NY 10010, USA}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\begin{abstract}
TODO:

\end{abstract}

\keywords{
    TODO
}

\section{Introduction}\label{sec:introduction}

\section{Observations and Data reduction}\label{sec:reduction}

See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example

TODO:
- Describe pipeline...

\section{Results}

\subsection{Catalog ...}\label{sec:}

\subsection{Physical properties of genuine co-moving pairs}\label{sec:}

\subsection{Galactic orbits of the co-moving pairs}\label{sec:}

- Orbits of true pairs in the Galaxy, any halo or thick-disk?

\section{Conclusions }

\acknowledgements
...

\bibliographystyle{aasjournal}
\bibliography{refs}

\end{document}

### 2017-04-07 scripts/identify_wavelengths.py
more gaussian process bitches

@@ -23,7 +23,7 @@ from scipy.interpolate import InterpolatedUnivariateSpline
 # Package
 from comoving_rv.log import logger
 from comoving_rv.longslit import voigt_polynomial
-from comoving_rv.longslit.wavelength import fit_spec_line
+from comoving_rv.longslit.wavelength import fit_spec_line_GP, gp_to_fit_pars
 
 class GUIWavelengthSolver(object):
 
@@ -57,12 +57,12 @@ class GUIWavelengthSolver(object):
         if init_map is None:
             self._map_dict['wavel'] = []
             self._map_dict['pixel'] = []
-            self._map_dict['pixel_err'] = []
+            # self._map_dict['pixel_err'] = []
 
         else:
             self._map_dict['wavel'] = init_map['wavelength']
             self._map_dict['pixel'] = init_map['pixel']
-            self._map_dict['pixel_err'] = np.zeros_like(init_map['pixel']) # TODO: hack
+            # self._map_dict['pixel_err'] = np.zeros_like(init_map['pixel']) # TODO: hack
 
         self._line_std_G = None
         self._line_fwhm_L = None
@@ -161,7 +161,8 @@ class GUIWavelengthSolver(object):
             wave_val = self.line_list[idx]
             self._done_wavel_idx.append(idx)
 
-        line_props, line_cov = self.get_line_props(xmin, xmax)
+        # line_props, line_cov = self.get_line_props(xmin, xmax)
+        line_props,_ = self.get_line_props(xmin, xmax)
         if line_props is None:
             return
 
@@ -173,7 +174,7 @@ class GUIWavelengthSolver(object):
 
         self._map_dict['wavel'].append(wave_val)
         self._map_dict['pixel'].append(line_props['x0'])
-        self._map_dict['pixel_err'].append(np.sqrt(line_cov[1,1]))
+        # self._map_dict['pixel_err'].append(np.sqrt(line_cov[1,1]))
 
     def get_line_props(self, xmin, xmax, **kwargs):
         i1 = int(np.floor(xmin))
@@ -188,8 +189,8 @@ class GUIWavelengthSolver(object):
             flux_ivar = None
 
         try:
-            line_props,line_cov = fit_spec_line(pix, flux, flux_ivar, n_bg_coef=3,
-                                                return_cov=True, **kwargs)
+            gp = fit_spec_line_GP(pix, flux, flux_ivar, n_bg_coef=1, absorp_emiss=1.,
[…]

### 2017-04-07 comoving_rv/longslit/wavelength.py
change some defaults with gp

@@ -15,7 +15,7 @@ from .models import voigt_polynomial
 __all__ = ['fit_spec_line', 'fit_spec_line_GP']
 
 def par_dict_to_list(p):
-    return [p['amp'], p['x0'], p['std_G'], p['fwhm_L']] + p['bg'].tolist()
+    return [p['amp'], p['x0'], p['std_G'], p['fwhm_L']] + p['bg_coef'].tolist()
 
 def get_init_guess(x, flux, ivar,
                    amp0=None, x0=None, std_G0=None, fwhm_L0=None,
@@ -75,7 +75,7 @@ def get_init_guess(x, flux, ivar,
     p0['x0'] = x0
     p0['std_G'] = std_G0
     p0['fwhm_L'] = fwhm_L0
-    p0['bg'] = bg0
+    p0['bg_coef'] = bg0
     return p0
 
 # ---
@@ -177,9 +177,40 @@ class MeanModel(Model):
                                 [getattr(self, "bg{}".format(i))
                                  for i in range(self._n_bg_coef)])
 
+def gp_to_fit_pars(gp, absorp_emiss):
+    """
+    Given a GP instance, return a parameter dictionary for the
+    mean model parameters only.
+    """
+    fit_pars = OrderedDict()
+    for k,v in gp.get_parameter_dict().items():
+        if 'mean' not in k:
+            continue
+
+        k = k[5:] # remove 'mean:'
+        if k.startswith('ln'):
+            if 'amp' in k:
+                fit_pars[k[3:]] = absorp_emiss * np.exp(v)
+            else:
+                fit_pars[k[3:]] = np.exp(v)
+
+        elif k.startswith('bg'):
+            if 'bg_coef' not in fit_pars:
+                fit_pars['bg_coef'] = []
+            fit_pars['bg_coef'].append(v)
+
+        else:
+            fit_pars[k] = v
+
+    if 'std_G' not in fit_pars:
+        fit_pars['std_G'] = 1E-10
+
+    return fit_pars
+
 def fit_spec_line_GP(x, flux, flux_ivar=None,
                      amp0=None, x0=None, std_G0=None, fwhm_L0=None,
                      bg0=None, n_bg_coef=2, target_x=None,
+                     log_sigma0=None, log_rho0=None, # GP parameters
                      absorp_emiss=1., minimize_kw=None):
 
[…]

### 2017-05-17 scripts/test_solve_velocity.py
whatevs

@@ -3,8 +3,6 @@
 """
 TODO:
 - n1.0073 Halpha is emission
-- Save results to a database file of some kind...sqlite? yaml file?
-
 """
 
 # Standard library
@@ -15,7 +13,6 @@ import glob
 
 # Third-party
 from astropy.constants import c
-import astropy.coordinates as coord
 import astropy.units as u
 from astropy.io import fits
 from astropy.table import Table, Column
@@ -28,34 +25,8 @@ import schwimmbad
 
 # Project
 from comoving_rv.log import logger
-from comoving_rv.longslit.wavelength import fit_spec_line_GP
+from comoving_rv.longslit.fitting import fit_spec_line_GP
 from comoving_rv.longslit.models import voigt_polynomial
-
-def get_fit_pars(gp):
-    fit_pars = OrderedDict()
-    for k,v in gp.get_parameter_dict().items():
-        if 'mean' not in k:
-            continue
-
-        k = k[5:] # remove 'mean:'
-        if k.startswith('ln'):
-            if 'amp' in k:
-                fit_pars[k[3:]] = -np.exp(v)
-            else:
-                fit_pars[k[3:]] = np.exp(v)
-
-        elif k.startswith('bg'):
-            if 'bg_coef' not in fit_pars:
-                fit_pars['bg_coef'] = []
-            fit_pars['bg_coef'].append(v)
-
-        else:
-            fit_pars[k] = v
-
-    if 'std_G' not in fit_pars:
-        fit_pars['std_G'] = 1E-10
-
-    return fit_pars
 
 def log_probability(params, gp, flux_data):
     gp.set_parameter_vector(params)
@@ -77,11 +48,14 @@ def log_probability(params, gp, flux_data):
     return ll + lp
 
 def main(overwrite=False):
-    night = 'n5'
+    night = 'n1' # HACK TODO
 
     plot_path = path.abspath('../plots/')
     root_path = path.abspath('../data/mdm-spring-2017/processed/')
-    table_path = path.join(root_path, 'rough_velocity.ecsv')
+    table_path = path.join(root_path, 'velocity.ecsv')
+
+    # air wavelength of Halpha -- wavelength calibration from comp lamp is done
+    #   at air wavelengths, so this is where Halpha should be, right?
     Halpha = 6562.8 * u.angstrom
 
     if not path.exists(plot_path):
@@ -103,15 +77,7 @@ def main(overwrite=False):
         logger.debug('Table exists, reading ({})'.format(table_path))
[…]

### 2017-05-17 scripts/test_solve_velocity.py
remove test script

@@ -1,327 +1,0 @@
-# coding: utf-8
-
-"""
-TODO:
-- n1.0073 Halpha is emission
-"""
-
-# Standard library
-import os
-from os import path
-from collections import OrderedDict
-import glob
-
-# Third-party
-from astropy.constants import c
-import astropy.units as u
-from astropy.io import fits
-from astropy.table import Table, Column
-import numpy as np
-import matplotlib.pyplot as plt
-plt.style.use('apw-notebook')
-import emcee
-import corner
-import schwimmbad
-
-# Project
-from comoving_rv.log import logger
-from comoving_rv.longslit.fitting import fit_spec_line_GP
-from comoving_rv.longslit.models import voigt_polynomial
-
-def log_probability(params, gp, flux_data):
-    gp.set_parameter_vector(params)
-    lp = gp.log_prior()
-    if not np.isfinite(lp):
-        return -np.inf
-
-    # HACK: Gaussian prior on log(rho)
-    var = 1.
-    lp += -0.5*(params[1]-1)**2/var - 0.5*np.log(2*np.pi*var)
-
-    if params[4] < -10. or params[5] < -10.:
-        return -np.inf
-
-    ll = gp.log_likelihood(flux_data)
-    if not np.isfinite(ll):
-        return -np.inf
-
-    return ll + lp
-
-def main(overwrite=False):
-    night = 'n1' # HACK TODO
-
-    plot_path = path.abspath('../plots/')
-    root_path = path.abspath('../data/mdm-spring-2017/processed/')
-    table_path = path.join(root_path, 'velocity.ecsv')
-
-    # air wavelength of Halpha -- wavelength calibration from comp lamp is done
-    #   at air wavelengths, so this is where Halpha should be, right?
-    Halpha = 6562.8 * u.angstrom
-
-    if not path.exists(plot_path):
-        os.makedirs(plot_path, exist_ok=True)
-
-    if not path.exists(table_path):
-        logger.debug('Creaing table at {}'.format(table_path))
-        tbl_init = [Column(name='object_name', dtype='|S30', data=[], length=0),
-                    Column(name='group_id', dtype=int, length=0),
-                    Column(name='filename', dtype='|S128', data=[], length=0),
-                    Column(name='wave0', dtype=float, unit=u.angstrom, length=0),
-                    Column(name='rv', dtype=float, unit=u.km/u.s, length=0),
-                    Column(name='rv_precision', dtype=float, unit=u.km/u.s, length=0)]
-        velocity_tbl = Table(tbl_init)
[…]

### 2017-05-17 comoving_rv/longslit/wavelength.py
ad fit all lines func

@@ -1,8 +1,3 @@
-"""
-TODO:
--
-"""
-
 # Third-party
 import numpy as np
 from numpy.polynomial.polynomial import polyval
@@ -12,7 +7,60 @@ from celerite import terms, GP
 # Project
 from comoving_rv.log import logger
 
-__all__ = ['MeanModel', 'GPModel']
+__all__ = ['fit_all_lines', 'MeanModel', 'GPModel']
+
+def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels, half_width=5):
+    """
+    Given a wavelength guess (a list of rough pixel-wavelength correspondences),
+    measure more precise line centroids to then fit for a wavelength
+    calibratrion function.
+
+    Parameters
+    ----------
+    pixels : array_like
+        Pixel array for a 1D comp. lamp spectrum.
+    flux : array_like
+        Flux array for a 1D comp. lamp spectrum.
+    flux_ivar : array_like
+        Inverse-variance array for a 1D comp. lamp spectrum.
+    line_waves : array_like
+        List of wavelength values for emission lines in a comp. lamp spectrum.
+    line_pixels : array_like
+        List of pixel centroids for emission lines in a comp. lamp spectrum.
+    half_width : int
+        Number of pixels on either side of line
+    """
+
+    _idx = np.argsort(line_waves)
+    wvln = np.array(line_waves)[_idx]
+    pixl = np.array(line_pixels)[_idx]
+
+    fit_centroids = []
+
+    for pix_ctr,wave in zip(pixl, wvln):
+
+        logger.debug("Fitting line at predicted pix={:.2f}, λ={:.2f}"
+                     .format(pix_ctr, wave))
+
+        # indices for region around line
+        i1 = int(np.floor(pix_ctr-half_width))
+        i2 = int(np.ceil(pix_ctr+half_width)) + 1
+
+        # recenter window
+        i0 = i1 + flux[i1:i2].argmax()
+        i1 = int(np.floor(i0-half_width))
+        i2 = int(np.ceil(i0+half_width)) + 1
+
+        _pixl = pixels[i1:i2]
+        _flux = flux[i1:i2]
+        _ivar = flux_ivar[i1:i2]
+
+        # instead of doing anything fancy (e.g., fitting a profile), just
+        # estimate the mean...
+        x0 = np.sum(_pixl * _flux**2 * _ivar) / np.sum(_flux**2 * _ivar)
[…]

### 2017-05-17 comoving_rv/velocity.py
hard code KPNO location

@@ -4,7 +4,9 @@ import astropy.units as u
 
 __all__ = ['bary_vel_corr']
 
-kitt_peak = coord.EarthLocation.of_site('KPNO')
+# kitt_peak = coord.EarthLocation.of_site('KPNO')
+# TODO: HACK: approximate
+kitt_peak = coord.EarthLocation(lon=-111.6150*u.degree, lat=31.95*u.degree)
 
 def bary_vel_corr(time, skycoord, location=None):
     """

### 2017-05-17 scripts/wavelength_calibrate.py
oops moved function to package

@@ -21,7 +21,7 @@ from scipy.optimize import minimize
 # Project
 from comoving_rv.log import logger
 from comoving_rv.longslit import GlobImageFileCollection
-from comoving_rv.longslit.wavelength import GPModel
+from comoving_rv.longslit.wavelength import GPModel, fit_all_lines
 
 # ----------------------------------------------------------------------------
 # Settings, or arbitrary choices / magic numbers
@@ -29,57 +29,6 @@ from comoving_rv.longslit.wavelength import GPModel
 n_bg_coef = 2 # linear
 #
 # ----------------------------------------------------------------------------
-
-def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels):
-    """
-    Given a wavelength guess (a list of rough pixel-wavelength correspondences),
-    measure more precise line centroids to then fit for a wavelength
-    calibratrion function.
-
-    Parameters
-    ----------
-    pixels : array_like
-        Pixel array for a 1D comp. lamp spectrum.
-    flux : array_like
-        Flux array for a 1D comp. lamp spectrum.
-    flux_ivar : array_like
-        Inverse-variance array for a 1D comp. lamp spectrum.
-    line_waves : array_like
-        List of wavelength values for emission lines in a comp. lamp spectrum.
-    line_pixels : array_like
-        List of pixel centroids for emission lines in a comp. lamp spectrum.
-    """
-
-    _idx = np.argsort(line_waves)
-    wvln = np.array(line_waves)[_idx]
-    pixl = np.array(line_pixels)[_idx]
-
-    fit_centroids = []
-
-    half_width = 5 # MAGIC NUMBER: number of pixels on either side of line
-    for pix_ctr,wave in zip(pixl, wvln):
-
-        logger.debug("Fitting line at predicted pix={:.2f}, λ={:.2f}"
-                     .format(pix_ctr, wave))
-
-        # indices for region around line
-        i1 = int(np.floor(pix_ctr-half_width))
-        i2 = int(np.ceil(pix_ctr+half_width))+1
-
-        # recenter window
-        i0 = i1 + flux[i1:i2].argmax()
-        i1 = int(np.floor(i0-half_width))
-        i2 = int(np.ceil(i0+half_width))+1
-
-        _pixl = pixels[i1:i2]
-        _flux = flux[i1:i2]
-
-        # instead of doing anything fancy (e.g., fitting a profile), just
-        # estimate the mean...
[…]

### 2017-05-17 paper/comoving-followup.tex
brainstorming intro and structure

@@
 
-\section{Observations and Data reduction}\label{sec:reduction}
+Outline:
+- Stars form in clusters, chemically homogeneous at some level
+- Clusters generally disrupt slowly through a process that preserves initial
+  clustering in conserved quantities but destroys phase coherence
+- Most significant process is probably tidal heating
+- Dissolved clusters are therefore useful dynamical tracers of mass distribution
+- In Galactic halo, globular clusters
+- In disk, hope is to use chemical signatures to connect stars, then can use
+  dynamics to reconstruct birth clusters, infer mass distribution / evolution
+- Clusters have advantage of many stars per cluster, but fewer clusters
+- Wide binary stars are little potentiometers themselves - only N=2, but are far
+  more numerous as a population
+- Long-term goal is to study the separation distribution of wide-binaries as a
+  function of age and Galactic orbit distribution (disk vs. halo)
+    - Is there a separation truncation?
+    - Can that be described just from Galactic tides, or need perturbers?
+- Also useful for stellar models: should have same age, chemistry, but need to
+  test this
+    - Chemical homogeneity tested with few open clusters - wide binaries another
+      test
+    - Need sample of confirmed widely-separated binaries, follow-up to get
+      detailed abundances
+    - Need way to verify that widely separated stars are physically associated,
+      via kinematics
+
+Most stars form in clusters with small dispersions in position, velocity, and
+(likely) small initial spreads in chemical abundances.
+Once a given cluster is visibly dispersing or fully dispersed, the former member
+stars retain chemical and kinematic signatures of their birth cluster
+
+, though
+the kinematic mapping is non-trivial because of phase-mixing.
+
+become
+useful
+
+Widely-separated co-moving stars are expected to form from the dissolution of
+birth star clusters
+
+\section{Data}\label{sec:data}
+
+\subsection{Sample selection}\label{sec:sample}
+
+Target sample taken from ... Oh et al.
+
+Briefly, ...
+
+TGAS roughly XX\% complete to 200 pc with parallax SNR cut.
+
+Oh catalog have a high likelihood-ratio cut, but that should be unbiased (right?)
+
+Randomly sample from that, subject to observational constraints.
+
+\subsection{Observations and Data reduction}\label{sec:reduction}
 
@@
 - Describe pipeline...
+
+- Comparing repeat visits on subsequent nights, systematic ~20 km/s
+    - see notebook "Compare RVs with previous measurements"
 

### 2017-05-17 scripts/solve_velocity.py
change table contents

@@ -378,8 +378,7 @@ def main(night_path, overwrite=False):
         if np.isnan(shift):
             shift = 0. * u.angstrom
 
-        centroid = centroid + shift
-        rv = (centroid - Halpha) / Halpha * c.to(u.km/u.s) + vbary
+        rv = (centroid + shift - Halpha) / Halpha * c.to(u.km/u.s) + vbary
         rv_err = centroid_err / Halpha * c.to(u.km/u.s)
         rv_err = np.sqrt(rv_err**2 + (10.*u.km/u.s)**2)
 
@@ -395,12 +394,12 @@ def main(night_path, overwrite=False):
                                   bary_rv_shift=vbary,
                                   sky_shift_flag=sky_flag,
                                   sky_wave_shift=wave_shifts,
-                                  rv=rv + vbary,
+                                  rv=rv,
                                   rv_err=rv_err))
 
         logger.info('{} [{}]: x0={x0:.3f} σ={err:.3f} rv={rv:.3f}'
                     .format(object_name, filebase, x0=centroid,
-                            err=centroid_err, rv=rv + vbary))
+                            err=centroid_err, rv=rv))
 
         velocity_tbl.write(table_path, format='fits', overwrite=True)
 

### 2017-05-17 scripts/solve_velocity.py
fix pep8 and add pool support

@@ -22,6 +22,7 @@ plt.style.use('apw-notebook')
 import emcee
 import corner
 import schwimmbad
+from schwimmbad import choose_pool
 
 # Project
 from comoving_rv.log import logger
@@ -49,7 +50,10 @@ def log_probability(params, gp, flux_data):
 
     return ll + lp
 
-def main(night_path, overwrite=False):
+def main(night_path, overwrite=False, pool=None):
+
+    if pool is None:
+        pool = schwimmbad.SerialPool()
 
     night_path = path.realpath(path.expanduser(night_path))
     if not path.exists(night_path):
@@ -128,7 +132,8 @@ def main(night_path, overwrite=False):
 
         if object_name in velocity_tbl['object_name']:
             if overwrite:
-                logger.debug('Object {} already done - overwriting!'.format(object_name))
+                logger.debug('Object {} already done - overwriting!'
+                             .format(object_name))
                 idx, = np.where(velocity_tbl['object_name'] == object_name)
                 for i in idx:
                     velocity_tbl.remove_row(i)
@@ -201,7 +206,8 @@ def main(night_path, overwrite=False):
         ax.set_ylabel('flux')
 
         fig.tight_layout()
-        fig.savefig(path.join(plot_path, '{}_maxlike.png'.format(filebase)), dpi=256)
+        fig.savefig(path.join(plot_path, '{}_maxlike.png'.format(filebase)),
+                    dpi=256)
         plt.close(fig)
         # ------------------------------------------------------------------------
 
@@ -216,33 +222,35 @@ def main(night_path, overwrite=False):
             initial[5] = -8.
         ndim, nwalkers = len(initial), 64
 
-        with schwimmbad.MultiPool() as pool:
-            # with schwimmbad.SerialPool() as pool:
-            sampler = emcee.EnsembleSampler(nwalkers, ndim, log_probability, pool=pool,
-                                            args=(gp, flux_data))
+        sampler = emcee.EnsembleSampler(nwalkers, ndim, log_probability,
+                                        pool=pool, args=(gp, flux_data))
 
-            logger.debug("Running burn-in...")
-            p0 = initial + 1e-6 * np.random.randn(nwalkers, ndim)
-            p0, lp, _ = sampler.run_mcmc(p0, 256)
+        logger.debug("Running burn-in...")
+        p0 = initial + 1e-6 * np.random.randn(nwalkers, ndim)
+        p0, lp, _ = sampler.run_mcmc(p0, 256)
 
[…]

### 2017-05-17 scripts/solve_velocity.py
store the precision, and do all objects not just HIP

@@ -127,8 +127,8 @@ def main(night_path, overwrite=False, pool=None):
         object_name = hdr['OBJECT']
 
         # HACK: for testing
-        if 'HIP' not in object_name:
-            continue
+        # if 'HIP' not in object_name:
+        #     continue
 
         if object_name in velocity_tbl['object_name']:
             if overwrite:
@@ -394,8 +394,7 @@ def main(night_path, overwrite=False, pool=None):
             shift = 0. * u.angstrom
 
         rv = (centroid + shift - Halpha) / Halpha * c.to(u.km/u.s) + vbary
-        rv_err = centroid_err / Halpha * c.to(u.km/u.s)
-        rv_err = np.sqrt(rv_err**2 + (10.*u.km/u.s)**2)
+        rv_err = centroid_err / Halpha * c.to(u.km/u.s) # formal precision
 
         # convert ra,dec to quantities
         ra = sc.ra.degree * u.deg

### 2017-05-17 scripts/test_skyline_shift.py
dunno

@@ -53,8 +53,8 @@ def main(overwrite=False):
     table_path = path.join(root_path, 'rough_velocity.ecsv')
 
     # config settings
-    sky_lines = [5577.3387, 6300.3] # [OI]
-    width = 100. # angstroms centered on line
+    sky_lines = [6300.304, 6363.776] # from: http://physics.nist.gov/PhysRefData/ASD/lines_form.html
+    width = 100. # window size in angstroms, centered on line
     absorp_emiss = 1. # all emission lines
     gp_color = "#ff7f0e"
 
@@ -127,7 +127,7 @@ def main(overwrite=False):
             dlam = pars['mean:x0'] - sky_line
             vshift = (dlam/sky_line*c).to(u.km/u.s).value
 
-            if ((pars['mean:ln_fwhm_L'] < 0 and pars['mean:ln_std_G'] < 0) or pars['mean:ln_amp'] > 10.):
+            if ((pars['mean:ln_fwhm_L'] < -0.5 and pars['mean:ln_std_G'] < (-0.5)) or pars['mean:ln_amp'] > 10.):
                 title = 'fucked'
             else:
                 title = '{:.2f}'.format(pars['mean:ln_amp'])

### 2017-05-17 scripts/test_skyline_shift.py
remove test script

@@ -1,203 +1,0 @@
-# coding: utf-8
-
-"""
-Major hack to add skyline shifts to the velocity table
-"""
-
-# Standard library
-from os import path
-from collections import OrderedDict
-
-# Third-party
-from astropy.constants import c
-import astropy.units as u
-from astropy.io import fits
-from astropy.table import Table
-import numpy as np
-import matplotlib.pyplot as plt
-plt.style.use('apw-notebook')
-
-# Project
-from comoving_rv.log import logger
-from comoving_rv.longslit.wavelength import fit_spec_line_GP
-
-def get_fit_pars(gp):
-    fit_pars = OrderedDict()
-    for k,v in gp.get_parameter_dict().items():
-        if 'mean' not in k:
-            continue
-
-        k = k[5:] # remove 'mean:'
-        if k.startswith('ln'):
-            if 'amp' in k:
-                fit_pars[k[3:]] = -np.exp(v)
-            else:
-                fit_pars[k[3:]] = np.exp(v)
-
-        elif k.startswith('bg'):
-            if 'bg_coef' not in fit_pars:
-                fit_pars['bg_coef'] = []
-            fit_pars['bg_coef'].append(v)
-
-        else:
-            fit_pars[k] = v
-
-    if 'std_G' not in fit_pars:
-        fit_pars['std_G'] = 1E-10
-
-    return fit_pars
-
-def main(overwrite=False):
-    plot_path = path.abspath('../plots/')
-    root_path = path.abspath('../data/mdm-spring-2017/processed/')
-    table_path = path.join(root_path, 'rough_velocity.ecsv')
-
-    # config settings
-    sky_lines = [6300.304, 6363.776] # from: http://physics.nist.gov/PhysRefData/ASD/lines_form.html
-    width = 100. # window size in angstroms, centered on line
-    absorp_emiss = 1. # all emission lines
-    gp_color = "#ff7f0e"
-
-    if not path.exists(plot_path) or not path.exists(table_path):
-        raise IOError("You must run test_solve_velocity first!")
-
-    logger.debug('Table exists, reading ({})'.format(table_path))
-    velocity_tbl = Table.read(table_path, format='ascii.ecsv', delimiter=',')
-    if 'skyline_shift' not in velocity_tbl.colnames:
-        velocity_tbl['skyline_shift'] = np.full(len(velocity_tbl), np.nan)
-
-    for row in velocity_tbl[velocity_tbl['object_name'] == '483-209']:
[…]

### 2017-05-18 scripts/wavelength_calibrate.py
whatever

@@ -3,7 +3,6 @@
 TODO:
 - Make the functions below part of a class so we can store diagnostic
     plots along the way (easily keep track of paths)
-- Need to do sky-spectrum adjustments here when we can
 """
 
 # Standard library
@@ -51,7 +50,7 @@ def generate_wavelength_model(comp_lamp_path, night_path, plot_path):
                                         '..', 'wavelength_guess.csv'))
     pix_wav = np.genfromtxt(guess_path, delimiter=',', names=True)
 
-    # fit line profiles to each emission line at the guessed positions of the lines
+    # get emission line centroids at the guessed positions of the lines
     pix_x0s = fit_all_lines(spec['pix'], spec['flux'], spec['ivar'],
                             pix_wav['wavelength'], pix_wav['pixel'])
 

### 2017-05-18 paper/comoving-followup.tex
notes

@@
 \newcommand{\gaia}{\project{Gaia}}
+\newcommand{\DR}[1]{\acronym{DR}1}
+\newcommand{\tgas}{\acronym{TGAS}}
 
@@
 \title{Co-moving stars in \textsl{Gaia DR1}:
-       a catalog of genuine co-moving star pairs from low-resolution spectroscopic follow-up}
+       a catalog of genuine co-moving star pairs from low-resolution
+       spectroscopic follow-up}
 
@@
 
-\author{David~W.~Hogg}
-\affiliation{todo}
-
 \author{David~N.~Spergel}
@@
              Princeton University, Princeton, NJ 08544, USA}
+
+% \author{David~W.~Hogg}
+% \affiliation{todo}
 
@@
       via kinematics
+- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\
 
@@
 \subsection{Sample selection}\label{sec:sample}
+
+The wide-separation co-moving stars presented later in this work were originally
+identified as candidate co-moving stars in a search of the \tgas\ catalog.
 
@@
 \acknowledgements
-...
+David W. Hogg (NYU)
 

### 2017-05-19 scripts/extract_1d.py
change path

@@ -74,13 +74,15 @@ def main(night_path, skip_list_file, mask_file, overwrite=False, plot=False):
         raise IOError("Path '{}' doesn't exist".format(night_path))
     logger.info("Reading data from path: {}".format(night_path))
 
-    base_path, name = path.split(night_path)
-    output_path = path.realpath(path.join(base_path, 'processed', name))
+    base_path, night_name = path.split(night_path)
+    data_path, run_name = path.split(base_path)
+    output_path = path.realpath(path.join(data_path, 'processed',
+                                          run_name, night_name))
     os.makedirs(output_path, exist_ok=True)
     logger.info("Saving processed files to path: {}".format(output_path))
 
     if plot: # if we're making plots
-        plot_path = path.realpath(path.join(base_path, 'processed', name, 'plots'))
+        plot_path = path.realpath(path.join(output_path, 'plots'))
         logger.debug("Will make and save plots to: {}".format(plot_path))
         os.makedirs(plot_path, exist_ok=True)
     else:

### 2017-05-26 scripts/injest_db.py
modify inject script to remove hacks

@@ -1,6 +1,7 @@
 # Standard library
 from os import path
 import glob
+from collections import OrderedDict
 
 # Third-party
 import astropy.coordinates as coord
@@ -68,29 +69,33 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
     session = Session()
 
     logger.debug("Loading SpectralLineInfo table")
-    lines = []
 
+    line_info = OrderedDict()
     # air wavelength of Halpha -- wavelength calibration from comp lamp is done
     #   at air wavelengths, so this is where Halpha should be, right?
-    lines.append(SpectralLineInfo(name='Halpha',
-                                  wavelength=6562.8*u.angstrom))
+    line_info['Halpha'] = 6562.8*u.angstrom
 
     # [OI] emission lines -- wavelengths from:
     #   http://physics.nist.gov/PhysRefData/ASD/lines_form.html
-    lines.append(SpectralLineInfo(name='[OI] 5577',
-                                  wavelength=5577.3387*u.angstrom))
-    lines.append(SpectralLineInfo(name='[OI] 6300',
-                                  wavelength=6300.304*u.angstrom))
-    lines.append(SpectralLineInfo(name='[OI] 6364',
-                                  wavelength=6363.776*u.angstrom))
-    session.add_all(lines)
-    session.commit()
+    line_info['[OI] 5577'] = 5577.3387*u.angstrom
+    line_info['[OI] 6300'] = 6300.304*u.angstrom
+    line_info['[OI] 6364'] = 6363.776*u.angstrom
+
+    for name, wvln in line_info.items():
+        n = session.query(SpectralLineInfo).filter(SpectralLineInfo.name == name).count()
+        if n == 0:
+            logger.debug('Loading line {0} at {1}'.format(name, wvln))
+            line = SpectralLineInfo(name=name, wavelength=wvln)
+            session.add(line)
+            session.commit()
 
     # Create an entry for this observing run
     data_path, run_name = path.split(run_root_path)
-    run = Run(name=run_name)
-    session.add(run)
-    session.commit()
+    n = session.query(Run).filter(Run.name == run_name).count()
+    if n == 0:
+        run = Run(name=run_name)
+        session.add(run)
+        session.commit()
 
     # Now we need to go through each processed night of data and load all of the
     # relevant observations of sources.
@@ -103,7 +108,7 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
[…]

### 2017-05-26 comoving_rv/db/model.py
fix filenames

@@ -86,18 +86,22 @@ class Observation(Base):
     tgas_source = relationship('TGASSource', cascade='all,delete-orphan',
                                backref='observation', single_parent=True)
 
+    def __repr__(self):
+        return ("<Observation {0.object}, {0.filename_raw}, {0.run.name}>"
+                .format(self))
+
     def path_raw(self, base_path):
-        p = path.join(base_path, self.run.name, 'n'.str(self.night))
+        p = path.join(base_path, self.run.name, 'n'+str(self.night))
         return path.join(p, self.filename_raw)
 
     def path_p(self, base_path):
         p = path.join(base_path, 'processed', self.run.name,
-                      'n'.str(self.night))
+                      'n'+str(self.night))
         return path.join(p, self.filename_p)
 
     def path_1d(self, base_path):
         p = path.join(base_path, 'processed', self.run.name,
-                      'n'.str(self.night))
+                      'n'+str(self.night))
         return path.join(p, self.filename_1d)
 
 
@@ -113,6 +117,16 @@ class SimbadInfo(Base):
     rv = Column('rv', VelocityType)
     rv_qual = Column('rv_qual', types.String)
     rv_bibcode = Column('rv_bibcode', types.String)
+
+    def __repr__(self):
+        names = []
+        pres = ['HD', 'HIP', 'TYC', '2MASS']
+        for pre,id_ in zip(pres, ['hd_id', 'hip_id', 'tyc_id', 'twomass_id']):
+            if getattr(self, id_) is not None:
+                names.append('{0} {1}'.format(pre, getattr(self, id_)))
+
+        return ("<SimbadInfo {0}>"
+                .format(', '.join(names)))
 
 class TGASSource(Base):
     __tablename__ = 'tgas_source'
@@ -167,7 +181,7 @@ class SpectralLineMeasurement(Base):
     amp_error = Column('amp_error', types.REAL)
 
     std_G = Column('std_G', types.REAL, nullable=False)
-    std_G_error = Column('std_G_error', types.REAL, nullable=False)
+    std_G_error = Column('std_G_error', types.REAL)
 
     hwhm_L = Column('hwhm_L', types.REAL)
     hwhm_L_error = Column('hwhm_L_error', types.REAL)

### 2017-05-26 scripts/solve_velocity.py
populate database not fits file

@@ -3,6 +3,7 @@
 """
 TODO:
 - n1.0073 Halpha is emission
+- Modify script so it writes to the database instead of a FITS file
 """
 
 # Standard library
@@ -26,9 +27,11 @@ from schwimmbad import choose_pool
 
 # Project
 from comoving_rv.log import logger
-from comoving_rv.longslit import GlobImageFileCollection, extract_region
+from comoving_rv.db import Session, Base, db_connect
+from comoving_rv.longslit import extract_region
 from comoving_rv.longslit.fitting import VoigtLineFitter, GaussianLineFitter
-from comoving_rv.velocity import bary_vel_corr, kitt_peak
+from comoving_rv.db.model import (Run, Observation, TGASSource, SimbadInfo,
+                                  SpectralLineMeasurement, SpectralLineInfo)
 
 def log_probability(params, gp, flux_data):
     gp.set_parameter_vector(params)
@@ -49,106 +52,66 @@ def log_probability(params, gp, flux_data):
 
     return ll + lp
 
-def main(night_path, overwrite=False, pool=None):
+def main(db_path, run_name, filename=None, overwrite=False, pool=None):
 
     if pool is None:
         pool = schwimmbad.SerialPool()
 
-    night_path = path.realpath(path.expanduser(night_path))
-    if not path.exists(night_path):
-        raise IOError("Path '{}' doesn't exist".format(night_path))
+    # connect to the database
+    engine = db_connect(db_path)
+    # engine.echo = True
+    logger.debug("Connected to database at '{}'".format(db_path))
 
-    if path.isdir(night_path):
-        data_file = None
-        logger.info("Reading data from path: {}".format(night_path))
+    # create a new session for interacting with the database
+    session = Session()
 
-    elif path.isfile(night_path):
-        data_file = night_path
-        base_path, name = path.split(night_path)
-        night_path = base_path
-        logger.info("Reading file: {}".format(data_file))
-
-    else:
-        raise RuntimeError("how?!")
-
-    plot_path = path.join(night_path, 'plots')
-    root_path = path.abspath(path.join(night_path, '..'))
-    table_path = path.join(root_path, 'velocity.fits')
-
-    # air wavelength of Halpha -- wavelength calibration from comp lamp is done
-    #   at air wavelengths, so this is where Halpha should be, right?
-    Halpha = 6562.8 * u.angstrom
-
[…]

### 2017-05-26 comoving_rv/db/model.py
reprs

@@ -197,9 +197,18 @@ class SpectralLineMeasurement(Base):
                      ForeignKey('spectral_line_info.id'))
     info = relationship('SpectralLineInfo')
 
+    def __repr__(self):
+        r = '<SpectralLineMeasurement {0} x0={1:.3f}>'.format(self.info.name,
+                                                              self.x0)
+        return r
+
 class SpectralLineInfo(Base):
     __tablename__ = 'spectral_line_info'
 
     id = Column(types.Integer, primary_key=True)
     name = Column('name', types.String, nullable=False)
     wavelength = Column('wavelength', WavelengthType, nullable=False)
+
+    def __repr__(self):
+        return '<SpectralLineInfo {0} @ {1}>'.format(self.name, self.wavelength)
+

### 2017-05-27 comoving_rv/db/model.py
get skycoord

@@ -10,6 +10,7 @@ from __future__ import division, print_function
 from os import path
 
 # Third-party
+import astropy.coordinates as coord
 import astropy.units as u
 from sqlalchemy import Column, types
 from sqlalchemy.schema import ForeignKey
@@ -168,6 +169,11 @@ class TGASSource(Base):
     b = Column('b', SexDegAngleType, nullable=False)
     ecl_lon = Column('ecl_lon', SexDegAngleType, nullable=False)
     ecl_lat = Column('ecl_lat', SexDegAngleType, nullable=False)
+
+    @property
+    def skycoord(self):
+        return coord.SkyCoord(ra=self.ra, dec=self.dec,
+                              distance=1000./self.parallax*u.pc)
 
 class SpectralLineMeasurement(Base):
     __tablename__ = 'spectral_line_measurement'

### 2017-05-30 paper/comoving-followup.tex
note to self
summarize previous paper

@@
 \title{Co-moving stars in \textsl{Gaia DR1}:
-       a catalog of genuine co-moving star pairs from low-resolution
+       a catalog of genuine comoving star pairs from low-resolution
        spectroscopic follow-up}
@@
 \begin{abstract}
-TODO:
-
+% Context
+Widely-separated, co-eval stars are [...] for studying the chemical homogeneity
+of birth, calibrating stellar atmosphere models across a range of [..], and for
+modeling the Galactic tidal field.
+Previous work using Hipparcos + literature RVs -> a few hundred genuine pairs
+out to 8 pc.
+% Aims
+Identified candidate pairs using the Tycho-Gaia Astrometric Solution catalog,
+Gaia DR1.
+Want to identify genuinely dynamically associated pairs for future
+high-resolution followup
+% Methods
+Low-resolution spectrograph to measure relative radial velocities to identify
+genuine pairs.
+% Results
+Of the XX pairs observed, we find YY ...
+% Conclusions
+Gaia DR2 -> how many do we expect?
 \end{abstract}
@@
 
+comoving not wide binary (don't know if they are bound)
+
 , though
@@
 
-Widely-separated co-moving stars are expected to form from the dissolution of
+Widely-separated comoving stars are expected to form from the dissolution of
 birth star clusters
@@
 
-The wide-separation co-moving stars presented later in this work were originally
-identified as candidate co-moving stars in a search of the \tgas\ catalog.
+In previous work, we have identified a sample of high-confidence, candidate
+comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
+the \tgas\ catalog (\citealt{tgas}).
+These comoving pairs were selected using a cut on a marginalized likelihood
+ratio computed for all nearly co-spatial pairs of stars.
+For a full description of the selection method, see \citealt{Oh:2017}.
+Briefly, likelihood 1, $L_1$, considers a model in which the full-space
+velocities of each star in the pair are identical and sampled from a disk-like,
+Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
+in which the full-space velocities of each star were drawn independently from
+the same velocity distribution used above; the likelihood ratio was taken as
+$L_1/L_2$.
+After a parallax signal-to-noise cut ($\pi/\sigma_\pi > 8$), we compute the
+likelihood ratio for all stars within $10~\pc$ of each target star from the
+\tgas\ catalog.
+The likelihoods appropriately take into account the reported covariances and
+uncertainties between the astrometric measurements in the \tgas\ catalog and
+naturally handle geometric or projection effects in comparing spherical velocity
+components for star pairs separated by large angles on the sky.
+After applying a conservative cut on the likelihood ratio motivated by computing
+the same likelihood ratio for random pairings of stars, we find 13,058
+comoving star pairs with physical separations between $1000~\au$ and $10~\pc$.
+Of these pairs, there are 10,606 unique sources, implying that a significant
+fraction of the identified pairs are members of larger groups.
 
-Target sample taken from ... Oh et al.
-
-Briefly, ...
-
-TGAS roughly XX\% complete to 200 pc with parallax SNR cut.
+TGAS $\approx$90\% complete to 200 pc with parallax SNR cut (\citealt{Bovy:2017}).
 
@@
 
+TODO:
+- Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).
+
 \subsection{Catalog ...}\label{sec:}
 
-\subsection{Physical properties of genuine co-moving pairs}\label{sec:}
+\subsection{Physical properties of genuine comoving pairs}\label{sec:}
 

### 2017-05-30 paper/comoving-followup.tex
sample paragraph

@@
 in which the full-space velocities of each star were drawn independently from
-the same velocity distribution used above; the likelihood ratio was taken as
+the same velocity distribution used above; the likelihood ratio is computed as
 $L_1/L_2$.
-After a parallax signal-to-noise cut ($\pi/\sigma_\pi > 8$), we compute the
-likelihood ratio for all stars within $10~\pc$ of each target star from the
-\tgas\ catalog.
+After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
+compute the likelihood ratio for all stars within $10~\pc$ of each target star
+from the \tgas\ catalog.
 The likelihoods appropriately take into account the reported covariances and
@@
 the same likelihood ratio for random pairings of stars, we find 13,058
-comoving star pairs with physical separations between $1000~\au$ and $10~\pc$.
+comoving star pairs with physical separations between $10^3~\au$ and $10~\pc$.
 Of these pairs, there are 10,606 unique sources, implying that a significant
 fraction of the identified pairs are members of larger groups.
+$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
+paired with other stars.
 
-TGAS $\approx$90\% complete to 200 pc with parallax SNR cut (\citealt{Bovy:2017}).
-
-Oh catalog have a high likelihood-ratio cut, but that should be unbiased (right?)
-
-Randomly sample from that, subject to observational constraints.
+To select candidate pairs for follow-up spectroscopic observations, we only
+consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
+distance; with the SNR cut used above and given the limitations of the \tgas\
+catalog itself, \tgas\ is $\approx$90\% complete to $200~\pc$
+(\citealt{Bovy:2017}).
+From this source target list, we randomly observed targets with airmass $\sec z
+< 1.5$.
 

### 2017-05-30 paper/comoving-followup.tex
omg the most boring words ever written

@@
 
-\author{Melissa~Ness}
-\affiliation{todo}
-
 \author{David~N.~Spergel}
@@
              Princeton University, Princeton, NJ 08544, USA}
+
+% \author{Melissa~Ness}
+% \affiliation{todo}
 
@@
 
+Spectra were obtained using the \project{Modspec} spectrograph mounted on the
+$2.4~{\rm m}$ Hiltner telescope at MDM
+observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
+\project{Modspec} was set up in long-slit mode with a 1'' slit, a $600~{\rm
+line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
+detector (``Echelle'').
+Only the central 300 pixel columns were read out from the CCD, and pixel rows
+$>1600$ (at the blue end) were later trimmed because of the poor quantum
+efficiency of the CCD at wavelengths $\lesssim 3600~{\rm \AA}$.
+The resolution and wavelength range were $\approx2~{\rm \AA}~{\rm pixel}^{-1}$
+from $\approx 3600$--$7200~{\rm \AA}$.
+At this resolution, to obtain high signal-to-noise spectra, most exposures were
+between $30$--$120~{\rm s}$ (15th and 85th percentiles).
+To maximize the number of observed pairs, we therefore chose to not take
+comparison lamp spectra at each pointing, which adds an overhead of $\approx
+90~{\rm s}$ per observation.
+To determine rough nightly wavelength solutions, spectra were obtained using
+Hg-Ne and Ne comparison lamps at the beginning and end of each night, and
+atmospheric emission lines were used to correct the wavelength solution on a
+per-object basis.
+
+We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
+2017-03-15, corresponding to 321 unique star pairs and $\approx 120$ calibration
+targets.
+The calibration targets have previously measured radial velocities and chemical
+abundances, which will be used for validation of our reduction pipeline and in
+future work to measure metallicities and stellar parameters for each observed
+source using \project{The Cannon} (\citealt{Ness:TODO}).
+The spectra were reduced and calibrated using a custom, publicly-available
+pipeline written in \python,\footnote{Released along with the code used to
+generate the figures in this document at
+\url{https://github.com/adrn/GaiaPairsFollowup}.} summarized below.
+
+For each source or comparison lamp observation, the two-dimensional image is
+bias corrected, flat-field corrected, and trimmed using standard routines
+implemented in \package{ccdproc} (\citealt{Craig:2015}).
+For comparison lamp observations, the central 100 pixel columns are extracted
+and summed along the spatial axis, weighted by the inverse-variances of each
+pixel.
+Source spectral traces were always placed in this region during observations,
+and the curvature of the wavelength solution over this region is negligible.
+For source observations, the 2D spectra are extracted to 1D using an
+empirically-estimated line spread function (LSF): at each of 16 pixel rows
+between indices 750 and 850 (i.e. around the effective center of the dispersion
+axis), a model for the LSF is fit to the 1D trace.
+The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
+profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
+half-width at half maximum, and the background is modeled using a linear
+polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept at
+the profile centroid, respectively.
+The maximum-likelihood LSF + background model parameters are estimated from each
+pixel row using \texttt{L-BFGS-B} optimization (\citealt{Zhu:1994}), implemented
+in the \package{scipy} package.
+From the 16 fits, the median profile width parameters, $\sigma_G$ and
+$\gamma_L$, are then taken to be the LSF width parameters.
+At each of 1600 rows in each source image, the LSF amplitude and centroid, and
+background slope and intercept at line centroid are then fit using the same
+procedure as above but with the width parameters fixed.
+The LSF amplitudes are stored as the source fluxes, and the background model
+intercepts are stored as the background (sky) fluxes.
+Occasionally, nearby sources fall in the slit; these are masked by hand in the
+2D images and do not bias the LSF extraction fits.
+
+\package{celerite} (\citealt{Foreman-Mackey:2017})
+
 See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example
@@
 \acknowledgements
-David W. Hogg (NYU)
+John Brewer (Yale),
+Jason Curtis (Columbia),
+James Davenport (WWU) for PyDIS,
+David W. Hogg (NYU),
+Melissa Ness (MPIA)
+
+This work is based on observations obtained at the MDM Observatory, operated by
+Dartmouth College, Columbia University, Ohio State University, Ohio University,
+and the University of Michigan.
+% The authors are pleased to acknowledge that the work reported on in this paper
+% was substantially performed at the TIGRESS high performance computer center at
+% Princeton University which is jointly supported by the Princeton Institute for
+% Computational Science and Engineering and the Princeton University Office of
+% Information Technology's Research Computing department.
+
+\software{
+The code used in this project is available from
+\url{https://github.com/adrn/GaiaPairsFollowup} under the MIT open-source
+software license.
+This research utilized the following open-source \python\ packages:
+    \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
+    \package{astroquery} (\citealt{Ginsburg:2016}),
+    \package{ccdproc} (\citealt{Craig:2015}),
+    \package{celerite} (\citealt{Foreman-Mackey:2017}),
+    \package{IPython} (\citealt{Perez:2007}),
+    \package{matplotlib} (\citealt{Hunter:2007}),
+    \package{numpy} (\citealt{Van-der-Walt:2011}),
+    \package{scipy} (\url{https://www.scipy.org/}),
+    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/})
+}
+
+\facility{MDM: Hiltner (Modspec)}
 

### 2017-05-31 paper/comoving-followup.tex
first figure
more text and shit

@@
 \documentclass[modern, letterpaper]{aastex61}
+% \documentclass[twocolumn]{aastex61}
 
@@
 \include{preamble}
+\graphicspath{{figures/}}
 
@@
 \newcommand{\tgas}{\acronym{TGAS}}
+\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
 
@@
 
-\subsection{Observations and Data reduction}\label{sec:reduction}
+\todo{color-magnitude diagram of all stars observed, color our targets different
+from the calibration targets}
+
+\subsection{Observations and data reduction}\label{sec:reduction}
 
@@
 The maximum-likelihood LSF + background model parameters are estimated from each
-pixel row using \texttt{L-BFGS-B} optimization (\citealt{Zhu:1994}), implemented
+pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented
 in the \package{scipy} package.
@@
 
-\package{celerite} (\citealt{Foreman-Mackey:2017})
+We start the wavelength calibration by interactively identifying the 12
+strongest, least blended emission lines in one of the extracted 1D comparison
+lamp spectra by comparing to \todo{NIST??}.
+Once a rough mapping from pixel to wavelength is determined, we then fit for the
+individual line centroids in each comparison lamp spectrum by extracting small,
+8 pixel-wide segments of spectrum around each identified emission line and
+fitting a model for the line profile and background to each individually.
+We use a Gaussian profile for each emission line with parameters: amplitude,
+$A$; centroid, $x_0$; and root-variance, $\sigma_G$.
+We found that nearby blended or low-amplitude lines (i.e. part of the
+``background'') sometimes bias the line centroid determination when using a
+simple background model.
+We therefore model the background flux around each comparison lamp emission line
+using a constant offset + Gaussian process (GP) with a Mat\'ern 3/2 kernel
+parametrized by amplitude parameter $\sigma_{3/2}$ and correlation scale
+$\rho_{3/2}$.
+We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
+likelihood, and maximize the (log-)likelihood again using \lbfgsb\ optimization.
+We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,
+$\sigma_{3/2}$, and $\rho_{3/2}$).
+Once we determine the pixel centroids for all identified emission lines for a
+given night's comparison lamp spectrum, we then fit an interpolating function to the pixel-centroid, wavelength values for each emission line.
+
+Typically, high-order standard polynomials or Chebyshev polynomials are used to
+fit for the wavelength dispersion function.
+We found that using such functions can lead to vastly incorrect wavelength
+values near either end of the pixel grid where the polynomial function is
+un- or weakly-constrained.
+We instead use a combined linear polynomial + GP to model the wavelength
+dispersion.
+We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
+scale parameter to log-space before maximizing the log-likelihood.
+\figurename~\ref{fig:wavelength-GP} shows an example of one such fit with the
+best-fit linear polynomial subtracted from the (pixel-centroid, wavelength)
+pairs (black points) and fit curve (darker line).
+The envelope around the best-fit curve (shaded region) shows the \todo{1$\sigma$
+blah blah GP...}
+When the GP variance is large, the pixel-to-wavelength mapping is uncertain; we
+store this variance and later use it to mask regions of the spectra with
+uncertain wavelength values.
+This wavelength calibration procedure produces a nightly model for the nonlinear
+dispersion of wavelength as a function of pixel value for each extracted 1D
+spectrum.
+\figurename~\ref{fig:sample-spectra} shows a random sample of 10 source spectra
+of comoving star pair members normalized to 1 at $\lambda = 5500~{\rm \AA}$ and
+shifted by index for visualization.
+From the appearances of these 10 alone, it seems there is a mix of metallicities
+and spectral types.
+Note that the spectra are not flux-calibrated, as we only care about measuring
+radial velocities from this sample.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{wavelength_gp.pdf}
+  \end{center}
+  \caption{%
+    \todo{Stuff}
+    \label{fig:wavelength-GP}}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{sample_spectra.pdf}
+  \end{center}
+  \caption{%
+    \todo{Stuff}
+    \label{fig:sample-spectra}}
+\end{figure}
+
+TODO: sky-line corrections to the spectra
+However, because of flexure in the spectrograph, each observation will have an
+additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm \AA}$,
+corresponding to a significant velocity offset at ${\rm H}\alpha$ of
+$\approx 20$--$100~\kms$.
+
+TODO: line measurements, model, gaussian process
 

### 2017-06-01 comoving_rv/db/model.py
new columns for 2mass

@@ -170,6 +170,13 @@ class TGASSource(Base):
     ecl_lon = Column('ecl_lon', SexDegAngleType, nullable=False)
     ecl_lat = Column('ecl_lat', SexDegAngleType, nullable=False)
 
+    J = Column('J', types.REAL)
+    J_err = Column('J_err', types.REAL)
+    H = Column('H', types.REAL)
+    H_err = Column('H_err', types.REAL)
+    Ks = Column('Ks', types.REAL)
+    Ks_err = Column('Ks_err', types.REAL)
+
     @property
     def skycoord(self):
         return coord.SkyCoord(ra=self.ra, dec=self.dec,

### 2017-06-01 scripts/injest_db.py

@@ -1,19 +1,27 @@
+"""
+TODO:
+- Get 2MASS magnitudes and load into tgas_source (see notebook
+  "Fill in 2MASS magnitudes") - see below
+"""
+
 # Standard library
 from os import path
 import glob
 from collections import OrderedDict
 
 # Third-party
+import numpy as np
+
 import astropy.coordinates as coord
 from astropy.time import Time
 import astropy.units as u
 from astropy.io import fits
 from astropy.table import Table
-from astroquery.simbad import Simbad
-Simbad.add_votable_fields('rv_value', 'rvz_qual', 'rvz_bibcode')
-import numpy as np
 from astropy.utils.console import ProgressBar
-# from tqdm import tqdm
+
+from astroquery.simbad import Simbad
+from astroquery.gaia import Gaia
+Simbad.add_votable_fields('rv_value', 'rvz_qual', 'rvz_bibcode')
 
 # Project
 from comoving_rv.db import Session, Base, db_connect
@@ -258,6 +266,29 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
                 for name in tgas.colnames:
                     if name in tgassource_columns:
                         tgas_kw[name] = tgas_row[name]
+
+                # TODO:
+                # query = """
+                # SELECT TOP 10 j_m, j_msigcom, h_m, h_msigcom, ks_m, ks_msigcom
+                # FROM gaiadr1.tmass_original_valid
+                # JOIN gaiadr1.tmass_best_neighbour USING (tmass_oid)
+                # JOIN gaiadr1.tgas_source USING (source_id)
+                # WHERE source_id = {0.source_id}
+                # """
+
+                # job = Gaia.launch_job(query.format(src), dump_to_file=False)
+                # res = job.get_results()
+
+                # if len(res) == 0:
+                #     print("No 2MASS data found for: {0}".format(src.source_id))
+
+                # elif len(res) == 1:
+                #     src.J = res['j_m'][0]
+                #     src.J_err = res['j_msigcom'][0]
+                #     src.H = res['h_m'][0]
+                #     src.H_err = res['h_msigcom'][0]
+                #     src.Ks = res['ks_m'][0]
+                #     src.Ks_err = res['ks_msigcom'][0]
 
                 tgas_source = TGASSource(**tgas_kw)
                 tgas_sources.append(tgas_source)

### 2017-06-01 paper/comoving-followup.tex
add note about cmd plot
fix label

@@
 from the calibration targets}
+\figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
+observed in this observing run (markers) plotted over the density of all \tgas\
+stars within $200~\pc$.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{sample_cmd.pdf}
+  \end{center}
+  \caption{%
+    \todo{Stuff}
+    \label{fig:sample-cmd}}
+\end{figure}
 
@@
 scale parameter to log-space before maximizing the log-likelihood.
-\figurename~\ref{fig:wavelength-GP} shows an example of one such fit with the
-best-fit linear polynomial subtracted from the (pixel-centroid, wavelength)
-pairs (black points) and fit curve (darker line).
+\figurename~\ref{fig:wavelength-GP} shows an example of the residuals from one such fit; black points show the (pixel-centroid, wavelength) pairs with the best-fit model subtracted, and  and fit curve (darker line).
 The envelope around the best-fit curve (shaded region) shows the \todo{1$\sigma$

### 2017-06-01 paper/comoving-followup.tex
words

@@
 scale parameter to log-space before maximizing the log-likelihood.
-\figurename~\ref{fig:wavelength-GP} shows an example of the residuals from one such fit; black points show the (pixel-centroid, wavelength) pairs with the best-fit model subtracted, and  and fit curve (darker line).
-The envelope around the best-fit curve (shaded region) shows the \todo{1$\sigma$
-blah blah GP...}
+\figurename~\ref{fig:wavelength-GP} shows an example of the residuals from one
+such fit; black points show the (pixel-centroid, wavelength) pairs with the
+best-fit model subtracted.
+The envelope around zero (shaded region) shows the \todo{1$\sigma$ blah blah
+GP...}
 When the GP variance is large, the pixel-to-wavelength mapping is uncertain; we

### 2017-06-01 paper/comoving-followup.tex
wavelength calib

@@
 best-fit model subtracted.
-The envelope around zero (shaded region) shows the \todo{1$\sigma$ blah blah
-GP...}
-When the GP variance is large, the pixel-to-wavelength mapping is uncertain; we
-store this variance and later use it to mask regions of the spectra with
-uncertain wavelength values.
+The envelope around zero (shaded region, orange) shows the \todo{1$\sigma$ blah
+blah GP...}
+When the GP variance is large, the pixel-to-wavelength mapping is uncertain.
+We therefore only use the pixel range from 250--1210, corresponding to a
+wavelength range of $\approx 5500$--$7400~{\rm \AA}$; grey shaded region shows
+the excluded pixel/wavelength ranges.
+
 This wavelength calibration procedure produces a nightly model for the nonlinear
@@
 spectrum.
+However, because of flexure in the spectrograph, each observation will have an
+additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm \AA}$,
+corresponding to a significant velocity offset at ${\rm H}\alpha$ of
+$\approx 20$--$100~\kms$.
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10 source spectra
@@
 TODO: sky-line corrections to the spectra
-However, because of flexure in the spectrograph, each observation will have an
-additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm \AA}$,
-corresponding to a significant velocity offset at ${\rm H}\alpha$ of
-$\approx 20$--$100~\kms$.
 

### 2017-06-01 paper/comoving-followup.tex
caption

@@
 
+\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}
+
 Most stars form in clusters with small dispersions in position, velocity, and
@@
   \caption{%
-    \todo{Stuff}
+    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (blue
+    histogram) and all comoving star pair members observed in this follow-up
+    effort (black points); lines (orange) connect comoving pairs of stars.
+    2MASS $J$-band magnitudes are obtained from the Gaia science
+    archive.
+    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
+    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
+    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
+    \citealt{Oh:2017}).
     \label{fig:sample-cmd}}
@@
 and the University of Michigan.
+This work has made use of data from the European Space Agency (ESA)
+mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
+the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
+\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
+for the DPAC has been provided by national institutions, in particular
+the institutions participating in the {\it Gaia} Multilateral Agreement.
 % The authors are pleased to acknowledge that the work reported on in this paper
@@
 }
+This work additionally used the Gaia science archive
+(\url{https://gea.esac.esa.int/archive/}).
 

### 2017-06-02 paper/comoving-followup.tex
omg sky line hacking

@@
 generate the figures in this document at
-\url{https://github.com/adrn/GaiaPairsFollowup}.} summarized below.
+\url{https://github.com/adrn/GaiaPairsFollowup}.} described below.
 
@@
 shifted by index for visualization.
-From the appearances of these 10 alone, it seems there is a mix of metallicities
-and spectral types.
+From the appearances of these 10 spectra, it seems there is a mix of
+metallicities and spectral types amongst the observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring

### 2017-06-03 scripts/solve_velocity.py
remove old imports

@@ -3,7 +3,6 @@
 """
 TODO:
 - n1.0073 Halpha is emission
-- Modify script so it writes to the database instead of a FITS file
 """
 
 # Standard library
@@ -11,12 +10,8 @@ import os
 from os import path
 
 # Third-party
-import astropy.coordinates as coord
-from astropy.time import Time
-from astropy.constants import c
 import astropy.units as u
-from astropy.io import fits
-from astropy.table import Table, Column
+from astropy.table import Table
 import numpy as np
 import matplotlib.pyplot as plt
 plt.style.use('apw-notebook')
@@ -27,11 +22,11 @@ from schwimmbad import choose_pool
 
 # Project
 from comoving_rv.log import logger
-from comoving_rv.db import Session, Base, db_connect
+from comoving_rv.db import Session, db_connect
 from comoving_rv.longslit import extract_region
 from comoving_rv.longslit.fitting import VoigtLineFitter, GaussianLineFitter
-from comoving_rv.db.model import (Run, Observation, TGASSource, SimbadInfo,
-                                  SpectralLineMeasurement, SpectralLineInfo)
+from comoving_rv.db.model import (SpectralLineMeasurement, SpectralLineInfo,
+                                  Run, Observation)
 
 def log_probability(params, gp, flux_data):
     gp.set_parameter_vector(params)

### 2017-06-03 comoving_rv/longslit/rv.py
paper stuff and empty rv placeholder

@@ -1,0 +1,5 @@
+# Standard library
+
+# Third-party
+import numpy as np
+

### 2017-06-03 paper/comoving-followup.tex
important word wrapping

@@
 uncertainties between the astrometric measurements in the \tgas\ catalog and
-naturally handle geometric or projection effects in comparing spherical velocity
-components for star pairs separated by large angles on the sky.
-After applying a conservative cut on the likelihood ratio motivated by computing
-the same likelihood ratio for random pairings of stars, we find 13,058
-comoving star pairs with physical separations between $10^3~\au$ and $10~\pc$.
+naturally handle geometric or projection effects in comparing spherical
+velocity components for star pairs separated by large angles on the sky.
+After applying a conservative cut on the likelihood ratio motivated by
+computing the same likelihood ratio for random pairings of stars, we find
+13,058 comoving star pairs with physical separations between $10^3~\au$ and
+$10~\pc$.
 Of these pairs, there are 10,606 unique sources, implying that a significant
@@
 
-\todo{color-magnitude diagram of all stars observed, color our targets different
-from the calibration targets}
+\todo{color-magnitude diagram of all stars observed, color our targets
+different from the calibration targets}
 \figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
@@
 We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
-2017-03-15, corresponding to 321 unique star pairs and $\approx 120$ calibration
-targets.
+2017-03-15, corresponding to 321 unique star pairs and $\approx 120$
+calibration targets.
 The calibration targets have previously measured radial velocities and chemical
@@
 half-width at half maximum, and the background is modeled using a linear
-polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept at
-the profile centroid, respectively.
-The maximum-likelihood LSF + background model parameters are estimated from each
-pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented
-in the \package{scipy} package.
+polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
+at the profile centroid, respectively.
+The maximum-likelihood LSF + background model parameters are estimated from
+each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
+the \package{scipy} package.
 From the 16 fits, the median profile width parameters, $\sigma_G$ and
@@
 lamp spectra by comparing to \todo{NIST??}.
-Once a rough mapping from pixel to wavelength is determined, we then fit for the
-individual line centroids in each comparison lamp spectrum by extracting small,
-8 pixel-wide segments of spectrum around each identified emission line and
-fitting a model for the line profile and background to each individually.
+Once a rough mapping from pixel to wavelength is determined, we then fit for
+the individual line centroids in each comparison lamp spectrum by extracting
+small, 8 pixel-wide segments of spectrum around each identified emission line
+and fitting a model for the line profile and background to each individually.
 We use a Gaussian profile for each emission line with parameters: amplitude,
@@
 simple background model.
-We therefore model the background flux around each comparison lamp emission line
-using a constant offset + Gaussian process (GP) with a Mat\'ern 3/2 kernel
+We therefore model the background flux around each comparison lamp emission
+line using a constant offset + Gaussian process (GP) with a Mat\'ern 3/2 kernel
 parametrized by amplitude parameter $\sigma_{3/2}$ and correlation scale
@@
 We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
-likelihood, and maximize the (log-)likelihood again using \lbfgsb\ optimization.
+likelihood, and maximize the (log-)likelihood again using \lbfgsb\
+optimization.
 We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,
@@
 Once we determine the pixel centroids for all identified emission lines for a
-given night's comparison lamp spectrum, we then fit an interpolating function to the pixel-centroid, wavelength values for each emission line.
+given night's comparison lamp spectrum, we then fit an interpolating function
+to the pixel-centroid, wavelength values for each emission line.
 
@@
 
-This wavelength calibration procedure produces a nightly model for the nonlinear
-dispersion of wavelength as a function of pixel value for each extracted 1D
-spectrum.
+This wavelength calibration procedure produces a nightly model for the
+nonlinear dispersion of wavelength as a function of pixel value for each
+extracted 1D spectrum.
 However, because of flexure in the spectrograph, each observation will have an
-additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm \AA}$,
-corresponding to a significant velocity offset at ${\rm H}\alpha$ of
+additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm
+\AA}$, corresponding to a significant velocity offset at ${\rm H}\alpha$ of
 $\approx 20$--$100~\kms$.
@@
 the institutions participating in the {\it Gaia} Multilateral Agreement.
-% The authors are pleased to acknowledge that the work reported on in this paper
-% was substantially performed at the TIGRESS high performance computer center at
-% Princeton University which is jointly supported by the Princeton Institute for
-% Computational Science and Engineering and the Princeton University Office of
-% Information Technology's Research Computing department.
+% The authors are pleased to acknowledge that the work reported on in this
+% paper was substantially performed at the TIGRESS high performance computer
+% center at Princeton University which is jointly supported by the Princeton
+% Institute for Computational Science and Engineering and the Princeton
+% University Office of Information Technology's Research Computing department.
 

### 2017-06-03 paper/comoving-followup.tex
add GP kernel refs

@@
   \caption{%
-    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (blue
-    histogram) and all comoving star pair members observed in this follow-up
-    effort (black points); lines (orange) connect comoving pairs of stars.
+    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
+    blue histogram) and all comoving star pair members observed in this
+    follow-up effort (black points); colored lines connect candidate comoving
+    star pairs.
     2MASS $J$-band magnitudes are obtained from the Gaia science
@@
 $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
-We found that nearby blended or low-amplitude lines (i.e. part of the
+We find that nearby blended or low-amplitude lines (i.e. part of the
 ``background'') sometimes bias the line centroid determination when using a
@@
 We therefore model the background flux around each comparison lamp emission
-line using a constant offset + Gaussian process (GP) with a Mat\'ern 3/2 kernel
-parametrized by amplitude parameter $\sigma_{3/2}$ and correlation scale
-$\rho_{3/2}$.
+line using a constant offset + Gaussian process (GP).
+For the GP model, we use a Mat\'ern 3/2 kernel function
+(\citealt{Matern:1986,Rasmussen:2005}) parametrized by amplitude parameter
+$\sigma_{3/2}$ and correlation scale $\rho_{3/2}$.
 We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
-likelihood, and maximize the (log-)likelihood again using \lbfgsb\
-optimization.
+likelihood for this model, and maximize the (log-)likelihood again using
+\lbfgsb\ optimization.
 We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,

### 2017-06-03 paper/comoving-followup.tex
smooth the spectra

@@
 However, because of flexure in the spectrograph, each observation will have an
-additional, unknown constant offset of, typically, $\approx 0.5$--$2~{\rm
-\AA}$, corresponding to a significant velocity offset at ${\rm H}\alpha$ of
-$\approx 20$--$100~\kms$.
+additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
+a significant velocity offset at ${\rm H}\alpha$ of $\approx 20$--$100~\kms$.
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10 source spectra

### 2017-06-03 paper/comoving-followup.tex
captions

@@
   \caption{%
-    \todo{Stuff}
+    A visualization of the pixel-to-wavelength model determined from a single
+    wavelength comparison lamp spectrum.
+    Black markers show the fit pixel centroids for known Hg or Ne emission
+    lines and the residual wavelength value between the linear + GP model and
+    the true emission line centroid.
+    The non-shaded region shows the section of spectrum where the comparison
+    lamp spectrum has sufficient line density to determine the wavelength
+    solution; shaded regions are masked from the resulting spectra.
+    Organge envelope shows the root-variance of the Gaussian process component
+    of the model, which can be interpreted as inherent uncertainties on the
+    wavelength calibration as a result of the model flexibility.
     \label{fig:wavelength-GP}}
@@
   \caption{%
-    \todo{Stuff}
+    A section of 10 randomly selected target spectra from this work.
+    Spectral fluxes are normalized to 1 at $\lambda = 5500~{\rm \AA}$ and the
+    offsets are chosen to separate the spectra for visualization.
+    The spectra are smoothed with a Gaussian filter with root-variance $\sigma
+    = 2~{\rm \AA}$.
     \label{fig:sample-spectra}}

### 2017-06-03 paper/comoving-followup.tex
starting to write words about sky line fixes

@@
 additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at ${\rm H}\alpha$ of $\approx 20$--$100~\kms$.
-\figurename~\ref{fig:sample-spectra} shows a random sample of 10 source spectra
-of comoving star pair members normalized to 1 at $\lambda = 5500~{\rm \AA}$ and
-shifted by index for visualization.
-From the appearances of these 10 spectra, it seems there is a mix of
-metallicities and spectral types amongst the observed targets.
+a significant velocity offset at ${\rm H}\alpha$ of $\approx 25$--$100~\kms$.
+\figurename~\ref{fig:sample-spectra} shows a random sample of 10
+wavelength-calibrated source spectra of comoving star pair members normalized
+to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by index for visualization.
+Suggested by the features in these 10 spectra, there is a mix of metallicities
+and spectral types amongst the observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring
@@
 \end{figure}
+
+As mentioned previously, comparison lamp spectra were not obtained at each
+telescope pointing.
+Depending primarily on the hour angle of the observation (because of the orientation of the spectrograph and CCD), there
 

### 2017-06-03 paper/comoving-followup.tex
sky line fitting

@@
   \begin{center}
-    \includegraphics[width=0.9\linewidth]{sample_cmd.pdf}
+    \includegraphics[width=\linewidth]{sample_cmd.pdf}
   \end{center}
@@
 and the curvature of the wavelength solution over this region is negligible.
-For source observations, the 2D spectra are extracted to 1D using an
-empirically-estimated line spread function (LSF): at each of 16 pixel rows
-between indices 750 and 850 (i.e. around the effective center of the dispersion
-axis), a model for the LSF is fit to the 1D trace.
+For source observations, the 2D spectra are extracted to 1D using an empirical
+estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
+rows between indices 750 and 850 (i.e. around the effective center of the
+dispersion axis), a model for the LSF is fit to the 1D trace.
 The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
@@
   \begin{center}
-    \includegraphics[width=0.9\linewidth]{wavelength_gp.pdf}
+    \includegraphics[width=\linewidth]{wavelength_gp.pdf}
   \end{center}
@@
 
-As mentioned previously, comparison lamp spectra were not obtained at each
-telescope pointing.
-Depending primarily on the hour angle of the observation (because of the orientation of the spectrograph and CCD), there
-
-TODO: sky-line corrections to the spectra
+As mentioned, comparison lamp spectra were not obtained at each telescope
+pointing.
+The absolute wavelength solution at each pointing can therefore be shifted
+(in pixels) relative to the nightly comparison lamp spectrum because of flexure
+in the instrument.
+We have found that this offset depends primarily on the hour angle of the
+observation because of the physical orientation of the spectrograph and CCD as
+mounted on the telescope.
+To correct each sourcce spectrum for this offset, we measure the centroids of
+the two prominent [OI] (sky) emission lines near $5577~{\rm \AA}$ and $6300~{\rm
+\AA}$ in the corresponding background spectrum derived from the spectral
+extraction (described above).
+We use a pixel-convolved Voigt profile as the emission line model (same
+parameters as above) and a Gaussian process with a Mat\'ern 3/2 kernel function
+for the background spectrum (same parameters as above).
+We again use \lbfgsb\ optimization and convert any scale parameter to log-space
+before maximizing the log-likelihood of this model to derive the line centroid
+in wavelength units.
+We compute wavelength offsets from true (air) values of the line centroids,
+$5577.XXX~{\rm \AA}$ and $6300.XXX~{\rm \AA}$ respectively, then convert these
+wavelength offsets to pixel offsets.
+We take the mean of the pixel offsets
+When one or both of the two sky lines is not present or the centroid fit fails,
+the spectrum \todo{flagged in some way, can have large absolute velocity
+systematic error.}
 
 TODO: line measurements, model, gaussian process
+  - per-spectrum precisions ~5-10 km/s
 
-See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example
+TODO: Comparing repeat visits on subsequent nights, absolute errors ~20 km/s
 
-TODO:
-- Describe pipeline...
-
-- Comparing repeat visits on subsequent nights, systematic ~20 km/s
-    - see notebook "Compare RVs with previous measurements"
+% See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example
 

### 2017-06-03 paper/comoving-followup.tex
TODOs for the future

@@
 
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{placeholder.pdf}
+%   \end{center}
+%   \caption{%
+%     \todo{Some summary plots about the MCMC fits...}
+%     \label{fig:Halpha-mcmc}}
+% \end{figure}
+
 TODO: Comparing repeat visits on subsequent nights, absolute errors ~20 km/s
+- How many of our targets already have measurements in Simbad?
+
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{placeholder.pdf}
+%   \end{center}
+%   \caption{%
+%     \todo{Plot showing my RV's vs. previously measured.}
+%     \label{fig:compare-previous}}
+% \end{figure}
 
@@
 
-TODO:
-- Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).
+\subsection{Catalog ...}\label{sec:}
 
-\subsection{Catalog ...}\label{sec:}
+TODO: 10 overlap with Shaya and Olling
+
+TODO: Do I want to do isochrone fitting for all of them?
 
@@
 
+TODO: Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).
+
+TODO: separation distribution - combine with Shaya and Olling sample!
+
+TODO: distribution of orbital poles in Galactic coordinates
+
 \subsection{Galactic orbits of the co-moving pairs}\label{sec:}
 
-- Orbits of true pairs in the Galaxy, any halo or thick-disk?
+TODO: Orbits of true pairs in the Galaxy, any halo or thick-disk?
+
+\section{Discussion}
+
+Abundances!
 

### 2017-06-04 paper/comoving-followup.tex
lots of words and spell check

@@
 \newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
+\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
 
@@
 Want to identify genuinely dynamically associated pairs for future
-high-resolution followup
+high-resolution follow-up
 % Methods
@@
 $\gamma_L$, are then taken to be the LSF width parameters.
-At each of 1600 rows in each source image, the LSF amplitude and centroid, and
-background slope and intercept at line centroid are then fit using the same
-procedure as above but with the width parameters fixed.
+At each of 1600 rows in each source spectrum image, the LSF amplitude and
+centroid, and background slope and intercept at line centroid are then fit using
+the same procedure as above but with the width parameters fixed.
 The LSF amplitudes are stored as the source fluxes, and the background model
@@
 and fitting a model for the line profile and background to each individually.
-We use a Gaussian profile for each emission line with parameters: amplitude,
-$A$; centroid, $x_0$; and root-variance, $\sigma_G$.
+We use a pixel-convolved Gaussian profile for each emission line with
+parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
 We find that nearby blended or low-amplitude lines (i.e. part of the
@@
 We therefore only use the pixel range from 250--1210, corresponding to a
-wavelength range of $\approx 5500$--$7400~{\rm \AA}$; grey shaded region shows
+wavelength range of $\approx 5500$--$7400~{\rm \AA}$; gray shaded region shows
 the excluded pixel/wavelength ranges.
@@
 additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at ${\rm H}\alpha$ of $\approx 25$--$100~\kms$.
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10
@@
     solution; shaded regions are masked from the resulting spectra.
-    Organge envelope shows the root-variance of the Gaussian process component
+    Orange envelope shows the root-variance of the Gaussian process component
     of the model, which can be interpreted as inherent uncertainties on the
@@
 We have found that this offset depends primarily on the hour angle of the
-observation because of the physical orientation of the spectrograph and CCD as
-mounted on the telescope.
-To correct each sourcce spectrum for this offset, we measure the centroids of
-the two prominent [OI] (sky) emission lines near $5577~{\rm \AA}$ and $6300~{\rm
+observation, likely because of the physical orientation of the spectrograph and
+CCD as mounted on the telescope.
+To correct each source spectrum for this offset, we measure the centroids of
+the two prominent [OI] sky emission lines near $5577~{\rm \AA}$ and $6300~{\rm
 \AA}$ in the corresponding background spectrum derived from the spectral
@@
 We compute wavelength offsets from true (air) values of the line centroids,
-$5577.XXX~{\rm \AA}$ and $6300.XXX~{\rm \AA}$ respectively, then convert these
+$5577.3387~{\rm \AA}$ and $6300.304~{\rm \AA}$ respectively,\footnote{\url{http:
+//physics.nist.gov/PhysRefData/ASD/lines_form.html}} then convert these
 wavelength offsets to pixel offsets.
-We take the mean of the pixel offsets
+We take the mean of the pixel offsets and apply the derived shift to the source
+spectrum before re-computing the wavelength grid using the nightly wavelength
+solution \todo{is this sensible?}.
 When one or both of the two sky lines is not present or the centroid fit fails,
@@
 
-TODO: line measurements, model, gaussian process
-  - per-spectrum precisions ~5-10 km/s
+To derive radial velocities for each source, we then measure the (wavelength)
+line centroid of \Ha\ in each corrected spectrum.
+We model the absorption line using a pixel-convolved Voigt profile and again use
+a Gaussian process for the background model (with the same parameterizations as
+described above).
+We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
+to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
+probability distribution over the 6 parameters: $\ln |A|$, $x_0$, $\ln\sigma_G$,
+$\ln\gamma_L$, $\ln\sigma_{3/2}$, and $\ln\rho_{3/2}$.
+The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
+an absorption feature) and is computed from the corresponding parameter as $A =
+-\exp(\ln |A|)$.
+We use uniform priors on all parameters except $\ln\rho_{3/2}$, for which we use
+a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
+\figurename~\ref{fig:Halpha-mcmc} shows \todo{mcmc shite}.
+With the resulting posterior samples, we compute the median and median absolute
+deviation (MAD) of the marginal distribution over line centroid, $x_0$.
+We use the MAD to estimate the standard deviation ($\sigma \approx 1.5~{\rm
+MAD}$, assuming the samples are Gaussian distributed) of the samples and find
+that the typical velocity \emph{precision} from measuring the centroid of \Ha\ is between $2$--$10~\kms$ \todo{double check numbers}.
+
+\begin{table}[ht]
+  \begin{center}
+    \begin{tabular}{ c | c }
+      \toprule
+        Parameter & Prior \\\toprule
+        $\ln A$ & $\mathcal{U}(2, 16)$\\
+        $x_0$ & $\mathcal{U}(\lambda_{\Ha}-16, \lambda_{\Ha}+16)$ [${\rm \AA}$]\\
+        $\ln\sigma_G$ & $\mathcal{U}(-8, 2)$ [${\rm \AA}$]\\
+        $\ln\gamma_L$ & $\mathcal{U}(-8, 2)$ [${\rm \AA}$]\\
+        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
+        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$ [${\rm \AA}$]\\
+      \bottomrule
+    \end{tabular}
+    \caption{
+      Priors used for \Ha\ line profile model parameters.
+      $\mathcal{U}(a, b)$ indicates a uniform prior between $a$ and $b$ and
+      $\mathcal{N}(\mu, \sigma^2)$ indicates a normal distribution with mean
+      $\mu$ and variance $\sigma^2$.
+      $\lambda_{\Ha}$ is the true \Ha\ line centroid in angstroms.
+      \label{tbl:prior-bounds}
+    }
+  \end{center}
+\end{table}
 
@@
 % \end{figure}
+
+Several of our targets and all of our calibration targets have previously
+measured radial velocities.
+We can therefore estimate our absolute velocity accuracy by comparing our
+derived radial velocities with previous measurements.
+\todo{Explain how we get previous measurements - Simbad or explicit catalogs?}
+\figurename~\ref{fig:compare-previous} shows the distribution of differences
+between our measurements and literature values \todo{uncertainties??}.
 

### 2017-06-04 scripts/fit_line_centroids.py
change prior and sampling length

@@ -38,7 +38,7 @@ def log_probability(params, gp, flux_data):
     var = 1.
     lp += -0.5*(params[1]-1)**2/var - 0.5*np.log(2*np.pi*var)
 
-    if params[4] < -10. or params[5] < -10.:
+    if (params[4] < -8. or params[4] > 2 or params[5] < -8. or params[5] > 2):
         return -np.inf
 
     ll = gp.log_likelihood(flux_data)
@@ -151,11 +151,11 @@ def main(db_path, run_name, filename=None, overwrite=False, pool=None):
         logger.debug("Running 2nd burn-in...")
         sampler.reset()
         p0 = p0[lp.argmax()] + 1e-3 * np.random.randn(nwalkers, ndim)
-        p0, lp, _ = sampler.run_mcmc(p0, 256)
+        p0, lp, _ = sampler.run_mcmc(p0, 512)
 
         logger.debug("Running production...")
         sampler.reset()
-        pos, lp, _ = sampler.run_mcmc(p0, 512)
+        pos, lp, _ = sampler.run_mcmc(p0, 1024)
 
         fit_kw = dict()
         for i,par_name in enumerate(lf.gp.get_parameter_names()):

### 2017-06-05 paper/comoving-followup.tex
todo
fix software section

@@
 a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
+\todo{Explain MCMC, emcee, steps, etc.}
 \figurename~\ref{fig:Halpha-mcmc} shows \todo{mcmc shite}.
@@
     \package{sqlalchemy} (\url{https://www.sqlalchemy.org/})
-}
 This work additionally used the Gaia science archive
 (\url{https://gea.esac.esa.int/archive/}).
+}
 

### 2017-06-05 paper/comoving-followup.tex
fix table

@@
 deviation (MAD) of the marginal distribution over line centroid, $x_0$.
-We use the MAD to estimate the standard deviation ($\sigma \approx 1.5~{\rm
+We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
 MAD}$, assuming the samples are Gaussian distributed) of the samples and find
@@
         $\ln A$ & $\mathcal{U}(2, 16)$\\
-        $x_0$ & $\mathcal{U}(\lambda_{\Ha}-16, \lambda_{\Ha}+16)$ [${\rm \AA}$]\\
-        $\ln\sigma_G$ & $\mathcal{U}(-8, 2)$ [${\rm \AA}$]\\
-        $\ln\gamma_L$ & $\mathcal{U}(-8, 2)$ [${\rm \AA}$]\\
+        $x_0$ & $\mathcal{U}(6547, 6579)$\\
+        $\ln\sigma_G$ & $\mathcal{U}(-8, 2)$\\
+        $\ln\gamma_L$ & $\mathcal{U}(-8, 2)$\\
         $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
-        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$ [${\rm \AA}$]\\
+        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$\\
       \bottomrule
@@
       $\mu$ and variance $\sigma^2$.
-      $\lambda_{\Ha}$ is the true \Ha\ line centroid in angstroms.
+      Wavelength units are all angstroms.
       \label{tbl:prior-bounds}

### 2017-06-05 comoving_rv/plot.py
add colors to module

@@ -1,0 +1,5 @@
+colors = dict()
+colors['line_marker'] = '#3182bd'
+colors['gp_model'] = '#ff7f0e'
+colors['not_black'] = '#333333'
+colors['fit'] = '#2ca25f'

### 2017-06-05 paper/comoving-followup.tex
emcee reference

@@
 line centroid of \Ha\ in each corrected spectrum.
-We model the absorption line using a pixel-convolved Voigt profile and again use
-a Gaussian process for the background model (with the same parameterizations as
-described above).
+We model the absorption line using a pixel-convolved Voigt profile and use a
+linear model + a Gaussian process for the background model.
 We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
 to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
-probability distribution over the 6 parameters: $\ln |A|$, $x_0$, $\ln\sigma_G$,
-$\ln\gamma_L$, $\ln\sigma_{3/2}$, and $\ln\rho_{3/2}$.
+probability distribution (pdf) over the 8 parameters: $\ln |A|$, $x_0$,
+$\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
+$\ln\rho_{3/2}$ where the linear background model is defined as $f(\lambda) =
+\alpha_1 + \alpha_2\,\lambda$.
 The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
@@
 a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
-\todo{Explain MCMC, emcee, steps, etc.}
-\figurename~\ref{fig:Halpha-mcmc} shows \todo{mcmc shite}.
+We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
+generate samples from the posterior pdf.
+We initially run the sampler for 128 steps from the maximum likelihood parameter
+values, then resample walker positions in a small ball around the median final
+walker positions and run again for 512 steps; together, these two steps
+constitute our burn-in phase.
+From the final walker positions after burn-in, we run again for 1024 steps and
+store these as our primary posterior samples.
+\figurename~\ref{fig:Halpha-mcmc} shows
+
 With the resulting posterior samples, we compute the median and median absolute
@@
         $x_0$ & $\mathcal{U}(6547, 6579)$\\
-        $\ln\sigma_G$ & $\mathcal{U}(-8, 2)$\\
-        $\ln\gamma_L$ & $\mathcal{U}(-8, 2)$\\
+        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$\\
+        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$\\
+        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$\\
+        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$\\
         $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
@@
 
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{placeholder.pdf}
-%   \end{center}
-%   \caption{%
-%     \todo{Some summary plots about the MCMC fits...}
-%     \label{fig:Halpha-mcmc}}
-% \end{figure}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{mcmc_trace.pdf}
+  \end{center}
+  \caption{%
+    \todo{Some summary plots about the MCMC fits...}
+    \label{fig:Halpha-mcmc}}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{example_fit.pdf}
+  \end{center}
+  \caption{%
+    \todo{More MCMC bullshit.}
+    \label{fig:compare-previous}}
+\end{figure}
 
@@
 %   \begin{center}
-%     \includegraphics[width=\linewidth]{placeholder.pdf}
+%     \includegraphics[width=\linewidth]{example_fit.pdf}
 %   \end{center}
@@
     \package{celerite} (\citealt{Foreman-Mackey:2017}),
+    \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),
     \package{IPython} (\citealt{Perez:2007}),

### 2017-06-05 paper/comoving-followup.tex
words
spacing around cmd

@@
 < 1.5$.
-
-\todo{color-magnitude diagram of all stars observed, color our targets
-different from the calibration targets}
 \figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_trace.pdf}
+    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
   \end{center}
@@
     \todo{Some summary plots about the MCMC fits...}
+    Small uncertainty in centroid = 2 km/s velocity precision here.
     \label{fig:Halpha-mcmc}}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{example_fit.pdf}
+    \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
   \end{center}
   \caption{%
-    \todo{More MCMC bullshit.}
+    Top shows background linear model + Voigt line model.
+
     \label{fig:compare-previous}}

### 2017-06-06 paper/comoving-followup.tex
fix period
omg methods section done

@@
 to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
-probability distribution (pdf) over the 8 parameters: $\ln |A|$, $x_0$,
+probability distribution (PDF) over the 8 parameters: $\ln |A|$, $x_0$,
 $\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
@@
 We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
-generate samples from the posterior pdf.
+generate samples from the posterior PDF.
 We initially run the sampler for 128 steps from the maximum likelihood parameter
@@
 From the final walker positions after burn-in, we run again for 1024 steps and
-store these as our primary posterior samples.
-\figurename~\ref{fig:Halpha-mcmc} shows
+store these as our posterior samples.
+\figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
+(\citealt{Foreman-Mackey:2017a}) of all projections of these posterior PDF
+samples for a randomly-chosen source (HD\todo{XXXX}, in this case).
+The marginal posterior distribution over $x_0$ can be converted to a
+distribution over radial velocities; in this case, the estimated (Gaussian
+standard deviation) precision is $\approx 2.5~\kms$.
+\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
+around \Ha\ used in the line modeling (black line) with uncertainties (grey
+bars).
+Top panel shows the linear polynomial + absorption line model computed for the
+median sample (dark green line); shaded region shows the 15th to 85th
+percentiles computed from all posterior samples.
+Bottom panel shows the residuals of the data from the median linear polynomial +
+absorption line model (black line) along with the Gaussian process computed from
+the median parameters (orange curve) used to model other structure in the
+vicinity of \Ha (which can be interpreted as correlated noise).
+These figures are produced for all source spectrum fits and visually inspected
+to ensure that the MCMC sampling appears to have converged for all cases.
 
-With the resulting posterior samples, we compute the median and median absolute
-deviation (MAD) of the marginal distribution over line centroid, $x_0$.
+For each source spectrum, with the resulting posterior samples, we compute the
+median and median absolute deviation (MAD) of the marginal distribution over
+line centroid, $x_0$.
 We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
 MAD}$, assuming the samples are Gaussian distributed) of the samples and find
-that the typical velocity \emph{precision} from measuring the centroid of \Ha\ is between $2$--$10~\kms$ \todo{double check numbers}.
+that the typical velocity \emph{precision} from measuring the centroid of \Ha\
+is between $2$--$8~\kms$ \todo{double check numbers}.
 
@@
       Wavelength units are all angstroms.
+      \todo{what are the flux units?}
       \label{tbl:prior-bounds}
@@
 
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
-  \end{center}
-  \caption{%
-    \todo{Some summary plots about the MCMC fits...}
-    Small uncertainty in centroid = 2 km/s velocity precision here.
-    \label{fig:Halpha-mcmc}}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
-  \end{center}
-  \caption{%
-    Top shows background linear model + Voigt line model.
-
-    \label{fig:compare-previous}}
-\end{figure}
-
-Several of our targets and all of our calibration targets have previously
-measured radial velocities.
-We can therefore estimate our absolute velocity accuracy by comparing our
-derived radial velocities with previous measurements.
-\todo{Explain how we get previous measurements - Simbad or explicit catalogs?}
-\figurename~\ref{fig:compare-previous} shows the distribution of differences
-between our measurements and literature values \todo{uncertainties??}.
-
-TODO: Comparing repeat visits on subsequent nights, absolute errors ~20 km/s
-- How many of our targets already have measurements in Simbad?
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{mcmc_corner.pdf}
+%   \end{center}
+%   \caption{%
+%     \todo{Some summary plots about the MCMC fits...}
+%     Small uncertainty in centroid = 2 km/s velocity precision here.
+%     \label{fig:Halpha-mcmc-corner}}
+% \end{figure}
 
@@
 %   \begin{center}
-%     \includegraphics[width=\linewidth]{example_fit.pdf}
+%     \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
+%   \end{center}
+%   \caption{%
+%     Top shows background linear model + Voigt line model.
+%     \label{fig:Halpha-mcmc-fit}}
+% \end{figure}
+
+\subsection{Validation}
+
+Several of our targets and all of our calibration targets have previously
+measured radial velocities (RVs).
+We can therefore estimate the absolute velocity uncertainties by comparing our
+derived radial velocities with high-quality prior measurements.
+Following previous work (\citealt{Shaya:2011}), we search for previous RV
+measurements (in decreasing order of priority) with:
+\begin{enumerate}
+  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
+  (\citealt{Nordstrom:2004}).
+  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
+  Data, the Second Version (\citealt{Kharchenko:2007}).
+  \item The General Catalogue of Mean Radial Velocities
+  (\citealt{Barbier-Brossat:2000}).
+  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
+  to retrieve records from the database, which do not contain uncertainties but
+  do contain a quality flag. We only retain RVs that have a quality flag set to
+  `A' and adopt a $5~\kms$ uncertainty for all such values.
+  \item
+\end{enumerate}
+We find that \todo{XXX} of our target stars have reliable prior measurements,
+along with the \todo{YYY} calibration targets.
+\figurename~\ref{fig:compare-previous} shows the distribution of velocity
+differences between our measurements and literature values.
+From this distribution, we compute the MAD and estimate the standard deviation
+to be $\approx 20~\kms$, which we adopt as our global systematic error for all
+absolute radial velocities.
+We note, however, that for computing \emph{relative} velocities for pairs of
+nearby stars, it is the RV precision that matters most.
+
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{compare_previous.pdf}
 %   \end{center}
@@
 % \end{figure}
-
-% See: http://iopscience.iop.org/article/10.1088/0004-6256/146/6/161/pdf for an example
 
@@
 the institutions participating in the {\it Gaia} Multilateral Agreement.
+This research has made use of the SIMBAD database, operated at CDS, Strasbourg,
+France.
 % The authors are pleased to acknowledge that the work reported on in this
@@
     \package{scipy} (\url{https://www.scipy.org/}),
-    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/})
+    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}).
 This work additionally used the Gaia science archive
-(\url{https://gea.esac.esa.int/archive/}).
+(\url{https://gea.esac.esa.int/archive/}), and the SIMBAD database
+(\citealt{Wenger:2000}).
 }

### 2017-06-06 paper/comoving-followup.tex
remove extra item
tODO about exoplanets

@@
   `A' and adopt a $5~\kms$ uncertainty for all such values.
-  \item
 \end{enumerate}
@@
 
+\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
+
 \subsection{Catalog ...}\label{sec:}
 
-TODO: 10 overlap with Shaya and Olling
+\todo{10 overlap with Shaya and Olling}
 
-TODO: Do I want to do isochrone fitting for all of them?
+\todo{Do I want to do isochrone fitting for all of them?}
 
@@
 
-TODO: Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).
+\todo{Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).}
 
-TODO: separation distribution - combine with Shaya and Olling sample!
+\todo{separation distribution - combine with Shaya and Olling sample!}
 
-TODO: distribution of orbital poles in Galactic coordinates
+\todo{distribution of orbital poles in Galactic coordinates}
 
@@
 
-TODO: Orbits of true pairs in the Galaxy, any halo or thick-disk?
+\todo{Orbits of true pairs in the Galaxy, any halo or thick-disk?}
 

### 2017-06-06 paper/comoving-followup.tex
no most
fix corner.py citation

@@
 \figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
-(\citealt{Foreman-Mackey:2017a}) of all projections of these posterior PDF
+(\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
 samples for a randomly-chosen source (HD\todo{XXXX}, in this case).
@@
 We note, however, that for computing \emph{relative} velocities for pairs of
-nearby stars, it is the RV precision that matters most.
+nearby stars, it is the RV precision that matters.
 
@@
     \package{celerite} (\citealt{Foreman-Mackey:2017}),
+    \package{corner} (\citealt{Foreman-Mackey:2016}),
     \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),

### 2017-06-07 comoving_rv/db/model.py
db model updated

@@ -82,14 +82,19 @@ class Observation(Base):
     simbad_info = relationship('SimbadInfo', cascade='all,delete-orphan',
                                backref='observation', single_parent=True)
 
+    prior_rv_id = Column('prior_rv_id', types.Integer,
+                         ForeignKey('prior_rv.id'))
+    prior_rv = relationship('PriorRV', cascade='all,delete-orphan',
+                            backref='observation', single_parent=True)
+
     tgas_source_id = Column('tgas_source_id', types.Integer,
                             ForeignKey('tgas_source.id'))
     tgas_source = relationship('TGASSource', cascade='all,delete-orphan',
                                backref='observation', single_parent=True)
 
     def __repr__(self):
-        return ("<Observation {0.object}, {0.filename_raw}, {0.run.name}>"
-                .format(self))
+        return ("<Observation {0.object} [{0.simbad_info.preferred_name}], "
+                "{0.filename_raw}, {0.run.name}>".format(self))
 
     def path_raw(self, base_path):
         p = path.join(base_path, self.run.name, 'n'+str(self.night))
@@ -115,10 +120,6 @@ class SimbadInfo(Base):
     tyc_id = Column('tyc_id', types.String)
     twomass_id = Column('twomass_id', types.String)
 
-    rv = Column('rv', VelocityType)
-    rv_qual = Column('rv_qual', types.String)
-    rv_bibcode = Column('rv_bibcode', types.String)
-
     def __repr__(self):
         names = []
         pres = ['HD', 'HIP', 'TYC', '2MASS']
@@ -126,8 +127,30 @@ class SimbadInfo(Base):
             if getattr(self, id_) is not None:
                 names.append('{0} {1}'.format(pre, getattr(self, id_)))
 
-        return ("<SimbadInfo {0}>"
-                .format(', '.join(names)))
+        return ("<SimbadInfo {0} [{1}]>"
+                .format(self.preferred_name, ', '.join(names)))
+
+    @property
+    def preferred_name(self):
+        pres = ['HD', 'HIP', 'TYC', '2MASS']
+        attrs = ['hd_id', 'hip_id', 'tyc_id', 'twomass_id']
+
+        for pre,id_ in zip(pres, attrs):
+            the_id = getattr(self, id_)
+            if the_id is not None:
+                return '{0} {1}'.format(pre, the_id)
+
+        return '{0} {1}'.format('SMOH', self.observation.object_name)
+
[…]

### 2017-06-07 scripts/injest_db.py
do gaia query when injesting

@@ -22,6 +22,13 @@ from astropy.utils.console import ProgressBar
 from astroquery.simbad import Simbad
 from astroquery.gaia import Gaia
 Simbad.add_votable_fields('rv_value', 'rvz_qual', 'rvz_bibcode')
+gaia_query = """
+SELECT TOP 10 j_m, j_msigcom, h_m, h_msigcom, ks_m, ks_msigcom
+FROM gaiadr1.tmass_original_valid
+JOIN gaiadr1.tmass_best_neighbour USING (tmass_oid)
+JOIN gaiadr1.tgas_source USING (source_id)
+WHERE source_id = {0}
+"""
 
 # Project
 from comoving_rv.db import Session, Base, db_connect
@@ -267,28 +274,21 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
                     if name in tgassource_columns:
                         tgas_kw[name] = tgas_row[name]
 
-                # TODO:
-                # query = """
-                # SELECT TOP 10 j_m, j_msigcom, h_m, h_msigcom, ks_m, ks_msigcom
-                # FROM gaiadr1.tmass_original_valid
-                # JOIN gaiadr1.tmass_best_neighbour USING (tmass_oid)
-                # JOIN gaiadr1.tgas_source USING (source_id)
-                # WHERE source_id = {0.source_id}
-                # """
+                job = Gaia.launch_job(gaia_query.format(tgas_kw['source_id']),
+                                      dump_to_file=False)
+                res = job.get_results()
 
-                # job = Gaia.launch_job(query.format(src), dump_to_file=False)
-                # res = job.get_results()
+                if len(res) == 0:
+                    logger.warning("No 2MASS data found for: {0}"
+                                   .format(tgas_kw['source_id']))
 
-                # if len(res) == 0:
-                #     print("No 2MASS data found for: {0}".format(src.source_id))
-
-                # elif len(res) == 1:
-                #     src.J = res['j_m'][0]
-                #     src.J_err = res['j_msigcom'][0]
-                #     src.H = res['h_m'][0]
-                #     src.H_err = res['h_msigcom'][0]
-                #     src.Ks = res['ks_m'][0]
-                #     src.Ks_err = res['ks_msigcom'][0]
+                elif len(res) == 1:
+                    tgas_kw['J'] = res['j_m'][0]
+                    tgas_kw['J_err'] = res['j_msigcom'][0]
[…]

### 2017-06-07 comoving_rv/db/literature_helpers.py
retrieve literature measurements using vizier or simbad

@@ -1,0 +1,238 @@
+"""
+Functions to help retrieving velocities from the literature via
+Vizier or Simbad.
+"""
+
+from astroquery.simbad import Simbad
+Simbad.add_votable_fields('rv_value', 'rvz_qual', 'rvz_bibcode')
+from astroquery.vizier import Vizier
+import numpy as np
+
+from ..log import logger
+
+__all__ = ['get_GCRV_rv', 'get_GCS_rv', 'get_CRVAD_rv', 'get_simbad_rv',
+           'get_best_rv']
+
+def get_GCS_rv(obs):
+    """
+    Try to retrieve a velocity from the Geneva-Copenhagen Survey.
+
+    Parameters
+    ----------
+    obs : `comoving_rv.db.model.Observation`
+        An observation instance.
+
+    Returns
+    -------
+    rv : float
+        Radial velocity (in km/s).
+    rv_err : float or None
+        Radial velocity error (in km/s).
+    rv_qual : str or None
+        Quality flag for the measurement.
+    bibcode : str
+        Bibliographic code for the source of the measurement.
+
+    or
+
+    ``None``
+
+    """
+    bibcode = '2004A&A...418..989N'
+
+    cat = Vizier(catalog="V/117/newcat", columns=['*', 'HIP', 'RVel', 'e_RVel'])
+
+    sinfo = obs.simbad_info
+    if sinfo.hd_id is not None:
+        tbl = cat.query_constraints(Name='HD {}'.format(sinfo.hd_id))
+        try:
+            row = tbl[0][0]
+        except IndexError:
+            return None
+
+    elif sinfo.hip_id is not None:
+        tbl = cat.query_constraints(HIP=int(sinfo.hip_id))
+        try:
+            row = tbl[0][0]
+        except IndexError:
+            return None
+
+    else:
+        return None
+
+    return float(row['RVel']), float(row['e_RVel']), None, bibcode
+
+def get_CRVAD_rv(obs):
+    """
+    Try to retrieve a velocity from the CRVAD.
+
+    Parameters
+    ----------
+    obs : `comoving_rv.db.model.Observation`
+        An observation instance.
+
+    Returns
+    -------
+    rv : float
+        Radial velocity (in km/s).
+    rv_err : float or None
+        Radial velocity error (in km/s).
+    rv_qual : str or None
+        Quality flag for the measurement.
+    bibcode : str
+        Bibliographic code for the source of the measurement.
+
+    or
+
+    ``None``
+
+    """
+    bibcode = '2007AN....328..889K'
[…]

### 2017-06-07 comoving_rv/db/__init__.py
retrieve prior rvs in inject script

@@ -1,2 +1,3 @@
 from .connect import Session, Base, db_connect
 from .model import *
+from literature_helpers import get_best_rv

### 2017-06-07 scripts/injest_db.py

@@ -31,9 +31,9 @@ WHERE source_id = {0}
 """
 
 # Project
-from comoving_rv.db import Session, Base, db_connect
+from comoving_rv.db import Session, Base, db_connect, get_best_rv
 from comoving_rv.db.model import (Run, Observation, TGASSource, SimbadInfo,
-                                  SpectralLineInfo)
+                                  SpectralLineInfo, PriorRV)
 from comoving_rv.velocity import bary_vel_corr, kitt_peak
 from comoving_rv.log import logger
 
@@ -130,6 +130,7 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
 
         observations = []
         tgas_sources = []
+        prior_rvs = []
 
         glob_pattr_1d = path.join(proc_night_path, '1d_*.fit')
         for path_1d in ProgressBar(glob.glob(glob_pattr_1d)):
@@ -298,28 +299,24 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
             else:
                 logger.log(1, 'TGAS row could not be found.')
 
-            # object_name is never None?
-            try:
-                result = Simbad.query_object(object_name)
-            except Exception as e:
-                logger.warning('Simbad query_object failed for "{0}" '
-                               'with error: {1}'
-                               .format(object_name, str(e)))
-                continue
-
-            if result is not None and not np.any(result['RV_VALUE'].mask):
-                k, = np.where(np.logical_not(result['RV_VALUE'].mask))
-                simbad_info.rv = float(result['RV_VALUE'][k]) * u.km/u.s
-                simbad_info.rv_qual = result['RVZ_QUAL'].astype(str)[k]
-                simbad_info.rv_bibcode = result['RVZ_BIBCODE'].astype(str)[k]
-
             obs.simbad_info = simbad_info
             observations.append(obs)
+
+            # retrieve a previous measurement from the literature
+            result = get_best_rv(obs)
+            if result is not None:
+                rv, rv_err, rv_qual, rv_bibcode = result
+
+                prv = PriorRV(rv=rv*u.km/u.s, err=rv_err*u.km/u.s,
+                              qual=rv_qual, bibcode=rv_bibcode)
+                obs.prior_rv = prv
+                prior_rvs.append(prv)
 
             logger.log(1, '-'*68)
 
         session.add_all(observations)
         session.add_all(tgas_sources)
+        session.add_all(prior_rvs)
         session.commit()
 
     # Last thing to do is cross-match with the Bensby catalog to

### 2017-06-08 comoving_rv/db/__init__.py
fix some backrefs

@@ -1,3 +1,3 @@
 from .connect import Session, Base, db_connect
 from .model import *
-from literature_helpers import get_best_rv
+from .literature_helpers import get_best_rv

### 2017-06-08 comoving_rv/db/literature_helpers.py

@@ -99,9 +99,13 @@ def get_CRVAD_rv(obs):
             return None
 
         err = float(row['e_RV'])
+        qual = str(row['q_RV'].astype(str)).lower()
+
         if np.isnan(err):
-            err = float(row['meRVel'])
-        return float(row['RV']), err, str(row['q_RV'].astype(str)), bibcode
+            if qual in ['a', 'b']:
+                err = 10.
+
+        return float(row['RV']), err, qual, bibcode
 
     return None
 
@@ -215,24 +219,26 @@ def get_best_rv(obs):
 
     """
 
+    sources = ['GCS', 'CRVAD', 'GCRV', 'SIMBAD']
+
     results = list()
-    for func in [get_GCS_rv, get_CRVAD_rv, get_GCRV_rv, get_simbad_rv]:
+    for i,func in enumerate([get_GCS_rv, get_CRVAD_rv, get_GCRV_rv, get_simbad_rv]):
         res = func(obs)
         if res is not None:
             err = res[1]
             if err is not None and not np.isnan(err) and err > 0 and err < 10: # it's a good measurement!
-                return res
+                return tuple(res) + (sources[i],)
 
             if res[1] is None and res[2] is None: # no error or quality flag
                 continue
 
-            results.append(res)
+            results.append(tuple(res) + (sources[i],))
 
     if len(results) == 0:
         return None
 
     for res in results:
-        if res[1] < 10 or res[2].lower() == 'a':
+        if (res[1] is not None and res[1] < 10) or (res[2] is not None and res[2].lower() == 'a'):
             return res
 
     return None

### 2017-06-08 comoving_rv/db/model.py

@@ -80,17 +80,20 @@ class Observation(Base):
     simbad_info_id = Column('simbad_info_id', types.Integer,
                             ForeignKey('simbad_info.id'))
     simbad_info = relationship('SimbadInfo', cascade='all,delete-orphan',
-                               backref='observation', single_parent=True)
+                               backref=backref('observation', uselist=False),
+                               single_parent=True)
 
     prior_rv_id = Column('prior_rv_id', types.Integer,
                          ForeignKey('prior_rv.id'))
     prior_rv = relationship('PriorRV', cascade='all,delete-orphan',
-                            backref='observation', single_parent=True)
+                            backref=backref('observation', uselist=False),
+                            single_parent=True)
 
     tgas_source_id = Column('tgas_source_id', types.Integer,
                             ForeignKey('tgas_source.id'))
     tgas_source = relationship('TGASSource', cascade='all,delete-orphan',
-                               backref='observation', single_parent=True)
+                               backref=backref('observation', uselist=False),
+                               single_parent=True)
 
     def __repr__(self):
         return ("<Observation {0.object} [{0.simbad_info.preferred_name}], "
@@ -140,7 +143,7 @@ class SimbadInfo(Base):
             if the_id is not None:
                 return '{0} {1}'.format(pre, the_id)
 
-        return '{0} {1}'.format('SMOH', self.observation.object_name)
+        return '{0} {1}'.format('SMOH', self.observation.object)
 
 class PriorRV(Base):
     __tablename__ = 'prior_rv'

### 2017-06-08 scripts/injest_db.py
query in injest

@@ -305,10 +305,11 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
             # retrieve a previous measurement from the literature
             result = get_best_rv(obs)
             if result is not None:
-                rv, rv_err, rv_qual, rv_bibcode = result
+                rv, rv_err, rv_qual, rv_bibcode, rv_source = result
 
                 prv = PriorRV(rv=rv*u.km/u.s, err=rv_err*u.km/u.s,
-                              qual=rv_qual, bibcode=rv_bibcode)
+                              qual=rv_qual, bibcode=rv_bibcode,
+                              source=rv_source)
                 obs.prior_rv = prv
                 prior_rvs.append(prv)
 

### 2017-06-28 comoving_rv/longslit/extract.py
use new fitter class to help with trace extraction

@@ -53,26 +53,21 @@ def fast_trace_fit(x, flux, ivar, lsf_pars,
         Number of terms in the background polynomial fit.
     """
 
-    # TODO: use the VoigtLineFitter class to help with this
-    #   - can use class methods to get params n' shit
-
     sort_idx = np.argsort(x)
     x = np.array(x)[sort_idx]
     flux = np.array(flux)[sort_idx]
-    if flux_ivar is not None:
-        flux_ivar = np.array(flux_ivar)[sort_idx]
+
+    if ivar is not None:
+        ivar = np.array(ivar)[sort_idx]
     else:
-        flux_ivar = np.ones_like(flux)
+        ivar = np.ones_like(flux)
 
-    # initial guess
-    p0 = get_init_guess(x=x, flux=flux, ivar=flux_ivar,
-                        amp0=amp0, x0=x0,
-                        std_G0=lsf_pars['std_G'], hwhm_L0=lsf_pars['hwhm_L'],
-                        bg0=bg0, n_bg_coef=n_bg_coef, target_x=target_x,
-                        absorp_emiss=1.)
+    lf = VoigtLineFitter(x, flux, ivar, absorp_emiss=1.)
+    init_params = lf.init_gp()
+    p0 = init_params[[2,3,6,7]]
+    p0[0] = np.exp(p0[0])
 
-    p0['std_G'] = lsf_pars['std_G']
-    p0['hwhm_L'] = lsf_pars['hwhm_L']
+    p0 = OrderedDict([(x,y) for x,y in zip(['amp', 'x0', 'bg0', 'bg1'], p0)])
 
     # shift x array so that line is approximately at 0
     _x = np.array(x, copy=True)
@@ -87,8 +82,8 @@ def fast_trace_fit(x, flux, ivar, lsf_pars,
     leastsq_kw.setdefault('xtol', 1e-10)
     leastsq_kw.setdefault('maxfev', 100000)
 
-    args = (x, flux, flux_ivar, lsf_pars)
-    p_opt,p_cov,*_,mesg,ier = leastsq(errfunc, par_dict_to_list(p0),
+    args = (x, flux, ivar, lsf_pars)
+    p_opt,p_cov,*_,mesg,ier = leastsq(errfunc, list(p0.values()),
                                       args=args, full_output=True, **leastsq_kw)
 
     s_sq = (errfunc(p_opt, *args)**2).sum() / (len(flux)-len(p0))
[…]

### 2017-06-28 comoving_rv/longslit/fitting/core.py

@@ -50,7 +50,7 @@ class LineFitter(object):
             self._err = None
 
         if absorp_emiss is None:
-            raise NotImplementedError("Sorry, can't do that yet...")
+            raise NotImplementedError("You must supply an absorp_emiss for now")
         self.absorp_emiss = float(absorp_emiss)
         self.target_x = target_x
 

### 2017-06-28 scripts/extract_1d.py

@@ -228,7 +228,9 @@ def main(night_path, skip_list_file, mask_file, overwrite=False, plot=False):
             ext = SourceCCDExtractor(filename=path.join(ic.location, fname),
                                      plot_path=plot_path, zscaler=zscaler,
                                      cmap=cmap, unit='adu', **ccd_props)
-            nccd = ext.process_raw_frame(pixel_mask_spec=pixel_mask_spec.get(fname, None),
+
+            _pix_mask = pixel_mask_spec.get(fname, None) if pixel_mask_spec is not None else None
+            nccd = ext.process_raw_frame(pixel_mask_spec=_pix_mask,
                                          master_bias=master_bias,
                                          master_flat=master_flat)
             nccd.write(new_fname, overwrite=overwrite)

### 2017-06-28 paper/comoving-followup.tex
uncomment figures

@@
              Princeton University, Princeton, NJ 08544, USA}
-\affiliation{To whom correspondence should be addressed:
-             \texttt{adrn@astro.princeton.edu}}
 
@@
 % \affiliation{todo}
+
+\correspondingauthor{Adrian M. Price-Whelan}
+\email{adrn@astro.princeton.edu}
 
@@
 \keywords{
-    TODO
+    binaries: spectroscopic
+    ---
+    methods: observational
+    ---
+    techniques: radial velocities
+    ---
+    catalogs
+    ---
+    stars: formation
 }
@@
 
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{mcmc_corner.pdf}
-%   \end{center}
-%   \caption{%
-%     \todo{Some summary plots about the MCMC fits...}
-%     Small uncertainty in centroid = 2 km/s velocity precision here.
-%     \label{fig:Halpha-mcmc-corner}}
-% \end{figure}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
+  \end{center}
+  \caption{%
+    \todo{Some summary plots about the MCMC fits...}
+    Small uncertainty in centroid = 2 km/s velocity precision here.
+    \label{fig:Halpha-mcmc-corner}}
+\end{figure}
 
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
-%   \end{center}
-%   \caption{%
-%     Top shows background linear model + Voigt line model.
-%     \label{fig:Halpha-mcmc-fit}}
-% \end{figure}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
+  \end{center}
+  \caption{%
+    Top shows background linear model + Voigt line model.
+    \label{fig:Halpha-mcmc-fit}}
+\end{figure}
 

### 2017-06-28 scripts/extract_1d.py
and save the damn file

@@ -262,7 +262,7 @@ def main(night_path, skip_list_file, mask_file, overwrite=False, plot=False):
             hdu1 = fits.table_to_hdu(tbl)
             hdulist = fits.HDUList([hdu0, hdu1])
 
-            # hdulist.writeto(fname_1d, overwrite=overwrite)
+            hdulist.writeto(fname_1d, overwrite=overwrite)
 
         del ext
 

### 2017-06-29 paper/comoving-followup.tex
microtype
first stab at abstract

@@
 \graphicspath{{figures/}}
+\usepackage{microtype}
 
@@
 
-\begin{document}
+\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me
 
-\title{Co-moving stars in \textsl{Gaia DR1}:
-       a catalog of genuine comoving star pairs from low-resolution
-       spectroscopic follow-up}
+\title{A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1}
+       and low-resolution spectroscopy}
 
@@
 % Context
-Widely-separated, co-eval stars are [...] for studying the chemical homogeneity
-of birth, calibrating stellar atmosphere models across a range of [..], and for
-modeling the Galactic tidal field.
-Previous work using Hipparcos + literature RVs -> a few hundred genuine pairs
-out to 8 pc.
+Widely-separated, coeval stars are powerful tools for testing star-formation
+model predictions, improving stellar atmosphere models, and measuring the
+mass distribution of the Milky Way.
+They vastly outnumber the Galactic open clusters, but are still expected to
+share a common origin and chemistry.
+The kinematic properties of bound wide binaries and their unbound complements
+serve as potentiometers for the smooth tidal field and perturbation spectrum in
+the Milky Way.
+Thus far, a few hundred genuine comoving star pairs with separations as large
+as $8~\pc$ have been discovered in the solar neighborhood.
 % Aims
-Identified candidate pairs using the Tycho-Gaia Astrometric Solution catalog,
-Gaia DR1.
-Want to identify genuinely dynamically associated pairs for future
-high-resolution follow-up
+In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
+using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
+catalog, a part of \gaia\ \DR{1}.
+However, without radial velocity measurements for the vast majority of the
+pairs, the contamination rate at large separations is expected to be
+considerable.
 % Methods
-Low-resolution spectrograph to measure relative radial velocities to identify
-genuine pairs.
+Here, we present results from our own radial-velocity survey of a sample of the
+candidate pairs to validate wide-separation comoving star pairs.
 % Results
-Of the XX pairs observed, we find YY ...
+Of the XX comoving pairs observed, we find YY ...
 % Conclusions
@@
 The spectra were reduced and calibrated using a custom, publicly-available
-pipeline written in \python,\footnote{Released along with the code used to
-generate the figures in this document at
-\url{https://github.com/adrn/GaiaPairsFollowup}.} described below.
+pipeline written in \python, described below.
 
@@
 
+- Stellar rotation of evolved-main sequence
+- Full catalog should contain Shaya and Olling, cross-match to RV surveys
+
 \section{Conclusions }

### 2017-06-29 paper/comoving-followup.tex
introduction

@@
 \newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
+\newcommand{\dex}{\ensuremath{{\rm dex}}}
 
@@
 Widely-separated, coeval stars are powerful tools for testing star-formation
-model predictions, improving stellar atmosphere models, and measuring the
+model predictions, improving stellar atmosphere models, and inferring the
 mass distribution of the Milky Way.
@@
 
+The majority of stars are thought to form in dynamically cold, chemically
+homogeneous clusters, at least relative to the kinematic and chemical
+dispersions observed amongst typical stars in the Galactic disk.
+Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
+progenitor gas cloud disperses quickly (cite), it is typically assumed that
+stars formed in the same birth cluster have the same age and initial chemistry
+(e.g., cite).
+These birth clusters are typically born un- or loosely-bound and therefore
+quickly succumb to the tidal field of the Milky Way; obvious dynamical
+associations between stars from the same cluster are then erased as the
+constituent stars mix into the general population of disk stars.
+However, the stars do preserve their initial associations in age, chemical
+abundance patterns, and dynamical invariants (e.g., orbital actions).
+It has therefore been long presumed that ``chemical tagging'' and chemodynamical
+models of the Galactic disk will allow for inferences about the formation and
+evolution of the Milky Way.
+That is, the chemical environment of a birth cluster is likely imprinted on its
+stellar children so that they may be re-associated with precisely measured
+abundance patterns (of $\gtrsim 10$ species).
+
+The feasibility of chemical tagging depends on the intrinsic chemical
+dispersions of stars formed from the same birth clusters, the precision of the
+abundance measurements, and the ability to tell that the chemical abundances of
+two stars are more similar than two randomly-paired Galactic disk stars.
+Recent work suggests that, at least for red giant stars, the abundance
+distributions of 20 elements in a handful of open clusters included in the
+\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
+with having intrinsic dispersions smaller than the abundance measurement
+precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
+In effect, for most elements, the intrinsic dispersions are unconstrained
+because of measurement noise, so the true intrinsic dispersions are not known.
+It was also recently shown that $\approx 0.3\%$ of random pairings of Galactic
+field red giant stars have abundance patterns that are indistinguishable at the
+same level (``doppelgangers''), even with 20 chemical species
+(\citealt{Ness:2017}).
+
+Are these properties of open clusters universal, or limited to the few clusters
+presently studied? And do the chemical properties of open clusters depend on the
+star formation rate or efficiency, which may evolve with time?
+Ideal tests of the chemical homogeneity assumptions would use many more open
+clusters with varied cluster masses, and member stars that span a large range in
+masses and evolutionary stages.
+However, because they disrupt quickly, the present-day open cluster population
+is biased towards younger or more massive clusters.
+
+An alternate tracer of the chemical properties mentioned above are
+widely-separated, coeval pairs or multiplets of stars.
+Summary of wide binary literature
+
+- Also useful for stellar models: should have same age, chemistry, but need to
+  test this
+    - Chemical homogeneity tested with few open clusters - wide binaries another
+      test
+    - Need sample of confirmed widely-separated binaries, follow-up to get
+      detailed abundances
+    - Need way to verify that widely separated stars are physically associated,
+      via kinematics
+- Tests of chemical enrichment due to planetary systems
+
+In this work, we present a catalog of genuine comoving pairs.
+We expect high-resolution follow-up to study chemistry...
+
+% Caution that prior results are for red giants, MS stars may have planetary
+% imprints that get erased as dredge-up and convective layer erases...
+
 Outline:
@@
     - Can that be described just from Galactic tides, or need perturbers?
-- Also useful for stellar models: should have same age, chemistry, but need to
-  test this
-    - Chemical homogeneity tested with few open clusters - wide binaries another
-      test
-    - Need sample of confirmed widely-separated binaries, follow-up to get
-      detailed abundances
-    - Need way to verify that widely separated stars are physically associated,
-      via kinematics
+
 - Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\
@@
 
-Most stars form in clusters with small dispersions in position, velocity, and
-(likely) small initial spreads in chemical abundances.
-Once a given cluster is visibly dispersing or fully dispersed, the former member
-stars retain chemical and kinematic signatures of their birth cluster
-
 comoving not wide binary (don't know if they are bound)
-
-, though
-the kinematic mapping is non-trivial because of phase-mixing.
-
-become
-useful
 
@@
 ratio computed for all nearly co-spatial pairs of stars.
-For a full description of the selection method, see \citealt{Oh:2017}.
 Briefly, likelihood 1, $L_1$, considers a model in which the full-space
@@
 paired with other stars.
+For a full description of the selection method, see \citealt{Oh:2017}.
 

### 2017-06-29 paper/comoving-followup.tex
more intro

@@
 % Conclusions
-Gaia DR2 -> how many do we expect?
+Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
+separation for a typical MS star can we determine they are bound assuming
+high-res spectra to get very precise RVs?
 \end{abstract}
@@
 progenitor gas cloud disperses quickly (cite), it is typically assumed that
-stars formed in the same birth cluster have the same age and initial chemistry
-(e.g., cite).
+stars formed in the same birth cluster have the same initial chemistry
+(e.g., cite + \citealt{Feng:2014}) and same ages.
 These birth clusters are typically born un- or loosely-bound and therefore
@@
 Are these properties of open clusters universal, or limited to the few clusters
-presently studied? And do the chemical properties of open clusters depend on the
-star formation rate or efficiency, which may evolve with time?
+presently studied?
+Do the chemical properties of open clusters depend on the star formation rate or
+efficiency, which may evolve with time?
 Ideal tests of the chemical homogeneity assumptions would use many more open
-clusters with varied cluster masses, and member stars that span a large range in
+clusters with varied cluster masses and member stars that span a large range in
 masses and evolutionary stages.
+\todo{Gaia will help \gaia\ (\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a})}
 However, because they disrupt quickly, the present-day open cluster population
@@
 An alternate tracer of the chemical properties mentioned above are
-widely-separated, coeval pairs or multiplets of stars.
-Summary of wide binary literature
+widely-separated ($>XX~\pc$), coeval pairs or multiplets of stars.
+These tracers include both bound and unbound wide-separation stellar systems; we
+refer to these systems collectively as ``comoving stars'' because, at large
+separations, with present-day and near-future velocity uncertainties, it is
+typically not possible to determine whether these systems are in fact bound.
+Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
+dissolving or disrupted open clusters.
+In any of these cases, it is likely that the stars formed from the same region
+of a birth cloud, yet are separated enough that there has not been any
+cross-pollution in the event of strong stellar winds or stellar evolution.
 
-- Also useful for stellar models: should have same age, chemistry, but need to
-  test this
-    - Chemical homogeneity tested with few open clusters - wide binaries another
-      test
+For chemical studies, open clusters have the advantage of having many stars per
+cluster, but there are many fewer clusters than comoving stars.
+Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
+surveys (cite rave, etc.), several hundred such systems are already known
+(\citealt{Shaya:2011}).
+In addition, over ten-thousand \emph{candidate} comoving stars have been
+discovered from astrometric data alone, using data from both the
+\project{Hipparcos} mission (\citealt{hipparcos?}) and the \gaia\ mission
+(\citealt{Oh:2017,Andrews:2017}).
 - Need sample of confirmed widely-separated binaries, follow-up to get
@@
   via kinematics
-- Tests of chemical enrichment due to planetary systems
 
-In this work, we present a catalog of genuine comoving pairs.
-We expect high-resolution follow-up to study chemistry...
+Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
+Also tests of chemical enrichment due to planetary systems
 
-% Caution that prior results are for red giants, MS stars may have planetary
-% imprints that get erased as dredge-up and convective layer erases...
-
-Outline:
-- Stars form in clusters, chemically homogeneous at some level
-- Clusters generally disrupt slowly through a process that preserves initial
-  clustering in conserved quantities but destroys phase coherence
-- Most significant process is probably tidal heating
-- Dissolved clusters are therefore useful dynamical tracers of mass distribution
-- In Galactic halo, globular clusters
-- In disk, hope is to use chemical signatures to connect stars, then can use
-  dynamics to reconstruct birth clusters, infer mass distribution / evolution
-- Clusters have advantage of many stars per cluster, but fewer clusters
-- Wide binary stars are little potentiometers themselves - only N=2, but are far
-  more numerous as a population
+Comoving stars are also sensitive potentiometers of the Galactic mass
+distribution.
 - Long-term goal is to study the separation distribution of wide-binaries as a
@@
 
-- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\
+In this work, we present a catalog of genuine comoving star pairs discovered
+through radial velocity follow-up of candidate comoving stars discovered from
+their astrometry alone (\citealt{Oh:2017}).
+We expect high-resolution follow-up to study chemistry...
 
-\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}
-
-comoving not wide binary (don't know if they are bound)
-
-Widely-separated comoving stars are expected to form from the dissolution of
-birth star clusters
+% Caution that prior results are for red giants, MS stars may have planetary
+% imprints that get erased as dredge-up and convective layer erases...
 

### 2017-07-03 paper/comoving-followup.tex
intro first pass done, woot

@@
 \newcommand{\gaia}{\project{Gaia}}
-\newcommand{\DR}[1]{\acronym{DR}1}
+\newcommand{\DR}[1]{\acronym{DR}#1}
 \newcommand{\tgas}{\acronym{TGAS}}
@@
 These birth clusters are typically born un- or loosely-bound and therefore
-quickly succumb to the tidal field of the Milky Way; obvious dynamical
-associations between stars from the same cluster are then erased as the
-constituent stars mix into the general population of disk stars.
+quickly succumb to the tidal field of the Milky Way.
+Obvious dynamical associations between stars from the same cluster are then
+erased as the constituent stars mix into the general population of disk stars.
 However, the stars do preserve their initial associations in age, chemical
@@
 models of the Galactic disk will allow for inferences about the formation and
-evolution of the Milky Way.
+evolution of the Milky Way (e.g., \citealt{todo}).
 That is, the chemical environment of a birth cluster is likely imprinted on its
-stellar children so that they may be re-associated with precisely measured
-abundance patterns (of $\gtrsim 10$ species).
+stellar children and preserved so that their initial association may be inferred
+with precisely measured abundance patterns (of $\gtrsim 10$ species).
 
@@
 because of measurement noise, so the true intrinsic dispersions are not known.
-It was also recently shown that $\approx 0.3\%$ of random pairings of Galactic
-field red giant stars have abundance patterns that are indistinguishable at the
-same level (``doppelgangers''), even with 20 chemical species
-(\citealt{Ness:2017}).
+This demonstration of chemical uniformity in open clusters is promising for
+chemical tagging.
+However, it was also recently shown that $\approx 0.3\%$ of random pairings of
+Galactic field red giant stars have abundance patterns that are
+indistinguishable at the same level (``doppelgangers''), even with 20 chemical
+species (\citealt{Ness:2017}).
 
@@
 masses and evolutionary stages.
-\todo{Gaia will help \gaia\ (\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a})}
+Precise kinematic and color data from the \gaia\ mission
+(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
+existing membership catalogs and will lead to discoveries of many more open
+clusters and young star associations (see, e.g., \citealt{Oh:2017}).
 However, because they disrupt quickly, the present-day open cluster population
@@
 refer to these systems collectively as ``comoving stars'' because, at large
-separations, with present-day and near-future velocity uncertainties, it is
-typically not possible to determine whether these systems are in fact bound.
+separations with present-day and near-future velocity uncertainties, it is
+typically not possible to determine whether these systems are truly bound.
 Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
@@
 discovered from astrometric data alone, using data from both the
-\project{Hipparcos} mission (\citealt{hipparcos?}) and the \gaia\ mission
-(\citealt{Oh:2017,Andrews:2017}).
-- Need sample of confirmed widely-separated binaries, follow-up to get
-  detailed abundances
-- Need way to verify that widely separated stars are physically associated,
-  via kinematics
-
-Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
-Also tests of chemical enrichment due to planetary systems
-
-Comoving stars are also sensitive potentiometers of the Galactic mass
-distribution.
-- Long-term goal is to study the separation distribution of wide-binaries as a
-  function of age and Galactic orbit distribution (disk vs. halo)
-    - Is there a separation truncation?
-    - Can that be described just from Galactic tides, or need perturbers?
+\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
+\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
+\citealt{Oh:2017,Andrews:2017}).
+Radial velocity follow-up is necessary to confirm pairs using these astrometric
+catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
+stars ($G < 12$).
+High-resolution spectroscopy is still needed to measure detailed chemical
+abundances of these pairs, which may present interesting surprises; one
+large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
+discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
+high-resolution, high signal-to-noise spectra display large abundance
+differences ($\approx 0.2~\dex$) that appear to have resulted from massive
+accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
+submitted).
 
 In this work, we present a catalog of genuine comoving star pairs discovered
-through radial velocity follow-up of candidate comoving stars discovered from
-their astrometry alone (\citealt{Oh:2017}).
-We expect high-resolution follow-up to study chemistry...
-
-% Caution that prior results are for red giants, MS stars may have planetary
-% imprints that get erased as dredge-up and convective layer erases...
+through radial velocity follow-up of candidate comoving stars selected from
+inferences using their astrometry alone (\citealt{Oh:2017}).
+This work serves as a pilot study for data expected in upcoming data releases
+from the \gaia\ mission.
+We describe the data, observations, and spectroscopic reduction in
+\sectionname~\ref{sec:data}.
+If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
+where we discuss our main results.
 
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 
@@
 \section{Discussion}
+
+Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
+Also tests of chemical enrichment due to planetary systems
+
+Comoving stars are also sensitive potentiometers of the Galactic mass
+distribution.
+- Long-term goal is to study the separation distribution of wide-binaries as a
+  function of age and Galactic orbit distribution (disk vs. halo)
+    - Is there a separation truncation?
+    - Can that be described just from Galactic tides, or need perturbers?
+
+% Caution that prior results are for red giants, MS stars may have planetary
+% imprints that get erased as dredge-up and convective layer erases...
 

### 2017-07-03 comoving_rv/db/model.py
do some insane corrections for sky lines

@@ -12,6 +12,7 @@ from os import path
 # Third-party
 import astropy.coordinates as coord
 import astropy.units as u
+import numpy as np
 from sqlalchemy import Column, types
 from sqlalchemy.schema import ForeignKey
 from sqlalchemy.orm import relationship, backref
@@ -23,7 +24,7 @@ from .custom_types import (QuantityTypeClassFactory, JDTimeType,
 from .numpy_adapt import * # just need to execute code
 
 __all__ = ['Run', 'Observation', 'SimbadInfo', 'TGASSource',
-           'SpectralLineMeasurement', 'SpectralLineInfo']
+           'SpectralLineMeasurement', 'SpectralLineInfo', 'RVMeasurement']
 
 VelocityType = QuantityTypeClassFactory(u.km/u.s)
 WavelengthType = QuantityTypeClassFactory(u.angstrom)
@@ -113,6 +114,9 @@ class Observation(Base):
                       'n'+str(self.night))
         return path.join(p, self.filename_1d)
 
+    @property
+    def utc_hour(self):
+        return np.sum(np.array(list(map(float, self.time_obs.split(':')))) / np.array([1., 60., 3600.]))
 
 class SimbadInfo(Base):
     __tablename__ = 'simbad_info'
@@ -251,3 +255,17 @@ class SpectralLineInfo(Base):
     def __repr__(self):
         return '<SpectralLineInfo {0} @ {1}>'.format(self.name, self.wavelength)
 
+class RVMeasurement(Base):
+    __tablename__ = 'rv_measurement'
+
+    id = Column(types.Integer, primary_key=True)
+    rv = Column('rv', VelocityType, nullable=False)
+    err = Column('err', VelocityType, nullable=False)
+
+    # Relationships
+    observation_id = Column('observation_id', types.Integer,
+                            ForeignKey('observation.id'))
+    observation = relationship('Observation', single_parent=True,
+                               backref=backref('measurements',
+                                               cascade='all,delete-orphan',
+                                               uselist=False))

### 2017-07-03 scripts/solve_velocity.py

@@ -1,0 +1,285 @@
+# coding: utf-8
+
+"""
+
+
+"""
+
+# Standard library
+import os
+from os import path
+
+# Third-party
+from astropy.constants import c
+import astropy.units as u
+import numpy as np
+import matplotlib.pyplot as plt
+plt.style.use('apw-notebook')
+import schwimmbad
+from schwimmbad import choose_pool
+
+# Project
+from comoving_rv.log import logger
+from comoving_rv.db import Session, db_connect
+from comoving_rv.db.model import (SpectralLineMeasurement, SpectralLineInfo,
+                                  Run, Observation, PriorRV, RVMeasurement)
+
+class RVCorrector(object):
+
+    def __init__(self, session, run_name):
+        self.session = session
+        self.run_name = str(run_name)
+
+        # get wavelength for Halpha
+        self.Halpha, = session.query(SpectralLineInfo.wavelength)\
+                              .filter(SpectralLineInfo.name == 'Halpha').one()
+
+        self._compute_offset_corrections()
+
+    def _compute_offset_corrections(self):
+        session = self.session
+        run_name = self.run_name
+
+        q = session.query(Observation).join(Run, SpectralLineMeasurement, PriorRV)
+        q = q.filter(Run.name == run_name)
+        q = q.filter(SpectralLineMeasurement.x0 != None)
+        q = q.filter(PriorRV.rv != None)
+        logger.debug('{0} observations with prior RV measurements'
+                     .format(q.distinct().count()))
+
+        # retrieve all observations with measured centroids and previous RV's
+        observations = q.all()
+
+        # What we do below is look at the residual offsets between applying a naïve
+        # sky-line correction and the true RV (with the barycentric velocity
+        # applied)
+
+        raw_offsets = np.zeros(len(observations)) * u.angstrom
+        all_sky_offsets = np.full((len(observations), 3), np.nan) * u.angstrom
+        true_rv = np.zeros(len(observations)) * u.km/u.s
+        obs_time = np.zeros(len(observations))
+        night_id = np.zeros(len(observations), dtype=int)
+        corrected_rv = np.zeros(len(observations)) * u.km/u.s
+
+        for i,obs in enumerate(observations):
+            # convert obstime into decimal hour
+            obs_time[i] = np.sum(np.array(list(map(float, obs.time_obs.split(':')))) / np.array([1., 60., 3600.]))
+
[…]

### 2017-07-03 comoving_rv/db/model.py
forgot to add columns

@@ -261,11 +261,12 @@ class RVMeasurement(Base):
     id = Column(types.Integer, primary_key=True)
     rv = Column('rv', VelocityType, nullable=False)
     err = Column('err', VelocityType, nullable=False)
+    flag = Column('flag', types.Integer, nullable=False)
 
     # Relationships
     observation_id = Column('observation_id', types.Integer,
                             ForeignKey('observation.id'))
     observation = relationship('Observation', single_parent=True,
-                               backref=backref('measurements',
+                               backref=backref('rv_measurement',
                                                cascade='all,delete-orphan',
                                                uselist=False))

### 2017-07-03 scripts/solve_velocity.py
protect against no line measurements

@@ -212,6 +212,11 @@ def main(db_path, run_name, overwrite=False, pool=None):
             raise RuntimeError('Multiple RV measurements found for object {0}'
                                .format(obs))
 
+        elif len(obs.measurements) == 0:
+            logger.debug('Observation {0} has no line measurements.'
+                         .format(obs))
+            continue
+
         corrected_rv, err, flag = rv_corr.get_corrected_rv(obs)
 
         # remove previous RV measurements

### 2017-07-04 comoving_rv/db/model.py
adding attributes and methods

@@ -118,6 +118,37 @@ class Observation(Base):
     def utc_hour(self):
         return np.sum(np.array(list(map(float, self.time_obs.split(':')))) / np.array([1., 60., 3600.]))
 
+    def tgas_star(self, with_rv=True):
+        """gwb.data.TGASData instance"""
+        from gwb.data import TGASStar
+
+        names = list(self.tgas_source.row_dict.keys())
+        dtype = dict(names=names, formats=['f8']*len(names))
+        arr = np.array([tuple([self.tgas_source.row_dict[k] for k in names])],
+                       dtype)
+
+        kw = dict()
+        if with_rv:
+            kw['rv'] = self.rv_measurement.rv
+            kw['rv_err'] = self.rv_measurement.err
+
+        return TGASStar(arr[0], **kw)
+
+    def icrs(self, with_rv=True):
+        kw = dict()
+        if with_rv:
+            if hasattr(with_rv, 'unit'):
+                rv = with_rv
+            else:
+                rv = self.rv_measurement.rv
+
+            kw['radial_velocity'] = rv
+
+        return coord.ICRS(ra=self.tgas_source.ra, dec=self.tgas_source.dec,
+                          distance=1000./self.tgas_source.parallax*u.pc,
+                          pm_ra_cosdec=self.tgas_source.pmra*u.mas/u.yr,
+                          pm_dec=self.tgas_source.pmdec*u.mas/u.yr, **kw)
+
 class SimbadInfo(Base):
     __tablename__ = 'simbad_info'
 
@@ -211,6 +242,24 @@ class TGASSource(Base):
     def skycoord(self):
         return coord.SkyCoord(ra=self.ra, dec=self.dec,
                               distance=1000./self.parallax*u.pc)
+
+    @property
+    def row_dict(self):
+        row = dict()
+        for k in self.__dict__:
+            if not k.startswith('_'):
+                v = getattr(self, k)
+
+                if hasattr(v, 'value'):
+                    row[k] = v.value
+
+                elif not isinstance(v, float):
+                    continue
+
+                else:
+                    row[k] = v
+
+        return row
 
 class SpectralLineMeasurement(Base):
     __tablename__ = 'spectral_line_measurement'

### 2017-07-04 comoving_rv/db/model.py
method to sample

@@ -149,6 +149,31 @@ class Observation(Base):
                           pm_ra_cosdec=self.tgas_source.pmra*u.mas/u.yr,
                           pm_dec=self.tgas_source.pmdec*u.mas/u.yr, **kw)
 
+    def icrs_samples(self, size=1, custom_rv=None):
+        y = np.zeros(6)
+        Cov = np.zeros((6,6))
+
+        y[:5] = [self.tgas_source.ra.value, self.tgas_source.dec.value,
+                 self.tgas_source.parallax, self.tgas_source.pmra,
+                 self.tgas_source.pmdec]
+        Cov[:5,:5] = self.tgas_star().get_cov()[:5,:5]
+
+        if custom_rv is None:
+            rv = self.rv_measurement.rv
+            rv_err = self.rv_measurement.err
+        else:
+            rv, rv_err = custom_rv
+
+        Cov[5, 5] = rv_err.value**2
+        y[5] = rv.value
+        samples = np.random.multivariate_normal(y, Cov, size=size)
+
+        return coord.ICRS(ra=samples[:,0]*u.deg, dec=samples[:,1]*u.deg,
+                          distance=1000./samples[:,2]*u.pc,
+                          pm_ra_cosdec=samples[:,3]*u.mas/u.yr,
+                          pm_dec=samples[:,4]*u.mas/u.yr,
+                          radial_velocity=samples[:,5]*u.km/u.s)
+
 class SimbadInfo(Base):
     __tablename__ = 'simbad_info'
 

### 2017-07-04 paper/comoving-followup.tex
new figure wooooooo
words

@@
 \newcommand{\dex}{\ensuremath{{\rm dex}}}
+\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
+\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
 
@@
 We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
-2017-03-15, corresponding to 321 unique star pairs and $\approx 120$
+2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
 calibration targets.
@@
 
-\subsection{Validation}
+\subsection{Validation of absolute radial velocities}
 
@@
 \end{enumerate}
-We find that \todo{XXX} of our target stars have reliable prior measurements,
-along with the \todo{YYY} calibration targets.
+% See: paper/figures/Compare with previous observations.ipynb
+We find that 155 of the observed stars (mainly calibration targets) have
+reliable prior measurements from the literature search described above.
 \figurename~\ref{fig:compare-previous} shows the distribution of velocity
 differences between our measurements and literature values.
-From this distribution, we compute the MAD and estimate the standard deviation
-to be $\approx 20~\kms$, which we adopt as our global systematic error for all
-absolute radial velocities.
+From this distribution, we compute the median absolute deviation (MAD) and
+estimate the standard deviation to be $\approx 20~\kms$, which we adopt as our
+global systematic error for all absolute radial velocities.
 We note, however, that for computing \emph{relative} velocities for pairs of
-nearby stars, it is the RV precision that matters.
+nearby stars, it is the RV precision (typically $\approx 5$--$10~\kms$) that
+matters.
 
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{compare_previous.pdf}
-%   \end{center}
-%   \caption{%
-%     \todo{Plot showing my RV's vs. previously measured.}
-%     \label{fig:compare-previous}}
-% \end{figure}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{rv-comparison.pdf}
+  \end{center}
+  \caption{%
+    \todo{Plot showing my RV's vs. previously measured.}
+    \label{fig:compare-previous}}
+\end{figure}
 
@@
 
-\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
+\subsection{Comoving pair catalog}\label{sec:catalog}
 
-\subsection{Catalog ...}\label{sec:}
+Of the 323 observed star pairs, 11 have at least one member with an unreliable
+(flagged) radial velocity measurement from the pipeline above.
+For the remaining 312 pairs, we use the measured radial velocities to re-compute
+the marginal likelihood ratio for the two hypotheses considered in previous work
+(see Appendix~B, \citealt{Oh:2017}).
+We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
+that the two stars are comoving with identical 3D velocity vectors (within some
+allowed small dispersion), and (2) assuming that the two stars have independent
+3D velocity vectors drawn from a prior velocity distribution.
+As with previous work, we set the small dispersion to the orbital velocity of a
+binary with the separation of the pair, and use a mixture of three Gaussian
+velocity components for the prior, meant to represent the thin disk, thick disk,
+and halo velocity distributions relative to the sun.
 
-\todo{10 overlap with Shaya and Olling}
+To compute the new likelihood ratio, we modify the expressions from
+\citet{Oh:2017} to include the radial velocity measurements.
+In the same notation, this requires the following changes:
+\begin{itemize}
+  \item The projection matrix, $\mat{M}$, now has to include the radial unit
+    vector
+    \begin{equation}
+      \mat{M} =
+        \left(
+          \begin{array}{c c c}
+            -\sin\alpha & \cos\alpha & 0 \\
+            -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
+            \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
+          \end{array}
+        \right)
+    \end{equation}
+    where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
+    declination) of a given star.
+  \item The data-space velocity vector, $\vec{y}$, for one star now becomes
+    \begin{equation}
+      \vec{y} =
+        \transpose{\left(
+          \begin{array}{c c c}
+            r_i\,\mu_{\alpha}^* &
+            r_i\,\mu_{\delta} &
+            v_{r}
+          \end{array}
+        \right)}
+    \end{equation}
+    where $\mu_\alpha^*$ is the proper motion in right ascension (including the
+    $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
+    $v_r$ is the radial velocity.
+  \item The covariance matrix becomes the block-diagonal matrix
+    \begin{equation}
+      \mat{\Sigma} = \left(
+        \begin{array}{c c}
+          r^2 \, \mat{C} & 0 \\
+          0 & \sigma_{v_r}^2
+        \end{array}
+      \right)
+    \end{equation}
+    where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
+    (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
+    uncertainty.
+\end{itemize}
+To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
+vector must become the 6-vector
+\begin{equation}
+  \vec{y} = \transpose{\left(
+    \begin{array}{c c}
+      \vec{y}_i &
+      \vec{y}_j
+    \end{array}
+    \right)}
+\end{equation}
+and the covariance matrix becomes the block-diagonal matrix
+\begin{equation}
+  \mat{\Sigma} = \left(
+    \begin{array}{c c}
+      \mat{\Sigma}_i & 0 \\
+      0 & \mat{\Sigma}_j
+    \end{array}
+  \right) \quad .
+\end{equation}
 
-\todo{Do I want to do isochrone fitting for all of them?}
+If the new likelihood ratio is larger than that computed from astrometry alone,
+the added radial velocity information likely supports the belief that the two
+stars are truly comoving in 3-space.
+Conversely, if the new likelihood ratio is smaller, the radial velocity data
+indicate that the pair is likely a false-positive and are not truly comoving.
+In practice, we compute the logarithm of the likelihood ratios,
+$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
+proper-motion-only log-FML-ratio for a given pair of stars as
+\llrold\, and the log-FML-ratio including the radial velocity data as
+\llrnew.
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 125 of
+the 312 comoving pairs satisfy this constraint.
+\figurename~\ref{fig:llr}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{llr-vs-dv.pdf}
+  \end{center}
+  \caption{%
+    FML
+    \label{fig:llr}}
+\end{figure}
+
+TODO: check on ipython notebook re-running with updated observation shite
 
@@
 
-\todo{Plot of magnitude of 3d velocity difference vs. 3d separation (inferred).}
+\todo{Plot of magnitude of 3d velocity difference vs. 3d separation?}
 
-\todo{separation distribution - combine with Shaya and Olling sample!}
+\todo{separation distribution - compare with Shaya and Olling sample!}
+\todo{10 overlap with Shaya and Olling}
 
@@
 \todo{Orbits of true pairs in the Galaxy, any halo or thick-disk?}
+
+\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
+
+\todo{Do I want to do isochrone fitting for all of them?}
 

### 2017-07-05 comoving_rv/db/model.py
add option for lutz-kelker

@@ -134,7 +134,7 @@ class Observation(Base):
 
         return TGASStar(arr[0], **kw)
 
-    def icrs(self, with_rv=True):
+    def icrs(self, with_rv=True, lutz_kelker=True):
         kw = dict()
         if with_rv:
             if hasattr(with_rv, 'unit'):
@@ -144,8 +144,9 @@ class Observation(Base):
 
             kw['radial_velocity'] = rv
 
+        dist = self.tgas_star(with_rv).get_distance(lutz_kelker)
         return coord.ICRS(ra=self.tgas_source.ra, dec=self.tgas_source.dec,
-                          distance=1000./self.tgas_source.parallax*u.pc,
+                          distance=dist,
                           pm_ra_cosdec=self.tgas_source.pmra*u.mas/u.yr,
                           pm_dec=self.tgas_source.pmdec*u.mas/u.yr, **kw)
 

### 2017-07-05 paper/comoving-followup.tex
words about new figures

@@
 \tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
-\citealt{Oh:2017,Andrews:2017}).
+\citealt{Oh:2017,Oelkers:2017,Andrews:2017}).
 Radial velocity follow-up is necessary to confirm pairs using these astrometric
@@
 (flagged) radial velocity measurement from the pipeline above.
-For the remaining 312 pairs, we use the measured radial velocities to re-compute
+For the remaining 311 pairs, we use the measured radial velocities to re-compute
 the marginal likelihood ratio for the two hypotheses considered in previous work
@@
 
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{llr-vs-dv.pdf}
+  \end{center}
+  \caption{%
+    FML
+    \label{fig:llr}}
+\end{figure}
+
 If the new likelihood ratio is larger than that computed from astrometry alone,
@@
 \llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 125 of
-the 312 comoving pairs satisfy this constraint.
-\figurename~\ref{fig:llr}
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 128 of
+the 311 comoving pairs satisfy this constraint.
+\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
+plotted against the change in log-likelihood-ratio for the two stars in all
+observed pairs.
+In detail, for visualization, we sample from the error distributions over proper
+motion, parallax, and radial velocity for each star individually and compute the
+magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
+samples.
+We plot the median of the resulting distribution over velocity-difference as
+dark markers (circles).
+The error bars show the 15th and 85th percentiles of the distributions for a few
+randomly chosen pairs.
+Note that pairs with large 3D velocity differences have large (negative) changes
+to their log-likelihood-ratio values.
+
+\subsection{Physical properties of genuine comoving pairs}\label{sec:separation}
+
+\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
+plotted against the 3D positional separation of the two stars in all observed
+pairs.
+We again plot the median of the velocity difference distribution, and here use
+the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
+the stars' \tgas\ parallax measurements to compute the 3D positions of the two
+stars.
+Small, grey markers are pairs for which $\llrnew < \llrold$, i.e. where the
+radial velocity measurements weaken or change our belief that a given pair is
+comoving.
+Larger, black markers are genuine comoving pairs ($\llrnew > \llrold$); note
+that many large-separation pairs have been ruled-out as comoving, but
+most small-separation pairs maintain their status as comoving.
+This is also apparent in \figurename~\ref{fig:separation}, which shows
+histograms of the spatial separation and sky separation for all pairs;
+background, grey histogram shows all observed pairs, and foreground, black
+histogram shows all pairs for which $\llrnew > \llrold$.
+
+At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
+genuinely comoving; interestingly, the separation distribution of the genuine
+pairs appears to be at least uniform or possibly growing in number as the
+separation increases.
+This strongly suggests the existence of a population of unbound wide binaries or
+dissolved open clusters, which maintain approximate spatial coherence for long
+times because of the small relative velocities at the time of disruption of the
+constituent stars (e.g., \citealt{Jiang:2010}).
+The kinematic properties of these unbound binaries and associations contain
+information about the tidal field and fluctuations in the mass distribution of
+the Milky Way.
+With \gaia\ \DR{2}, we will be able to study the statistical properties of these
+widely-separated but comoving stars at even larger separations ($>10~\pc$).
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{llr-vs-dv.pdf}
+    \includegraphics[width=\linewidth]{dx-dv.pdf}
   \end{center}
@@
     FML
-    \label{fig:llr}}
+    \label{fig:dx-dv}}
 \end{figure}
 
-TODO: check on ipython notebook re-running with updated observation shite
-
-\subsection{Physical properties of genuine comoving pairs}\label{sec:}
-
-\todo{Plot of magnitude of 3d velocity difference vs. 3d separation?}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{separation-hist.pdf}
+  \end{center}
+  \caption{%
+    FML
+    \label{fig:separation}}
+\end{figure}
 
@@
 \todo{10 overlap with Shaya and Olling}
-
-\todo{distribution of orbital poles in Galactic coordinates}
+\todo{RAVE pairs - both observed}
 
@@
 
-\section{Conclusions }
+\section{Conclusions}
 

### 2017-07-05 paper/comoving-followup.tex
second cmd

@@
 
-\subsection{Physical properties of genuine comoving pairs}\label{sec:separation}
+\subsection{Properties of genuine comoving pairs}\label{sec:genuine}
 
@@
 
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
+  \end{center}
+  \caption{%
+    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
+    the density of all \tgas\ sources within 200 pc; similar to
+    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
+    pairs.
+    Several interesting pairs that include main-sequence and giant component
+    stars are highlighted in the text
+    (\sectionname~\ref{sec:interesting-pairs}).
+    \label{fig:genuine-cmd}}
+\end{figure}
+
 \todo{separation distribution - compare with Shaya and Olling sample!}
@@
 
-\subsection{Galactic orbits of the co-moving pairs}\label{sec:}
+\subsection{Interesting comoving pairs}\label{sec:interesting-pairs}
 
-\todo{Orbits of true pairs in the Galaxy, any halo or thick-disk?}
-
-\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
-
-\todo{Do I want to do isochrone fitting for all of them?}
+\todo{Highlight the MS-Giant pairs}
+\todo{Highlight the binary sequence-main sequence pairs}
 
@@
 % imprints that get erased as dredge-up and convective layer erases...
+
+\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
+
+\todo{Do I want to do isochrone fitting for all of them?}
 

### 2017-07-05 paper/comoving-followup.tex
change section name
adjust figure sizes
one more todo on contamination rate

@@
   \begin{center}
-    \includegraphics[width=\linewidth]{wavelength_gp.pdf}
+    \includegraphics[width=0.8\linewidth]{wavelength_gp.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.9\linewidth]{sample_spectra.pdf}
+    \includegraphics[width=0.6\linewidth]{sample_spectra.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{llr-vs-dv.pdf}
+    \includegraphics[width=0.7\linewidth]{llr-vs-dv.pdf}
   \end{center}
@@
 
+\todo{Discuss contamination rate vs. separation for ratio cut of 6?}
 At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
@@
 
-\subsection{Interesting comoving pairs}\label{sec:interesting-pairs}
+\subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
 

### 2017-07-06 paper/comoving-followup.tex
title
abstract and TGAS ref

@@
 
-\title{A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1}
-       and low-resolution spectroscopy}
+\title{Spectroscopic confirmation of very-wide binaries and large-separation
+       comoving stars from \textsl{Gaia DR1}}
 
@@
 catalog, a part of \gaia\ \DR{1}.
-However, without radial velocity measurements for the vast majority of the
-pairs, the contamination rate at large separations is expected to be
-considerable.
+Without radial velocity measurements for the vast majority of the pairs, the
+contamination rate at large separations is significant ($\sim 50\%$).
 % Methods
@@
 % Results
-Of the XX comoving pairs observed, we find YY ...
+Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
+separations as large as $\approx 10~\pc$.
 % Conclusions
-Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
-separation for a typical MS star can we determine they are bound assuming
-high-res spectra to get very precise RVs?
+We find that the number of comoving star pairs with very wide separations
+appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
+suggesting the existence of a population of recently unbound wide binaries or
+stellar associations.
+% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
+% separation for a typical MS star can we determine they are bound assuming
+% high-res spectra to get very precise RVs?
 \end{abstract}
@@
 clusters with varied cluster masses and member stars that span a large range in
-masses and evolutionary stages.
+stellar masses and evolutionary stages.
 Precise kinematic and color data from the \gaia\ mission
@@
 comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
-the \tgas\ catalog (\citealt{tgas}).
+the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
 These comoving pairs were selected using a cut on a marginalized likelihood

### 2017-07-06 paper/comoving-followup.tex
clean up todos, captions

@@
 An alternate tracer of the chemical properties mentioned above are
-widely-separated ($>XX~\pc$), coeval pairs or multiplets of stars.
+widely-separated ($\gtrsim 0.01~\pc$), coeval pairs or multiplets of stars.
 These tracers include both bound and unbound wide-separation stellar systems; we
@@
 If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
-where we discuss our main results.
+and \sectionname~\ref{sec:discussion} where we present and discuss our main
+results.
 
@@
 in which the full-space velocities of each star were drawn independently from
-the same velocity distribution used above; the likelihood ratio is computed as
-$L_1/L_2$.
+the same velocity distribution used above. The computed likelihood ratio is
+actually the the ratio of fully marginalized likelihoods,
+$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
+distance for each star in a given pair.
 After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
@@
 velocity components for star pairs separated by large angles on the sky.
-After applying a conservative cut on the likelihood ratio motivated by
-computing the same likelihood ratio for random pairings of stars, we find
-13,058 comoving star pairs with physical separations between $10^3~\au$ and
-$10~\pc$.
+After applying a conservative cut on the likelihood ratio motivated by computing
+the same likelihood ratio for random pairings of stars ($\ln
+\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
+physical separations between $10^3~\au$ and $10~\pc$.
 Of these pairs, there are 10,606 unique sources, implying that a significant
@@
 stars within $200~\pc$.
+Candidate comoving pairs are connected by (colored) lines.
 
@@
 future work to measure metallicities and stellar parameters for each observed
-source using \project{The Cannon} (\citealt{Ness:TODO}).
+source using \project{The Cannon} (\citealt{Ness:2015}).
 The spectra were reduced and calibrated using a custom, publicly-available
@@
 strongest, least blended emission lines in one of the extracted 1D comparison
-lamp spectra by comparing to \todo{NIST??}.
-Once a rough mapping from pixel to wavelength is determined, we then fit for
-the individual line centroids in each comparison lamp spectrum by extracting
-small, 8 pixel-wide segments of spectrum around each identified emission line
-and fitting a model for the line profile and background to each individually.
+lamp spectra by comparing to standard HgNe line lists.\footnote{\url{http:
+//physics.nist.gov/PhysRefData/ASD/lines_form.html}}
+Once a rough mapping from pixel to wavelength is determined, we then fit for the
+individual line centroids in each comparison lamp spectrum by extracting small,
+8 pixel-wide segments of the comparison lamp spectrum around each identified
+emission line and fitting a model for the line profile and background to each
+individually.
 We use a pixel-convolved Gaussian profile for each emission line with
@@
 best-fit model subtracted.
-The envelope around zero (shaded region, orange) shows the \todo{1$\sigma$ blah
-blah GP...}
-When the GP variance is large, the pixel-to-wavelength mapping is uncertain.
+The envelope around zero (shaded region, orange) shows the variance of the GP
+model projected into the data space; when the GP variance is large, the
+pixel-to-wavelength mapping is uncertain.
 We therefore only use the pixel range from 250--1210, corresponding to a
@@
 additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
+below).
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10
@@
 to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by index for visualization.
-Suggested by the features in these 10 spectra, there is a mix of metallicities
-and spectral types amongst the observed targets.
+As suggested by the features in these 10 spectra, there is a mix of
+metallicities and spectral types amongst the observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring
@@
 spectrum before re-computing the wavelength grid using the nightly wavelength
-solution \todo{is this sensible?}.
+solution.
 When one or both of the two sky lines is not present or the centroid fit fails,
-the spectrum \todo{flagged in some way, can have large absolute velocity
-systematic error.}
+the spectrum is flagged as having a suspect wavelength calibration.
+In practice, most such sources are calibration targets, which are more likely to
+have very short exposure times and thus weak night-sky lines.
 
@@
 (\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
-samples for a randomly-chosen source (HD\todo{XXXX}, in this case).
+samples for a randomly-chosen source (HD 63408, in this case).
 The marginal posterior distribution over $x_0$ can be converted to a
-distribution over radial velocities; in this case, the estimated (Gaussian
+distribution over radial velocity; in this case, the estimated (Gaussian
 standard deviation) precision is $\approx 2.5~\kms$.
 \figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
-around \Ha\ used in the line modeling (black line) with uncertainties (grey
+around \Ha\ used in the line modeling (black line) with uncertainties (gray
 bars).
@@
 absorption line model (black line) along with the Gaussian process computed from
-the median parameters (orange curve) used to model other structure in the
-vicinity of \Ha (which can be interpreted as correlated noise).
+the median parameters (orange curve) used to model other correlated structure in
+the vicinity of \Ha (which can be interpreted as correlated noise).
 These figures are produced for all source spectrum fits and visually inspected
@@
 that the typical velocity \emph{precision} from measuring the centroid of \Ha\
-is between $2$--$8~\kms$ \todo{double check numbers}.
+is between $5$--$10~\kms$.
 
@@
   \caption{%
-    \todo{Some summary plots about the MCMC fits...}
-    Small uncertainty in centroid = 2 km/s velocity precision here.
+    A corner plot showing all projections of posterior PDF samples for the \Ha\
+    line model and the spectrum of the source HD 63408.
+    Note the small uncertainty in the line centroid measurement, $x_0$, which
+    corresponds to a $\approx 2~\kms$ velocity precision for this source.
+    Parameters are described in \sectionname~\ref{sec:reduction}, and prior PDFs
+    over the parameters are described in \tablename~\ref{tbl:prior-bounds}.
     \label{fig:Halpha-mcmc-corner}}
@@
   \caption{%
-    Top shows background linear model + Voigt line model.
+    Visualization of the \Ha\ absorption line model and fits for the source HD
+    63408 as a demonstration of the procedure used for all targets.
+    \emph{Top}:
+    The source spectrum in a small wavelength range around \Ha\ (black),
+    uncertainty in spectral flux (gray error bars).
+    Over-plotted is the line model + background computed from the median
+    posterior PDF sample (dark green line) and the 15th to 85th percentile
+    region computed from all posterior PDF samples (shaded green region).
+    \emph{Bottom}:
+    Source spectrum residuals computed by subtracting the median model (dark
+    green line in top panel) from the source spectrum.
+    The over-plotted line (orange) shows the Gaussian process background model
+    computed from the median posterior PDF sample.
     \label{fig:Halpha-mcmc-fit}}
@@
 stars.
-Small, grey markers are pairs for which $\llrnew < \llrold$, i.e. where the
+Small, gray markers are pairs for which $\llrnew < \llrold$, i.e. where the
 radial velocity measurements weaken or change our belief that a given pair is
@@
 histograms of the spatial separation and sky separation for all pairs;
-background, grey histogram shows all observed pairs, and foreground, black
+background, gray histogram shows all observed pairs, and foreground, black
 histogram shows all pairs for which $\llrnew > \llrold$.
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 

### 2017-07-06 paper/comoving-followup.tex
minor changes to figure 6

@@
 At each of 1600 rows in each source spectrum image, the LSF amplitude and
-centroid, and background slope and intercept at line centroid are then fit using
-the same procedure as above but with the width parameters fixed.
+centroid, and background slope and intercept at line centroid, are then fit
+using the same procedure as above but with the width parameters fixed.
 The LSF amplitudes are stored as the source fluxes, and the background model
@@
 
-\subsection{Validation of absolute radial velocities}
+\subsection{Validation of absolute radial velocities}\label{sec:rv-validation}
 
@@
   \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
-  to retrieve records from the database, which do not contain uncertainties but
-  do contain a quality flag. We only retain RVs that have a quality flag set to
-  `A' and adopt a $5~\kms$ uncertainty for all such values.
+  (\citealt{Ginsburg:2016}) to retrieve records from the database, which do not
+  contain uncertainties but do contain a quality flag. We only retain RVs that
+  have a quality flag set to `A' and adopt a $5~\kms$ uncertainty for all such
+  values.
 \end{enumerate}
@@
 differences between our measurements and literature values.
-From this distribution, we compute the median absolute deviation (MAD) and
-estimate the standard deviation to be $\approx 20~\kms$, which we adopt as our
-global systematic error for all absolute radial velocities.
+From this distribution (right panel), we compute the median absolute deviation
+(MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
+adopt as our global systematic error for all absolute radial velocities.
 We note, however, that for computing \emph{relative} velocities for pairs of
@@
   \caption{%
-    \todo{Plot showing my RV's vs. previously measured.}
+    Comparison of radial velocities measured from our observations with radial
+    velocities obtained from the literature search described in
+    \sectionname~\ref{sec:rv-validation}.
+    Left panel shows our values on the $x$-axis, ${\rm RV}$, against the
+    previous literature values on the $y$-axis, ${\rm RV}_{\rm lit}$, for 155
+    observed stars (mainly calibration targets) with previously measured radial
+    velocities.
+    Right panel shows a histogram of the differences, $\Delta{\rm RV} = {\rm RV}
+    - {\rm RV}_{\rm lit}$.
+    The estimated standard-deviation of the difference distribution implies a
+    global systematic uncertainty of $\approx 20~\kms$.
     \label{fig:compare-previous}}

### 2017-07-06 paper/comoving-followup.tex
number
update cmd figures to add titles

@@
 
-Of the 323 observed star pairs, 11 have at least one member with an unreliable
+Of the 323 observed star pairs, 12 have at least one member with an unreliable
 (flagged) radial velocity measurement from the pipeline above.
@@
 3D velocity vectors drawn from a prior velocity distribution.
-As with previous work, we set the small dispersion to the orbital velocity of a
-binary with the separation of the pair, and use a mixture of three Gaussian
-velocity components for the prior, meant to represent the thin disk, thick disk,
-and halo velocity distributions relative to the sun.
+We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
+a mixture of three Gaussian velocity components for the prior, meant to
+represent the thin disk, thick disk, and halo velocity distributions relative to
+the sun.
 
@@
   \caption{%
-    FML
+    Point estimate of the 3D velocity difference between two stars in a comoving
+    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
+    log-likelihood-ratio computed when including radial velocity data.
+    Error bars are shown for a few randomly-chosen, representative pairs and
+    represent the 15th and 85th percentiles of velocity differences computed
+    from samples from the error distributions of parallax, proper motion
+    components, and radial velocity.
+    The likelihood ratio increases (i.e. the belief that the stars are comoving
+    becomes stronger) when the 3D velocity difference is small.
     \label{fig:llr}}
@@
 If the new likelihood ratio is larger than that computed from astrometry alone,
-the added radial velocity information likely supports the belief that the two
-stars are truly comoving in 3-space.
+the added radial velocity information supports the belief that the two stars are
+truly comoving in 3-space.
 Conversely, if the new likelihood ratio is smaller, the radial velocity data
@@
 proper-motion-only log-FML-ratio for a given pair of stars as
-\llrold\, and the log-FML-ratio including the radial velocity data as
-\llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 128 of
+\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$---128 of
 the 311 comoving pairs satisfy this constraint.
 \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
-plotted against the change in log-likelihood-ratio for the two stars in all
+plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
 observed pairs.
@@
 samples.
-We plot the median of the resulting distribution over velocity-difference as
+We plot the median of the resulting distribution of velocity-differences as
 dark markers (circles).
@@
 
-\todo{Discuss contamination rate vs. separation for ratio cut of 6?}
 At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
 genuinely comoving; interestingly, the separation distribution of the genuine
-pairs appears to be at least uniform or possibly growing in number as the
-separation increases.
+pairs appears to be uniform or possibly growing in number with increasing
+separation.
 This strongly suggests the existence of a population of unbound wide binaries or
@@
   \caption{%
-    FML
+    Point estimate of the 3D velocity difference between two stars in a comoving
+    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
+    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
+    The distances used in the point estimates are computed using a
+    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
     \label{fig:dx-dv}}
@@
   \caption{%
-    FML
+    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
+    the stars in all observed comoving star pairs (gray), and only those
+    identified as genuinely comoving (black). Note that the contamination rate
+    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
+    significant number of genuine comoving pairs at these very wide separations.
+    \emph{Right}: Distribution of sky separations of all comoving star pairs
+    (gray) and only those identified as genuinely comoving (black).
     \label{fig:separation}}
@@
 
-\todo{separation distribution - compare with Shaya and Olling sample!}
-\todo{10 overlap with Shaya and Olling}
-\todo{RAVE pairs - both observed}
+\subsection{Combined with RAVE}\label{sec:rave}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
+  \end{center}
+  \caption{%
+    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
+    the stars in all observed comoving star pairs (gray), and only those
+    identified as genuinely comoving (black). Note that the contamination rate
+    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
+    significant number of genuine comoving pairs at these very wide separations.
+    \emph{Right}: Distribution of sky separations of all comoving star pairs
+    (gray) and only those identified as genuinely comoving (black).
+    \label{fig:separation}}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
+  \end{center}
+  \caption{%
+    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
+    the density of all \tgas\ sources within 200 pc; similar to
+    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
+    pairs.
+    Several interesting pairs that include main-sequence and giant component
+    stars are highlighted in the text
+    (\sectionname~\ref{sec:interesting-pairs}).
+    \label{fig:genuine-cmd}}
+\end{figure}
 

### 2017-07-07 comoving_rv/db/model.py
add new table to map new group id to old one and get pair of observations

@@ -24,7 +24,8 @@ from .custom_types import (QuantityTypeClassFactory, JDTimeType,
 from .numpy_adapt import * # just need to execute code
 
 __all__ = ['Run', 'Observation', 'SimbadInfo', 'TGASSource',
-           'SpectralLineMeasurement', 'SpectralLineInfo', 'RVMeasurement']
+           'SpectralLineMeasurement', 'SpectralLineInfo', 'RVMeasurement',
+           'GroupToObservations']
 
 VelocityType = QuantityTypeClassFactory(u.km/u.s)
 WavelengthType = QuantityTypeClassFactory(u.angstrom)
@@ -345,3 +346,24 @@ class RVMeasurement(Base):
                                backref=backref('rv_measurement',
                                                cascade='all,delete-orphan',
                                                uselist=False))
+
+class GroupToObservations(Base):
+    __tablename__ = 'group_to_observations'
+
+    id = Column(types.Integer, primary_key=True)
+    group_id = Column(types.Integer, nullable=False)
+    new_group_id = Column(types.Integer, nullable=False)
+
+    observation1_id = Column('observation1_id', types.Integer,
+                             ForeignKey('observation.id'))
+    # observation1 = relationship('Observation', single_parent=True)
+
+    observation2_id = Column('observation2_id', types.Integer,
+                             ForeignKey('observation.id'))
+    # observation2 = relationship('Observation', single_parent=True)
+
+    def __repr__(self):
+        r = ('<GroupToObservations old group id={0}, new group id={1}, [{2}, {3}]>'
+             .format(self.group_id, self.new_group_id, self.observation1_id,
+                     self.observation2_id))
+        return r

### 2017-07-07 paper/comoving-followup.tex
add new table to map new group id to old one and get pair of observations
sketch discussion and conclusion

@@
 
+\subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
+
+\todo{Highlight the G subgiant-A pair}
+\todo{Highlight a binary sequence-main sequence pair}
+\todo{Highlight the largest mag difference MS pair}
+
 \subsection{Combined with RAVE}\label{sec:rave}
@@
 
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
-  \end{center}
-  \caption{%
-    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
-    the density of all \tgas\ sources within 200 pc; similar to
-    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
-    pairs.
-    Several interesting pairs that include main-sequence and giant component
-    stars are highlighted in the text
-    (\sectionname~\ref{sec:interesting-pairs}).
-    \label{fig:genuine-cmd}}
-\end{figure}
-
-\subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
-
-\todo{Highlight the MS-Giant pairs}
-\todo{Highlight the binary sequence-main sequence pairs}
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
+%   \end{center}
+%   \caption{%
+%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
+%     the density of all \tgas\ sources within 200 pc; similar to
+%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
+%     pairs.
+%     Several interesting pairs that include main-sequence and giant component
+%     stars are highlighted in the text
+%     (\sectionname~\ref{sec:interesting-pairs}).
+%     \label{fig:genuine-cmd}}
+% \end{figure}
 
@@
 
-Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
+Abundance of large-separation pairs
+
+Need precise chemical information
+
+But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?
+
+Going forward, will have to use hierarchical mixture models to study population of singles in conjunction
+
+Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
 Also tests of chemical enrichment due to planetary systems
@@
 
-\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
-
 \todo{Do I want to do isochrone fitting for all of them?}
 
-Abundances!
+Abundances - future work, Coronado in prep.!
 
-- Stellar rotation of evolved-main sequence
-- Full catalog should contain Shaya and Olling, cross-match to RV surveys
+Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>
 
 \section{Conclusions}
+
+Low-resolution spectroscopic follow-up of comoving pairs.
+
+Using relative RV's, confirm a sample of very wide separation comoving star pairs.
+
+Either wide binaries, unbound binaries, or dissolving clusters
+
+Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution
+
+Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.
 

### 2017-07-09 paper/comoving-followup.tex
add names

@@
     significant number of genuine comoving pairs at these very wide separations.
-    \emph{Right}: Distribution of sky separations of all comoving star pairs
-    (gray) and only those identified as genuinely comoving (black).
+    \emph{Right}: Distribution of chord lengths (tangential separations) for all
+    comoving star pairs (gray) and only those identified as genuinely comoving
+    (black).
+    Noise in the parallax measurements makes the 3D separation distribution
+    appear roughly uniform.
+    Even the distribution in chord length appears to be approximately linear for
+    large separations ($\gtrsim 0.1~\pc$).
     \label{fig:separation}}
@@
 
+We naively expect that comoving, coeval stars should lie along lines parallel to
+the main sequence in color-magnitude space. The majority of our star pairs
+appear to follow this expectation, with the exception of a few visible outliers.
+For example, there appear to be at least a few main-sequence--giant-branch
+pairs, and a few pairs that appear to cross the main-sequence.
+We highlight a few of these pairs below.
+
+\subsubsection{HD 95088 \& HD 92278}
+
 \todo{Highlight the G subgiant-A pair}
+
+\subsubsection{TYC 2486-1699-1 \& TYC 2477-430-1}
+
 \todo{Highlight a binary sequence-main sequence pair}
+
+\subsubsection{TYC 3792-658-1 \& HD 54318}
+
 \todo{Highlight the largest mag difference MS pair}
@@
   \caption{%
-    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
-    the stars in all observed comoving star pairs (gray), and only those
-    identified as genuinely comoving (black). Note that the contamination rate
-    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
-    significant number of genuine comoving pairs at these very wide separations.
-    \emph{Right}: Distribution of sky separations of all comoving star pairs
-    (gray) and only those identified as genuinely comoving (black).
-    \label{fig:separation}}
+    Same as \figurename~\ref{fig:separation}, but for our sample of genuine
+    comoving stars combined with all genuine comoving stars identified from the
+    RAVE overlap.
+    \label{fig:separation-RAVE}}
 \end{figure}

### 2017-07-10 paper/comoving-followup.tex
todos

@@
 
-\subsubsection{TYC 2486-1699-1 \& TYC 2477-430-1}
+\subsubsection{HD 151090 \& HIP 81988}
 
-\todo{Highlight a binary sequence-main sequence pair}
+\todo{G-subgiant - K-dwarf - well known high proper-motion comoving pair}
 
-\subsubsection{TYC 3792-658-1 \& HD 54318}
+\subsubsection{HD 54318 \& TYC 3792-658-1}
 
-\todo{Highlight the largest mag difference MS pair}
+\todo{Highlight the largest mag difference MS pair - K-dwarf - B-star}
 
@@
 
-Abundances - future work, Coronado in prep.!
+Abundances - future work, Coronado in prep.! But: throw shade on abundances
+because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
+2017b
 

### 2017-07-11 paper/comoving-followup.tex
Add smoh comments

--- 2017-07-07 paper/comoving-followup.tex
@@
 \newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
+
+\definecolor{mediumpersianblue}{rgb}{0.0, 0.4, 0.65}
+\newcommand{\smoh}[1]{\textcolor{mediumpersianblue}{SO:#1}}
 
@@
 % Context
-Widely-separated, coeval stars are powerful tools for testing star-formation
+Widely-separated ($\gtrsim X$~pc), coeval stars are powerful tools for testing star-formation
 model predictions, improving stellar atmosphere models, and inferring the
 mass distribution of the Milky Way.
+\smoh{The following sentence seems to be in the air while I get what you mean.}
 They vastly outnumber the Galactic open clusters, but are still expected to
 share a common origin and chemistry.
-The kinematic properties of bound wide binaries and their unbound complements
-serve as potentiometers for the smooth tidal field and perturbation spectrum in
+\smoh{How about sticking with the terminology ("comoving (star) pair") we want settled?}
+The kinematic properties of comoving pairs including bound wide binaries
+serve as potentiometers for the smooth tidal field and perturbation spectrum of
 the Milky Way.
+\smoh{If you are referring to Shaya \& Olling, how about adding 'using HIPPARCOS'?}
 Thus far, a few hundred genuine comoving star pairs with separations as large
@@
 Without radial velocity measurements for the vast majority of the pairs, the
+\smoh{I prefer "false-positive rate" to "contamination rate" because it conveys
+the meaning more directly.}
 contamination rate at large separations is significant ($\sim 50\%$).
 % Methods
-Here, we present results from our own radial-velocity survey of a sample of the
+\smoh{I think some details of telescope/instrument or just R~2000 should be mentioned here.}
+Here, we present results from our own low-resolution radial-velocity survey of a sample of the
 candidate pairs to validate wide-separation comoving star pairs.
@@
 Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
-separations as large as $\approx 10~\pc$.
+separations as large as $\approx 10~\pc$, our original search limit.
 % Conclusions
@@
 The majority of stars are thought to form in dynamically cold, chemically
-homogeneous clusters, at least relative to the kinematic and chemical
-dispersions observed amongst typical stars in the Galactic disk.
+homogeneous clusters, relative to the kinematic and chemical
+dispersions observed amongst typical field stars in the Galactic disk.
 Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
@@
 quickly succumb to the tidal field of the Milky Way.
-Obvious dynamical associations between stars from the same cluster are then
+The initial kinematic associations between stars from the same cluster are then
 erased as the constituent stars mix into the general population of disk stars.
@@
 
-The feasibility of chemical tagging depends on the intrinsic chemical
-dispersions of stars formed from the same birth clusters, the precision of the
-abundance measurements, and the ability to tell that the chemical abundances of
-two stars are more similar than two randomly-paired Galactic disk stars.
-Recent work suggests that, at least for red giant stars, the abundance
+The feasibility of chemical tagging depends on not only on the precision of
+the abundance measurements but also on the intrinsic chemical
+dispersions of stars formed from the same birth clusters,
+and the ability to tell that the chemical abundances of
+two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
+Recent works suggest that, at least for red giant stars, the abundance
 distributions of 20 elements in a handful of open clusters included in the
@@
 precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
+\smoh{I find the following sentence out of place. Doesn't it mean that
+  the intrinsic dispersion is effectively zero rather than ``unconstrained''?}
 In effect, for most elements, the intrinsic dispersions are unconstrained
@@
 clusters and young star associations (see, e.g., \citealt{Oh:2017}).
-However, because they disrupt quickly, the present-day open cluster population
+However, because they disrupt quickly (\smoh{a quantitative time scale?},
+the present-day open cluster population
 is biased towards younger or more massive clusters.
@@
 the same velocity distribution used above. The computed likelihood ratio is
-actually the the ratio of fully marginalized likelihoods,
+actually the ratio of fully marginalized likelihoods,
 $\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
@@
 from the \tgas\ catalog.
+\smoh{To be complete, we should mention $\Delta v_\perp < 10$~km/s cut.
+  Or I think it's also an option to cut on much of these details consistently.}
 The likelihoods appropriately take into account the reported covariances and
 uncertainties between the astrometric measurements in the \tgas\ catalog and
-naturally handle geometric or projection effects in comparing spherical
+naturally handle geometric or projection effects in comparing spherical \smoh{?}
 velocity components for star pairs separated by large angles on the sky.
@@
 observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
-\project{Modspec} was set up in long-slit mode with a 1'' slit, a $600~{\rm
+\project{Modspec} was set up in long-slit mode with a 1\arcsec\ slit, a $600~{\rm
 line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
@@
 between $30$--$120~{\rm s}$ (15th and 85th percentiles).
-To maximize the number of observed pairs, we therefore chose to not take
-comparison lamp spectra at each pointing, which adds an overhead of $\approx
+To maximize the number of observed pairs, we chose not to take
+comparison lamp spectra at each pointing as it adds an overhead of $\approx
 90~{\rm s}$ per observation.
-To determine rough nightly wavelength solutions, spectra were obtained using
-Hg-Ne and Ne comparison lamps at the beginning and end of each night, and
-atmospheric emission lines were used to correct the wavelength solution on a
+To determine rough nightly wavelength solutions, calibration spectra were obtained using
+Hg-Ne and Ne comparison lamps at the beginning and end of each night.
+Additionally, atmospheric emission lines were used to correct the wavelength solution on a
 per-object basis.
@@
 calibration targets.
+\smoh{How about ``The calibration targets serve two purposes. First, ...
+  just so that it is more explicit, plain, and direct, and also ties in
+  with how we chose the calibration targets?}
 The calibration targets have previously measured radial velocities and chemical
@@
 pipeline written in \python, described below.
+\smoh{How the calibration targets were chosen is missing here.}
 
+\smoh{Throughout this paragraph, consider changing e.g.,
+  ``16 pixel rows'' to ``16 pixels in wavelength direction''?
+  Or explicitly mention somewhere (perhaps at the start of this section)
+  that x (along rows) is spatial and y (along columns) is spectral direction?}
 For each source or comparison lamp observation, the two-dimensional image is
@@
 and the curvature of the wavelength solution over this region is negligible.
-For source observations, the 2D spectra are extracted to 1D using an empirical
+For source observations, 1D spectra are extracted using an empirical
 estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
@@
 wavelength-calibrated source spectra of comoving star pair members normalized
-to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by index for visualization.
+to 1 at $\lambda = 5500~{\rm \AA}$ and \smoh{shifted by index (confusing)} for visualization.
 As suggested by the features in these 10 spectra, there is a mix of
@@
 3D velocity vectors drawn from a prior velocity distribution.
-We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
-a mixture of three Gaussian velocity components for the prior, meant to
-represent the thin disk, thick disk, and halo velocity distributions relative to
-the sun.
+As with previous work, we set the small dispersion to the orbital velocity of a
+binary with the separation of the pair, and use a mixture of three Gaussian
+velocity components for the prior, meant to represent the thin disk, thick disk,
+and halo velocity distributions relative to the sun.
 
@@
   \caption{%
-    Point estimate of the 3D velocity difference between two stars in a comoving
-    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
-    log-likelihood-ratio computed when including radial velocity data.
-    Error bars are shown for a few randomly-chosen, representative pairs and
-    represent the 15th and 85th percentiles of velocity differences computed
-    from samples from the error distributions of parallax, proper motion
-    components, and radial velocity.
-    The likelihood ratio increases (i.e. the belief that the stars are comoving
-    becomes stronger) when the 3D velocity difference is small.
+    FML
     \label{fig:llr}}
@@
 If the new likelihood ratio is larger than that computed from astrometry alone,
-the added radial velocity information supports the belief that the two stars are
-truly comoving in 3-space.
+the added radial velocity information likely supports the belief that the two
+stars are truly comoving in 3-space.
 Conversely, if the new likelihood ratio is smaller, the radial velocity data
@@
 proper-motion-only log-FML-ratio for a given pair of stars as
-\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$---128 of
+\llrold\, and the log-FML-ratio including the radial velocity data as
+\llrnew.
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 128 of
 the 311 comoving pairs satisfy this constraint.
 \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
-plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
+plotted against the change in log-likelihood-ratio for the two stars in all
 observed pairs.
@@
 samples.
-We plot the median of the resulting distribution of velocity-differences as
+We plot the median of the resulting distribution over velocity-difference as
 dark markers (circles).
@@
 
+\todo{Discuss contamination rate vs. separation for ratio cut of 6?}
 At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
 genuinely comoving; interestingly, the separation distribution of the genuine
-pairs appears to be uniform or possibly growing in number with increasing
-separation.
+pairs appears to be at least uniform or possibly growing in number as the
+separation increases.
 This strongly suggests the existence of a population of unbound wide binaries or
@@
   \caption{%
-    Point estimate of the 3D velocity difference between two stars in a comoving
-    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
-    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
-    The distances used in the point estimates are computed using a
-    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
+    FML
     \label{fig:dx-dv}}
@@
   \caption{%
-    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
-    the stars in all observed comoving star pairs (gray), and only those
-    identified as genuinely comoving (black). Note that the contamination rate
-    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
-    significant number of genuine comoving pairs at these very wide separations.
-    \emph{Right}: Distribution of sky separations of all comoving star pairs
-    (gray) and only those identified as genuinely comoving (black).
+    FML
     \label{fig:separation}}
@@
 
+\todo{separation distribution - compare with Shaya and Olling sample!}
+\todo{10 overlap with Shaya and Olling}
+\todo{RAVE pairs - both observed}
+
 \subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
 
-\todo{Highlight the G subgiant-A pair}
-\todo{Highlight a binary sequence-main sequence pair}
-\todo{Highlight the largest mag difference MS pair}
-
-\subsection{Combined with RAVE}\label{sec:rave}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
-  \end{center}
-  \caption{%
-    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
-    the stars in all observed comoving star pairs (gray), and only those
-    identified as genuinely comoving (black). Note that the contamination rate
-    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
-    significant number of genuine comoving pairs at these very wide separations.
-    \emph{Right}: Distribution of sky separations of all comoving star pairs
-    (gray) and only those identified as genuinely comoving (black).
-    \label{fig:separation}}
-\end{figure}
-
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
-%   \end{center}
-%   \caption{%
-%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
-%     the density of all \tgas\ sources within 200 pc; similar to
-%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
-%     pairs.
-%     Several interesting pairs that include main-sequence and giant component
-%     stars are highlighted in the text
-%     (\sectionname~\ref{sec:interesting-pairs}).
-%     \label{fig:genuine-cmd}}
-% \end{figure}
+\todo{Highlight the MS-Giant pairs}
+\todo{Highlight the binary sequence-main sequence pairs}
 
@@
 
-Abundance of large-separation pairs
-
-Need precise chemical information
-
-But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?
-
-Going forward, will have to use hierarchical mixture models to study population of singles in conjunction
-
-Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
+Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
 Also tests of chemical enrichment due to planetary systems
@@
 
+\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
+
 \todo{Do I want to do isochrone fitting for all of them?}
 
-Abundances - future work, Coronado in prep.!
+Abundances!
 
-Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>
+- Stellar rotation of evolved-main sequence
+- Full catalog should contain Shaya and Olling, cross-match to RV surveys
 
 \section{Conclusions}
-
-Low-resolution spectroscopic follow-up of comoving pairs.
-
-Using relative RV's, confirm a sample of very wide separation comoving star pairs.
-
-Either wide binaries, unbound binaries, or dissolving clusters
-
-Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution
-
-Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.
 

### 2017-07-11 paper/comoving-followup.tex

@@
   \caption{%
-    FML
+    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
+    the stars in all observed comoving star pairs (gray), and only those
+    identified as genuinely comoving (black). Note that the contamination rate
+    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
+    significant number of genuine comoving pairs at these very wide separations.
+    \emph{Right}: Distribution of chord lengths (tangential separations) for all
+    comoving star pairs (gray) and only those identified as genuinely comoving
+    (black).
+    Noise in the parallax measurements makes the 3D separation distribution
+    appear roughly uniform.
+    Even the distribution in chord length appears to be approximately linear for
+    large separations ($\gtrsim 0.1~\pc$).
     \label{fig:separation}}
@@
 
-\todo{separation distribution - compare with Shaya and Olling sample!}
-\todo{10 overlap with Shaya and Olling}
-\todo{RAVE pairs - both observed}
-
 \subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
 
-\todo{Highlight the MS-Giant pairs}
-\todo{Highlight the binary sequence-main sequence pairs}
+We naively expect that comoving, coeval stars should lie along lines parallel to
+the main sequence in color-magnitude space. The majority of our star pairs
+appear to follow this expectation, with the exception of a few visible outliers.
+For example, there appear to be at least a few main-sequence--giant-branch
+pairs, and a few pairs that appear to cross the main-sequence.
+We highlight a few of these pairs below.
+
+\subsubsection{HD 95088 \& HD 92278}
+
+\todo{Highlight the G subgiant-A pair}
+
+\subsubsection{HD 151090 \& HIP 81988}
+
+\todo{G-subgiant - K-dwarf - well known high proper-motion comoving pair}
+
+\subsubsection{HD 54318 \& TYC 3792-658-1}
+
+\todo{Highlight the largest mag difference MS pair - K-dwarf - B-star}
+
+\subsection{Combined with RAVE}\label{sec:rave}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
+  \end{center}
+  \caption{%
+    Same as \figurename~\ref{fig:separation}, but for our sample of genuine
+    comoving stars combined with all genuine comoving stars identified from the
+    RAVE overlap.
+    \label{fig:separation-RAVE}}
+\end{figure}
+
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
+%   \end{center}
+%   \caption{%
+%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
+%     the density of all \tgas\ sources within 200 pc; similar to
+%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
+%     pairs.
+%     Several interesting pairs that include main-sequence and giant component
+%     stars are highlighted in the text
+%     (\sectionname~\ref{sec:interesting-pairs}).
+%     \label{fig:genuine-cmd}}
+% \end{figure}
 
@@
 
-\todo{Cross-match with exoplanets catalog http://exoplanets.org/csv-files/exoplanets.csv}
-
 \todo{Do I want to do isochrone fitting for all of them?}
 
-Abundances!
+Abundances - future work, Coronado in prep.! But: throw shade on abundances
+because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
+2017b
 
-- Stellar rotation of evolved-main sequence
-- Full catalog should contain Shaya and Olling, cross-match to RV surveys
+Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>
 
 \section{Conclusions}
+
+Low-resolution spectroscopic follow-up of comoving pairs.
+
+Using relative RV's, confirm a sample of very wide separation comoving star pairs.
+
+Either wide binaries, unbound binaries, or dissolving clusters
+
+Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution
+
+Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.
 

### 2017-07-11 paper/comoving-followup.tex
smoh comments

@@
 \newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
+\newcommand{\npairsobs}{311}
+\newcommand{\ncomoving}{127}
 
@@
 % Context
-Widely-separated ($\gtrsim X$~pc), coeval stars are powerful tools for testing star-formation
-model predictions, improving stellar atmosphere models, and inferring the
-mass distribution of the Milky Way.
-\smoh{The following sentence seems to be in the air while I get what you mean.}
-They vastly outnumber the Galactic open clusters, but are still expected to
-share a common origin and chemistry.
-\smoh{How about sticking with the terminology ("comoving (star) pair") we want settled?}
-The kinematic properties of comoving pairs including bound wide binaries
-serve as potentiometers for the smooth tidal field and perturbation spectrum of
-the Milky Way.
-\smoh{If you are referring to Shaya \& Olling, how about adding 'using HIPPARCOS'?}
+Widely-separated ($\gtrsim 0.01$~pc), coeval stars are powerful tools for
+testing star-formation model predictions, improving stellar atmosphere models,
+and inferring the mass distribution of the Milky Way.
+They likely outnumber the Galactic open clusters by a large margin, but are
+still expected to share a common origin and chemistry.
+The kinematic properties of comoving pairs---including bound wide binaries---
+serve as potentiometers for the smooth tidal field and mass perturbation
+spectrum of the Milky Way.
 Thus far, a few hundred genuine comoving star pairs with separations as large
-as $8~\pc$ have been discovered in the solar neighborhood.
+as $8~\pc$ have been discovered in the solar neighborhood using HIPPARCOS.
 % Aims
@@
 Without radial velocity measurements for the vast majority of the pairs, the
-\smoh{I prefer "false-positive rate" to "contamination rate" because it conveys
-the meaning more directly.}
-contamination rate at large separations is significant ($\sim 50\%$).
+false-positive rate at large separations is significant ($\sim 50\%$).
 % Methods
-\smoh{I think some details of telescope/instrument or just R~2000 should be mentioned here.}
-Here, we present results from our own low-resolution radial-velocity survey of a sample of the
-candidate pairs to validate wide-separation comoving star pairs.
+Here, we present results from our own low-resolution radial-velocity survey of a
+sample of the candidate pairs to validate wide-separation comoving star pairs.
 % Results
-Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
-separations as large as $\approx 10~\pc$, our original search limit.
+Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
+pairs with separations as large as $\approx 10~\pc$, our original search limit.
 % Conclusions
@@
 with having intrinsic dispersions smaller than the abundance measurement
-precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
-\smoh{I find the following sentence out of place. Doesn't it mean that
-  the intrinsic dispersion is effectively zero rather than ``unconstrained''?}
-In effect, for most elements, the intrinsic dispersions are unconstrained
-because of measurement noise, so the true intrinsic dispersions are not known.
+precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
+intrinsic dispersions are not known.
 This demonstration of chemical uniformity in open clusters is promising for
@@
 distance for each star in a given pair.
-After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
-compute the likelihood ratio for all stars within $10~\pc$ of each target star
+After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
+we compute the likelihood ratio for all stars within $10~\pc$ and with a
+difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each target star
 from the \tgas\ catalog.
-\smoh{To be complete, we should mention $\Delta v_\perp < 10$~km/s cut.
-  Or I think it's also an option to cut on much of these details consistently.}
 The likelihoods appropriately take into account the reported covariances and
 uncertainties between the astrometric measurements in the \tgas\ catalog and
-naturally handle geometric or projection effects in comparing spherical \smoh{?}
-velocity components for star pairs separated by large angles on the sky.
+naturally handle geometric or projection effects in comparing spherical
+velocities for star pairs separated by large angles on the sky.
 After applying a conservative cut on the likelihood ratio motivated by computing
@@
 calibration targets.
-\smoh{How about ``The calibration targets serve two purposes. First, ...
-  just so that it is more explicit, plain, and direct, and also ties in
-  with how we chose the calibration targets?}
-The calibration targets have previously measured radial velocities and chemical
-abundances, which will be used for validation of our reduction pipeline and in
-future work to measure metallicities and stellar parameters for each observed
-source using \project{The Cannon} (\citealt{Ness:2015}).
+For the purposes of this work, all of the calibration targets have previously measured radial velocities and were selected from high-resolution spectroscopic
+surveys \citealt{Bensby:2014,Pinsonneault:2014}; these stars are used to
+validate our radial velocity measurements.
+These stars also have measured chemical abundances, and will be used in future
+work as a training set for \project{The Cannon} (\citealt{Ness:2015}) to infer
+metallicities and stellar parameters for our observed sources.
 The spectra were reduced and calibrated using a custom, publicly-available
 pipeline written in \python, described below.
-\smoh{How the calibration targets were chosen is missing here.}
 
-\smoh{Throughout this paragraph, consider changing e.g.,
-  ``16 pixel rows'' to ``16 pixels in wavelength direction''?
-  Or explicitly mention somewhere (perhaps at the start of this section)
-  that x (along rows) is spatial and y (along columns) is spectral direction?}
 For each source or comparison lamp observation, the two-dimensional image is
@@
 implemented in \package{ccdproc} (\citealt{Craig:2015}).
+The 2D spectra are oriented such that the spatial direction is along a row of
+pixels, and the dispersion direction is along a column of pixels.
 For comparison lamp observations, the central 100 pixel columns are extracted
@@
 wavelength-calibrated source spectra of comoving star pair members normalized
-to 1 at $\lambda = 5500~{\rm \AA}$ and \smoh{shifted by index (confusing)} for visualization.
+to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by an arbitrary offset for
+visualization.
 As suggested by the features in these 10 spectra, there is a mix of

### 2017-07-11 paper/comoving-followup.tex
intro

@@
 % Context
-Widely-separated ($\gtrsim 0.01$~pc), coeval stars are powerful tools for
+Widely-separated ($\gtrsim 0.1$~pc), coeval stars are powerful tools for
 testing star-formation model predictions, improving stellar atmosphere models,
@@
 
-The majority of stars are thought to form in dynamically cold, chemically
-homogeneous clusters, relative to the kinematic and chemical
-dispersions observed amongst typical field stars in the Galactic disk.
-Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
-progenitor gas cloud disperses quickly (cite), it is typically assumed that
-stars formed in the same birth cluster have the same initial chemistry
-(e.g., cite + \citealt{Feng:2014}) and same ages.
-These birth clusters are typically born un- or loosely-bound and therefore
-quickly succumb to the tidal field of the Milky Way.
-The initial kinematic associations between stars from the same cluster are then
-erased as the constituent stars mix into the general population of disk stars.
-However, the stars do preserve their initial associations in age, chemical
-abundance patterns, and dynamical invariants (e.g., orbital actions).
-It has therefore been long presumed that ``chemical tagging'' and chemodynamical
-models of the Galactic disk will allow for inferences about the formation and
-evolution of the Milky Way (e.g., \citealt{todo}).
-That is, the chemical environment of a birth cluster is likely imprinted on its
-stellar children and preserved so that their initial association may be inferred
-with precisely measured abundance patterns (of $\gtrsim 10$ species).
-
-The feasibility of chemical tagging depends on not only on the precision of
-the abundance measurements but also on the intrinsic chemical
-dispersions of stars formed from the same birth clusters,
-and the ability to tell that the chemical abundances of
-two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
-Recent works suggest that, at least for red giant stars, the abundance
-distributions of 20 elements in a handful of open clusters included in the
-\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
-with having intrinsic dispersions smaller than the abundance measurement
-precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
-intrinsic dispersions are not known.
-This demonstration of chemical uniformity in open clusters is promising for
-chemical tagging.
-However, it was also recently shown that $\approx 0.3\%$ of random pairings of
-Galactic field red giant stars have abundance patterns that are
-indistinguishable at the same level (``doppelgangers''), even with 20 chemical
-species (\citealt{Ness:2017}).
-
-Are these properties of open clusters universal, or limited to the few clusters
-presently studied?
-Do the chemical properties of open clusters depend on the star formation rate or
-efficiency, which may evolve with time?
-Ideal tests of the chemical homogeneity assumptions would use many more open
-clusters with varied cluster masses and member stars that span a large range in
-stellar masses and evolutionary stages.
-Precise kinematic and color data from the \gaia\ mission
-(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
-existing membership catalogs and will lead to discoveries of many more open
-clusters and young star associations (see, e.g., \citealt{Oh:2017}).
-However, because they disrupt quickly (\smoh{a quantitative time scale?},
-the present-day open cluster population
-is biased towards younger or more massive clusters.
-
-An alternate tracer of the chemical properties mentioned above are
-widely-separated ($\gtrsim 0.01~\pc$), coeval pairs or multiplets of stars.
-These tracers include both bound and unbound wide-separation stellar systems; we
-refer to these systems collectively as ``comoving stars'' because, at large
-separations with present-day and near-future velocity uncertainties, it is
-typically not possible to determine whether these systems are truly bound.
-Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
-dissolving or disrupted open clusters.
-In any of these cases, it is likely that the stars formed from the same region
-of a birth cloud, yet are separated enough that there has not been any
-cross-pollution in the event of strong stellar winds or stellar evolution.
-
-For chemical studies, open clusters have the advantage of having many stars per
-cluster, but there are many fewer clusters than comoving stars.
-Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
-surveys (cite rave, etc.), several hundred such systems are already known
-(\citealt{Shaya:2011}).
-In addition, over ten-thousand \emph{candidate} comoving stars have been
-discovered from astrometric data alone, using data from both the
-\project{Hipparcos} mission (\citealt{ESA:1997}---\citealt{Shaya:2011}) and
-\tgas\ catalog (\citealt{Gaia-Collaboration:2016a}---
-\citealt{Oh:2017,Oelkers:2017,Andrews:2017}).
-Radial velocity follow-up is necessary to confirm pairs using these astrometric
-catalogs, though \gaia\ \DR{2} will provide radial velocities for the brightest
-stars ($G < 12$).
-High-resolution spectroscopy is still needed to measure detailed chemical
-abundances of these pairs, which may present interesting surprises; one
-large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
-discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
-high-resolution, high signal-to-noise spectra display large abundance
-differences ($\approx 0.2~\dex$) that appear to have resulted from massive
-accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
-submitted).
-
-In this work, we present a catalog of genuine comoving star pairs discovered
+Widely-separated, coeval pairs or multiplets of stars --- i.e. wide binaries,
+multiplets, or former binaries or multiplets --- are important tracers of many
+processes that formed and continue to shape the Milky Way.
+For example, they have been used to place limits on the mass distribution and
+population of massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to
+study the dynamical outcomes of star formation processes (e.g.,
+\citealt{Reipurth:2012}), and to calibrate stellar models and inferences across
+the main sequence (e.g., \citealt{Brewer:2016}).
+These examples represent a small subset of a much broader range of applications
+(see, e.g., \citealt{Chaname:2007}).
+
+At smaller separations ($\lesssim 0.1~\pc$), comoving pairs of stars are likely
+bound wide binaries and are therefore coeval.
+At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
+tenuously bound stars, recently disrupted binaries, moving groups or dynamical
+features, and the final products of dissolving stellar associations or open
+clusters.
+However, even at separations of $\sim 0.1~\pc$, we typically lack the precision
+to determine the binding energy of a system and therefore can't determine its
+true dynamical state; because of this ambiguity, we refer to both bound and
+unbound systems collectively as ``comoving stars.''
+
+Comoving stars are identified by searching for approximately cospatial stars
+that appear to be comoving either in full-space or in sky-projected velocities.
+Primarily thanks to large astrometric surveys and catalogs
+(\citealt{ESA:1997,Lepine:2005,Gaia-Collaboration:2016}) and radial-velocity
+surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving pairs
+(\citealt{Shaya:2011}) and over 10,000 candidate systems
+(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,Oelkers:2017,
+Andrews:2017}) have already been discovered.
+Radial velocity follow-up is still needed to confirm the majority of these
+pairs, though \gaia\ \DR{2} will provide radial velocities for the brightest
+stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).
+
+It has long been recognized that the observed period or separation distribution
+of binary stars encapsulates information about star formation and dynamical
+processes of the Galaxy (\citealt{Opik:1924}).
+[If the initial distribution power law, break and steepens from perturbations]
+[a bit of theory about wide binaries...predictions for wide binary separations?]
+(Weinberg87)
+
+Using candidate comoving pairs selected based on their proper motions alone,
+several groups have measured the separation distribution of comoving stars at
+angular separations $\Delta \theta \lesssim 1^\circ$ where the false-positive
+rate is expected to be lower (\citealt{Chaname:2004,Lepine:2007,Sesar:2008});
+these are likely predominantly wide binaries.
+In most cases, the angular or projected separation distribution is fit using a
+decreasing power-law, and the power-law index, $\alpha$, is found to be $\alpha
+\approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
+0.65~\pc$ at a distance of $150~\pc$).
+Tentative reports of a steepening or cutoff of the separation distribution at
+larger separations have been claimed (e.g., \citealt{Yoo:2004,Quinn:2009}), but
+these claims are limited by small numbers of confirmed pairs at large
+separations.
+
+While the separation distribution of wide binaries appears to and is expected to
+fall off rapidly with increasing separation, predictions for the separation
+distribution of \emph{comoving stars} are less clear.
+Early work suggested ionized pairs don't matter much...
+[But including tidal field...] (jiang and tremaine)
+[But still haven't included perturbers...]
+[Tentative evidence of this population in (shaya and olling)]
+[Need larger samples of confirmed large-separation pairs]
+
+In this work, we present a catalog of confirmed comoving star pairs discovered
 through radial velocity follow-up of candidate comoving stars selected from
@@
 results.
+
+% The majority of stars are thought to form in dynamically cold, chemically
+% homogeneous clusters, relative to the kinematic and chemical
+% dispersions observed amongst typical field stars in the Galactic disk.
+% Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
+% progenitor gas cloud disperses quickly (cite), it is typically assumed that
+% stars formed in the same birth cluster have the same initial chemistry
+% (e.g., cite + \citealt{Feng:2014}) and same ages.
+% These birth clusters are typically born un- or loosely-bound and therefore
+% quickly succumb to the tidal field of the Milky Way.
+% The initial kinematic associations between stars from the same cluster are then
+% erased as the constituent stars mix into the general population of disk stars.
+% However, the stars do preserve their initial associations in age, chemical
+% abundance patterns, and dynamical invariants (e.g., orbital actions).
+% It has therefore been long presumed that ``chemical tagging'' and chemodynamical
+% models of the Galactic disk will allow for inferences about the formation and
+% evolution of the Milky Way (e.g., \citealt{todo}).
+% That is, the chemical environment of a birth cluster is likely imprinted on its
+% stellar children and preserved so that their initial association may be inferred
+% with precisely measured abundance patterns (of $\gtrsim 10$ species).
+
+% The feasibility of chemical tagging depends on not only on the precision of
+% the abundance measurements but also on the intrinsic chemical
+% dispersions of stars formed from the same birth clusters,
+% and the ability to tell that the chemical abundances of
+% two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
+% Recent works suggest that, at least for red giant stars, the abundance
+% distributions of 20 elements in a handful of open clusters included in the
+% \acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
+% with having intrinsic dispersions smaller than the abundance measurement
+% precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
+% intrinsic dispersions are not known.
+% This demonstration of chemical uniformity in open clusters is promising for
+% chemical tagging.
+% However, it was also recently shown that $\approx 0.3\%$ of random pairings of
+% Galactic field red giant stars have abundance patterns that are
+% indistinguishable at the same level (``doppelgangers''), even with 20 chemical
+% species (\citealt{Ness:2017}).
+
+% Are these properties of open clusters universal, or limited to the few clusters
+% presently studied?
+% Do the chemical properties of open clusters depend on the star formation rate or
+% efficiency, which may evolve with time?
+% Ideal tests of the chemical homogeneity assumptions would use many more open
+% clusters with varied cluster masses and member stars that span a large range in
+% stellar masses and evolutionary stages.
+% Precise kinematic and color data from the \gaia\ mission
+% (\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
+% existing membership catalogs and will lead to discoveries of many more open
+% clusters and young star associations (see, e.g., \citealt{Oh:2017}).
+% However, because they disrupt quickly (\smoh{a quantitative time scale?},
+% the present-day open cluster population
+% is biased towards younger or more massive clusters.
+
+% High-resolution spectroscopy is still needed to measure detailed chemical
+% abundances of these pairs, which may present interesting surprises; one
+% large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
+% discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
+% high-resolution, high signal-to-noise spectra display large abundance
+% differences ($\approx 0.2~\dex$) that appear to have resulted from massive
+% accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
+% submitted).
 

### 2017-07-12 paper/comoving-followup.tex
crushed the intro re-write. fixes #5

@@
 processes that formed and continue to shape the Milky Way.
-For example, they have been used to place limits on the mass distribution and
-population of massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to
-study the dynamical outcomes of star formation processes (e.g.,
-\citealt{Reipurth:2012}), and to calibrate stellar models and inferences across
-the main sequence (e.g., \citealt{Brewer:2016}).
+They have been used to place limits on the mass distribution and population of
+massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to study the
+dynamical outcomes of star formation processes (e.g., \citealt{Reipurth:2012}),
+and to calibrate stellar models and inferences across the main sequence (e.g.,
+\citealt{Brewer:2016}).
 These examples represent a small subset of a much broader range of applications
@@
 At smaller separations ($\lesssim 0.1~\pc$), comoving pairs of stars are likely
-bound wide binaries and are therefore coeval.
+bound wide binaries.
+It has long been recognized that the observed period or separation distribution
+of these low-binding-energy binaries encapsulates information about the
+Galactic mass distribution (\citealt{Opik:1924,Oort:1950,Bahcall:1985}).
+In particular, early simulations of populations of wide binaries showed that
+the expected separation distribution, $p(s)$, for $0.01 < s < 1~\pc$ tends to
+follow a power law, $p(s) \propto s^{-\alpha}$ for $\alpha \approx 1$--$2$,
+with a possible steep break that depends on the details of the population of
+massive perturbers in the Galaxy, i.e. giant molecular clouds (GMCs), black holes, etc. (\citealt{Weinberg:1987}).
+
 At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
-tenuously bound stars, recently disrupted binaries, moving groups or dynamical
-features, and the final products of dissolving stellar associations or open
+tenuously bound stars, recently disrupted binaries, moving groups or resonant
+features, and the final products of dissolving stellar associations or
 clusters.
@@
 true dynamical state; because of this ambiguity, we refer to both bound and
-unbound systems collectively as ``comoving stars.''
-
-Comoving stars are identified by searching for approximately cospatial stars
-that appear to be comoving either in full-space or in sky-projected velocities.
+unbound widely-separated ($\gtrsim 10^{-3}~\pc$) systems collectively as
+``comoving'' systems.
+
 Primarily thanks to large astrometric surveys and catalogs
 (\citealt{ESA:1997,Lepine:2005,Gaia-Collaboration:2016}) and radial-velocity
-surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving pairs
-(\citealt{Shaya:2011}) and over 10,000 candidate systems
-(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,Oelkers:2017,
-Andrews:2017}) have already been discovered.
-Radial velocity follow-up is still needed to confirm the majority of these
-pairs, though \gaia\ \DR{2} will provide radial velocities for the brightest
-stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).
-
-It has long been recognized that the observed period or separation distribution
-of binary stars encapsulates information about star formation and dynamical
-processes of the Galaxy (\citealt{Opik:1924}).
-[If the initial distribution power law, break and steepens from perturbations]
-[a bit of theory about wide binaries...predictions for wide binary separations?]
-(Weinberg87)
-
-Using candidate comoving pairs selected based on their proper motions alone,
-several groups have measured the separation distribution of comoving stars at
-angular separations $\Delta \theta \lesssim 1^\circ$ where the false-positive
-rate is expected to be lower (\citealt{Chaname:2004,Lepine:2007,Sesar:2008});
-these are likely predominantly wide binaries.
+surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving star
+pairs (\citealt{Shaya:2011}) and over 10,000 candidate systems
+(\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
+Oelkers:2017, Andrews:2017}) have been discovered over the last several decades (see also \citealt{Chaname:2007} and references therein).
+These pairs are identified by searching for approximately cospatial stars that
+appear to be comoving either in sky-projected velocities --- we refer to these
+as ``comoving pair candidates'' --- or using full-space velocity information --- we refer to these as ``comoving pairs.''
+Radial velocity follow-up is still needed to confirm the majority of the known
+candidates, though \gaia\ \DR{2} will provide radial velocities for the
+brightest stars ($G < 12$; \citealt{Gaia-Collaboration:2016}).
+
+With large samples of candidate pairs, several groups have measured the
+separation distribution of wide binaries at angular separations $\Delta \theta
+\lesssim 1^\circ$ where the false-positive rate is expected to be lower
+and the samples are likely predominantly wide binaries
+(\citealt{Chaname:2004,Lepine:2007,Sesar:2008}).
 In most cases, the angular or projected separation distribution is fit using a
@@
 \approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
-0.65~\pc$ at a distance of $150~\pc$).
+0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations.
 Tentative reports of a steepening or cutoff of the separation distribution at
@@
 
-While the separation distribution of wide binaries appears to and is expected to
-fall off rapidly with increasing separation, predictions for the separation
-distribution of \emph{comoving stars} are less clear.
-Early work suggested ionized pairs don't matter much...
-[But including tidal field...] (jiang and tremaine)
-[But still haven't included perturbers...]
-[Tentative evidence of this population in (shaya and olling)]
-[Need larger samples of confirmed large-separation pairs]
-
+Most measurements of the separation distribution have neglected the possible
+existence of larger separation ($\gtrsim 1~\pc$) unbound or ionized pairs.
+Early theoretical work either explicitly ignored these stars (e.g.,
+\citealt{Weinberg:1987}), or suggested that unbound stars would drift and no
+longer remain spatially coherent within one Galactic orbit (e.g., see Appendix
+in \citealt{Yoo:2004}).
+More recent simulations of the disruption and evolution of wide binary
+populations include an approximation of the Galactic tidal field and instead
+predict that ionized binaries could appear as a second peak or component in the
+separation distribution of comoving stars at much larger separations (10--
+$1000~\pc$; \citealt{Jiang:2010}).
+Though these simulations don't include perturbers like GMCs or black holes,
+this at least indicates that a simple, monotonic separation distribution may
+be too simple of a model for the separation distribution of comoving stars in
+our Galaxy.
+
+Indeed, observational work that followed appears to find tentative evidence for
+the existence of a second peak in the separation distribution of
+radial-velocity-confirmed comoving stars (\citealt{Shaya:2011}).
+Using a ... todo: describe their paper
+
+To verify the existence of this unbound population of very wide-separation
+comoving pairs, we will need to identify a large sample of confirmed comoving
+stars.
+This will already be possible for the brightest stars with \gaia\ \DR{2}, but
+obtaining even larger samples will require significant efforts to obtain radial
+velocities for candidate comoving pairs.
 In this work, we present a catalog of confirmed comoving star pairs discovered
-through radial velocity follow-up of candidate comoving stars selected from
-inferences using their astrometry alone (\citealt{Oh:2017}).
-This work serves as a pilot study for data expected in upcoming data releases
-from the \gaia\ mission.
+through low-resolution spectroscopic follow-up of candidate comoving stars
+selected from inferences using astrometric measurements from the \tgas\ catalog
+alone (\citealt{Oh:2017}).
+
 We describe the data, observations, and spectroscopic reduction in

### 2017-07-12 paper/comoving-followup.tex
shaya olling text

@@
 Thus far, a few hundred genuine comoving star pairs with separations as large
-as $8~\pc$ have been discovered in the solar neighborhood using HIPPARCOS.
+as $8~\pc$ have been discovered in the solar neighborhood using
+\project{Hipparcos}.
 % Aims
@@
 massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to study the
-dynamical outcomes of star formation processes (e.g., \citealt{Reipurth:2012}),
+physics of star formation (e.g., \citealt{Parker:2009,Reipurth:2012}),
 and to calibrate stellar models and inferences across the main sequence (e.g.,
@@
 our Galaxy.
+In addition, the distribution of unbound but comoving stars should be very
+sensitive to diffusive processes that determine the rate at which the pairs
+separate.
 
@@
 radial-velocity-confirmed comoving stars (\citealt{Shaya:2011}).
-Using a ... todo: describe their paper
+A key aspect of this work was to move from searching for common proper motion
+pairs to performing a statistical inference of the difference in full-space
+velocity vector for pairs of stars.
+At separations of $1$--$10~\pc$, star pairs in the local solar neighborhood
+subtend large angles on the sky; therefore, even if the two component stars were
+moving with the same 3D velocity, their proper motions and radial velocities
+would be different because of geometric effects.
+By compiling radial velocities from the literature for \project{Hipparcos}
+stars, they applied this method to search for all comoving star pairs with
+separations larger than $0.01~\pc$ within $100~\pc$ of the sun.
+They find $\sim 200$ high probability comoving pairs with a separation
+distribution that falls of until $\sim 1~\pc$, where the number of pairs then
+begins rising with separations.
 
 To verify the existence of this unbound population of very wide-separation
-comoving pairs, we will need to identify a large sample of confirmed comoving
-stars.
+comoving pairs, we need to identify a large sample of confirmed comoving stars.
 This will already be possible for the brightest stars with \gaia\ \DR{2}, but

### 2017-07-12 paper/comoving-followup.tex
discussion

@@
 results.
-
-% The majority of stars are thought to form in dynamically cold, chemically
-% homogeneous clusters, relative to the kinematic and chemical
-% dispersions observed amongst typical field stars in the Galactic disk.
-% Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
-% progenitor gas cloud disperses quickly (cite), it is typically assumed that
-% stars formed in the same birth cluster have the same initial chemistry
-% (e.g., cite + \citealt{Feng:2014}) and same ages.
-% These birth clusters are typically born un- or loosely-bound and therefore
-% quickly succumb to the tidal field of the Milky Way.
-% The initial kinematic associations between stars from the same cluster are then
-% erased as the constituent stars mix into the general population of disk stars.
-% However, the stars do preserve their initial associations in age, chemical
-% abundance patterns, and dynamical invariants (e.g., orbital actions).
-% It has therefore been long presumed that ``chemical tagging'' and chemodynamical
-% models of the Galactic disk will allow for inferences about the formation and
-% evolution of the Milky Way (e.g., \citealt{todo}).
-% That is, the chemical environment of a birth cluster is likely imprinted on its
-% stellar children and preserved so that their initial association may be inferred
-% with precisely measured abundance patterns (of $\gtrsim 10$ species).
-
-% The feasibility of chemical tagging depends on not only on the precision of
-% the abundance measurements but also on the intrinsic chemical
-% dispersions of stars formed from the same birth clusters,
-% and the ability to tell that the chemical abundances of
-% two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
-% Recent works suggest that, at least for red giant stars, the abundance
-% distributions of 20 elements in a handful of open clusters included in the
-% \acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
-% with having intrinsic dispersions smaller than the abundance measurement
-% precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
-% intrinsic dispersions are not known.
-% This demonstration of chemical uniformity in open clusters is promising for
-% chemical tagging.
-% However, it was also recently shown that $\approx 0.3\%$ of random pairings of
-% Galactic field red giant stars have abundance patterns that are
-% indistinguishable at the same level (``doppelgangers''), even with 20 chemical
-% species (\citealt{Ness:2017}).
-
-% Are these properties of open clusters universal, or limited to the few clusters
-% presently studied?
-% Do the chemical properties of open clusters depend on the star formation rate or
-% efficiency, which may evolve with time?
-% Ideal tests of the chemical homogeneity assumptions would use many more open
-% clusters with varied cluster masses and member stars that span a large range in
-% stellar masses and evolutionary stages.
-% Precise kinematic and color data from the \gaia\ mission
-% (\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
-% existing membership catalogs and will lead to discoveries of many more open
-% clusters and young star associations (see, e.g., \citealt{Oh:2017}).
-% However, because they disrupt quickly (\smoh{a quantitative time scale?},
-% the present-day open cluster population
-% is biased towards younger or more massive clusters.
-
-% High-resolution spectroscopy is still needed to measure detailed chemical
-% abundances of these pairs, which may present interesting surprises; one
-% large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
-% discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
-% high-resolution, high signal-to-noise spectra display large abundance
-% differences ($\approx 0.2~\dex$) that appear to have resulted from massive
-% accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
-% submitted).
 
@@
 \llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$---128 of
-the 311 comoving pairs satisfy this constraint.
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
+\ncomoving of the \npairsobs\ comoving pairs satisfy this constraint.
 \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
@@
     stars are highlighted in the text
-    (\sectionname~\ref{sec:interesting-pairs}).
+    (\sectionname~\ref{sec:interesting-pairs} and
+    \figurename~\ref{fig:highlighted-cmd}).
     \label{fig:genuine-cmd}}
@@
 pairs, and a few pairs that appear to cross the main-sequence.
-We highlight a few of these pairs below.
-
-\subsubsection{HD 95088 \& HD 92278}
-
-\todo{Highlight the G subgiant-A pair}
-
-\subsubsection{HD 151090 \& HIP 81988}
-
-\todo{G-subgiant - K-dwarf - well known high proper-motion comoving pair}
-
-\subsubsection{HD 54318 \& TYC 3792-658-1}
-
-\todo{Highlight the largest mag difference MS pair - K-dwarf - B-star}
+We highlight a few of these pairs below and in
+\figurename~\ref{fig:highlighted-cmd}.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{highlighted_cmd.pdf}
+  \end{center}
+  \caption{%
+    Same as \figurename~\ref{fig:genuine-cmd}, but only three highlighted
+    comoving star pairs are shown.
+    \label{fig:highlighted-cmd}}
+\end{figure}
+
+\subsubsection{HD 95088 -- HD 92278}
+
+HD 95088 HD 92278
+Angular sep.: 4.49 deg
+Chord length: 8.66 pc
+3D sep.: 9.53 pc
+parallaxes: 9.00 +/- 0.50, 9.32 +/- 0.51
+
+This pair is separated by $8.7~\pc$ in chord length ($4.5^\circ$ on the sky) and
+have consistent distances within their uncertainties of $\approx 110~\pc$.
+From visual classification of the stars' location in the color--magnitude
+diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
+comprised of an A-type dwarf and a slightly less massive G-type subgiant.
+If this is indeed the case, the two stars could therefore not be coeval, as the
+more massive A star has not begun its post-main-sequence evolution.
+One possible explanation is that this pair is a false-positive in three
+dimensions, in which case they are unassociated but just appear to be comoving
+with small relative velocity at present day.
+
+Other possibilities stem from the possibility that the comoving pair is the
+outcome of an interaction between more stars.
+For example, an interaction between two wide binaries could result in an
+exchange of members, a hardening of one resulting binary, and a widening of the
+other binary (see, e.g., \citealt{Nathan??}).
+Or, the initial system could have been an unstable three-body system that
+resulted in a tight, inner binary and a wide outer binary;
+if the inner binary were close enough, it could have merged and formed a blue
+straggler star (e.g., \citealt{??}), leaving the wide binary untouched.
+
+These explanations are all speculative at the moment; high-resolution
+spectroscopy of these stars --- and precise measurements of the stellar
+parameters and chemical abundances --- may help to distinguish these
+possibilities.
+
+\subsubsection{HD 79247 -- TYC 3807-1439-1}
+
+This pair is separated by $XX~\pc$ in chord length ($XX^\circ$ on the sky) and
+have consistent distances within their uncertainties of $\approx XX~\pc$.
+From visual classification of the stars' location in the color--magnitude
+diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
+comprised of a G-type subgiant and a G-type dwarf.
+
+\subsubsection{HD 54318 -- TYC 3792-658-1}
+
+TYC 3792-658-1 HD 54318
+Angular sep.: 2.51 deg
+Chord length: 8.45 pc
+3D sep.: 8.73 pc
+parallaxes: 5.20 +/- 0.23, 5.30 +/- 0.32
+
+This pair is separated by $8.5~\pc$ in chord length ($2.5^\circ$ on the sky) and
+have consistent distances within their uncertainties of $\approx 190~\pc$.
+
+From visual classification of the stars' location in the color--magnitude
+diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
+comprised of a late-A-/early-B-type dwarf and a G-type dwarf.
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
-  \end{center}
-  \caption{%
-    Same as \figurename~\ref{fig:separation}, but for our sample of genuine
-    comoving stars combined with all genuine comoving stars identified from the
-    RAVE overlap.
+    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
+  \end{center}
+  \caption{%
     \label{fig:separation-RAVE}}
@@
 
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
-%   \end{center}
-%   \caption{%
-%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
-%     the density of all \tgas\ sources within 200 pc; similar to
-%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
-%     pairs.
-%     Several interesting pairs that include main-sequence and giant component
-%     stars are highlighted in the text
-%     (\sectionname~\ref{sec:interesting-pairs}).
-%     \label{fig:genuine-cmd}}
-% \end{figure}
+Caution about selection criteria, false-positives at large separation...
 
@@
 
-\todo{Do I want to do isochrone fitting for all of them?}
-
 Abundances - future work, Coronado in prep.! But: throw shade on abundances
@@
 
+% The majority of stars are thought to form in dynamically cold, chemically
+% homogeneous clusters, relative to the kinematic and chemical
+% dispersions observed amongst typical field stars in the Galactic disk.
+% Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
+% progenitor gas cloud disperses quickly (cite), it is typically assumed that
+% stars formed in the same birth cluster have the same initial chemistry
+% (e.g., cite + \citealt{Feng:2014}) and same ages.
+% These birth clusters are typically born un- or loosely-bound and therefore
+% quickly succumb to the tidal field of the Milky Way.
+% The initial kinematic associations between stars from the same cluster are then
+% erased as the constituent stars mix into the general population of disk stars.
+% However, the stars do preserve their initial associations in age, chemical
+% abundance patterns, and dynamical invariants (e.g., orbital actions).
+% It has therefore been long presumed that ``chemical tagging'' and chemodynamical
+% models of the Galactic disk will allow for inferences about the formation and
+% evolution of the Milky Way (e.g., \citealt{todo}).
+% That is, the chemical environment of a birth cluster is likely imprinted on its
+% stellar children and preserved so that their initial association may be inferred
+% with precisely measured abundance patterns (of $\gtrsim 10$ species).
+
+% The feasibility of chemical tagging depends on not only on the precision of
+% the abundance measurements but also on the intrinsic chemical
+% dispersions of stars formed from the same birth clusters,
+% and the ability to tell that the chemical abundances of
+% two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
+% Recent works suggest that, at least for red giant stars, the abundance
+% distributions of 20 elements in a handful of open clusters included in the
+% \acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
+% with having intrinsic dispersions smaller than the abundance measurement
+% precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}), though the true
+% intrinsic dispersions are not known.
+% This demonstration of chemical uniformity in open clusters is promising for
+% chemical tagging.
+% However, it was also recently shown that $\approx 0.3\%$ of random pairings of
+% Galactic field red giant stars have abundance patterns that are
+% indistinguishable at the same level (``doppelgangers''), even with 20 chemical
+% species (\citealt{Ness:2017}).
+
+% Are these properties of open clusters universal, or limited to the few clusters
+% presently studied?
+% Do the chemical properties of open clusters depend on the star formation rate or
+% efficiency, which may evolve with time?
+% Ideal tests of the chemical homogeneity assumptions would use many more open
+% clusters with varied cluster masses and member stars that span a large range in
+% stellar masses and evolutionary stages.
+% Precise kinematic and color data from the \gaia\ mission
+% (\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
+% existing membership catalogs and will lead to discoveries of many more open
+% clusters and young star associations (see, e.g., \citealt{Oh:2017}).
+% However, because they disrupt quickly (\smoh{a quantitative time scale?},
+% the present-day open cluster population
+% is biased towards younger or more massive clusters.
+
+% High-resolution spectroscopy is still needed to measure detailed chemical
+% abundances of these pairs, which may present interesting surprises; one
+% large-separation ($\approx 0.6~\pc$) comoving pair of solar-type stars
+% discovered in previous work (\citealt{Oh:2017}) with serendipitously observed
+% high-resolution, high signal-to-noise spectra display large abundance
+% differences ($\approx 0.2~\dex$) that appear to have resulted from massive
+% accretion of planetary (rocky) material onto one of the two stars (Oh et al.,
+% submitted).
+
 \section{Conclusions}
@@
 \acknowledgements
+
+Dave Zurek, Keith, Ruth, Kelle
 John Brewer (Yale),

### 2017-07-12 paper/figures/fit_isochrones.py
figure tweaks

@@ -1,0 +1,174 @@
+"""
+Fit isochrones to the stars in the 3 highlighted pairs
+
+"""
+
+# Standard library
+from collections import OrderedDict
+from os import path
+import os
+
+# Third-party
+from astropy.table import Table
+import matplotlib.pyplot as plt
+import numpy as np
+import emcee
+from tqdm import tqdm
+
+# Package
+from comoving_rv.log import logger
+from comoving_rv.db import Session, Base, db_connect
+from comoving_rv.db.model import (Run, Observation, TGASSource, SimbadInfo,
+                                  GroupToObservations, SpectralLineInfo,
+                                  SpectralLineMeasurement, RVMeasurement)
+
+from isochrones import StarModel
+# from isochrones.mist import MIST_Isochrone
+# iso = MIST_Isochrone() # interpolation issues with MIST isochrones
+from isochrones.dartmouth import Dartmouth_Isochrone
+iso = Dartmouth_Isochrone()
+
+def main():
+    # TODO: bad, hard-coded...
+    # base_path = '/Volumes/ProjectData/gaia-comoving-followup/'
+    base_path = '../../data/'
+    db_path = path.join(base_path, 'db.sqlite')
+    engine = db_connect(db_path)
+    session = Session()
+
+    chain_path = path.abspath('./isochrone_chains')
+    os.makedirs(chain_path, exist_ok=True)
+
+    # Check out the bottom of "Color-magnitude diagram.ipynb":
+    interesting_group_ids = [1500, 1229, 974]
+
+    all_photometry = OrderedDict([
+        ('1500-8455', OrderedDict([('J', (6.8379998, 0.021)),
+                                   ('H', (6.4640002, 0.017000001)),
+                                   ('K', (6.3369999, 0.017999999)),
+                                   ('W1', (6.2950001, 0.093000002)),
+                                   ('W2', (6.2490001, 0.026000001)),
+                                   ('W3', (6.3330002, 0.015)),
+                                   ('B', (9.5950003, 0.022)),
+                                   ('V', (8.5120001, 0.014))])),
+        ('1500-1804', OrderedDict([('J', (6.9039998, 0.041000001)),
+                                   ('H', (6.8559999, 0.027000001)),
+                                   ('K', (6.7989998, 0.017000001)),
+                                   ('W1', (6.803, 0.064999998)),
+                                   ('W2', (6.7600002, 0.018999999)),
+                                   ('W3', (6.8270001, 0.016000001)),
+                                   ('B', (7.4980001, 0.015)),
[…]

### 2017-07-13 paper/comoving-followup.tex
conclusions - BOOM

@@
 \usepackage{microtype}
+\usepackage{multirow}
 
@@
 \newcommand{\dex}{\ensuremath{{\rm dex}}}
+\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
+\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
 \newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
 \newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
+
 \newcommand{\npairsobs}{311}
 \newcommand{\ncomoving}{127}
+\newcommand{\nraveoverlap}{154}
+\newcommand{\nravecomoving}{59}
+\newcommand{\ncomovingtotal}{186}
 
@@
 spectrum of the Milky Way.
-Thus far, a few hundred genuine comoving star pairs with separations as large
+Thus far, a few hundred confirmed comoving star pairs with separations as large
 as $8~\pc$ have been discovered in the solar neighborhood using
@@
 with a possible steep break that depends on the details of the population of
-massive perturbers in the Galaxy, i.e. giant molecular clouds (GMCs), black holes, etc. (\citealt{Weinberg:1987}).
+massive perturbers in the Galaxy, i.e. giant molecular clouds (GMCs), black
+holes, etc. (\citealt{Weinberg:1987}).
 
@@
 (\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
-Oelkers:2017, Andrews:2017}) have been discovered over the last several decades (see also \citealt{Chaname:2007} and references therein).
+Oelkers:2017, Andrews:2017}) have been discovered over the last several decades
+(see also \citealt{Chaname:2007} and references therein).
 These pairs are identified by searching for approximately cospatial stars that
 appear to be comoving either in sky-projected velocities --- we refer to these
-as ``comoving pair candidates'' --- or using full-space velocity information --- we refer to these as ``comoving pairs.''
+as ``comoving pair candidates'' --- or using full-space velocity information ---
+we refer to these as ``comoving pairs.''
 Radial velocity follow-up is still needed to confirm the majority of the known
@@
 observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
-\project{Modspec} was set up in long-slit mode with a 1\arcsec\ slit, a $600~{\rm
-line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
-detector (``Echelle'').
+\project{Modspec} was set up in long-slit mode with a 1\arcsec\ slit, a
+$600~{\rm line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe
+CCD detector (``Echelle'').
 Only the central 300 pixel columns were read out from the CCD, and pixel rows
@@
 90~{\rm s}$ per observation.
-To determine rough nightly wavelength solutions, calibration spectra were obtained using
-Hg-Ne and Ne comparison lamps at the beginning and end of each night.
-Additionally, atmospheric emission lines were used to correct the wavelength solution on a
-per-object basis.
+To determine rough nightly wavelength solutions, calibration spectra were
+obtained using Hg-Ne and Ne comparison lamps at the beginning and end of each
+night.
+Additionally, atmospheric emission lines were used to correct the wavelength
+solution on a per-object basis.
 
@@
 calibration targets.
-For the purposes of this work, all of the calibration targets have previously measured radial velocities and were selected from high-resolution spectroscopic
+For the purposes of this work, all of the calibration targets have previously
+measured radial velocities and were selected from high-resolution spectroscopic
 surveys \citealt{Bensby:2014,Pinsonneault:2014}; these stars are used to
@@
 proper-motion-only log-FML-ratio for a given pair of stars as
-\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
+\llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
+radial velocity data as \llrnew.
 We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
@@
 
-\subsection{Properties of genuine comoving pairs}\label{sec:genuine}
+\subsection{Confirmed comoving pairs}\label{sec:genuine}
 
@@
 plotted against the 3D positional separation of the two stars in all observed
-pairs.
+pairs from this work.
 We again plot the median of the velocity difference distribution, and here use
@@
 comoving.
-Larger, black markers are genuine comoving pairs ($\llrnew > \llrold$); note
+Larger, black markers are confirmed comoving pairs ($\llrnew > \llrold$); note
 that many large-separation pairs have been ruled-out as comoving, but
@@
 At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
-genuinely comoving; interestingly, the separation distribution of the genuine
-pairs appears to be uniform or possibly growing in number with increasing
-separation.
+confirmed to be comoving; interestingly, the separation distribution of the
+genuine pairs appears to be uniform or possibly growing in number with
+increasing separation.
 This strongly suggests the existence of a population of unbound wide binaries or
@@
     the stars in all observed comoving star pairs (gray), and only those
-    identified as genuinely comoving (black). Note that the contamination rate
-    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
-    significant number of genuine comoving pairs at these very wide separations.
+    confirmed as comoving (black).
+    Note that the false-positive rate is $\approx 60\%$ at large separations
+    ($>1~\pc$), but there are still a significant number of comoving pairs at
+    these very wide separations.
     \emph{Right}: Distribution of chord lengths (tangential separations) for all
-    comoving star pairs (gray) and only those identified as genuinely comoving
-    (black).
+    comoving star pairs (gray) and only those identified as comoving with 3D
+    velocity information (black).
     Noise in the parallax measurements makes the 3D separation distribution
@@
   \caption{%
-    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
+    Color-magnitude diagram of confirmed comoving star pairs plotted over
     the density of all \tgas\ sources within 200 pc; similar to
-    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
-    pairs.
+    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
     Several interesting pairs that include main-sequence and giant component
@@
 
-\subsection{Highlighted comoving pairs}\label{sec:interesting-pairs}
+An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have radial
+velocities (for each star) reported in \DR{5} of the Radial Velocity Experiment
+(RAVE; \citealt{Kunder:2017}).
+Using the same procedure described in \sectionname~\ref{sec:catalog}, we
+identify confirmed comoving pairs using the RAVE velocities for all pairs within
+$200~\pc$ and find a similar false-positive rate at large separations ($\approx
+38\%$ are confirmed).
+Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
+that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
+We plot the confirmed comoving pairs identified through radial velocity
+follow-up in this work along with the RAVE-confirmed pairs in
+\figurename~\ref{fig:separation-RAVE}.
+Top histogram shows the separation (chord length) distribution for the joint
+sample, scatter plot shows log-chord-length and the pair distances, and right
+histogram shows the distance distribution of the confirmed pairs.
+At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
+of our method drops, as expected given the hard cut in log-FML-ratio applied to
+define the candidate sample.
+Reassuringly, there do not appear to be any trends of the chord length
+distribution with distance.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
+  \end{center}
+  \caption{%
+    \emph{Top}: Distribution of chord lengths for all confirmed comoving star
+      pairs, either from the radial velocity follow-up presented in this work or
+      from RAVE \DR{5} radial velocities.
+      At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log)
+      appears uniform, consistent with past measurements of the wide binary
+      separation distribution.
+      At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to
+      be an underabundance of comoving pairs, but the number of comoving pairs
+      then begins to increase again out to our sample separation limit of
+      $10~\pc$.
+    \emph{Middle}: Chord length plotted against mean distance for each comoving
+      pair identified from this work (circles, blue) or from RAVE overlap
+      (squares, orange).
+    \emph{Right}: Distance distribution for all confirmed comoving pairs.
+      Note that at distances $\gtrsim 150~\pc$, our detection efficiency appears
+      to decline.
+    \label{fig:separation-RAVE}}
+\end{figure}
+
+\subsection{A few highlighted comoving pairs}\label{sec:interesting-pairs}
 
@@
 
+We compile Tycho 2 ($B$, $V$; \citealt{derp}), 2MASS ($J$, $H$, $K_{\rm s}$;
+\citealt{derp}), and WISE ($W_1$, $W_2$, $W_3$; \citealt{derp}) photometry for
+each star in the pairs below and fit isochrones to the stars individually to
+measure their stellar parameters.
+We use the \package{isochrones} \python\ package (\citealt{Morton:2015}) to fit
+the photometric and parallax information (and uncertainties) for each star to
+infer the \logg, \teff, and stellar mass; these values are reported in
+\tablename~\ref{tbl:highlighted-pairs}.
+In brief, we use the model grids from the Dartmouth Stellar Evolution Database
+(\citealt{Dotter:2008}) and the built-in interpolation in \package{isochrones}
+to compute the posterior probability of a given set of stellar parameters given
+the data (photometry and parallax).
+We generate samples (in stellar parameters) from this posterior pdf using the
+ensemble MCMC sampler implemented in \package{emcee}
+(\citealt{Foreman-Mackey:2013}) using 128 walkers.
+We run the sampler for 256 initial steps from random initial positions, compute
+the median position of the walkers in parameter-space and re-initialize the
+walkers in a small Gaussian ball around this median position to run for another
+1024 burn-in steps.
+We then discard all but the last position and re-run for a final 8192 steps,
+saving only every 128th step.
+The stellar parameter values listed below are computed by taking the median
+posterior sample, and the uncertainties are estimated by computing the median
+absolute deviation, ${\rm MAD}$, and converting to a standard deviation under
+the assumption of Gaussianity ($\sigma \approx 1.5 \times {\rm MAD}$).
+
+\begin{table}[ht]
+  \begin{center}
+    \begin{tabular}{ c | c | c | c | c }
+      \toprule
+        Star name & \logg\ & \teff\ [K] & mass [$M_\odot$] &
+          chord length [pc]  \\
+        \toprule
+        HD 95088 & $3.7 \pm 0.02$ & $5303 \pm 74$ & $1.490 \pm 0.002$ &
+          \multirow{2}{*}{$8.6 \pm 0.8$}\\
+        HD 92278 & $4.2 \pm 0.05$ & $8519 \pm 300$ & $1.8 \pm 0.1$ &\\
+        \midrule
+        HD 79247 & $3.61 \pm 0.02$ & $5075 \pm 60$ & $1.150 \pm 0.004$ &
+          \multirow{2}{*}{$6.8 \pm 0.3$}\\
+        TYC 3807-1439-1 & $4.50 \pm 0.04$ & $5409 \pm 80$ & $0.88 \pm 0.05$ &\\
+        \midrule
+        HD 54318 & $4.15 \pm 0.04$ & $10030 \pm 345$ & $2.4 \pm 0.1$ &
+          \multirow{2}{*}{$8.4 \pm 1.1$}\\
+        TYC 3792-658-1 & $4.54 \pm 0.03$ & $5462 \pm 139$ & $0.89 \pm 0.06$ &\\
+      \bottomrule
+    \end{tabular}
+    \caption{
+      Stellar parameters inferred using isochrone-fitting for the three pairs
+      highlighted in \sectionname~\ref{sec:interesting-pairs}.
+      \label{tbl:highlighted-pairs}
+    }
+  \end{center}
+\end{table}
+
 \subsubsection{HD 95088 -- HD 92278}
 
-HD 95088 HD 92278
-Angular sep.: 4.49 deg
-Chord length: 8.66 pc
-3D sep.: 9.53 pc
-parallaxes: 9.00 +/- 0.50, 9.32 +/- 0.51
-
-This pair is separated by $8.7~\pc$ in chord length ($4.5^\circ$ on the sky) and
-have consistent distances within their uncertainties of $\approx 110~\pc$.
-From visual classification of the stars' location in the color--magnitude
-diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
-comprised of an A-type dwarf and a slightly less massive G-type subgiant.
-If this is indeed the case, the two stars could therefore not be coeval, as the
+From their surface gravities and temperatures, the component stars of this pair
+appear to be an A-type dwarf (HD 92278) and a slightly less massive G-type
+subgiant (HD 95088).
+This pair is separated by $4.5^\circ$ on the sky (corresponding to $\approx
+8.6~\pc$ in chord length, see \tablename~\ref{tbl:highlighted-pairs}) and have
+consistent distances within their uncertainties of $\approx 110 \pm 5~\pc$.
+At first glance, it seems unlikely that the two stars could be coeval, as the
 more massive A star has not begun its post-main-sequence evolution.
@@
 
-Other possibilities stem from the possibility that the comoving pair is the
-outcome of an interaction between more stars.
+Other speculative explanations for this pair stem from the possibility that it
+is the outcome of an interaction between even more stars.
 For example, an interaction between two wide binaries could result in an
@@
 Or, the initial system could have been an unstable three-body system that
-resulted in a tight, inner binary and a wide outer binary;
+resulted in a tight inner binary and a wide outer binary;
 if the inner binary were close enough, it could have merged and formed a blue
@@
 
-These explanations are all speculative at the moment; high-resolution
-spectroscopy of these stars --- and precise measurements of the stellar
-parameters and chemical abundances --- may help to distinguish these
-possibilities.
+High-resolution spectroscopy of these stars --- and precise measurements of the
+stellar parameters and chemical abundances --- may help to distinguish these
+scenarios.
 
@@
 
-This pair is separated by $XX~\pc$ in chord length ($XX^\circ$ on the sky) and
-have consistent distances within their uncertainties of $\approx XX~\pc$.
-From visual classification of the stars' location in the color--magnitude
-diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
-comprised of a G-type subgiant and a G-type dwarf.
+The components of this pair appear to be a K-type subgiant (HD 79247) and a
+G-type dwarf (TYC 3807-1439-1).
+This pair is separated by $4^\circ$ on the sky or $6.8~\pc$ in chord length and
+have an apparent radial separation of $\approx 3~\pc$, corresponding to a 3D
+separation of $\approx 9.8~\pc$ at a distance of $\approx 100~\pc$.
+Coeval stars at different phases of stellar evolution are extremely valuable for
+calibrating and testing stellar evolution models (e.g., \citealt{Torres:2013}).
+With a larger sample of confirmed comoving pairs, identified pairs that contain
+both an evolved star and a main sequence star --- such as this pair --- will be
+of great interest.
 
@@
 
-TYC 3792-658-1 HD 54318
-Angular sep.: 2.51 deg
-Chord length: 8.45 pc
-3D sep.: 8.73 pc
-parallaxes: 5.20 +/- 0.23, 5.30 +/- 0.32
-
-This pair is separated by $8.5~\pc$ in chord length ($2.5^\circ$ on the sky) and
-have consistent distances within their uncertainties of $\approx 190~\pc$.
-
-From visual classification of the stars' location in the color--magnitude
-diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
-comprised of a late-A-/early-B-type dwarf and a G-type dwarf.
-
-\subsection{Combined with RAVE}\label{sec:rave}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
-  \end{center}
-  \caption{%
-    \label{fig:separation-RAVE}}
-\end{figure}
-
-Caution about selection criteria, false-positives at large separation...
+The components of this pair appear to be an A- or B-type dwarf (HD 54318) and a
+G-type dwarf (TYC 3792-658-1), the largest magnitude difference between all of
+the confirmed comoving pairs.
+This pair is separated by $2.5^\circ$ on the sky or $8.5~\pc$ in chord length
+and have a consistent distance of $\approx 190~\pc$ within their uncertainties.
+The mass ratio distribution of comoving star pairs will help constrain star
+formation models (e.g., \citealt{Raghavan:2010}).
 
 \section{Discussion} \label{sec:discussion}
-
-Abundance of large-separation pairs
 
@@
 
+% Low-resolution spectroscopic follow-up of comoving pairs.
+
+% Using relative RV's, confirm a sample of very wide separation comoving star pairs.
+
+% Either wide binaries, unbound binaries, or dissolving clusters
+
+% Need much larger sample with no cuts on separation or signal-to-noise to study
+% true separation distribution
+
 \section{Conclusions}
 
-Low-resolution spectroscopic follow-up of comoving pairs.
-
-Using relative RV's, confirm a sample of very wide separation comoving star pairs.
-
-Either wide binaries, unbound binaries, or dissolving clusters
-
-Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution
-
-Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.
+The information contained in a large sample of widely-separated, comoving star
+pairs spans from the dynamics of stars and dark matter in the Galaxy, the
+fundamentals of star formation, and the stability of planetary systems.
+Through our own low-resolution spectroscopic follow-up and a cross-match with
+the RAVE \DR{5} catalog, we have identified \ncomovingtotal\ comoving star pairs
+by combining these radial velocity measurements with astrometry from the \tgas\
+catalog.
+The separation distribution of these spectroscopically-confirmed comoving star
+pairs presented shows an abundance of comoving pairs at separations between
+$1$--$10~\pc$ (\figurename~\ref{fig:separation-RAVE}), the imposed limit in our
+initial search (\citealt{Oh:2017}).
+This result is further tentative evidence (see also \citealt{Shaya:2011}) for
+the existence of a population of unbound comoving stars that represent the
+ultimate fate of wide binaries and moving groups that slowly dissociate due to
+weak perturbations from stars (\citealt{Jiang:2010}) and other massive
+perturbers in the Galaxy.
+With future data releases from the \gaia\ mission combined with large
+spectroscopic surveys, the anticipated catalogs of confirmed comoving star pairs
+will offer a new ``stereo'' view on the evolution and state of our Galaxy.
 
@@
 
-Dave Zurek, Keith, Ruth, Kelle
+It is a pleasure to thank
+Ruth Angus (Columbia),
 John Brewer (Yale),
 Jason Curtis (Columbia),
-James Davenport (WWU) for PyDIS,
+Keith Hawkins (Columbia),
 David W. Hogg (NYU),
-Melissa Ness (MPIA)
-
+Melissa Ness (MPIA), and
+Dave Zurek (AMNH)
+for useful discussions and advice.
+We thank James Davenport (WWU) for releasing his open source spectroscopic
+reduction code, \package{PyDIS}, which served as inspiration for the pipeline
+used in this work.
 This work is based on observations obtained at the MDM Observatory, operated by

### 2017-07-13 paper/comoving-followup.tex
columns of data tables

@@
 \newcommand{\tgas}{\acronym{TGAS}}
+\newcommand{\tmass}{\acronym{2MASS}}
 \newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
@@
 
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[width=0.8\linewidth]{sample_cmd.pdf}
+  \end{center}
+  \caption{%
+    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
+    blue histogram) and all comoving star pair members observed in this
+    follow-up effort (black points); colored lines connect candidate comoving
+    star pairs.
+    \tmass\ (\citealt{Skrutskie:2006}) $J$-band magnitudes are obtained
+    from the Gaia science archive.
+    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
+    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
+    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
+    \citealt{Oh:2017}).
+    \label{fig:sample-cmd}}
+\end{figure}
+
 To select candidate pairs for follow-up spectroscopic observations, we only
@@
 Candidate comoving pairs are connected by (colored) lines.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{sample_cmd.pdf}
-  \end{center}
-  \caption{%
-    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
-    blue histogram) and all comoving star pair members observed in this
-    follow-up effort (black points); colored lines connect candidate comoving
-    star pairs.
-    2MASS $J$-band magnitudes are obtained from the Gaia science
-    archive.
-    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
-    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
-    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
-    \citealt{Oh:2017}).
-    \label{fig:sample-cmd}}
-\end{figure}
 
@@
 have very short exposure times and thus weak night-sky lines.
+
+\begin{table}[ht]
+  \begin{center}
+    \begin{tabular}{ c | c }
+      \toprule
+        Parameter & Prior \\\toprule
+        $\ln A$ & $\mathcal{U}(2, 16)$\\
+        $x_0$ & $\mathcal{U}(6547, 6579)$\\
+        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$\\
+        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$\\
+        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$\\
+        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$\\
+        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
+        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$\\
+      \bottomrule
+    \end{tabular}
+    \caption{
+      Priors used for \Ha\ line profile model parameters.
+      $\mathcal{U}(a, b)$ indicates a uniform prior between $a$ and $b$ and
+      $\mathcal{N}(\mu, \sigma^2)$ indicates a normal distribution with mean
+      $\mu$ and variance $\sigma^2$.
+      Wavelength units are all angstroms.
+      \todo{what are the flux units?}
+      \label{tbl:prior-bounds}
+    }
+  \end{center}
+\end{table}
 
@@
 is between $5$--$10~\kms$.
-
-\begin{table}[ht]
-  \begin{center}
-    \begin{tabular}{ c | c }
-      \toprule
-        Parameter & Prior \\\toprule
-        $\ln A$ & $\mathcal{U}(2, 16)$\\
-        $x_0$ & $\mathcal{U}(6547, 6579)$\\
-        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$\\
-        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$\\
-        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$\\
-        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$\\
-        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
-        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$\\
-      \bottomrule
-    \end{tabular}
-    \caption{
-      Priors used for \Ha\ line profile model parameters.
-      $\mathcal{U}(a, b)$ indicates a uniform prior between $a$ and $b$ and
-      $\mathcal{N}(\mu, \sigma^2)$ indicates a normal distribution with mean
-      $\mu$ and variance $\sigma^2$.
-      Wavelength units are all angstroms.
-      \todo{what are the flux units?}
-      \label{tbl:prior-bounds}
-    }
-  \end{center}
-\end{table}
-
-\begin{figure}[htbp]
+We report the measured absolute radial velocities along with compiled sky
+position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
+$J$-band magnitudes for all of our comoving star targets in the ``stars'' data
+table available with this \documentname; see \tablename~\ref{tbl:data-stars} for
+a detailed description of the content of this table.
+
+\begin{figure}[!hp]
   \begin{center}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
+    \includegraphics[width=0.7\linewidth]{mcmc_example_fit.pdf}
   \end{center}
@@
 \end{figure}
+
+\begin{table}[htb]
+    \centering
+    \caption{Description of columns in data table containing results from radial
+    velocity follow-up.} \label{tbl:data-stars}
+    \begin{tabular}{l|l|l}
+        \toprule
+        Column Name    & Unit & Description\\
+        \midrule
+        \texttt{Oh17\_id}         &      & Row index in candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
+        \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
+        \texttt{name}             &      & HD, \project{Hipparcos}, or \project{Tycho-2} identifier\\
+        \texttt{ra}               & deg  & Right ascension from \tgas\\
+        \texttt{dec}              & deg  & Declination from \tgas\\
+        \texttt{parallax}         & mas  & Parallax from \tgas\\
+        \texttt{distance}         & pc   & Distance estimate using inverse parallax from \tgas\\
+        \texttt{G}                & mag  & \gaia\ $G$-band magnitudes\\
+        \texttt{J}                & mag  & \tmass\ $J$-band magnitudes\\
+        \texttt{rv}               & \kms & Radial velocity measured in this work\\
+        \texttt{rv\_err}          & \kms & Radial velocity error\\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+% \clearpage
 
@@
 
+The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
+all observed candidate comoving pairs are provided with the associated ``pairs''
+table; see \tablename~\ref{tbl:data-pairs} for a detailed description of the
+content of this table.
+
+\begin{table}[htb]
+    \centering
+    \caption{Description of columns in data table containing results from radial
+    velocity follow-up.} \label{tbl:data-stars}
+    \begin{tabular}{l|l|l}
+        \toprule
+        Column Name    & Unit & Description\\
+        \midrule
+        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
+        \texttt{R\_mu}            &      & Log-likelihood-ratio computed with only astrometric information\\
+        \texttt{R\_rv}            &      & Log-likelihood-ratio computed with radial velocities as well\\
+        \texttt{dx\_50}           & \pc  & Median of 3D spatial separation error distribution\\
+        \texttt{dv\_50}           & \kms & Median of 3D velocity difference error distribution\\
+        \texttt{dx\_15}           & \pc  & 15th percentile of 3D spatial separation error distribution\\
+        \texttt{dx\_85}           & \pc  & 85th percentile of 3D spatial separation error distribution\\
+        \texttt{dv\_15}           & \kms & 15th percentile of 3D velocity difference error distribution\\
+        \texttt{dv\_85}           & \kms & 85th percentile of 3D velocity difference error distribution\\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
 \subsection{Confirmed comoving pairs}\label{sec:genuine}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
+    \includegraphics[width=0.8\linewidth]{genuine_cmd.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{highlighted_cmd.pdf}
+    \includegraphics[width=0.8\linewidth]{highlighted_cmd.pdf}
   \end{center}
@@
 
-We compile Tycho 2 ($B$, $V$; \citealt{derp}), 2MASS ($J$, $H$, $K_{\rm s}$;
-\citealt{derp}), and WISE ($W_1$, $W_2$, $W_3$; \citealt{derp}) photometry for
-each star in the pairs below and fit isochrones to the stars individually to
-measure their stellar parameters.
+We compile \project{Tycho-2} ($B$, $V$; \citealt{Hog:2000}), \tmass\
+($J$, $H$, $K_{\rm s}$; \citealt{Skrutskie:2006}), and \acronym{WISE} ($W_1$,
+$W_2$, $W_3$; \citealt{Wright:2010}) photometry for each star in the pairs below
+and fit isochrones to the stars individually to measure their stellar
+parameters.
 We use the \package{isochrones} \python\ package (\citealt{Morton:2015}) to fit
@@
 
+\subsection{Caveats} \label{sec:caveats}
+
+Need much larger sample with no cuts on separation or signal-to-noise to study
+true separation distribution
+
 Need precise chemical information
@@
 Also tests of chemical enrichment due to planetary systems
-
-Comoving stars are also sensitive potentiometers of the Galactic mass
-distribution.
-- Long-term goal is to study the separation distribution of wide-binaries as a
-  function of age and Galactic orbit distribution (disk vs. halo)
-    - Is there a separation truncation?
-    - Can that be described just from Galactic tides, or need perturbers?
-
-% Caution that prior results are for red giants, MS stars may have planetary
-% imprints that get erased as dredge-up and convective layer erases...
 
@@
 
-Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>
+% \subsection{Prospects for chemical tagging} \label{sec:chemical-tagging}
 
@@
 
-% Low-resolution spectroscopic follow-up of comoving pairs.
-
-% Using relative RV's, confirm a sample of very wide separation comoving star pairs.
-
-% Either wide binaries, unbound binaries, or dissolving clusters
-
-% Need much larger sample with no cuts on separation or signal-to-noise to study
-% true separation distribution
+% \subsection{Prospects for Galactic structure} \label{sec:dynamics}
+
+\subsection{Future spectroscopic surveys} \label{sec:future-rvs}
+
+DISCO / AS4, 4MOST, WEAVE, DESI, PFS
 

### 2017-07-25 paper/comoving-followup.tex
Comments on Sec 1 and 2

@@
 \title{Spectroscopic confirmation of very-wide binaries and large-separation
-       comoving stars from \textsl{Gaia DR1}}
+       comoving pairs from \textsl{Gaia DR1}}
 
@@
 % Aims
-In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
+In previous work, we have identified \smoh{$\sim 10,000$: this number is probably wrong; need to check} candidate comoving star pairs
 using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
@@
 At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
-tenuously bound stars, recently disrupted binaries, moving groups or resonant
+tenuously bound stars such as recently disrupted binaries, moving groups or resonant
 features, and the final products of dissolving stellar associations or
@@
 surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving star
-pairs (\citealt{Shaya:2011}) and over 10,000 candidate systems
+pairs (\citealt{Shaya:2011}) and over \smoh{check: 10,000} candidate systems
 (\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
@@
 separation distribution of wide binaries at angular separations $\Delta \theta
-\lesssim 1^\circ$ where the false-positive rate is expected to be lower
+\lesssim 1^\circ$ where the false-positive rate is expected to be lower \smoh{value?}
 and the samples are likely predominantly wide binaries
@@
 \approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
-0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations.
+0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations \smoh{which? citation needed}.
 Tentative reports of a steepening or cutoff of the separation distribution at
@@
 Though these simulations don't include perturbers like GMCs or black holes,
-this at least indicates that a simple, monotonic separation distribution may
+they at least indicate that \smoh{the phase-space coherence of wide binaries
+  may linger on for a longer period of time than previously thought, and thus,
+  a monotonically decreasing separation distribution may
 be too simple of a model for the separation distribution of comoving stars in
-our Galaxy.
+our Galaxy.}
 In addition, the distribution of unbound but comoving stars should be very
@@
 moving with the same 3D velocity, their proper motions and radial velocities
-would be different because of geometric effects.
+can be different due to projections.
 By compiling radial velocities from the literature for \project{Hipparcos}
@@
 ratio computed for all nearly co-spatial pairs of stars.
-Briefly, likelihood 1, $L_1$, considers a model in which the full-space
-velocities of each star in the pair are identical and sampled from a disk-like,
-Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
-in which the full-space velocities of each star were drawn independently from
-the same velocity distribution used above. The computed likelihood ratio is
-actually the ratio of fully marginalized likelihoods,
-$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
-distance for each star in a given pair.
+Briefly, we perform a model selection between
+two cases: that the velocities of the two stars in a pair are
+identical (case 1) or that they are indepndent (case 2).
+We marginalize over the (unknown) true velocity and distance of each star
+to obtain the ratio of fully marginalized likelihoods,
+$\mathcal{L}_1/\mathcal{L}_2$.
+The likelihoods appropriately take into account the reported uncertainties and
+covariances of the astrometric measurements in the \tgas\ catalog
+as well as projection effect in comparing spherical
+velocities for star pairs separated by large angles on the sky.
 After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
-we compute the likelihood ratio for all stars within $10~\pc$ and with a
-difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each target star
-from the \tgas\ catalog.
-The likelihoods appropriately take into account the reported covariances and
-uncertainties between the astrometric measurements in the \tgas\ catalog and
-naturally handle geometric or projection effects in comparing spherical
-velocities for star pairs separated by large angles on the sky.
+we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
+with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each
+target star from the \tgas\ catalog.
 After applying a conservative cut on the likelihood ratio motivated by computing
 the same likelihood ratio for random pairings of stars ($\ln
-\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
+\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
 physical separations between $10^3~\au$ and $10~\pc$.
@@
 fraction of the identified pairs are members of larger groups.
-$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
-paired with other stars.
+However, we have also found a large number ($\approx$4,000) of mutually
+exclusively-linked pairs in increasing number with increasing separation at
+$\sim 1$~\pc.
 For a full description of the selection method, see \citealt{Oh:2017}.
@@
 from $\approx 3600$--$7200~{\rm \AA}$.
-At this resolution, to obtain high signal-to-noise spectra, most exposures were
+At this resolution, to obtain high signal-to-noise \smoh{value?} spectra, most exposures were
 between $30$--$120~{\rm s}$ (15th and 85th percentiles).
@@
 We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
-2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
+2017-03-15, corresponding to 323 unique pairs and $\approx 120$
 calibration targets.
-For the purposes of this work, all of the calibration targets have previously
-measured radial velocities and were selected from high-resolution spectroscopic
-surveys \citealt{Bensby:2014,Pinsonneault:2014}; these stars are used to
-validate our radial velocity measurements.
+All calibration targets have previously measured radial velocities from
+high-resolution spectroscopic surveys (\citealt{Bensby:2014,Pinsonneault:2014})
+in order to validate our radial velocity measurements.
 These stars also have measured chemical abundances, and will be used in future
-work as a training set for \project{The Cannon} (\citealt{Ness:2015}) to infer
+work as a training set for a method of transferring chemical abundance labels,
+\project{The Cannon} (\citealt{Ness:2015}), to infer
 metallicities and stellar parameters for our observed sources.
@@
 
+% this paragraph can be simplified
 For each source or comparison lamp observation, the two-dimensional image is
@@
 For source observations, 1D spectra are extracted using an empirical
-estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
-rows between indices 750 and 850 (i.e. around the effective center of the
-dispersion axis), a model for the LSF is fit to the 1D trace.
+estimate of the line spread function (LSF).
 The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
@@
 The maximum-likelihood LSF + background model parameters are estimated from
-each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
+{\bf each pixel row} using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
 the \package{scipy} package.
-From the 16 fits, the median profile width parameters, $\sigma_G$ and
+We first determine the width parameters, $\sigma_G$ and $\gamma_L$
+from the fits to
+\smoh{16 pixel rows - Is it every 16 rows (meaning total/16=1600/16=100 fits),
+  or 16 rows spread across the entire 1600 rows? I think you mean the latter,
+  but this combined with bold-faced text is a little confusing. }
+between indices 750 and 850 (i.e. around the effective center of the
+dispersion axis), a model for the LSF is fit to the 1D trace.
+{\bf From the 16 fits}, the median profile width parameters, $\sigma_G$ and
 $\gamma_L$, are then taken to be the LSF width parameters.
-At each of 1600 rows in each source spectrum image, the LSF amplitude and
+Then, {\bf at each of 1600 rows} in each source spectrum image, the LSF amplitude and
 centroid, and background slope and intercept at line centroid, are then fit
@@
 //physics.nist.gov/PhysRefData/ASD/lines_form.html}}
-Once a rough mapping from pixel to wavelength is determined, we then fit for the
-individual line centroids in each comparison lamp spectrum by extracting small,
-8 pixel-wide segments of the comparison lamp spectrum around each identified
-emission line and fitting a model for the line profile and background to each
-individually.
-We use a pixel-convolved Gaussian profile for each emission line with
+Once a rough mapping from pixel to wavelength is determined, we fit
+8 pixel-wide segments around each identified emission line
+for the individual line centroids in each comparison lamp spectrum.
+We use a {\bf pixel-convolved} Gaussian profile for each emission line with
 parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
@@
 ``background'') sometimes bias the line centroid determination when using a
-simple background model.
+simple {\bf linear? quadratic?} background model.
 We therefore model the background flux around each comparison lamp emission
-line using a constant offset + Gaussian process (GP).
+line using a constant offset + Gaussian process (GP)
+{\bf in order to ..?}.
 For the GP model, we use a Mat\'ern 3/2 kernel function
@@
 un- or weakly-constrained.
-We instead use a combined linear polynomial + GP to model the wavelength
-dispersion.
+\smoh{We instead use a combined linear polynomial + GP to model the wavelength
+  dispersion. - what is each component supposed to capture?}
 We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
@@
 the excluded pixel/wavelength ranges.
-
-This wavelength calibration procedure produces a nightly model for the
-nonlinear dispersion of wavelength as a function of pixel value for each
-extracted 1D spectrum.
-However, because of flexure in the spectrograph, each observation will have an
-additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
-below).
+\smoh{The wavelength solution is not well constrained no matter what (due to
+  lack of data), and using GP model instead of polynomials does not change
+  that. Instead, it seems to me that GP model informs us of the regions to
+  discard.  Thus, technically, I think ``We instead use ...'' is not correctly
+  justified. Modify at will. Obviously quite a minor point..}
+
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10
@@
 visualization.
-As suggested by the features in these 10 spectra, there is a mix of
-metallicities and spectral types amongst the observed targets.
+As suggested by the features ({\bf the depth of the absorption features?}) in
+these 10 spectra, there is a mix of metallicities and spectral types amongst
+the observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring
@@
 
+The wavelength calibration procedure described above produces a nightly model
+for the nonlinear dispersion of wavelength as a function of pixel value for
+each extracted 1D spectrum.
+However, because of flexure in the spectrograph, each observation will have an
+additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
+below).
 As mentioned, comparison lamp spectra were not obtained at each telescope
@@
       \todo{what are the flux units?}
+      \smoh{A column of brief description of parameters would be useful.}
       \label{tbl:prior-bounds}
@@
 
+\smoh{Long paragraph - may break up at the discussion of an example case (bold-faced)}
 To derive radial velocities for each source, we then measure the (wavelength)
@@
 store these as our posterior samples.
+{\bf
 \figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
 (\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
-samples for a randomly-chosen source (HD 63408, in this case).
+samples for a randomly-chosen source (HD 63408, in this case).}
 The marginal posterior distribution over $x_0$ can be converted to a
@@
         \midrule
-        \texttt{Oh17\_id}         &      & Row index in candidate catalog (\citealt{Oh:2017})\\
-        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_id}         &      & Star row index from the candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_group\_id}  &      & Group ID from the candidate catalog (\citealt{Oh:2017})\\
         \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
@@
     \caption{Description of columns in data table containing results from radial
-    velocity follow-up.} \label{tbl:data-stars}
+      velocity follow-up.} \label{tbl:data-pairs}
     \begin{tabular}{l|l|l}

### 2017-07-25 paper/comoving-followup.tex
No need since paragraphs are combined

@@
 additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
-below).
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
 As mentioned, comparison lamp spectra were not obtained at each telescope

### 2017-07-25 paper/comoving-followup.tex
stuff

@@
 % Aims
-In previous work, we have identified \smoh{$\sim 10,000$: this number is probably wrong; need to check} candidate comoving star pairs
+In previous work, we have identified $\sim 4,000$ candidate comoving star pairs
 using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
@@
 clusters.
-However, even at separations of $\sim 0.1~\pc$, we typically lack the precision
+However, even at separations of $\sim 0.1~\pc$, we typically lack the precision {\bf in what?}
 to determine the binding energy of a system and therefore can't determine its
@@
 We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
-\ncomoving of the \npairsobs\ comoving pairs satisfy this constraint.
+\ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
 \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference

### 2017-07-27 paper/comoving-followup.tex

@@
 the excluded pixel/wavelength ranges.
-\smoh{The wavelength solution is not well constrained no matter what (due to
-  lack of data), and using GP model instead of polynomials does not change
-  that. Instead, it seems to me that GP model informs us of the regions to
-  discard.  Thus, technically, I think ``We instead use ...'' is not correctly
-  justified. Modify at will. Obviously quite a minor point..}
+\smoh{The wavelength solution is not well constrained at either end no matter
+  what (due to lack of data), and using GP model instead of polynomials does
+  not change that. Instead, it seems to me that GP model informs us of the
+  regions to discard.  Thus, technically, I think ``We instead use ...'' is not
+  correctly justified. Modify at will. Obviously quite a minor point..}
 

### 2017-07-27 paper/comoving-followup.tex
rave

--- 2017-07-25 paper/comoving-followup.tex
@@
 \title{Spectroscopic confirmation of very-wide binaries and large-separation
-       comoving pairs from \textsl{Gaia DR1}}
+       comoving stars from \textsl{Gaia DR1}}
 
@@
 % Aims
-In previous work, we have identified \smoh{$\sim 10,000$: this number is probably wrong; need to check} candidate comoving star pairs
+In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
 using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
@@
 At larger separations ($\gtrsim 0.1~\pc$), comoving pairs are likely a mix of
-tenuously bound stars such as recently disrupted binaries, moving groups or resonant
+tenuously bound stars, recently disrupted binaries, moving groups or resonant
 features, and the final products of dissolving stellar associations or
@@
 surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving star
-pairs (\citealt{Shaya:2011}) and over \smoh{check: 10,000} candidate systems
+pairs (\citealt{Shaya:2011}) and over 10,000 candidate systems
 (\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
@@
 separation distribution of wide binaries at angular separations $\Delta \theta
-\lesssim 1^\circ$ where the false-positive rate is expected to be lower \smoh{value?}
+\lesssim 1^\circ$ where the false-positive rate is expected to be lower
 and the samples are likely predominantly wide binaries
@@
 \approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
-0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations \smoh{which? citation needed}.
+0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations.
 Tentative reports of a steepening or cutoff of the separation distribution at
@@
 Though these simulations don't include perturbers like GMCs or black holes,
-they at least indicate that \smoh{the phase-space coherence of wide binaries
-  may linger on for a longer period of time than previously thought, and thus,
-  a monotonically decreasing separation distribution may
+this at least indicates that a simple, monotonic separation distribution may
 be too simple of a model for the separation distribution of comoving stars in
-our Galaxy.}
+our Galaxy.
 In addition, the distribution of unbound but comoving stars should be very
@@
 moving with the same 3D velocity, their proper motions and radial velocities
-can be different due to projections.
+would be different because of geometric effects.
 By compiling radial velocities from the literature for \project{Hipparcos}
@@
 ratio computed for all nearly co-spatial pairs of stars.
-Briefly, we perform a model selection between
-two cases: that the velocities of the two stars in a pair are
-identical (case 1) or that they are indepndent (case 2).
-We marginalize over the (unknown) true velocity and distance of each star
-to obtain the ratio of fully marginalized likelihoods,
-$\mathcal{L}_1/\mathcal{L}_2$.
-The likelihoods appropriately take into account the reported uncertainties and
-covariances of the astrometric measurements in the \tgas\ catalog
-as well as projection effect in comparing spherical
+Briefly, likelihood 1, $L_1$, considers a model in which the full-space
+velocities of each star in the pair are identical and sampled from a disk-like,
+Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
+in which the full-space velocities of each star were drawn independently from
+the same velocity distribution used above. The computed likelihood ratio is
+actually the ratio of fully marginalized likelihoods,
+$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
+distance for each star in a given pair.
+After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
+we compute the likelihood ratio for all stars within $10~\pc$ and with a
+difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each target star
+from the \tgas\ catalog.
+The likelihoods appropriately take into account the reported covariances and
+uncertainties between the astrometric measurements in the \tgas\ catalog and
+naturally handle geometric or projection effects in comparing spherical
 velocities for star pairs separated by large angles on the sky.
-After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
-we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
-with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each
-target star from the \tgas\ catalog.
 After applying a conservative cut on the likelihood ratio motivated by computing
 the same likelihood ratio for random pairings of stars ($\ln
-\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
+\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
 physical separations between $10^3~\au$ and $10~\pc$.
@@
 fraction of the identified pairs are members of larger groups.
-However, we have also found a large number ($\approx$4,000) of mutually
-exclusively-linked pairs in increasing number with increasing separation at
-$\sim 1$~\pc.
+$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
+paired with other stars.
 For a full description of the selection method, see \citealt{Oh:2017}.
@@
 from $\approx 3600$--$7200~{\rm \AA}$.
-At this resolution, to obtain high signal-to-noise \smoh{value?} spectra, most exposures were
+At this resolution, to obtain high signal-to-noise spectra, most exposures were
 between $30$--$120~{\rm s}$ (15th and 85th percentiles).
@@
 We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
-2017-03-15, corresponding to 323 unique pairs and $\approx 120$
+2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
 calibration targets.
-All calibration targets have previously measured radial velocities from
-high-resolution spectroscopic surveys (\citealt{Bensby:2014,Pinsonneault:2014})
-in order to validate our radial velocity measurements.
+For the purposes of this work, all of the calibration targets have previously
+measured radial velocities and were selected from high-resolution spectroscopic
+surveys \citealt{Bensby:2014,Pinsonneault:2014}; these stars are used to
+validate our radial velocity measurements.
 These stars also have measured chemical abundances, and will be used in future
-work as a training set for a method of transferring chemical abundance labels,
-\project{The Cannon} (\citealt{Ness:2015}), to infer
+work as a training set for \project{The Cannon} (\citealt{Ness:2015}) to infer
 metallicities and stellar parameters for our observed sources.
@@
 
-% this paragraph can be simplified
 For each source or comparison lamp observation, the two-dimensional image is
@@
 For source observations, 1D spectra are extracted using an empirical
-estimate of the line spread function (LSF).
+estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
+rows between indices 750 and 850 (i.e. around the effective center of the
+dispersion axis), a model for the LSF is fit to the 1D trace.
 The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
@@
 The maximum-likelihood LSF + background model parameters are estimated from
-{\bf each pixel row} using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
+each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
 the \package{scipy} package.
-We first determine the width parameters, $\sigma_G$ and $\gamma_L$
-from the fits to
-\smoh{16 pixel rows - Is it every 16 rows (meaning total/16=1600/16=100 fits),
-  or 16 rows spread across the entire 1600 rows? I think you mean the latter,
-  but this combined with bold-faced text is a little confusing. }
-between indices 750 and 850 (i.e. around the effective center of the
-dispersion axis), a model for the LSF is fit to the 1D trace.
-{\bf From the 16 fits}, the median profile width parameters, $\sigma_G$ and
+From the 16 fits, the median profile width parameters, $\sigma_G$ and
 $\gamma_L$, are then taken to be the LSF width parameters.
-Then, {\bf at each of 1600 rows} in each source spectrum image, the LSF amplitude and
+At each of 1600 rows in each source spectrum image, the LSF amplitude and
 centroid, and background slope and intercept at line centroid, are then fit
@@
 //physics.nist.gov/PhysRefData/ASD/lines_form.html}}
-Once a rough mapping from pixel to wavelength is determined, we fit
-8 pixel-wide segments around each identified emission line
-for the individual line centroids in each comparison lamp spectrum.
-We use a {\bf pixel-convolved} Gaussian profile for each emission line with
+Once a rough mapping from pixel to wavelength is determined, we then fit for the
+individual line centroids in each comparison lamp spectrum by extracting small,
+8 pixel-wide segments of the comparison lamp spectrum around each identified
+emission line and fitting a model for the line profile and background to each
+individually.
+We use a pixel-convolved Gaussian profile for each emission line with
 parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
@@
 ``background'') sometimes bias the line centroid determination when using a
-simple {\bf linear? quadratic?} background model.
+simple background model.
 We therefore model the background flux around each comparison lamp emission
-line using a constant offset + Gaussian process (GP)
-{\bf in order to ..?}.
+line using a constant offset + Gaussian process (GP).
 For the GP model, we use a Mat\'ern 3/2 kernel function
@@
 un- or weakly-constrained.
-\smoh{We instead use a combined linear polynomial + GP to model the wavelength
-  dispersion. - what is each component supposed to capture?}
+We instead use a combined linear polynomial + GP to model the wavelength
+dispersion.
 We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
@@
 the excluded pixel/wavelength ranges.
-\smoh{The wavelength solution is not well constrained no matter what (due to
-  lack of data), and using GP model instead of polynomials does not change
-  that. Instead, it seems to me that GP model informs us of the regions to
-  discard.  Thus, technically, I think ``We instead use ...'' is not correctly
-  justified. Modify at will. Obviously quite a minor point..}
-
+
+This wavelength calibration procedure produces a nightly model for the
+nonlinear dispersion of wavelength as a function of pixel value for each
+extracted 1D spectrum.
+However, because of flexure in the spectrograph, each observation will have an
+additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
+below).
 \figurename~\ref{fig:sample-spectra} shows a random sample of 10
@@
 visualization.
-As suggested by the features ({\bf the depth of the absorption features?}) in
-these 10 spectra, there is a mix of metallicities and spectral types amongst
-the observed targets.
+As suggested by the features in these 10 spectra, there is a mix of
+metallicities and spectral types amongst the observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring
@@
 
-The wavelength calibration procedure described above produces a nightly model
-for the nonlinear dispersion of wavelength as a function of pixel value for
-each extracted 1D spectrum.
-However, because of flexure in the spectrograph, each observation will have an
-additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
 As mentioned, comparison lamp spectra were not obtained at each telescope
@@
       \todo{what are the flux units?}
-      \smoh{A column of brief description of parameters would be useful.}
       \label{tbl:prior-bounds}
@@
 
-\smoh{Long paragraph - may break up at the discussion of an example case (bold-faced)}
 To derive radial velocities for each source, we then measure the (wavelength)
@@
 store these as our posterior samples.
-{\bf
 \figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
 (\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
-samples for a randomly-chosen source (HD 63408, in this case).}
+samples for a randomly-chosen source (HD 63408, in this case).
 The marginal posterior distribution over $x_0$ can be converted to a
@@
         \midrule
-        \texttt{Oh17\_id}         &      & Star row index from the candidate catalog (\citealt{Oh:2017})\\
-        \texttt{Oh17\_group\_id}  &      & Group ID from the candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_id}         &      & Row index in candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
         \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
@@
     \caption{Description of columns in data table containing results from radial
-      velocity follow-up.} \label{tbl:data-pairs}
+    velocity follow-up.} \label{tbl:data-stars}
     \begin{tabular}{l|l|l}
@@
 
-An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have radial
-velocities (for each star) reported in \DR{5} of the Radial Velocity Experiment
-(RAVE; \citealt{Kunder:2017}).
+An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
+previously-measured radial velocities (for each star) reported in \DR{5} of the
+Radial Velocity Experiment (RAVE; \citealt{Kunder:2017}).
 Using the same procedure described in \sectionname~\ref{sec:catalog}, we

### 2017-07-27 scripts/get_all_photometry.py
script to download all photometry for all db pairs

@@ -1,0 +1,84 @@
+# Standard library
+from collections import OrderedDict
+from os import path
+import sys
+
+# Third-party
+from astroquery.gaia import Gaia
+import numpy as np
+
+# Package
+from comoving_rv.log import logger
+from comoving_rv.db import Session, Base, db_connect
+from comoving_rv.db.model import (Run, Observation, TGASSource, SimbadInfo,
+                                  GroupToObservations, SpectralLineInfo,
+                                  SpectralLineMeasurement, RVMeasurement,
+                                  Photometry)
+
+from stellarparameters import get_photometry
+
+result_columns = ['bt_mag',
+                  'vt_mag',
+                  'e_bt_mag',
+                  'e_vt_mag',
+                  'j_m',
+                  'j_msigcom',
+                  'h_m',
+                  'h_msigcom',
+                  'ks_m',
+                  'ks_msigcom',
+                  'w1mpro',
+                  'w1mpro_error',
+                  'w2mpro',
+                  'w2mpro_error',
+                  'w3mpro',
+                  'w3mpro_error',
+                  'w4mpro',
+                  'w4mpro_error',
+                  'u_mag',
+                  'u_mag_error',
+                  'g_mag',
+                  'g_mag_error',
+                  'r_mag',
+                  'r_mag_error',
+                  'i_mag',
+                  'i_mag_error',
+                  'z_mag',
+                  'z_mag_error']
+
+def main():
+    # TODO: bad, hard-coded...
+    # base_path = '/Volumes/ProjectData/gaia-comoving-followup/'
+    base_path = '../data/'
+    db_path = path.join(base_path, 'db.sqlite')
+    engine = db_connect(db_path)
+    session = Session()
+
+    credentials = dict(user='apricewh', password='7Gy2^otQNj6FKz6')
+    Gaia.login(**credentials)
+
+    for obs in session.query(Observation).all():
+        q = session.query(Photometry).join(Observation).filter(Observation.id == obs.id).count()
+        if q > 0:
+            logger.debug('Photometry already exists')
+            continue
+
+        if obs.tgas_source is None:
+            continue
+
+        tgas_source_id = obs.tgas_source.source_id
+        res = get_photometry(tgas_source_id)
+
+        phot_kw = dict()
+        for col in result_columns:
+            phot_kw[col] = res[col]
+
+        phot = Photometry(**phot_kw)
+        phot.observation = obs
+        session.add(phot)
[…]

### 2017-07-28 paper/comoving-followup.tex
abstract

@@
 
-\title{Spectroscopic confirmation of very-wide binaries and large-separation
-       comoving pairs from \textsl{Gaia DR1}}
+\title{Spectroscopic confirmation of very-wide stellar binaries and
+       large-separation comoving pairs from \textsl{Gaia DR1}}
 
@@
 % Context
-Widely-separated ($\gtrsim 0.1$~pc), coeval stars are powerful tools for
-testing star-formation model predictions, improving stellar atmosphere models,
-and inferring the mass distribution of the Milky Way.
-They likely outnumber the Galactic open clusters by a large margin, but are
-still expected to share a common origin and chemistry.
-The kinematic properties of comoving pairs---including bound wide binaries---
-serve as potentiometers for the smooth tidal field and mass perturbation
-spectrum of the Milky Way.
-Thus far, a few hundred confirmed comoving star pairs with separations as large
-as $8~\pc$ have been discovered in the solar neighborhood using
-\project{Hipparcos}.
+Members of widely-separated stellar binaries ($\gtrsim 0.1$~pc) or loosely bound
+stellar associations and clusters are born with small spreads in kinematics and
+chemistry, but disrupt and dissociate on timescales comparable to their orbital
+time within the Milky Way.
+The kinematic properties of comoving star pairs---bound wide binaries or unbound
+but coeval stars---are therefore sensitive to both the smooth tidal field and
+mass perturbation spectrum of the Milky Way.
 % Aims
-In previous work, we have identified $\sim 4,000$ candidate comoving star pairs
-using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
-catalog, a part of \gaia\ \DR{1}.
-Without radial velocity measurements for the vast majority of the pairs, the
-false-positive rate at large separations is significant ($\sim 50\%$).
+% Thus far, a few hundred confirmed comoving pairs with separations as large as
+% $8~\pc$ have been discovered in the solar neighborhood using
+% \project{Hipparcos} astrometry combined with radial velocity surveys.
+In previous work, we identified $>4,000$ candidate comoving pairs using only
+astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog and
+find an abundance of pairs at separations $\gtrsim 1~\pc$.
+Without radial velocity measurements for the vast majority of these pairs, the
+false-positive rate at these separations is significant ($\sim 50\%$).
 % Methods
 Here, we present results from our own low-resolution radial-velocity survey of a
-sample of the candidate pairs to validate wide-separation comoving star pairs.
+random sample of the candidate pairs to identify and validate true comoving star
+pairs.
 % Results
@@
 pairs with separations as large as $\approx 10~\pc$, our original search limit.
+From the smallest separation pairs, $\approx 10^{-3}~\pc$, out to $\approx
+0.5~\pc$, we find that the number of confirmed comoving star pairs decreases
+with increasing separation, as expected.
+From $\approx 0.5~\pc$ to $10~\pc$, the separation distribution appears
+approximately uniform.
 % Conclusions
-We find that the number of comoving star pairs with very wide separations
-appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
-suggesting the existence of a population of recently unbound wide binaries or
-stellar associations.
-% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
-% separation for a typical MS star can we determine they are bound assuming
-% high-res spectra to get very precise RVs?
+We therefore confirm the discovery of a population of comoving star pairs at
+very large separations, strongly suggesting an abundance of disrupted wide
+binaries and stellar associations in the Galactic disk.
 \end{abstract}
+
+% The confirmed discovery of this population of coeval, comoving stars will
+% enable future dynamical tests of dark matter and XX of star formation XX.
+% as predicted from simulations of the disruption and evolution of wide binary
+% stars. Gaia DR2 -> how many do we expect? Can we say anything? Out to what
+% distance and separation for a typical MS star can we determine they are bound
+% assuming high-res spectra to get very precise RVs?
 

### 2017-07-28 paper/comoving-followup.tex
abstract again

@@
 but coeval stars---are therefore sensitive to both the smooth tidal field and
-mass perturbation spectrum of the Milky Way.
+gravitational perturbation spectrum of the Milky Way.
 % Aims
@@
 pairs with separations as large as $\approx 10~\pc$, our original search limit.
-From the smallest separation pairs, $\approx 10^{-3}~\pc$, out to $\approx
-0.5~\pc$, we find that the number of confirmed comoving star pairs decreases
-with increasing separation, as expected.
-From $\approx 0.5~\pc$ to $10~\pc$, the separation distribution appears
-approximately uniform.
+At separations of $10^{-3}~\pc$ to $0.5~\pc$,
+the number of confirmed pairs per separation decreases with increasing
+separation (number per log separation is uniform).
+From $0.5~\pc$ to $10~\pc$, the number per separation is approximately
+uniform (number per log separation increases).
 % Conclusions

### 2017-07-28 paper/comoving-followup.tex
address smoh comments

@@
 \definecolor{mediumpersianblue}{rgb}{0.0, 0.4, 0.65}
-\newcommand{\smoh}[1]{\textcolor{mediumpersianblue}{SO:#1}}
+\newcommand{\smoh}[1]{\textcolor{red}{SO:#1}}
 
@@
 pairs with separations as large as $\approx 10~\pc$, our original search limit.
-At separations of $10^{-3}~\pc$ to $0.5~\pc$,
+At separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
 the number of confirmed pairs per separation decreases with increasing
 separation (number per log separation is uniform).
-From $0.5~\pc$ to $10~\pc$, the number per separation is approximately
+From $\approx 0.5~\pc$ to $10~\pc$, the number per separation is approximately
 uniform (number per log separation increases).
@@
 clusters.
-However, even at separations of $\sim 0.1~\pc$, we typically lack the precision {\bf in what?}
-to determine the binding energy of a system and therefore can't determine its
-true dynamical state; because of this ambiguity, we refer to both bound and
-unbound widely-separated ($\gtrsim 10^{-3}~\pc$) systems collectively as
-``comoving'' systems.
+However, even at separations of $\sim 0.1~\pc$, we typically lack the velocity
+or distance precision to determine the binding energy of a system and therefore
+can't determine its true dynamical state; because of this ambiguity, we refer to
+both bound and unbound widely-separated ($\gtrsim 10^{-3}~\pc$) systems
+collectively as ``comoving'' systems.
 
@@
 surveys (\citealt{Steinmetz:2006}), several hundred confirmed comoving star
-pairs (\citealt{Shaya:2011}) and over \smoh{check: 10,000} candidate systems
+pairs (\citealt{Shaya:2011}) and over 4,000 candidate systems
 (\citealt{Gould:2004,Lepine:2007,Tokovinin:2012,Allen:2014,Oh:2017,
@@
 \approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
-0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations \smoh{which? citation needed}.
+0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations
+(e.g., \citealt{Weinberg:1987}).
 Tentative reports of a steepening or cutoff of the separation distribution at
@@
 $1000~\pc$; \citealt{Jiang:2010}).
-Though these simulations don't include perturbers like GMCs or black holes,
-they at least indicate that \smoh{the phase-space coherence of wide binaries
-  may linger on for a longer period of time than previously thought, and thus,
-  a monotonically decreasing separation distribution may
-be too simple of a model for the separation distribution of comoving stars in
-our Galaxy.}
+Though these simulations don't include perturbers like GMCs or black holes, they
+at least indicate that disrupted wide binaries may remain coherent for a longer
+period of time than previously thought, and thus, a monotonically decreasing
+separation distribution may be too simple of a model for the separation
+distribution of comoving stars in our Galaxy.
 In addition, the distribution of unbound but comoving stars should be very
@@
 from $\approx 3600$--$7200~{\rm \AA}$.
-At this resolution, to obtain high signal-to-noise \smoh{value?} spectra, most exposures were
-between $30$--$120~{\rm s}$ (15th and 85th percentiles).
+At this resolution, to obtain high signal-to-noise ($>100$) spectra, most
+exposures were between $30$--$120~{\rm s}$ (15th and 85th percentiles).
 To maximize the number of observed pairs, we chose not to take
@@
 The maximum-likelihood LSF + background model parameters are estimated from
-{\bf each pixel row} using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
+each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
 the \package{scipy} package.
-We first determine the width parameters, $\sigma_G$ and $\gamma_L$
-from the fits to
-\smoh{16 pixel rows - Is it every 16 rows (meaning total/16=1600/16=100 fits),
-  or 16 rows spread across the entire 1600 rows? I think you mean the latter,
-  but this combined with bold-faced text is a little confusing. }
-between indices 750 and 850 (i.e. around the effective center of the
-dispersion axis), a model for the LSF is fit to the 1D trace.
-{\bf From the 16 fits}, the median profile width parameters, $\sigma_G$ and
+We first determine the LSF width parameters, $\sigma_G$ and $\gamma_L$ from fits
+to 16 pixel rows evenly spaced between indices 750 and 850 (i.e. around the
+effective center of the dispersion axis).
+From these 16 fits, the median profile width parameters, $\sigma_G$ and
 $\gamma_L$, are then taken to be the LSF width parameters.
-Then, {\bf at each of 1600 rows} in each source spectrum image, the LSF amplitude and
+Then, at each of 1600 rows in each source spectrum image, the LSF amplitude and
 centroid, and background slope and intercept at line centroid, are then fit
@@
 for the individual line centroids in each comparison lamp spectrum.
-We use a {\bf pixel-convolved} Gaussian profile for each emission line with
+We use a pixel-convolved Gaussian profile for each emission line with
 parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
@@
 ``background'') sometimes bias the line centroid determination when using a
-simple {\bf linear? quadratic?} background model.
+simple (i.e. low-order polynomial) background model.
 We therefore model the background flux around each comparison lamp emission
-line using a constant offset + Gaussian process (GP)
-{\bf in order to ..?}.
+line using a constant offset + Gaussian process (GP) in order to capture any
+unmasked structure from nearby sources.
 For the GP model, we use a Mat\'ern 3/2 kernel function
@@
 values near either end of the pixel grid where the polynomial function is
-un- or weakly-constrained.
-\smoh{We instead use a combined linear polynomial + GP to model the wavelength
-  dispersion. - what is each component supposed to capture?}
+un- or weakly-constrained, and can easily over-fit to poorly centroided emission
+lines.
+We instead use a combined linear polynomial + GP to model the wavelength
+dispersion to flexibly account for nonlinear behavior in the pixel to wavelength
+mapping.
 We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
@@
 the excluded pixel/wavelength ranges.
-\smoh{The wavelength solution is not well constrained at either end no matter
-  what (due to lack of data), and using GP model instead of polynomials does
-  not change that. Instead, it seems to me that GP model informs us of the
-  regions to discard.  Thus, technically, I think ``We instead use ...'' is not
-  correctly justified. Modify at will. Obviously quite a minor point..}
 
@@
 visualization.
-As suggested by the features ({\bf the depth of the absorption features?}) in
-these 10 spectra, there is a mix of metallicities and spectral types amongst
-the observed targets.
+As suggested by the features (the number and depths of absorption features) in
+these 10 spectra, there is a mix of metallicities and spectral types amongst the
+observed targets.
 Note that the spectra are not flux-calibrated, as we only care about measuring
@@
 store these as our posterior samples.
-{\bf
+
 \figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
 (\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
-samples for a randomly-chosen source (HD 63408, in this case).}
+samples for a randomly-chosen source (HD 63408, in this case).
 The marginal posterior distribution over $x_0$ can be converted to a
@@
 
+\todo{TESS-HERMES}
 An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have

### 2017-07-28 paper/comoving-followup.tex
figures and shuffle notebooks

@@
 
-% this paragraph can be simplified
+\subsubsection{1D spectrum extraction}
+
 For each source or comparison lamp observation, the two-dimensional image is
@@
 2D images and do not bias the LSF extraction fits.
+
+\subsubsection{Wavelength calibration}
 
@@
 
+\subsubsection{Radial velocity measurement}
+
 To derive radial velocities for each source, we then measure the (wavelength)
@@
 
-\figurename~\ref{fig:Halpha-mcmc-corner} shows a \package{corner} plot
-(\citealt{Foreman-Mackey:2016}) of all projections of these posterior PDF
-samples for a randomly-chosen source (HD 63408, in this case).
+\figurename~\ref{fig:Halpha-mcmc-corner} shows a corner plot of all projections
+of these posterior PDF samples for a randomly-chosen source (HD 63408, in this
+case).
 The marginal posterior distribution over $x_0$ can be converted to a
@@
 the median parameters (orange curve) used to model other correlated structure in
-the vicinity of \Ha (which can be interpreted as correlated noise).
+the vicinity of \Ha\ (which can be interpreted as correlated noise).
 These figures are produced for all source spectrum fits and visually inspected
@@
 position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
-$J$-band magnitudes for all of our comoving star targets in the ``stars'' data
-table available with this \documentname; see \tablename~\ref{tbl:data-stars} for
-a detailed description of the content of this table.
+$J$-band magnitudes for all of our comoving star targets in
+\tablename~\ref{tbl:data-stars} available with this \documentname\ (see table
+caption for a detailed description of the contents).
 
@@
 derived radial velocities with high-quality prior measurements.
-Following previous work (\citealt{Shaya:2011}), we search for previous RV
-measurements (in decreasing order of priority) with:
+For calibration sources, we use radial velocity values reported in the source
+catalogs (\citealt{Bensby:2014,Pinsonneault:2014}).
+For all other sources, following previous work (\citealt{Shaya:2011}), we search
+for previous RV measurements (in decreasing order of priority) with:
 \begin{enumerate}
@@
 The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
-all observed candidate comoving pairs are provided with the associated ``pairs''
-table; see \tablename~\ref{tbl:data-pairs} for a detailed description of the
-content of this table.
+all observed candidate comoving pairs are provided with the associated
+\tablename~\ref{tbl:data-pairs} table (see table caption for a detailed
+description of the content).
 
@@
 
-Need much larger sample with no cuts on separation or signal-to-noise to study
-true separation distribution
-
-Need precise chemical information
-
-But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?
-
-Going forward, will have to use hierarchical mixture models to study population of singles in conjunction
-
-Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
+The results presented in this work demonstrate that even at separations larger
+than a parsec, a substantial fraction of comoving pair candidates selected from
+astrometry alone are confirmed with three-dimensional velocity information.
+This is exciting because it confirms the existence of what is likely a
+population of nearly-coherent but unbound coeval stars.
+It is also encouraging for the prospect of selecting large samples of comoving
+pairs purely with astrometric data from future data releases of the \gaia\
+mission (radial velocity measurements will only be released for bright stars, $G
+< 12$).
+However, in this work, we do not claim to report the true number or density of
+large-separation comoving pairs.
+To model the abundance and formation of comoving pairs, it will be crucial to
+select larger samples of pairs without hard cuts on separation, tangential
+velocity, signal-to-noise, or likelihood ratio, as were done with the input
+sample used in this work.
+
+% \subsection{Prospects for chemical tagging} \label{sec:chemical-tagging}
+
+% Need precise chemical information
+
+% But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?
+
+% Going forward, will have to use hierarchical mixture models to study population of singles in conjunction
+
+% Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
 Also tests of chemical enrichment due to planetary systems
 
-Abundances - future work, Coronado in prep.! But: throw shade on abundances
-because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
-2017b
-
-% \subsection{Prospects for chemical tagging} \label{sec:chemical-tagging}
+% Abundances - future work, Coronado in prep.! But: throw shade on abundances
+% because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
+% 2017b
 

### 2017-07-29 paper/comoving-followup.tex
fixing some other issues

@@
   \begin{center}
-    \begin{tabular}{ c | c }
+    \begin{tabular}{ c | c | l }
       \toprule
-        Parameter & Prior \\\toprule
-        $\ln A$ & $\mathcal{U}(2, 16)$\\
-        $x_0$ & $\mathcal{U}(6547, 6579)$\\
-        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$\\
-        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$\\
-        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$\\
-        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$\\
-        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$\\
-        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$\\
+        Parameter & Prior & Description \\\toprule
+        $\ln A$ & $\mathcal{U}(2, 16)$ & Logarithm of line amplitude
+          [${\rm photons}$]\\
+        $x_0$ & $\mathcal{U}(6547, 6579)$ & Line wavelength centroid
+          [${\rm \AA}$]\\
+        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$ & Logarithm of Gaussian standard
+          deviation of line profile [${\rm \AA}$]\\
+        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$ & Logarithm of Lorentzian HWHM of
+          line profile\\
+        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$ & Background model offset at line
+          centroid [${\rm photons}$]\\
+        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$ & Background model slope
+          [${\rm photons}~{\rm \AA}^{-1}$]\\
+        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$ & Logarithm of Gaussian process
+          amplitude [${\rm photons}$]\\
+        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$ & Logarithm of Gaussian process
+          correlation scale [${\rm \AA}$]\\
       \bottomrule
@@
       $\mu$ and variance $\sigma^2$.
-      Wavelength units are all angstroms.
-      \todo{what are the flux units?}
-      \smoh{A column of brief description of parameters would be useful.}
+      Flux units are photons, i.e. the spectra are not flux-calibrated.
       \label{tbl:prior-bounds}
@@
 
-\todo{TESS-HERMES}
 An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
@@
 
-\subsection{Caveats} \label{sec:caveats}
+% \subsection{Caveats and additional validation} \label{sec:caveats}
 
@@
 sample used in this work.
+It will also be important to quantify the contamination rate of (3D) comoving
+pair lookalikes, especially at large separations.
+
+We do not try to estimate a contamination rate in this work, but instead use
+metallicities and radial velocities from 61 candidate comoving pairs that are in
+\DR{3} of the \acronym{LAMOST} survey (\citealt{Deng:2012}) as an additional
+validation test.
+\figurename~\ref{fig:lamost} shows these pairs plotted as the difference in
+radial velocity and difference in iron abundance with points colored by the
+chord length separation of each pair.
+There is a clear over-density of pairs with abundance and radial-velocity
+differences close to zero.
+From this figure it is also clear that the false positives are predominantly the
+largest separation pairs (i.e. are the darkest markers), but a number of
+$>1~\pc$ pairs appear to have small radial-velocity differences and small
+differences in iron abundances, making them even more likely to be genuine
+pairs.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{lamost.pdf}
+  \end{center}
+  \caption{%
+    Stuff
+    \label{fig:lamost}}
+\end{figure}
+
+Going forward, definitive confirmation of comoving star pairs may require
+precise measurements of chemical abundances combined with precise radial
+velocities, as demonstrated with the \acronym{LAMOST} data.
+However, we would also like to \emph{test} assumptions about chemical
+homogeneity, rather than simply assuming that all true comoving pairs must have
+identical abundances.
+For example, the outcomes of planet formation may lead to large or appreciable
+abundance differences between coeval stars (e.g., \citealt{yo}, Oh et al.,
+submitted).
+We therefore caution that---to study the chemical properties of comoving pairs
+and star formation in general---the comoving pairs should be selected from their
+kinematics alone.
 
 % \subsection{Prospects for chemical tagging} \label{sec:chemical-tagging}
-
-% Need precise chemical information
-
-% But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?
-
-% Going forward, will have to use hierarchical mixture models to study population of singles in conjunction
-
-% Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc.
-Also tests of chemical enrichment due to planetary systems
-
-% Abundances - future work, Coronado in prep.! But: throw shade on abundances
-% because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
-% 2017b
 
@@
 
-% \subsection{Prospects for Galactic structure} \label{sec:dynamics}
-
-\subsection{Future spectroscopic surveys} \label{sec:future-rvs}
-
-DISCO / AS4, 4MOST, WEAVE, DESI, PFS
+% \subsection{Future spectroscopic surveys} \label{sec:future-rvs}
+
+% DISCO / AS4, 4MOST, WEAVE, DESI, PFS
 
@@
 % University Office of Information Technology's Research Computing department.
+Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic
+Telescope LAMOST) is a National Major Scientific Project built by the Chinese
+Academy of Sciences. Funding for the project has been provided by the National
+Development and Reform Commission. LAMOST is operated and managed by the
+National Astronomical Observatories, Chinese Academy of Sciences.
 

### 2017-07-29 paper/comoving-followup.tex
caption

@@
   \caption{%
-    Stuff
+    Differences in metallicity and radial velocity for candidate comoving pair
+    member stars from the \acronym{LAMOST} survey.
+    Points represent each pair, colored by the separation (chord length), and
+    error bars represent the uncertainties taken from the \acronym{LAMOST}
+    \DR{3} catalog.
+    There is a clear over-density of pairs with near-zero differences in
+    metallicity and radial velocity even at large ($>1~\pc$) separations, adding
+    even more confidence to the existence of comoving star pairs with these
+    extreme separations.
     \label{fig:lamost}}

### 2017-07-29 paper/comoving-followup.tex
refs

@@
 exchange of members, a hardening of one resulting binary, and a widening of the
-other binary (see, e.g., \citealt{Nathan??}).
+other binary (see, e.g., \citealt{Leigh:2016}).
 Or, the initial system could have been an unstable three-body system that
@@
 if the inner binary were close enough, it could have merged and formed a blue
-straggler star (e.g., \citealt{??}), leaving the wide binary untouched.
+straggler star (e.g., \citealt{Naoz:2014}), leaving the wide binary untouched.
 
@@
 will offer a new ``stereo'' view on the evolution and state of our Galaxy.
+
+\todo{Reiterate that monotonically decreasing separation distribution is too
+simple}
 

### 2017-07-29 paper/comoving-followup.tex
another ref

@@
 For example, the outcomes of planet formation may lead to large or appreciable
-abundance differences between coeval stars (e.g., \citealt{yo}, Oh et al.,
-submitted).
+abundance differences between coeval stars (e.g., \citealt{Pinsonneault:2001},
+Oh et al., submitted).
 We therefore caution that---to study the chemical properties of comoving pairs

### 2017-08-14 scripts/injest_db.py
for ruth

@@ -170,6 +170,8 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
 
             # get group id from object name
             if '-' in hdr['OBJECT']:
+                # Per APW and SMOH's convention
+
                 split_name = hdr['OBJECT'].split('-')
                 kw['group_id'] = int(split_name[0])
 
@@ -204,6 +206,37 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
                 else:
                     logger.log(1, 'simbad names for this object could not be '
                                'retrieved')
+
+            elif hdr['OBJECT'].startswith('k') or hdr['OBJECT'][0].isdigit():
+                # Assume it's a KIC number - per Ruth and Dan's convention
+
+                if hdr['OBJECT'].startswith('k'):
+                    object_name = 'KIC {0}'.format(hdr['OBJECT'][1:])
+
+                else:
+                    object_name = 'KIC {0}'.format(hdr['OBJECT'])
+
+                # query Simbad to get all possible names for this target
+                logger.log(1, 'common name: {0}'.format(object_name))
+
+                try:
+                    all_ids = Simbad.query_objectids(object_name)['ID'].astype(str)
+                except Exception as e:
+                    logger.warning('Simbad query_objectids failed for "{0}" '
+                                   'with error: {1}'
+                                   .format(object_name, str(e)))
+                    all_ids = []
+
+                logger.log(1, 'this is a KIC object')
+
+                if len(all_ids) > 0:
+                    logger.log(1, 'other names for this object: {0}'
+                               .format(', '.join(all_ids)))
+                else:
+                    logger.log(1, 'simbad names for this object could not be '
+                               'retrieved')
+
+                result_table = Simbad.query_object(object_name)
 
             else:
                 object_name = hdr['OBJECT']

### 2017-08-14 paper/comoving-followup.tex
final edits

@@
 
-\author{Adrian~M.~Price-Whelan}
+\author[0000-0003-0872-7098]{Adrian~M.~Price-Whelan}
 \affiliation{Department of Astrophysical Sciences,
@@
 
-\author{Semyeong~Oh}
+\author[0000-0001-7790-5308]{Semyeong~Oh}
 \affiliation{Department of Astrophysical Sciences,
@@
 The spectra were reduced and calibrated using a custom, publicly-available
-pipeline written in \python, described below.
-
-\subsubsection{1D spectrum extraction}
-
-For each source or comparison lamp observation, the two-dimensional image is
-bias corrected, flat-field corrected, and trimmed using standard routines
-implemented in \package{ccdproc} (\citealt{Craig:2015}).
-The 2D spectra are oriented such that the spatial direction is along a row of
-pixels, and the dispersion direction is along a column of pixels.
-For comparison lamp observations, the central 100 pixel columns are extracted
-and summed along the spatial axis, weighted by the inverse-variances of each
-pixel.
-Source spectral traces were always placed in this region during observations,
-and the curvature of the wavelength solution over this region is negligible.
-For source observations, 1D spectra are extracted using an empirical
-estimate of the line spread function (LSF).
-The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
-profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
-half-width at half maximum, and the background is modeled using a linear
-polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
-at the profile centroid, respectively.
-The maximum-likelihood LSF + background model parameters are estimated from
-each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
-the \package{scipy} package.
-We first determine the LSF width parameters, $\sigma_G$ and $\gamma_L$ from fits
-to 16 pixel rows evenly spaced between indices 750 and 850 (i.e. around the
-effective center of the dispersion axis).
-From these 16 fits, the median profile width parameters, $\sigma_G$ and
-$\gamma_L$, are then taken to be the LSF width parameters.
-Then, at each of 1600 rows in each source spectrum image, the LSF amplitude and
-centroid, and background slope and intercept at line centroid, are then fit
-using the same procedure as above but with the width parameters fixed.
-The LSF amplitudes are stored as the source fluxes, and the background model
-intercepts are stored as the background (sky) fluxes.
-Occasionally, nearby sources fall in the slit; these are masked by hand in the
-2D images and do not bias the LSF extraction fits.
-
-\subsubsection{Wavelength calibration}
-
-We start the wavelength calibration by interactively identifying the 12
-strongest, least blended emission lines in one of the extracted 1D comparison
-lamp spectra by comparing to standard HgNe line lists.\footnote{\url{http:
-//physics.nist.gov/PhysRefData/ASD/lines_form.html}}
-Once a rough mapping from pixel to wavelength is determined, we fit
-8 pixel-wide segments around each identified emission line
-for the individual line centroids in each comparison lamp spectrum.
-We use a pixel-convolved Gaussian profile for each emission line with
-parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
-We find that nearby blended or low-amplitude lines (i.e. part of the
-``background'') sometimes bias the line centroid determination when using a
-simple (i.e. low-order polynomial) background model.
-We therefore model the background flux around each comparison lamp emission
-line using a constant offset + Gaussian process (GP) in order to capture any
-unmasked structure from nearby sources.
-For the GP model, we use a Mat\'ern 3/2 kernel function
-(\citealt{Matern:1986,Rasmussen:2005}) parametrized by amplitude parameter
-$\sigma_{3/2}$ and correlation scale $\rho_{3/2}$.
-We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
-likelihood for this model, and maximize the (log-)likelihood again using
-\lbfgsb\ optimization.
-We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,
-$\sigma_{3/2}$, and $\rho_{3/2}$).
-Once we determine the pixel centroids for all identified emission lines for a
-given night's comparison lamp spectrum, we then fit an interpolating function
-to the pixel-centroid, wavelength values for each emission line.
-
-Typically, high-order standard polynomials or Chebyshev polynomials are used to
-fit for the wavelength dispersion function.
-We found that using such functions can lead to vastly incorrect wavelength
-values near either end of the pixel grid where the polynomial function is
-un- or weakly-constrained, and can easily over-fit to poorly centroided emission
-lines.
-We instead use a combined linear polynomial + GP to model the wavelength
-dispersion to flexibly account for nonlinear behavior in the pixel to wavelength
-mapping.
-We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
-scale parameter to log-space before maximizing the log-likelihood.
-\figurename~\ref{fig:wavelength-GP} shows an example of the residuals from one
-such fit; black points show the (pixel-centroid, wavelength) pairs with the
-best-fit model subtracted.
-The envelope around zero (shaded region, orange) shows the variance of the GP
-model projected into the data space; when the GP variance is large, the
-pixel-to-wavelength mapping is uncertain.
-We therefore only use the pixel range from 250--1210, corresponding to a
-wavelength range of $\approx 5500$--$7400~{\rm \AA}$; gray shaded region shows
-the excluded pixel/wavelength ranges.
-
-\figurename~\ref{fig:sample-spectra} shows a random sample of 10
-wavelength-calibrated source spectra of comoving star pair members normalized
-to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by an arbitrary offset for
-visualization.
-As suggested by the features (the number and depths of absorption features) in
-these 10 spectra, there is a mix of metallicities and spectral types amongst the
-observed targets.
-Note that the spectra are not flux-calibrated, as we only care about measuring
-radial velocities from this sample.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.8\linewidth]{wavelength_gp.pdf}
-  \end{center}
-  \caption{%
-    A visualization of the pixel-to-wavelength model determined from a single
-    wavelength comparison lamp spectrum.
-    Black markers show the fit pixel centroids for known Hg or Ne emission
-    lines and the residual wavelength value between the linear + GP model and
-    the true emission line centroid.
-    The non-shaded region shows the section of spectrum where the comparison
-    lamp spectrum has sufficient line density to determine the wavelength
-    solution; shaded regions are masked from the resulting spectra.
-    Orange envelope shows the root-variance of the Gaussian process component
-    of the model, which can be interpreted as inherent uncertainties on the
-    wavelength calibration as a result of the model flexibility.
-    \label{fig:wavelength-GP}}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.6\linewidth]{sample_spectra.pdf}
-  \end{center}
-  \caption{%
-    A section of 10 randomly selected target spectra from this work.
-    Spectral fluxes are normalized to 1 at $\lambda = 5500~{\rm \AA}$ and the
-    offsets are chosen to separate the spectra for visualization.
-    The spectra are smoothed with a Gaussian filter with root-variance $\sigma
-    = 2~{\rm \AA}$.
-    \label{fig:sample-spectra}}
-\end{figure}
-
-The wavelength calibration procedure described above produces a nightly model
-for the nonlinear dispersion of wavelength as a function of pixel value for
-each extracted 1D spectrum.
-However, because of flexure in the spectrograph, each observation will have an
-additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
-a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
-As mentioned, comparison lamp spectra were not obtained at each telescope
-pointing.
-The absolute wavelength solution at each pointing can therefore be shifted
-(in pixels) relative to the nightly comparison lamp spectrum because of flexure
-in the instrument.
-We have found that this offset depends primarily on the hour angle of the
-observation, likely because of the physical orientation of the spectrograph and
-CCD as mounted on the telescope.
-To correct each source spectrum for this offset, we measure the centroids of
-the two prominent [OI] sky emission lines near $5577~{\rm \AA}$ and $6300~{\rm
-\AA}$ in the corresponding background spectrum derived from the spectral
-extraction (described above).
-We use a pixel-convolved Voigt profile as the emission line model (same
-parameters as above) and a Gaussian process with a Mat\'ern 3/2 kernel function
-for the background spectrum (same parameters as above).
-We again use \lbfgsb\ optimization and convert any scale parameter to log-space
-before maximizing the log-likelihood of this model to derive the line centroid
-in wavelength units.
-We compute wavelength offsets from true (air) values of the line centroids,
-$5577.3387~{\rm \AA}$ and $6300.304~{\rm \AA}$ respectively,\footnote{\url{http:
-//physics.nist.gov/PhysRefData/ASD/lines_form.html}} then convert these
-wavelength offsets to pixel offsets.
-We take the mean of the pixel offsets and apply the derived shift to the source
-spectrum before re-computing the wavelength grid using the nightly wavelength
-solution.
-When one or both of the two sky lines is not present or the centroid fit fails,
-the spectrum is flagged as having a suspect wavelength calibration.
-In practice, most such sources are calibration targets, which are more likely to
-have very short exposure times and thus weak night-sky lines.
-
-\begin{table}[ht]
-  \begin{center}
-    \begin{tabular}{ c | c | l }
-      \toprule
-        Parameter & Prior & Description \\\toprule
-        $\ln A$ & $\mathcal{U}(2, 16)$ & Logarithm of line amplitude
-          [${\rm photons}$]\\
-        $x_0$ & $\mathcal{U}(6547, 6579)$ & Line wavelength centroid
-          [${\rm \AA}$]\\
-        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$ & Logarithm of Gaussian standard
-          deviation of line profile [${\rm \AA}$]\\
-        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$ & Logarithm of Lorentzian HWHM of
-          line profile\\
-        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$ & Background model offset at line
-          centroid [${\rm photons}$]\\
-        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$ & Background model slope
-          [${\rm photons}~{\rm \AA}^{-1}$]\\
-        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$ & Logarithm of Gaussian process
-          amplitude [${\rm photons}$]\\
-        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$ & Logarithm of Gaussian process
-          correlation scale [${\rm \AA}$]\\
-      \bottomrule
-    \end{tabular}
-    \caption{
-      Priors used for \Ha\ line profile model parameters.
-      $\mathcal{U}(a, b)$ indicates a uniform prior between $a$ and $b$ and
-      $\mathcal{N}(\mu, \sigma^2)$ indicates a normal distribution with mean
-      $\mu$ and variance $\sigma^2$.
-      Flux units are photons, i.e. the spectra are not flux-calibrated.
-      \label{tbl:prior-bounds}
-    }
-  \end{center}
-\end{table}
-
-\subsubsection{Radial velocity measurement}
-
-To derive radial velocities for each source, we then measure the (wavelength)
-line centroid of \Ha\ in each corrected spectrum.
-We model the absorption line using a pixel-convolved Voigt profile and use a
-linear model + a Gaussian process for the background model.
-We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
-to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
-probability distribution (PDF) over the 8 parameters: $\ln |A|$, $x_0$,
-$\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
-$\ln\rho_{3/2}$ where the linear background model is defined as $f(\lambda) =
-\alpha_1 + \alpha_2\,\lambda$.
-The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
-an absorption feature) and is computed from the corresponding parameter as $A =
--\exp(\ln |A|)$.
-We use uniform priors on all parameters except $\ln\rho_{3/2}$, for which we use
-a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
-We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
-generate samples from the posterior PDF.
-We initially run the sampler for 128 steps from the maximum likelihood parameter
-values, then resample walker positions in a small ball around the median final
-walker positions and run again for 512 steps; together, these two steps
-constitute our burn-in phase.
-From the final walker positions after burn-in, we run again for 1024 steps and
-store these as our posterior samples.
-
-\figurename~\ref{fig:Halpha-mcmc-corner} shows a corner plot of all projections
-of these posterior PDF samples for a randomly-chosen source (HD 63408, in this
-case).
-The marginal posterior distribution over $x_0$ can be converted to a
-distribution over radial velocity; in this case, the estimated (Gaussian
-standard deviation) precision is $\approx 2.5~\kms$.
-\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
-around \Ha\ used in the line modeling (black line) with uncertainties (gray
-bars).
-Top panel shows the linear polynomial + absorption line model computed for the
-median sample (dark green line); shaded region shows the 15th to 85th
-percentiles computed from all posterior samples.
-Bottom panel shows the residuals of the data from the median linear polynomial +
-absorption line model (black line) along with the Gaussian process computed from
-the median parameters (orange curve) used to model other correlated structure in
-the vicinity of \Ha\ (which can be interpreted as correlated noise).
-These figures are produced for all source spectrum fits and visually inspected
-to ensure that the MCMC sampling appears to have converged for all cases.
-
-For each source spectrum, with the resulting posterior samples, we compute the
-median and median absolute deviation (MAD) of the marginal distribution over
-line centroid, $x_0$.
-We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
-MAD}$, assuming the samples are Gaussian distributed) of the samples and find
-that the typical velocity \emph{precision} from measuring the centroid of \Ha\
-is between $5$--$10~\kms$.
-We report the measured absolute radial velocities along with compiled sky
-position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
-$J$-band magnitudes for all of our comoving star targets in
-\tablename~\ref{tbl:data-stars} available with this \documentname\ (see table
-caption for a detailed description of the contents).
-
-\begin{figure}[!hp]
-  \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
-  \end{center}
-  \caption{%
-    A corner plot showing all projections of posterior PDF samples for the \Ha\
-    line model and the spectrum of the source HD 63408.
-    Note the small uncertainty in the line centroid measurement, $x_0$, which
-    corresponds to a $\approx 2~\kms$ velocity precision for this source.
-    Parameters are described in \sectionname~\ref{sec:reduction}, and prior PDFs
-    over the parameters are described in \tablename~\ref{tbl:prior-bounds}.
-    \label{fig:Halpha-mcmc-corner}}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.7\linewidth]{mcmc_example_fit.pdf}
-  \end{center}
-  \caption{%
-    Visualization of the \Ha\ absorption line model and fits for the source HD
-    63408 as a demonstration of the procedure used for all targets.
-    \emph{Top}:
-    The source spectrum in a small wavelength range around \Ha\ (black),
-    uncertainty in spectral flux (gray error bars).
-    Over-plotted is the line model + background computed from the median
-    posterior PDF sample (dark green line) and the 15th to 85th percentile
-    region computed from all posterior PDF samples (shaded green region).
-    \emph{Bottom}:
-    Source spectrum residuals computed by subtracting the median model (dark
-    green line in top panel) from the source spectrum.
-    The over-plotted line (orange) shows the Gaussian process background model
-    computed from the median posterior PDF sample.
-    \label{fig:Halpha-mcmc-fit}}
-\end{figure}
-
-\begin{table}[htb]
-    \centering
-    \caption{Description of columns in data table containing results from radial
-    velocity follow-up.} \label{tbl:data-stars}
-    \begin{tabular}{l|l|l}
-        \toprule
-        Column Name    & Unit & Description\\
-        \midrule
-        \texttt{Oh17\_id}         &      & Star row index from the candidate catalog (\citealt{Oh:2017})\\
-        \texttt{Oh17\_group\_id}  &      & Group ID from the candidate catalog (\citealt{Oh:2017})\\
-        \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
-        \texttt{name}             &      & HD, \project{Hipparcos}, or \project{Tycho-2} identifier\\
-        \texttt{ra}               & deg  & Right ascension from \tgas\\
-        \texttt{dec}              & deg  & Declination from \tgas\\
-        \texttt{parallax}         & mas  & Parallax from \tgas\\
-        \texttt{distance}         & pc   & Distance estimate using inverse parallax from \tgas\\
-        \texttt{G}                & mag  & \gaia\ $G$-band magnitudes\\
-        \texttt{J}                & mag  & \tmass\ $J$-band magnitudes\\
-        \texttt{rv}               & \kms & Radial velocity measured in this work\\
-        \texttt{rv\_err}          & \kms & Radial velocity error\\
-        \bottomrule
-    \end{tabular}
-\end{table}
-
-% \clearpage
+pipeline written in \python; see Appendices~\ref{sec:1d-ext}--\ref{sec:rv-meas}
+for a full description of the data reduction procedure.
 
@@
 catalog.
-The separation distribution of these spectroscopically-confirmed comoving star
+From a cross-match to the \acronym{LAMOST} survey, we find that candidate
+comoving star pairs with small differences in radial velocities also have small
+differences in iron abundance, even at large separations ($>1~\pc$).
+The separation distribution of the spectroscopically-confirmed comoving star
 pairs presented shows an abundance of comoving pairs at separations between
@@
 perturbers in the Galaxy.
+From this work it appears that modeling the population of wide binaries or
+comoving stars with a monotonically decreasing separation distribution---as is commonly done---too simple to also represent the population of unbound systems.
 With future data releases from the \gaia\ mission combined with large
@@
 will offer a new ``stereo'' view on the evolution and state of our Galaxy.
-
-\todo{Reiterate that monotonically decreasing separation distribution is too
-simple}
 
@@
 France.
-% The authors are pleased to acknowledge that the work reported on in this
-% paper was substantially performed at the TIGRESS high performance computer
-% center at Princeton University which is jointly supported by the Princeton
-% Institute for Computational Science and Engineering and the Princeton
-% University Office of Information Technology's Research Computing department.
+This project was developed in part at the 2017 Heidelberg Gaia Sprint, hosted by the Max-Planck-Institut für Astronomie, Heidelberg.
+
 Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic
@@
 
+\appendix
+
+\section{1D spectrum extraction} \label{sec:1d-ext}
+
+For each source or comparison lamp observation, the two-dimensional image is
+bias corrected, flat-field corrected, and trimmed using standard routines
+implemented in \package{ccdproc} (\citealt{Craig:2015}).
+The 2D spectra are oriented such that the spatial direction is along a row of
+pixels, and the dispersion direction is along a column of pixels.
+For comparison lamp observations, the central 100 pixel columns are extracted
+and summed along the spatial axis, weighted by the inverse-variances of each
+pixel.
+Source spectral traces were always placed in this region during observations,
+and the curvature of the wavelength solution over this region is negligible.
+For source observations, 1D spectra are extracted using an empirical
+estimate of the line spread function (LSF).
+The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
+profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
+half-width at half maximum, and the background is modeled using a linear
+polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
+at the profile centroid, respectively.
+The maximum-likelihood LSF + background model parameters are estimated from
+each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
+the \package{scipy} package.
+We first determine the LSF width parameters, $\sigma_G$ and $\gamma_L$ from fits
+to 16 pixel rows evenly spaced between indices 750 and 850 (i.e. around the
+effective center of the dispersion axis).
+From these 16 fits, the median profile width parameters, $\sigma_G$ and
+$\gamma_L$, are then taken to be the LSF width parameters.
+Then, at each of 1600 rows in each source spectrum image, the LSF amplitude and
+centroid, and background slope and intercept at line centroid, are then fit
+using the same procedure as above but with the width parameters fixed.
+The LSF amplitudes are stored as the source fluxes, and the background model
+intercepts are stored as the background (sky) fluxes.
+Occasionally, nearby sources fall in the slit; these are masked by hand in the
+2D images and do not bias the LSF extraction fits.
+
+\section{Wavelength calibration}
+
+We start the wavelength calibration by interactively identifying the 12
+strongest, least blended emission lines in one of the extracted 1D comparison
+lamp spectra by comparing to standard HgNe line lists.\footnote{\url{http:
+//physics.nist.gov/PhysRefData/ASD/lines_form.html}}
+Once a rough mapping from pixel to wavelength is determined, we fit
+8 pixel-wide segments around each identified emission line
+for the individual line centroids in each comparison lamp spectrum.
+We use a pixel-convolved Gaussian profile for each emission line with
+parameters: amplitude, $A$; centroid, $x_0$; and root-variance, $\sigma_G$.
+We find that nearby blended or low-amplitude lines (i.e. part of the
+``background'') sometimes bias the line centroid determination when using a
+simple (i.e. low-order polynomial) background model.
+We therefore model the background flux around each comparison lamp emission
+line using a constant offset + Gaussian process (GP) in order to capture any
+unmasked structure from nearby sources.
+For the GP model, we use a Mat\'ern 3/2 kernel function
+(\citealt{Matern:1986,Rasmussen:2005}) parametrized by amplitude parameter
+$\sigma_{3/2}$ and correlation scale $\rho_{3/2}$.
+We use \package{celerite} (\citealt{Foreman-Mackey:2017}) to evaluate the full
+likelihood for this model, and maximize the (log-)likelihood again using
+\lbfgsb\ optimization.
+We optimize over the logarithm of any scale parameters (i.e. $A$, $\sigma_G$,
+$\sigma_{3/2}$, and $\rho_{3/2}$).
+Once we determine the pixel centroids for all identified emission lines for a
+given night's comparison lamp spectrum, we then fit an interpolating function
+to the pixel-centroid, wavelength values for each emission line.
+
+Typically, high-order standard polynomials or Chebyshev polynomials are used to
+fit for the wavelength dispersion function.
+We found that using such functions can lead to vastly incorrect wavelength
+values near either end of the pixel grid where the polynomial function is
+un- or weakly-constrained, and can easily over-fit to poorly centroided emission
+lines.
+We instead use a combined linear polynomial + GP to model the wavelength
+dispersion to flexibly account for nonlinear behavior in the pixel to wavelength
+mapping.
+We again use a Mat\'ern 3/2 kernel, use \lbfgsb\ optimization, and convert any
+scale parameter to log-space before maximizing the log-likelihood.
+\figurename~\ref{fig:wavelength-GP} shows an example of the residuals from one
+such fit; black points show the (pixel-centroid, wavelength) pairs with the
+best-fit model subtracted.
+The envelope around zero (shaded region, orange) shows the variance of the GP
+model projected into the data space; when the GP variance is large, the
+pixel-to-wavelength mapping is uncertain.
+We therefore only use the pixel range from 250--1210, corresponding to a
+wavelength range of $\approx 5500$--$7400~{\rm \AA}$; gray shaded region shows
+the excluded pixel/wavelength ranges.
+
+\figurename~\ref{fig:sample-spectra} shows a random sample of 10
+wavelength-calibrated source spectra of comoving star pair members normalized
+to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by an arbitrary offset for
+visualization.
+As suggested by the features (the number and depths of absorption features) in
+these 10 spectra, there is a mix of metallicities and spectral types amongst the
+observed targets.
+Note that the spectra are not flux-calibrated, as we only care about measuring
+radial velocities from this sample.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.8\linewidth]{wavelength_gp.pdf}
+  \end{center}
+  \caption{%
+    A visualization of the pixel-to-wavelength model determined from a single
+    wavelength comparison lamp spectrum.
+    Black markers show the fit pixel centroids for known Hg or Ne emission
+    lines and the residual wavelength value between the linear + GP model and
+    the true emission line centroid.
+    The non-shaded region shows the section of spectrum where the comparison
+    lamp spectrum has sufficient line density to determine the wavelength
+    solution; shaded regions are masked from the resulting spectra.
+    Orange envelope shows the root-variance of the Gaussian process component
+    of the model, which can be interpreted as inherent uncertainties on the
+    wavelength calibration as a result of the model flexibility.
+    \label{fig:wavelength-GP}}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.6\linewidth]{sample_spectra.pdf}
+  \end{center}
+  \caption{%
+    A section of 10 randomly selected target spectra from this work.
+    Spectral fluxes are normalized to 1 at $\lambda = 5500~{\rm \AA}$ and the
+    offsets are chosen to separate the spectra for visualization.
+    The spectra are smoothed with a Gaussian filter with root-variance $\sigma
+    = 2~{\rm \AA}$.
+    \label{fig:sample-spectra}}
+\end{figure}
+
+The wavelength calibration procedure described above produces a nightly model
+for the nonlinear dispersion of wavelength as a function of pixel value for
+each extracted 1D spectrum.
+However, because of flexure in the spectrograph, each observation will have an
+additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
+a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$.
+As mentioned, comparison lamp spectra were not obtained at each telescope
+pointing.
+The absolute wavelength solution at each pointing can therefore be shifted
+(in pixels) relative to the nightly comparison lamp spectrum because of flexure
+in the instrument.
+We have found that this offset depends primarily on the hour angle of the
+observation, likely because of the physical orientation of the spectrograph and
+CCD as mounted on the telescope.
+To correct each source spectrum for this offset, we measure the centroids of
+the two prominent [OI] sky emission lines near $5577~{\rm \AA}$ and $6300~{\rm
+\AA}$ in the corresponding background spectrum derived from the spectral
+extraction (described above).
+We use a pixel-convolved Voigt profile as the emission line model (same
+parameters as above) and a Gaussian process with a Mat\'ern 3/2 kernel function
+for the background spectrum (same parameters as above).
+We again use \lbfgsb\ optimization and convert any scale parameter to log-space
+before maximizing the log-likelihood of this model to derive the line centroid
+in wavelength units.
+We compute wavelength offsets from true (air) values of the line centroids,
+$5577.3387~{\rm \AA}$ and $6300.304~{\rm \AA}$ respectively,\footnote{\url{http:
+//physics.nist.gov/PhysRefData/ASD/lines_form.html}} then convert these
+wavelength offsets to pixel offsets.
+We take the mean of the pixel offsets and apply the derived shift to the source
+spectrum before re-computing the wavelength grid using the nightly wavelength
+solution.
+When one or both of the two sky lines is not present or the centroid fit fails,
+the spectrum is flagged as having a suspect wavelength calibration.
+In practice, most such sources are calibration targets, which are more likely to
+have very short exposure times and thus weak night-sky lines.
+
+\begin{table}[ht]
+  \begin{center}
+    \begin{tabular}{ c | c | l }
+      \toprule
+        Parameter & Prior & Description \\\toprule
+        $\ln A$ & $\mathcal{U}(2, 16)$ & Logarithm of line amplitude
+          [${\rm photons}$]\\
+        $x_0$ & $\mathcal{U}(6547, 6579)$ & Line wavelength centroid
+          [${\rm \AA}$]\\
+        $\ln\sigma_G$ & $\mathcal{U}(-4, 2)$ & Logarithm of Gaussian standard
+          deviation of line profile [${\rm \AA}$]\\
+        $\ln\gamma_L$ & $\mathcal{U}(-4, 2)$ & Logarithm of Lorentzian HWHM of
+          line profile\\
+        $\alpha_1$ & $\mathcal{U}(0, 10^{16})$ & Background model offset at line
+          centroid [${\rm photons}$]\\
+        $\alpha_2$ & $\mathcal{U}(-10^8, 10^8)$ & Background model slope
+          [${\rm photons}~{\rm \AA}^{-1}$]\\
+        $\ln\sigma_{3/2}$ & $\mathcal{U}(-8, 8)$ & Logarithm of Gaussian process
+          amplitude [${\rm photons}$]\\
+        $\ln\rho_{3/2}$ & $\mathcal{N}(1, 1)$ & Logarithm of Gaussian process
+          correlation scale [${\rm \AA}$]\\
+      \bottomrule
+    \end{tabular}
+    \caption{
+      Priors used for \Ha\ line profile model parameters.
+      $\mathcal{U}(a, b)$ indicates a uniform prior between $a$ and $b$ and
+      $\mathcal{N}(\mu, \sigma^2)$ indicates a normal distribution with mean
+      $\mu$ and variance $\sigma^2$.
+      Flux units are photons, i.e. the spectra are not flux-calibrated.
+      \label{tbl:prior-bounds}
+    }
+  \end{center}
+\end{table}
+
+\section{Radial velocity measurement} \label{sec:rv-meas}
+
+To derive radial velocities for each source, we then measure the (wavelength)
+line centroid of \Ha\ in each corrected spectrum.
+We model the absorption line using a pixel-convolved Voigt profile and use a
+linear model + a Gaussian process for the background model.
+We now use \lbfgsb\ optimization and the resulting maximum-likelihood parameters
+to initialize a Markov Chain Monte Carlo (MCMC) sampling of the posterior
+probability distribution (PDF) over the 8 parameters: $\ln |A|$, $x_0$,
+$\ln\sigma_G$, $\ln\gamma_L$, $\alpha_1$, $\alpha_2$, $\ln\sigma_{3/2}$, and
+$\ln\rho_{3/2}$ where the linear background model is defined as $f(\lambda) =
+\alpha_1 + \alpha_2\,\lambda$.
+The linear amplitude, $A$, is assumed to be negative (i.e. \Ha\ is assumed to be
+an absorption feature) and is computed from the corresponding parameter as $A =
+-\exp(\ln |A|)$.
+We use uniform priors on all parameters except $\ln\rho_{3/2}$, for which we use
+a Gaussian; prior parameters are specified in \tablename~\ref{tbl:prior-bounds}.
+We use \package{emcee} (\citealt{Foreman-Mackey:2013}) with 64 walkers to
+generate samples from the posterior PDF.
+We initially run the sampler for 128 steps from the maximum likelihood parameter
+values, then resample walker positions in a small ball around the median final
+walker positions and run again for 512 steps; together, these two steps
+constitute our burn-in phase.
+From the final walker positions after burn-in, we run again for 1024 steps and
+store these as our posterior samples.
+
+\figurename~\ref{fig:Halpha-mcmc-corner} shows a corner plot of all projections
+of these posterior PDF samples for a randomly-chosen source (HD 63408, in this
+case).
+The marginal posterior distribution over $x_0$ can be converted to a
+distribution over radial velocity; in this case, the estimated (Gaussian
+standard deviation) precision is $\approx 2.5~\kms$.
+\figurename~\ref{fig:Halpha-mcmc-fit} shows the segment of the source spectrum
+around \Ha\ used in the line modeling (black line) with uncertainties (gray
+bars).
+Top panel shows the linear polynomial + absorption line model computed for the
+median sample (dark green line); shaded region shows the 15th to 85th
+percentiles computed from all posterior samples.
+Bottom panel shows the residuals of the data from the median linear polynomial +
+absorption line model (black line) along with the Gaussian process computed from
+the median parameters (orange curve) used to model other correlated structure in
+the vicinity of \Ha\ (which can be interpreted as correlated noise).
+These figures are produced for all source spectrum fits and visually inspected
+to ensure that the MCMC sampling appears to have converged for all cases.
+
+For each source spectrum, with the resulting posterior samples, we compute the
+median and median absolute deviation (MAD) of the marginal distribution over
+line centroid, $x_0$.
+We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
+MAD}$, assuming the samples are Gaussian distributed) of the samples and find
+that the typical velocity \emph{precision} from measuring the centroid of \Ha\
+is between $5$--$10~\kms$.
+We report the measured absolute radial velocities along with compiled sky
+position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
+$J$-band magnitudes for all of our comoving star targets in
+\tablename~\ref{tbl:data-stars} available with this \documentname\ (see table
+caption for a detailed description of the contents).
+
+\begin{figure}[!hp]
+  \begin{center}
+    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
+  \end{center}
+  \caption{%
+    A corner plot showing all projections of posterior PDF samples for the \Ha\
+    line model and the spectrum of the source HD 63408.
+    Note the small uncertainty in the line centroid measurement, $x_0$, which
+    corresponds to a $\approx 2~\kms$ velocity precision for this source.
+    Parameters are described in \sectionname~\ref{sec:reduction}, and prior PDFs
+    over the parameters are described in \tablename~\ref{tbl:prior-bounds}.
+    \label{fig:Halpha-mcmc-corner}}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.7\linewidth]{mcmc_example_fit.pdf}
+  \end{center}
+  \caption{%
+    Visualization of the \Ha\ absorption line model and fits for the source HD
+    63408 as a demonstration of the procedure used for all targets.
+    \emph{Top}:
+    The source spectrum in a small wavelength range around \Ha\ (black),
+    uncertainty in spectral flux (gray error bars).
+    Over-plotted is the line model + background computed from the median
+    posterior PDF sample (dark green line) and the 15th to 85th percentile
+    region computed from all posterior PDF samples (shaded green region).
+    \emph{Bottom}:
+    Source spectrum residuals computed by subtracting the median model (dark
+    green line in top panel) from the source spectrum.
+    The over-plotted line (orange) shows the Gaussian process background model
+    computed from the median posterior PDF sample.
+    \label{fig:Halpha-mcmc-fit}}
+\end{figure}
+
+\begin{table}[htb]
+    \centering
+    \caption{Description of columns in data table containing results from radial
+    velocity follow-up.} \label{tbl:data-stars}
+    \begin{tabular}{l|l|l}
+        \toprule
+        Column Name    & Unit & Description\\
+        \midrule
+        \texttt{Oh17\_id}         &      & Star row index from the candidate catalog (\citealt{Oh:2017})\\
+        \texttt{Oh17\_group\_id}  &      & Group ID from the candidate catalog (\citealt{Oh:2017})\\
+        \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
+        \texttt{name}             &      & HD, \project{Hipparcos}, or \project{Tycho-2} identifier\\
+        \texttt{ra}               & deg  & Right ascension from \tgas\\
+        \texttt{dec}              & deg  & Declination from \tgas\\
+        \texttt{parallax}         & mas  & Parallax from \tgas\\
+        \texttt{distance}         & pc   & Distance estimate using inverse parallax from \tgas\\
+        \texttt{G}                & mag  & \gaia\ $G$-band magnitudes\\
+        \texttt{J}                & mag  & \tmass\ $J$-band magnitudes\\
+        \texttt{rv}               & \kms & Radial velocity measured in this work\\
+        \texttt{rv\_err}          & \kms & Radial velocity error\\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+% \clearpage
+
 \bibliographystyle{aasjournal}

### 2017-08-15 paper/comoving-followup.tex
Some spergel comments - preflafter

@@
 astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog and
-find an abundance of pairs at separations $\gtrsim 1~\pc$.
+find a large number of pairs with separations $\gtrsim 1~\pc$.
 Without radial velocity measurements for the vast majority of these pairs, the
@@
 They have been used to place limits on the mass distribution and population of
-massive perturbers in the Galaxy (e.g., \citealt{Yoo:2004}), to study the
-physics of star formation (e.g., \citealt{Parker:2009,Reipurth:2012}),
-and to calibrate stellar models and inferences across the main sequence (e.g.,
-\citealt{Brewer:2016}).
+massive perturbers in the Galaxy (e.g.,
+\citealt{Yoo:2004,Monroy-Rodriguez:2014}), to study the physics of star
+formation (e.g., \citealt{Parker:2009,Reipurth:2012}), and to calibrate stellar
+models and inferences across the main sequence (e.g., \citealt{Brewer:2016}).
 These examples represent a small subset of a much broader range of applications
@@
 Briefly, we perform a model selection between
-two cases: that the velocities of the two stars in a pair are
+two hypotheses: that the velocities of the two stars in a pair are
 identical (case 1) or that they are indepndent (case 2).
@@
 comparison lamp spectra at each pointing as it adds an overhead of $\approx
-90~{\rm s}$ per observation.
+90~{\rm s}$ per observation; we are primarily interested in relative velocities
+between pair members and therefore don't need precise absolute calibrations of
+their radial velocities.
 To determine rough nightly wavelength solutions, calibration spectra were
@@
   \begin{center}
-    \includegraphics[width=0.7\linewidth]{llr-vs-dv.pdf}
+    \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
   \end{center}
@@
 radial velocity data as \llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
+We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
 \ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
@@
 description of the content).
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.7\linewidth]{dx-dv.pdf}
+  \end{center}
+  \caption{%
+    Point estimate of the 3D velocity difference between two stars in a comoving
+    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
+    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
+    The distances used in the point estimates are computed using a
+    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
+    \label{fig:dx-dv}}
+\end{figure}
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{dx-dv.pdf}
-  \end{center}
-  \caption{%
-    Point estimate of the 3D velocity difference between two stars in a comoving
-    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
-    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
-    The distances used in the point estimates are computed using a
-    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
-    \label{fig:dx-dv}}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
     \includegraphics[width=\linewidth]{separation-hist.pdf}
@@
 
-We naively expect that comoving, coeval stars should lie along lines parallel to
-the main sequence in color-magnitude space. The majority of our star pairs
-appear to follow this expectation, with the exception of a few visible outliers.
+We naively expect that comoving, coeval stars should lie along isochrones
+color-magnitude space.
+The majority of our star pairs appear to follow this expectation, with the
+exception of a few visible outliers.
 For example, there appear to be at least a few main-sequence--giant-branch
@@
 David W. Hogg (NYU),
-Melissa Ness (MPIA), and
-Dave Zurek (AMNH)
+Nathan Leigh (AMNH),
+Melissa Ness (MPIA),
+Dave Zurek (AMNH),
 for useful discussions and advice.
+We also thank all attendees of the ``Stars'' group meeting at the Center for
+Computational Astrophysics at the Flatiron Institute.
+The Flatiron Institute is supported by the Simons Foundation.
+
 We thank James Davenport (WWU) for releasing his open source spectroscopic
@@
 France.
-This project was developed in part at the 2017 Heidelberg Gaia Sprint, hosted by the Max-Planck-Institut für Astronomie, Heidelberg.
+This project was developed in part at the 2017 Heidelberg Gaia Sprint, hosted by
+the Max-Planck-Institut für Astronomie, Heidelberg.
 

### 2017-08-15 paper/comoving-followup.tex
figure shuffle

@@
 
-\subsection{Sample selection}\label{sec:sample}
-
-In previous work, we have identified a sample of high-confidence, candidate
-comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
-the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
-These comoving pairs were selected using a cut on a marginalized likelihood
-ratio computed for all nearly co-spatial pairs of stars.
-Briefly, we perform a model selection between
-two hypotheses: that the velocities of the two stars in a pair are
-identical (case 1) or that they are indepndent (case 2).
-We marginalize over the (unknown) true velocity and distance of each star
-to obtain the ratio of fully marginalized likelihoods,
-$\mathcal{L}_1/\mathcal{L}_2$.
-The likelihoods appropriately take into account the reported uncertainties and
-covariances of the astrometric measurements in the \tgas\ catalog
-as well as projection effect in comparing spherical
-velocities for star pairs separated by large angles on the sky.
-After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
-we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
-with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each
-target star from the \tgas\ catalog.
-After applying a conservative cut on the likelihood ratio motivated by computing
-the same likelihood ratio for random pairings of stars ($\ln
-\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
-physical separations between $10^3~\au$ and $10~\pc$.
-Of these pairs, there are 10,606 unique sources, implying that a significant
-fraction of the identified pairs are members of larger groups.
-However, we have also found a large number ($\approx$4,000) of mutually
-exclusively-linked pairs in increasing number with increasing separation at
-$\sim 1$~\pc.
-For a full description of the selection method, see \citealt{Oh:2017}.
-
-\begin{figure}[ht!]
+\begin{figure}[htb!]
   \begin{center}
@@
 \end{figure}
+
+\subsection{Sample selection}\label{sec:sample}
+
+In previous work, we have identified a sample of high-confidence, candidate
+comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
+the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
+These comoving pairs were selected using a cut on a marginalized likelihood
+ratio computed for all nearly co-spatial pairs of stars.
+Briefly, we perform a model selection between
+two hypotheses: that the velocities of the two stars in a pair are
+identical (case 1) or that they are indepndent (case 2).
+We marginalize over the (unknown) true velocity and distance of each star
+to obtain the ratio of fully marginalized likelihoods,
+$\mathcal{L}_1/\mathcal{L}_2$.
+The likelihoods appropriately take into account the reported uncertainties and
+covariances of the astrometric measurements in the \tgas\ catalog
+as well as projection effect in comparing spherical
+velocities for star pairs separated by large angles on the sky.
+
+After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
+we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
+with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each
+target star from the \tgas\ catalog.
+After applying a conservative cut on the likelihood ratio motivated by computing
+the same likelihood ratio for random pairings of stars ($\ln
+\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
+physical separations between $10^3~\au$ and $10~\pc$.
+Of these pairs, there are 10,606 unique sources, implying that a significant
+fraction of the identified pairs are members of larger groups.
+However, we have also found a large number ($\approx$4,000) of mutually
+exclusively-linked pairs in increasing number with increasing separation at
+$\sim 1$~\pc.
+For a full description of the selection method, see \citealt{Oh:2017}.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htb]
   \begin{center}
-    \includegraphics[width=\linewidth]{rv-comparison.pdf}
+    \includegraphics[width=0.85\linewidth]{rv-comparison.pdf}
   \end{center}
@@
 The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
-all observed candidate comoving pairs are provided with the associated
-\tablename~\ref{tbl:data-pairs} table (see table caption for a detailed
-description of the content).
-
-\begin{figure}[htbp]
+all observed candidate comoving pairs are provided with the associated table
+(\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).
+
+\begin{figure}[htb]
   \begin{center}
-    \includegraphics[width=0.7\linewidth]{dx-dv.pdf}
+    \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
   \end{center}
@@
 \end{figure}
-
-\begin{table}[htb]
-    \centering
-    \caption{Description of columns in data table containing results from radial
-      velocity follow-up.} \label{tbl:data-pairs}
-    \begin{tabular}{l|l|l}
-        \toprule
-        Column Name    & Unit & Description\\
-        \midrule
-        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
-        \texttt{R\_mu}            &      & Log-likelihood-ratio computed with only astrometric information\\
-        \texttt{R\_rv}            &      & Log-likelihood-ratio computed with radial velocities as well\\
-        \texttt{dx\_50}           & \pc  & Median of 3D spatial separation error distribution\\
-        \texttt{dv\_50}           & \kms & Median of 3D velocity difference error distribution\\
-        \texttt{dx\_15}           & \pc  & 15th percentile of 3D spatial separation error distribution\\
-        \texttt{dx\_85}           & \pc  & 85th percentile of 3D spatial separation error distribution\\
-        \texttt{dv\_15}           & \kms & 15th percentile of 3D velocity difference error distribution\\
-        \texttt{dv\_85}           & \kms & 85th percentile of 3D velocity difference error distribution\\
-        \bottomrule
-    \end{tabular}
-\end{table}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htb]
   \begin{center}
@@
 
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.8\linewidth]{genuine_cmd.pdf}
-  \end{center}
-  \caption{%
-    Color-magnitude diagram of confirmed comoving star pairs plotted over
-    the density of all \tgas\ sources within 200 pc; similar to
-    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
-    Several interesting pairs that include main-sequence and giant component
-    stars are highlighted in the text
-    (\sectionname~\ref{sec:interesting-pairs} and
-    \figurename~\ref{fig:highlighted-cmd}).
-    \label{fig:genuine-cmd}}
-\end{figure}
-
-An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
-previously-measured radial velocities (for each star) reported in \DR{5} of the
-Radial Velocity Experiment (RAVE; \citealt{Kunder:2017}).
-Using the same procedure described in \sectionname~\ref{sec:catalog}, we
-identify confirmed comoving pairs using the RAVE velocities for all pairs within
-$200~\pc$ and find a similar false-positive rate at large separations ($\approx
-38\%$ are confirmed).
-Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
-that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
-We plot the confirmed comoving pairs identified through radial velocity
-follow-up in this work along with the RAVE-confirmed pairs in
-\figurename~\ref{fig:separation-RAVE}.
-Top histogram shows the separation (chord length) distribution for the joint
-sample, scatter plot shows log-chord-length and the pair distances, and right
-histogram shows the distance distribution of the confirmed pairs.
-At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
-of our method drops, as expected given the hard cut in log-FML-ratio applied to
-define the candidate sample.
-Reassuringly, there do not appear to be any trends of the chord length
-distribution with distance.
-
-\begin{figure}[htbp]
+\begin{figure}[p]
   \begin{center}
@@
 
-\subsection{A few highlighted comoving pairs}\label{sec:interesting-pairs}
-
-We naively expect that comoving, coeval stars should lie along isochrones
-color-magnitude space.
-The majority of our star pairs appear to follow this expectation, with the
-exception of a few visible outliers.
+An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
+previously-measured radial velocities (for each star) reported in \DR{5} of the
+Radial Velocity Experiment (RAVE; \citealt{Kunder:2017}).
+Using the same procedure described in \sectionname~\ref{sec:catalog}, we
+identify confirmed comoving pairs using the RAVE velocities for all pairs within
+$200~\pc$ and find a similar false-positive rate at large separations ($\approx
+38\%$ are confirmed).
+Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
+that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
+We plot the confirmed comoving pairs identified through radial velocity
+follow-up in this work along with the RAVE-confirmed pairs in
+\figurename~\ref{fig:separation-RAVE}.
+Top histogram shows the separation (chord length) distribution for the joint
+sample, scatter plot shows log-chord-length and the pair distances, and right
+histogram shows the distance distribution of the confirmed pairs.
+At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
+of our method drops, as expected given the hard cut in log-FML-ratio applied to
+define the candidate sample.
+Reassuringly, there do not appear to be any trends of the chord length
+distribution with distance.
+
+\figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
+diagram for all confirmed comoving pairs.
+As would be expected if all of the pairs are coeval, the majority of pairs lie
+along isochrones in color-magnitude space, with the exception of a few visible
+outliers.
 For example, there appear to be at least a few main-sequence--giant-branch
 pairs, and a few pairs that appear to cross the main-sequence.
-We highlight a few of these pairs below and in
-\figurename~\ref{fig:highlighted-cmd}.
+We highlight a few of these pairs in the next section
+(\sectionname~\ref{sec:interesting-pairs}).
+
+\clearpage
 
@@
   \begin{center}
-    \includegraphics[width=0.8\linewidth]{highlighted_cmd.pdf}
+    \includegraphics[width=\linewidth]{genuine_highlighted_cmd.pdf}
   \end{center}
   \caption{%
-    Same as \figurename~\ref{fig:genuine-cmd}, but only three highlighted
-    comoving star pairs are shown.
-    \label{fig:highlighted-cmd}}
+    {\it Left}: Color-magnitude diagram of confirmed comoving star pairs plotted
+    over the density of all \tgas\ sources within 200 pc; similar to
+    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
+    Several interesting pairs that include main-sequence and giant component
+    stars are highlighted in the text (\sectionname~\ref{sec:interesting-pairs})
+    and in the right panel.
+    {\it Right}: Same as left panel, but only three highlighted comoving star
+    pairs are shown.
+    \label{fig:genuine-highlighted-cmd}}
 \end{figure}
 
+\subsection{A few highlighted comoving pairs}\label{sec:interesting-pairs}
+
+\figurename~\ref{fig:genuine-highlighted-cmd} (right) shows a few highlighted
+pairs selected from the color-magnitude diagram of confirmed comoving stars.
 We compile \project{Tycho-2} ($B$, $V$; \citealt{Hog:2000}), \tmass\
@@
 \appendix
+
+\section{Data products} \label{sec:table}
+
+See the caption for \tablename~\ref{tbl:data-pairs} for a full description of
+the columns in the associated data file.
+
+\begin{table}[htb]
+    \centering
+    \caption{Description of columns in data table containing results from radial
+      velocity follow-up.} \label{tbl:data-pairs}
+    \begin{tabular}{l|l|l}
+        \toprule
+        Column Name    & Unit & Description\\
+        \midrule
+        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
+        \texttt{R\_mu}            &      & Log-likelihood-ratio computed with only astrometric information\\
+        \texttt{R\_rv}            &      & Log-likelihood-ratio computed with radial velocities as well\\
+        \texttt{dx\_50}           & \pc  & Median of 3D spatial separation error distribution\\
+        \texttt{dv\_50}           & \kms & Median of 3D velocity difference error distribution\\
+        \texttt{dx\_15}           & \pc  & 15th percentile of 3D spatial separation error distribution\\
+        \texttt{dx\_85}           & \pc  & 85th percentile of 3D spatial separation error distribution\\
+        \texttt{dv\_15}           & \kms & 15th percentile of 3D velocity difference error distribution\\
+        \texttt{dv\_85}           & \kms & 85th percentile of 3D velocity difference error distribution\\
+        \bottomrule
+    \end{tabular}
+\end{table}
 

### 2017-08-20 paper/comoving-followup.tex
figure updates and shit

@@
 
-\author{David~N.~Spergel}
+\author[0000-0002-5151-0006]{David~N.~Spergel}
 \affiliation{Flatiron Institute,
@@
 % Methods
-Here, we present results from our own low-resolution radial-velocity survey of a
+Here we present results from our own low-resolution radial-velocity survey of a
 random sample of the candidate pairs to identify and validate true comoving star
@@
 % Conclusions
-We therefore confirm the discovery of a population of comoving star pairs at
-very large separations, strongly suggesting an abundance of disrupted wide
-binaries and stellar associations in the Galactic disk.
+We confirm the discovery of a population of comoving star pairs at very large
+separations, suggesting that disrupted wide binaries and stellar associations
+remain approximately spatially coherent.
 \end{abstract}
@@
 
-Widely-separated, coeval pairs or multiplets of stars --- i.e. wide binaries,
+Widely-separated, coeval star pairs or multiples --- i.e. wide binaries,
 multiplets, or former binaries or multiplets --- are important tracers of many
@@
 as ``comoving pair candidates'' --- or using full-space velocity information ---
-we refer to these as ``comoving pairs.''
+we refer to these as ``confirmed comoving pairs.''
 Radial velocity follow-up is still needed to confirm the majority of the known
@@
 predict that ionized binaries could appear as a second peak or component in the
-separation distribution of comoving stars at much larger separations (10--
-$1000~\pc$; \citealt{Jiang:2010}).
+separation distribution of comoving stars at much larger separations
+(10--$1000~\pc$; \citealt{Jiang:2010}).
 Though these simulations don't include perturbers like GMCs or black holes, they
 at least indicate that disrupted wide binaries may remain coherent for a longer
-period of time than previously thought, and thus, a monotonically decreasing
-separation distribution may be too simple of a model for the separation
-distribution of comoving stars in our Galaxy.
+period of time than previously thought.
+Thus, a monotonically decreasing separation distribution may be too simple of a
+model for the separation distribution of comoving stars in our Galaxy.
 In addition, the distribution of unbound but comoving stars should be very
@@
 the existence of a second peak in the separation distribution of
-radial-velocity-confirmed comoving stars (\citealt{Shaya:2011}).
+radial-velocity-confirmed comoving stars (\citealt{Shaya:2011}; hereafter SO11).
 A key aspect of this work was to move from searching for common proper motion
@@
 At separations of $1$--$10~\pc$, star pairs in the local solar neighborhood
-subtend large angles on the sky; therefore, even if the two component stars were
-moving with the same 3D velocity, their proper motions and radial velocities
-can be different due to projections.
+subtend large angles on the sky: even if the two component stars were moving
+with the same 3D velocity, their proper motions and radial velocities can be
+different.
 By compiling radial velocities from the literature for \project{Hipparcos}
-stars, they applied this method to search for all comoving star pairs with
+stars, SO11 applied this method to search for all comoving star pairs with
 separations larger than $0.01~\pc$ within $100~\pc$ of the sun.
@@
 distribution that falls of until $\sim 1~\pc$, where the number of pairs then
-begins rising with separations.
-
-To verify the existence of this unbound population of very wide-separation
-comoving pairs, we need to identify a large sample of confirmed comoving stars.
+begins rising in bins of log-separation.
+
+To verify the existence of a population of very wide-separation comoving pairs,
+we first need to identify a large sample of confirmed comoving stars.
 This will already be possible for the brightest stars with \gaia\ \DR{2}, but
@@
 
-We describe the data, observations, and spectroscopic reduction in
-\sectionname~\ref{sec:data}.
-If you are uninterested in these details, skip to \sectionname~\ref{sec:results}
-and \sectionname~\ref{sec:discussion} where we present and discuss our main
-results.
+We describe the data and observations in \sectionname~\ref{sec:data}.
+Details of the spectroscopic reduction and velocity measurements are described
+in the Appendix.
+We present our main results in \sectionname~\ref{sec:results} and
+\sectionname~\ref{sec:discussion}.
 
@@
 ratio computed for all nearly co-spatial pairs of stars.
-Briefly, we perform a model selection between
-two hypotheses: that the velocities of the two stars in a pair are
-identical (case 1) or that they are indepndent (case 2).
-We marginalize over the (unknown) true velocity and distance of each star
-to obtain the ratio of fully marginalized likelihoods,
+We performed a model selection between two hypotheses: that the velocities of
+the two stars in a pair are identical (case 1) or that they are indepndent (case
+2).
+We marginalized over the (unknown) true velocity and distance of each star to
+obtain the ratio of fully marginalized likelihoods,
 $\mathcal{L}_1/\mathcal{L}_2$.
@@
 
-After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$),
+After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$,
 we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
-with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ of each
-target star from the \tgas\ catalog.
+with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ for each
+star in the \tgas\ catalog.
 After applying a conservative cut on the likelihood ratio motivated by computing
-the same likelihood ratio for random pairings of stars ($\ln
+the same likelihood ratio for random pairings of stars, $\ln
 \mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
@@
 fraction of the identified pairs are members of larger groups.
-However, we have also found a large number ($\approx$4,000) of mutually
-exclusively-linked pairs in increasing number with increasing separation at
-$\sim 1$~\pc.
+However, we have also found a large number ($\approx$4,000) of
+mutually-exclusive linked pairs that increase in number with increasing
+log-separation from $\sim 1$~\pc.
 For a full description of the selection method, see \citealt{Oh:2017}.
@@
 consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
-distance; with the SNR cut used above and given the limitations of the \tgas\
-catalog itself, \tgas\ is $\approx$90\% complete to $200~\pc$
-(\citealt{Bovy:2017}).
+distance.
+With the SNR cut used above and given the limitations of the \tgas\ catalog
+itself, \tgas\ is $\approx$90\% complete to $200~\pc$ (\citealt{Bovy:2017}).
 From this source target list, we randomly observed targets with airmass $\sec z
-< 1.5$.
+< 1.5$ during our observing run.
 \figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
@@
 from $\approx 3600$--$7200~{\rm \AA}$.
-At this resolution, to obtain high signal-to-noise ($>100$) spectra, most
-exposures were between $30$--$120~{\rm s}$ (15th and 85th percentiles).
+At this resolution, to obtain high signal-to-noise ($>100$ per pixel) spectra,
+most exposures were between $30$--$120~{\rm s}$.
 To maximize the number of observed pairs, we chose not to take
 comparison lamp spectra at each pointing as it adds an overhead of $\approx
-90~{\rm s}$ per observation; we are primarily interested in relative velocities
-between pair members and therefore don't need precise absolute calibrations of
-their radial velocities.
+90~{\rm s}$ per observation.
+We are primarily interested in relative velocities between pair members and
+therefore don't need precise absolute calibrations of their radial velocities.
 To determine rough nightly wavelength solutions, calibration spectra were
@@
 These stars also have measured chemical abundances, and will be used in future
-work as a training set for a method of transferring chemical abundance labels,
-\project{The Cannon} (\citealt{Ness:2015}), to infer
+work as a training set for \project{The Cannon} (\citealt{Ness:2015}) to infer
 metallicities and stellar parameters for our observed sources.
@@
 
-\begin{figure}[htb]
+\begin{figure}[!ht]
   \begin{center}
@@
 Of the 323 observed star pairs, 12 have at least one member with an unreliable
-(flagged) radial velocity measurement from the pipeline above.
+(flagged) radial velocity measurement from the spectroscopic reduction pipeline.
 For the remaining 311 pairs, we use the measured radial velocities to re-compute
@@
 represent the thin disk, thick disk, and halo velocity distributions relative to
-the sun.
+the sun (see \citealt{Oh:2017}).
 
@@
 to their log-likelihood-ratio values.
+Some pairs with large velocity differences also have small (near zero) changes
+to the log-likelihood ratio: these are pairs where each member has a velocity
+that is an outlier with respect to the prior velocity distribution.
 
@@
 most small-separation pairs maintain their status as comoving.
-This is also apparent in \figurename~\ref{fig:separation}, which shows
-histograms of the spatial separation and sky separation for all pairs;
-background, gray histogram shows all observed pairs, and foreground, black
-histogram shows all pairs for which $\llrnew > \llrold$.
-
-At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
+
+\figurename~\ref{fig:separation} shows histograms of the spatial separation and
+sky separation for all pairs; background, gray histogram shows all observed
+pairs, and foreground, black histogram shows all pairs for which $\llrnew >
+\llrold$.
+In this and subsequent figures, we use the phrase ``tangential separation'' (and
+symbol $s_{\rm tan}$) to refer to the chord length separation between the two
+pairs assuming the two stars are at the same distance (the smaller distance of
+the two, $d_{\rm min}$):
+\begin{equation}
+    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
+\end{equation}
+where $\theta$ is the angular extent of the pair.
+
+At sky separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
 confirmed to be comoving; interestingly, the separation distribution of the
@@
     these very wide separations.
-    \emph{Right}: Distribution of chord lengths (tangential separations) for all
-    comoving star pairs (gray) and only those identified as comoving with 3D
-    velocity information (black).
+    \emph{Right}: Distribution of tangential separations for all comoving star
+    pairs (gray) and only those identified as comoving with 3D velocity
+    information (black).
     Noise in the parallax measurements makes the 3D separation distribution
     appear roughly uniform.
-    Even the distribution in chord length appears to be approximately linear for
-    large separations ($\gtrsim 0.1~\pc$).
+    Even the distribution in tangential separation appears to be approximately
+    linear for large separations ($\gtrsim 0.1~\pc$).
     \label{fig:separation}}
@@
   \caption{%
-    \emph{Top}: Distribution of chord lengths for all confirmed comoving star
-      pairs, either from the radial velocity follow-up presented in this work or
-      from RAVE \DR{5} radial velocities.
+    \emph{Top}: Distribution of tangential separation for all confirmed comoving
+      star pairs, either from the radial velocity follow-up presented in this
+      work or from RAVE \DR{5} radial velocities.
       At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log)
@@
       $10~\pc$.
-    \emph{Middle}: Chord length plotted against mean distance for each comoving
-      pair identified from this work (circles, blue) or from RAVE overlap
-      (squares, orange).
+    \emph{Middle}: Tangential separation plotted against mean distance for each
+      comoving pair identified from this work (circles, blue) or from RAVE
+      overlap (squares, orange).
     \emph{Right}: Distance distribution for all confirmed comoving pairs.
@@
 \figurename~\ref{fig:separation-RAVE}.
-Top histogram shows the separation (chord length) distribution for the joint
-sample, scatter plot shows log-chord-length and the pair distances, and right
+Top histogram shows the tangential separation distribution for the joint
+sample, scatter plot shows log-separation and the pair distances, and right
 histogram shows the distance distribution of the confirmed pairs.
@@
 define the candidate sample.
-Reassuringly, there do not appear to be any trends of the chord length
+Reassuringly, there do not appear to be any trends of the separation
 distribution with distance.
@@
         Star name & \logg\ & \teff\ [K] & mass [$M_\odot$] &
-          chord length [pc]  \\
+          tangential separation [pc]  \\
         \toprule
@@
 This pair is separated by $4.5^\circ$ on the sky (corresponding to $\approx
-8.6~\pc$ in chord length, see \tablename~\ref{tbl:highlighted-pairs}) and have
-consistent distances within their uncertainties of $\approx 110 \pm 5~\pc$.
+8.6~\pc$ in tangential separation, see \tablename~\ref{tbl:highlighted-pairs})
+and have consistent distances within their uncertainties of $\approx 110 \pm
+5~\pc$.
 At first glance, it seems unlikely that the two stars could be coeval, as the
@@
 G-type dwarf (TYC 3807-1439-1).
-This pair is separated by $4^\circ$ on the sky or $6.8~\pc$ in chord length and
-have an apparent radial separation of $\approx 3~\pc$, corresponding to a 3D
-separation of $\approx 9.8~\pc$ at a distance of $\approx 100~\pc$.
+This pair is separated by $4^\circ$ on the sky or $6.8~\pc$ in tangential
+separation and have an apparent radial separation of $\approx 3~\pc$,
+corresponding to a 3D separation of $\approx 9.8~\pc$ at a distance of $\approx
+100~\pc$.
 Coeval stars at different phases of stellar evolution are extremely valuable for
@@
 the confirmed comoving pairs.
-This pair is separated by $2.5^\circ$ on the sky or $8.5~\pc$ in chord length
-and have a consistent distance of $\approx 190~\pc$ within their uncertainties.
+This pair is separated by $2.5^\circ$ on the sky or $8.5~\pc$ in tangential
+separation and have a consistent distance of $\approx 190~\pc$ within their
+uncertainties.
 The mass ratio distribution of comoving star pairs will help constrain star
@@
 radial velocity and difference in iron abundance with points colored by the
-chord length separation of each pair.
+tangential separation of each pair.
 There is a clear over-density of pairs with abundance and radial-velocity
@@
     member stars from the \acronym{LAMOST} survey.
-    Points represent each pair, colored by the separation (chord length), and
+    Points represent each pair, colored by the tangential separation, and
     error bars represent the uncertainties taken from the \acronym{LAMOST}
@@
 \appendix
-
-\section{Data products} \label{sec:table}
-
-See the caption for \tablename~\ref{tbl:data-pairs} for a full description of
-the columns in the associated data file.
-
-\begin{table}[htb]
-    \centering
-    \caption{Description of columns in data table containing results from radial
-      velocity follow-up.} \label{tbl:data-pairs}
-    \begin{tabular}{l|l|l}
-        \toprule
-        Column Name    & Unit & Description\\
-        \midrule
-        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
-        \texttt{R\_mu}            &      & Log-likelihood-ratio computed with only astrometric information\\
-        \texttt{R\_rv}            &      & Log-likelihood-ratio computed with radial velocities as well\\
-        \texttt{dx\_50}           & \pc  & Median of 3D spatial separation error distribution\\
-        \texttt{dv\_50}           & \kms & Median of 3D velocity difference error distribution\\
-        \texttt{dx\_15}           & \pc  & 15th percentile of 3D spatial separation error distribution\\
-        \texttt{dx\_85}           & \pc  & 85th percentile of 3D spatial separation error distribution\\
-        \texttt{dv\_15}           & \kms & 15th percentile of 3D velocity difference error distribution\\
-        \texttt{dv\_85}           & \kms & 85th percentile of 3D velocity difference error distribution\\
-        \bottomrule
-    \end{tabular}
-\end{table}
 
@@
 
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.8\linewidth]{wavelength_gp.pdf}
+  \end{center}
+  \caption{%
+    A visualization of the pixel-to-wavelength model determined from a single
+    wavelength comparison lamp spectrum.
+    Black markers show the fit pixel centroids for known Hg or Ne emission
+    lines and the residual wavelength value between the linear + GP model and
+    the true emission line centroid.
+    The non-shaded region shows the section of spectrum where the comparison
+    lamp spectrum has sufficient line density to determine the wavelength
+    solution; shaded regions are masked from the resulting spectra.
+    Orange envelope shows the root-variance of the Gaussian process component
+    of the model, which can be interpreted as inherent uncertainties on the
+    wavelength calibration as a result of the model flexibility.
+    \label{fig:wavelength-GP}}
+\end{figure}
+
 Typically, high-order standard polynomials or Chebyshev polynomials are used to
@@
 radial velocities from this sample.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.8\linewidth]{wavelength_gp.pdf}
-  \end{center}
-  \caption{%
-    A visualization of the pixel-to-wavelength model determined from a single
-    wavelength comparison lamp spectrum.
-    Black markers show the fit pixel centroids for known Hg or Ne emission
-    lines and the residual wavelength value between the linear + GP model and
-    the true emission line centroid.
-    The non-shaded region shows the section of spectrum where the comparison
-    lamp spectrum has sufficient line density to determine the wavelength
-    solution; shaded regions are masked from the resulting spectra.
-    Orange envelope shows the root-variance of the Gaussian process component
-    of the model, which can be interpreted as inherent uncertainties on the
-    wavelength calibration as a result of the model flexibility.
-    \label{fig:wavelength-GP}}
-\end{figure}
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
+    \includegraphics[width=0.75\linewidth]{mcmc_corner.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.7\linewidth]{mcmc_example_fit.pdf}
+    \includegraphics[width=0.5\linewidth]{mcmc_example_fit.pdf}
   \end{center}
@@
 
-\begin{table}[htb]
+\section{Data products} \label{sec:table}
+
+See tables below for full descriptions of the columns in the associated data
+files.
+
+\begin{table}[ht]
     \centering
-    \caption{Description of columns in data table containing results from radial
-    velocity follow-up.} \label{tbl:data-stars}
+    \caption{Description of columns in data table containing absolute radial
+    velocity measurements and summary information from the \tgas\ catalog.}
+    \label{tbl:data-stars}
     \begin{tabular}{l|l|l}
@@
         \midrule
-        \texttt{Oh17\_id}         &      & Star row index from the candidate catalog (\citealt{Oh:2017})\\
         \texttt{Oh17\_group\_id}  &      & Group ID from the candidate catalog (\citealt{Oh:2017})\\
@@
 
-% \clearpage
+\begin{table}[h]
+    \centering
+    \caption{Description of columns in data table containing separations and
+    relative radial velocities for all observed groups, along with likelihood
+    ratios computed with and without relative radial velocities.}
+    \label{tbl:data-pairs}
+    \begin{tabular}{l|l|l}
+        \toprule
+        Column Name    & Unit & Description\\
+        \midrule
+        \texttt{Oh17\_group\_id}    &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
+        \texttt{R\_mu}              &      & Log-likelihood-ratio computed with only astrometric information\\
+        \texttt{R\_rv}              &      & Log-likelihood-ratio computed with radial velocities as well\\
+        \texttt{sep\_tan}           & \pc  & Median of tangential separation error distribution\\
+        \texttt{sep\_tan\_err}      & \pc  & Uncertainty in tangential separation\\
+        \texttt{relative\_rv}       & \kms & Relative radial velocity between the two stars\\
+        \texttt{relative\_rv\_err}  & \kms & Uncertainty in relative radial velocity\\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\clearpage
 

### 2017-09-11 paper/comoving-followup.tex
between

@@
 pairs with separations as large as $\approx 10~\pc$, our original search limit.
-At separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
+Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
 the number of confirmed pairs per separation decreases with increasing

### 2017-10-01 scripts/mixture_model.py
some improvements

@@ -1,3 +1,7 @@
+# Standard library
+from math import log
+import sys
+
 # Third-party
 from astropy.io import fits
 import astropy.units as u
@@ -15,7 +19,10 @@ from gwb.data import TGASData
 pc_mas_yr_per_km_s = (1*u.km/u.s).to(u.pc*u.mas/u.yr, u.dimensionless_angles()).value
 km_s_per_pc_mas_yr = 1/pc_mas_yr_per_km_s
 
-def get_icrs_samples(data, size=1):
+def get_icrs_samples(data, size=1, seed=None):
+    if seed is not None:
+        np.random.seed(seed)
+
     all_samples = []
     for i in range(len(data)):
         star = data[i]
@@ -62,24 +69,26 @@ class MixtureModel:
 
         self._field_vdisp = field_vdisp.to(u.km/u.s).value
 
-        icrs1 = get_icrs_samples(data1, n_dv_samples)
-        icrs2 = get_icrs_samples(data2, n_dv_samples)
+        self._icrs1 = get_icrs_samples(data1, n_dv_samples)
+        self._icrs2 = get_icrs_samples(data2, n_dv_samples)
 
         # We can pre-compute the tangent basis matrices given the sky positions
-        # of each star
-        self.M1 = get_tangent_basis(icrs1[:,0,0], icrs1[:,0,1])
-        self.M2 = get_tangent_basis(icrs2[:,0,0], icrs2[:,0,1])
+        # of each star. We transpose it using swapaxes()
+        self.M1 = np.swapaxes(get_tangent_basis(self._icrs1[:,0,0],
+                                                self._icrs1[:,0,1]), 1, 2)
+        self.M2 = np.swapaxes(get_tangent_basis(self._icrs2[:,0,0],
+                                                self._icrs2[:,0,1]), 1, 2)
 
         # We can also pre-compute sample vectors in proper motion components
         # and in radial velocity and make units consistent. To get velocity
         # samples we just need to multiply in the distances
-        self._v1_samples = np.array([icrs1[...,3] * km_s_per_pc_mas_yr,
-                                     icrs1[...,4] * km_s_per_pc_mas_yr,
-                                     icrs1[...,5]])
+        self._v1_samples = np.array([self._icrs1[...,3].T * km_s_per_pc_mas_yr,
+                                     self._icrs1[...,4].T * km_s_per_pc_mas_yr,
[…]

### 2017-10-01 scripts/mixture_model.py
sampling

@@ -11,6 +11,7 @@ from scipy.misc import logsumexp
 from scipy.integrate import simps
 from scipy.stats import norm
 from tqdm import tqdm
+import emcee
 
 # Project
 from gwb.coords import get_tangent_basis
@@ -161,10 +162,31 @@ class MixtureModel:
 
         return np.logaddexp(lls1, lls2), (lls1, lls2)
 
+    def ln_prior(self, p):
+        f = p[0]
 
-def main(data1, data2):
-    mm = MixtureModel(data1, data2, field_vdisp=25.*u.km/u.s)
+        if f <= 0 or f >= 1:
+            return -np.inf
 
+        return 0.
+
+    def ln_posterior(self, p):
+        lp = self.ln_prior(p)
+        if not np.isfinite(lp):
+            return -np.inf, None
+
+        ll, blobs = self.ln_likelihood(p)
+        if np.any(np.logical_not(np.isfinite(ll))):
+            return -np.inf, None
+
+        return lp + ll.sum(), blobs
+
+    def __call__(self, p):
+        print(p[0])
+        return self.ln_posterior(p)
+
+def plot_posterior(mm):
+    # Test: plot the posterior curve
     lls = []
     fs = np.linspace(0.15, 0.7, 32)
     for f in tqdm(fs):
@@ -176,7 +198,37 @@ def main(data1, data2):
     plt.plot(fs, np.exp(lls - lls.max()))
     plt.show()
 
+def run_emcee(model, pool):
+
+    n_walkers = 28
+    p0 = np.random.normal(0.5, 1E-3, size=(n_walkers, 1))
+    sampler = emcee.EnsembleSampler(n_walkers, 1, model)
+    sampler.run_mcmc(p0, 1024)
+
+    np.save('../data/sampler.chain', sampler.chain)
+    np.save('../data/sampler.blobs', sampler.blobs)
+
 if __name__ == "__main__":
+    import schwimmbad
+    from argparse import ArgumentParser
+
+    # Define parser object
+    parser = ArgumentParser(description="")
+    parser.add_argument('--mpi', action='store_true', default=False,
+                        dest='mpi')
+
+    args = parser.parse_args()
+
+    if args.mpi:
+        pool = schwimmbad.MPIPool()
+
+        if not pool.is_master():
+            pool.wait()
+            sys.exit(0)
+
+    else:
[…]

### 2017-10-01 scripts/mixture_model.py
close pool when done

@@ -182,7 +182,7 @@ class MixtureModel:
         return lp + ll.sum(), blobs
 
     def __call__(self, p):
-        print(p[0])
+        # print(p[0])
         return self.ln_posterior(p)
 
 def plot_posterior(mm):
@@ -201,9 +201,10 @@ def plot_posterior(mm):
 def run_emcee(model, pool):
 
     n_walkers = 28
+    n_steps = 1024
     p0 = np.random.normal(0.5, 1E-3, size=(n_walkers, 1))
     sampler = emcee.EnsembleSampler(n_walkers, 1, model)
-    sampler.run_mcmc(p0, 1024)
+    _ = sampler.run_mcmc(p0, n_steps)
 
     np.save('../data/sampler.chain', sampler.chain)
     np.save('../data/sampler.blobs', sampler.blobs)
@@ -249,5 +250,8 @@ if __name__ == "__main__":
 
     mm = MixtureModel(data1, data2, field_vdisp=25.*u.km/u.s)
     # plot_posterior(data1, data2)
-
+    print("Model created - starting sampling")
     run_emcee(mm, pool)
+    
+    pool.close()
+    sys.exit(0)

### 2017-10-01 scripts/mixture_model.py
add batches

@@ -202,12 +202,21 @@ def run_emcee(model, pool):
 
     n_walkers = 28
     n_steps = 1024
+    n_batches = 8
+
     p0 = np.random.normal(0.5, 1E-3, size=(n_walkers, 1))
     sampler = emcee.EnsembleSampler(n_walkers, 1, model)
-    _ = sampler.run_mcmc(p0, n_steps)
 
-    np.save('../data/sampler.chain', sampler.chain)
-    np.save('../data/sampler.blobs', sampler.blobs)
+    for batch in range(n_batches):
+        pos, *_ = sampler.run_mcmc(p0, n_steps // n_batches)
+        p0 = pos
+
+        np.save('../data/sampler_chain{0}.npy'.format(batch),
+                sampler.chain)
+        np.save('../data/sampler_blobs{0}.npy'.format(batch),
+                sampler.blobs)
+
+        sampler.reset()
 
 if __name__ == "__main__":
     import schwimmbad
@@ -252,6 +261,6 @@ if __name__ == "__main__":
     # plot_posterior(data1, data2)
     print("Model created - starting sampling")
     run_emcee(mm, pool)
-    
+
     pool.close()
     sys.exit(0)

### 2017-10-01 scripts/mixture_model.py
use mpipool from emcee

@@ -12,6 +12,7 @@ from scipy.integrate import simps
 from scipy.stats import norm
 from tqdm import tqdm
 import emcee
+from emcee.utils import MPIPool
 
 # Project
 from gwb.coords import get_tangent_basis
@@ -230,7 +231,7 @@ if __name__ == "__main__":
     args = parser.parse_args()
 
     if args.mpi:
-        pool = schwimmbad.MPIPool()
+        pool = MPIPool()
 
         if not pool.is_master():
             pool.wait()

### 2017-10-01 scripts/mixture_model.py
print batch

@@ -209,6 +209,8 @@ def run_emcee(model, pool):
     sampler = emcee.EnsembleSampler(n_walkers, 1, model)
 
     for batch in range(n_batches):
+        print("Batch: {0}".format(batch))
+
         pos, *_ = sampler.run_mcmc(p0, n_steps // n_batches)
         p0 = pos
 

### 2017-10-01 scripts/mixture_model.py
more print

@@ -251,6 +251,8 @@ if __name__ == "__main__":
     # data2 = TGASData(_tbl2, rv=_tbl2['RV']*u.km/u.s,
     #                  rv_err=_tbl2['RV_err']*u.km/u.s)
 
+    print("Loading data...")
+
     # Load real data
     _tbl1 = fits.getdata('../data/tgas_apw1.fits')
     data1 = TGASData(_tbl1, rv=_tbl1['RV']*u.km/u.s,
@@ -260,8 +262,11 @@ if __name__ == "__main__":
     data2 = TGASData(_tbl2, rv=_tbl2['RV']*u.km/u.s,
                      rv_err=_tbl2['RV_err']*u.km/u.s)
 
+    print("Data loaded, creating model...")
+
     mm = MixtureModel(data1, data2, field_vdisp=25.*u.km/u.s)
     # plot_posterior(data1, data2)
+
     print("Model created - starting sampling")
     run_emcee(mm, pool)
 

### 2017-10-02 scripts/mixture_model.py
also analyze the chains

@@ -1,5 +1,6 @@
 # Standard library
 from math import log
+import os
 import sys
 
 # Third-party
@@ -199,7 +200,7 @@ def plot_posterior(mm):
     plt.plot(fs, np.exp(lls - lls.max()))
     plt.show()
 
-def run_emcee(model, pool):
+def run_emcee(model, pool, chain_file, blobs_file):
 
     n_walkers = 28
     n_steps = 1024
@@ -221,6 +222,47 @@ def run_emcee(model, pool):
 
         sampler.reset()
 
+    # Now collect all the individual files into one...
+    chains = []
+    blobs = []
+    for batch in range(n_batches):
+        chains.append(np.load('../data/sampler_chain{0}.npy'.format(batch)))
+        blobs.append(np.load('../data/sampler_blobs{0}.npy'.format(batch)))
+
+    chain = np.hstack(chains)
+    blobs = np.vstack(blobs)
+    np.save(chain_file, chain)
+    np.save(blobs_file, blobs)
+
+    # Now clean up / delete the files
+    for batch in range(n_batches):
+        os.remove('../data/sampler_chain{0}.npy'.format(batch))
+        os.remove('../data/sampler_blobs{0}.npy'.format(batch))
+
+def analyze_chain(chain, blobs):
+    # MAGIC NUMBER: index after which walkers are converged
+    ix = 256
+    trim_chain = chain[:,ix:]
+    trim_blobs = blobs[ix:]
+
+    # Downsample chains because correlation
+    flat_f = np.vstack(trim_chain[:,::8])[:,0]
+    med_f = np.median(flat_f)
+    std_f = 1.5 * np.median(np.abs(flat_f - med_f))
+    print('f = {0:.2f} +/- {1:.2f}'.format(med_f, std_f))
+
+    # Now we compute the per-pair probability
+    norm = 0.0
+    post_prob = np.zeros(blobs.shape[-1])
+    for i in range(trim_chain.shape[1]): # steps
+        for j in range(trim_chain.shape[0]): # walkers
+            ll_fg, ll_bg = trim_blobs[i][j]
+            post_prob += np.exp(ll_fg - np.logaddexp(ll_fg, ll_bg))
+            norm += 1
+    post_prob /= norm
+
+    np.save('../data/pair_probs.npy', post_prob)
+
 if __name__ == "__main__":
     import schwimmbad
     from argparse import ArgumentParser
[…]

### 2017-10-02 paper/comoving-followup.tex
für
placeholder section on mixture model

@@
 This project was developed in part at the 2017 Heidelberg Gaia Sprint, hosted by
-the Max-Planck-Institut für Astronomie, Heidelberg.
+the Max-Planck-Institut f\"ur Astronomie, Heidelberg.
 
@@
 
+\section{Mixture model...}\label{sec:TODO}
+
+% TODO: assumptions:
+% - Same velocity dispersion for all ages. Could have an age-dependent
+%   dispersion
+% -
+With the new radial velocity measurements, we construct a statistical model to
+determine the probability that a given candidate comoving pair is truly
+comoving.
+We don't model the positional separation distribution and instead construct a
+model purely in velocity-space.
+The data are therefore proper motions in the ICRS reference frame, $\bs{\mu}$,
+and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric radial
+velocity, $v_r$, for each star in a candidate comoving pair; we assume that the
+sky position for each star $(\alpha, \delta)$ is known perfectly.
+For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
+\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
+radial velocity data for star 1 in the pair.
+To simplify some expressions, we additionally ignore any correlations between
+the observed parallax and the proper motion components (but keep the covariance
+structure between the two proper motion components).
+
+Our model is expressed in terms of the magnitude of the true (unobserved) 3D
+velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
+and the (global) fraction of pairs that are comoving, $f$.
+For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
+f)$, we use a mixture model with two components
+\begin{eqnarray}
+    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
+\end{eqnarray}
+where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
+pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.\footnote{
+This is actually the marginal likelihood under a model in which every candidate
+comoving pair has an additional latent parameter: a binary flag, $q_n$, that
+specifies which mixture component it belongs too
+(\citealt{Hogg:2010,Foreman-Mackey:2014}). We later use this fact to compute the
+probability that each pair is truly comoving, or, more precisely, belongs to the
+mixture component $p_{\rm co}$).}
+The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
+to be the probability of the magnitude of the 3D separation of two vectors under
+the assumption that the two vectors were drawn independently from an isotropic,
+3D Gaussian with dispersion $\sigma$.
+That is,
+\begin{eqnarray}
+    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
+        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
+    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
+    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
+    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
+\end{eqnarray}
+where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
+\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
+covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
+
+Under this model, the joint probability of the velocity data for a given pair
+$\bs{\eta}_1$ and $\bs{\eta}_2$, and the latent parameters $\bs{v}_1, \bs{v}_2,
+\psi$, conditional on the unobserved true distances $d_1$ and $d_2$ (we handle
+these separately below), is
+\begin{equation}
+p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f) =
+    p(\bs{\eta}_1 \given d_1, \bs{v}_1) \,
+    p(\bs{\eta}_2, v_{r,2} \given d_2, \bs{v}_2) \,
+    p(\bs{v}_1, \bs{v}_2 \given \psi) \, p(\psi \given f)\label{eq:full-ll}
+\end{equation}
+where
+\begin{eqnarray}
+    p(\bs{v}_1, \bs{v}_2 \given \psi) &=& \delta(|\bs{v}_1-\bs{v}_2| - \psi) \quad .
+\end{eqnarray}
+
+Because we are only interested in inferring the fraction of stars consistent
+with being comoving, we marginalize this expression over the true 3D velocities
+and velocity difference, obtaining the velocity-marginalized, per-pair
+likelihood
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) =
+    \int \dd^3\bs{v}_1 \, \dd^3\bs{v}_2 \, \dd \psi \,
+    p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f)
+    \quad .
+\end{equation}
+This expression is still conditioned on the true distances, $d_1$ and $d_2$, but
+we observe the parallaxes $\varpi_1$ and $\varpi_2$.
+To remove the final per-pair latent parameters in this model, we therefore also
+have to marginalize the likelihood over these distance parameters to obtain the
+velocity-and-distance marginalized likelihood
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2 \given f) =
+        \int \dd d_1 \, \dd d_2 \,
+        p(\varpi_1 \given d_1) \, p(\varpi_2 \given d_2) \,
+        p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \,
+        p(d_1) \, p(d_2)
+\end{equation}
+where we assume a uniform space-density prior for the distances
+(\citealt{Bailer-Jones:2015aa}) with a maximum distance of 1~kpc.
+
+The posterior probability over the fraction of true comoving star pairs, $f$, is
+then given by
+\begin{equation}
+    p(f \given \{\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2\}_{N}) \propto
+        p(f) \, \prod_n^N
+        p(\bs{\eta}_{1,n}, \bs{\eta}_{2,n}, \varpi_{1,n}, \varpi_{2,n} \given f)
+\end{equation}
+where $N$ is the number of candidate comoving star pairs and the index $n$
+specifies the data for the $n$th pair.
+
+We approximate all of the marginalization integrals using numerical procedures.
+Given distances $d_1$ and $d_2$, we can approximate the velocity marginalization
+using Monte Carlo integration:
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \approx
+    \frac{1}{K} \, \sum^K_k \,
+    p(\psi^{(k)} \given f) \quad .
+\end{equation}
+In this expression, the $k$-th sample $\psi^{(k)}$ is defined as
+\begin{eqnarray}
+    \psi^{(k)} &=& |\bs{v}_1^{(k)} - \bs{v}_2^{(k)}| \\
+    \bs{v}_1^{(k)} &=& \mat{M}(\alpha_1, \delta_1) \cdot
+        \left(\begin{array}{c}
+            d_1\,\mu_{\alpha^*, 1}^{(k)} \\
+            d_1\,\mu_{\delta, 1}^{(k)} \\
+            v_{r, 1}^{(k)}
+        \end{array}\right)\\
+    \left(\begin{array}{c}
+        \mu_{\alpha^*, 1}^{(k)} \\
+        \mu_{\delta, 1}^{(k)} \\
+        v_{r, 1}^{(k)}
+    \end{array}\right) &\sim& \mathcal{N}(\bs{\eta}_1, \mat{C}_1)
+\end{eqnarray}
+where $\mat{C}_1$ is the covariance matrix of the \gaia\ data provided by the
+\tgas\ catalog with an additional (uncorrelated) diagonal term added for the
+uncertainty in the radial velocity measurement, and the rotation matrix
+$\mat{M}(\alpha, \delta)$ transforms from the local, tangent-space coordinates
+at $(\alpha, \delta)$ to the ICRS (see, e.g., Section~3 of \citealt{Oh:2017}).
+We perform the outer integral over the two distances, $d_1$ and $d_2$, using
+Simpson's rule on a 2D grid of distances that, in each dimension, spans the
+domain $\left[ (\varpi + 3\sigma_{\varpi})^{-1}, (\varpi -
+3\sigma_{\varpi})^{-1} \right]$, where $\sigma_{\varpi}$ is the parallax
+uncertainty reported in the \tgas\ catalog for each star.
+From convergence tests, we have found that it is sufficient to use 1024 samples
+in the Monte Carlo integral and 5 by 5 distance grid points.
+
+\clearpage
+
 \bibliographystyle{aasjournal}

### 2017-10-02 paper/comoving-followup.tex
dispersions

@@
 
-\section{Results} \label{sec:results}
-
-\subsection{Comoving pair catalog}\label{sec:catalog}
+\section{Methods: Comoving pair probabilities}\label{sec:methods}
+
+% \subsection{Comoving pair probabilities} \label{sec:pair-probs}
+
+% TODO: assumptions:
+% - Same velocity dispersion for all ages. Could have an age-dependent
+%   dispersion
+% -
 
@@
 (flagged) radial velocity measurement from the spectroscopic reduction pipeline.
-For the remaining 311 pairs, we use the measured radial velocities to re-compute
-the marginal likelihood ratio for the two hypotheses considered in previous work
-(see Appendix~B, \citealt{Oh:2017}).
+For the remaining 311 pairs, we construct a statistical model to determine the
+probability that a given candidate comoving pair is truly comoving.
+For this inference, the data are the proper motions in the ICRS reference frame,
+$\bs{\mu}$, and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric
+radial velocity, $v_r$, for each star in a candidate comoving pair; we assume
+that the sky position for each star $(\alpha, \delta)$ is known perfectly.
+For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
+\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
+radial velocity data for star 1 in the pair.
+To simplify some expressions, we additionally ignore any correlations between
+the observed parallax and the proper motion components (but keep the covariance
+structure between the two proper motion components).
+
+Our model is expressed in terms of the magnitude of the true (unobserved) 3D
+velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
+and the (global) fraction of pairs that are comoving, $f$.
+For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
+f)$, we use a mixture model with two components
+\begin{eqnarray}
+    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
+\end{eqnarray}
+where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
+pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.
+Using a mixture model and inferring the mixture weight is a common trick to
+avoid explicitly having to sample over the per-pair probabilities: this
+expression is equivalent to the marginal likelihood under a model in which every
+candidate comoving pair has an additional latent parameter (a binary flag,
+$q_n$) that specifies which mixture component it belongs to
+(\citealt{Hogg:2010,Foreman-Mackey:2014}).
+We later use this fact to compute the probability that each pair is truly
+comoving, or, more precisely, belongs to the mixture component $p_{\rm co}$.
+
+The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
+to be the probability of the magnitude of the 3D separation of two vectors under
+the assumption that the two vectors were drawn independently from an isotropic,
+3D Gaussian with dispersion $\sigma$.
+That is,
+\begin{eqnarray}
+    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
+        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
+    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
+    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
+    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
+\end{eqnarray}
+where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
+\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
+covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
+
+Under this model, the joint probability of the velocity data for a given pair
+$\bs{\eta}_1$ and $\bs{\eta}_2$, and the latent parameters $\bs{v}_1, \bs{v}_2,
+\psi$, conditional on the unobserved true distances $d_1$ and $d_2$ (we handle
+these separately below), is
+\begin{equation}
+p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f) =
+    p(\bs{\eta}_1 \given d_1, \bs{v}_1) \,
+    p(\bs{\eta}_2, v_{r,2} \given d_2, \bs{v}_2) \,
+    p(\bs{v}_1, \bs{v}_2 \given \psi) \, p(\psi \given f)\label{eq:full-ll}
+\end{equation}
+where
+\begin{eqnarray}
+    p(\bs{v}_1, \bs{v}_2 \given \psi) &=& \delta(|\bs{v}_1-\bs{v}_2| - \psi) \quad .
+\end{eqnarray}
+
+Because we are only interested in inferring the fraction of stars consistent
+with being comoving, we marginalize this expression over the true 3D velocities
+and velocity difference, obtaining the velocity-marginalized, per-pair
+likelihood
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) =
+    \int \dd^3\bs{v}_1 \, \dd^3\bs{v}_2 \, \dd \psi \,
+    p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f)
+    \quad .
+\end{equation}
+This expression is still conditioned on the true distances, $d_1$ and $d_2$, but
+we observe the parallaxes $\varpi_1$ and $\varpi_2$.
+To remove the final per-pair latent parameters in this model, we therefore also
+have to marginalize the likelihood over these distance parameters to obtain the
+velocity-and-distance marginalized likelihood
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2 \given f) =
+        \int \dd d_1 \, \dd d_2 \,
+        p(\varpi_1 \given d_1) \, p(\varpi_2 \given d_2) \,
+        p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \,
+        p(d_1) \, p(d_2)
+\end{equation}
+where we assume a uniform space-density prior for the distances
+(\citealt{Bailer-Jones:2015aa}) with a maximum distance of 1~kpc.
+
+The posterior probability over the fraction of true comoving star pairs, $f$, is
+then given by
+\begin{equation}
+    p(f \given \{\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2\}_{N}) \propto
+        p(f) \, \prod_n^N
+        p(\bs{\eta}_{1,n}, \bs{\eta}_{2,n}, \varpi_{1,n}, \varpi_{2,n} \given f)
+\end{equation}
+where $N$ is the number of candidate comoving star pairs and the index $n$
+specifies the data for the $n$th pair.
+
+We approximate all of the marginalization integrals using numerical procedures.
+Given distances $d_1$ and $d_2$, we can approximate the velocity marginalization
+using Monte Carlo integration:
+\begin{equation}
+    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \approx
+    \frac{1}{K} \, \sum^K_k \,
+    p(\psi^{(k)} \given f) \quad .
+\end{equation}
+In this expression, the $k$-th sample $\psi^{(k)}$ is defined as
+\begin{eqnarray}
+    \psi^{(k)} &=& |\bs{v}_1^{(k)} - \bs{v}_2^{(k)}| \\
+    \bs{v}_1^{(k)} &=& \mat{M}(\alpha_1, \delta_1) \cdot
+        \left(\begin{array}{c}
+            d_1\,\mu_{\alpha^*, 1}^{(k)} \\
+            d_1\,\mu_{\delta, 1}^{(k)} \\
+            v_{r, 1}^{(k)}
+        \end{array}\right)\\
+    \left(\begin{array}{c}
+        \mu_{\alpha^*, 1}^{(k)} \\
+        \mu_{\delta, 1}^{(k)} \\
+        v_{r, 1}^{(k)}
+    \end{array}\right) &\sim& \mathcal{N}(\bs{\eta}_1, \mat{C}_1)
+\end{eqnarray}
+where $\mat{C}_1$ is the covariance matrix of the \gaia\ data provided by the
+\tgas\ catalog with an additional (uncorrelated) diagonal term added for the
+uncertainty in the radial velocity measurement, and the rotation matrix
+$\mat{M}(\alpha, \delta)$ transforms from the local, tangent-space coordinates
+at $(\alpha, \delta)$ to the ICRS (see, e.g., Section~3 of \citealt{Oh:2017}).
+We perform the outer integral over the two distances, $d_1$ and $d_2$, using
+Simpson's rule on a 2D grid of distances that, in each dimension, spans the
+domain $\left[ (\varpi + 3\sigma_{\varpi})^{-1}, (\varpi -
+3\sigma_{\varpi})^{-1} \right]$, where $\sigma_{\varpi}$ is the parallax
+uncertainty reported in the \tgas\ catalog for each star.
+From convergence tests, we have found that it is sufficient to use 1024 samples
+in the Monte Carlo integral and 5 by 5 distance grid points.
+
+Dispersion ... Nordstrom 2004
+
+\section{Old section on FMLs}
 We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
@@
 
+\section{Results} \label{sec:results}
+
 \subsection{Confirmed comoving pairs}\label{sec:genuine}
@@
 
-\section{Mixture model...}\label{sec:TODO}
-
-% TODO: assumptions:
-% - Same velocity dispersion for all ages. Could have an age-dependent
-%   dispersion
-% -
-With the new radial velocity measurements, we construct a statistical model to
-determine the probability that a given candidate comoving pair is truly
-comoving.
-We don't model the positional separation distribution and instead construct a
-model purely in velocity-space.
-The data are therefore proper motions in the ICRS reference frame, $\bs{\mu}$,
-and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric radial
-velocity, $v_r$, for each star in a candidate comoving pair; we assume that the
-sky position for each star $(\alpha, \delta)$ is known perfectly.
-For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
-\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
-radial velocity data for star 1 in the pair.
-To simplify some expressions, we additionally ignore any correlations between
-the observed parallax and the proper motion components (but keep the covariance
-structure between the two proper motion components).
-
-Our model is expressed in terms of the magnitude of the true (unobserved) 3D
-velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
-and the (global) fraction of pairs that are comoving, $f$.
-For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
-f)$, we use a mixture model with two components
-\begin{eqnarray}
-    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
-\end{eqnarray}
-where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
-pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.\footnote{
-This is actually the marginal likelihood under a model in which every candidate
-comoving pair has an additional latent parameter: a binary flag, $q_n$, that
-specifies which mixture component it belongs too
-(\citealt{Hogg:2010,Foreman-Mackey:2014}). We later use this fact to compute the
-probability that each pair is truly comoving, or, more precisely, belongs to the
-mixture component $p_{\rm co}$).}
-The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
-to be the probability of the magnitude of the 3D separation of two vectors under
-the assumption that the two vectors were drawn independently from an isotropic,
-3D Gaussian with dispersion $\sigma$.
-That is,
-\begin{eqnarray}
-    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
-        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
-    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
-    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
-    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
-\end{eqnarray}
-where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
-\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
-covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
-
-Under this model, the joint probability of the velocity data for a given pair
-$\bs{\eta}_1$ and $\bs{\eta}_2$, and the latent parameters $\bs{v}_1, \bs{v}_2,
-\psi$, conditional on the unobserved true distances $d_1$ and $d_2$ (we handle
-these separately below), is
-\begin{equation}
-p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f) =
-    p(\bs{\eta}_1 \given d_1, \bs{v}_1) \,
-    p(\bs{\eta}_2, v_{r,2} \given d_2, \bs{v}_2) \,
-    p(\bs{v}_1, \bs{v}_2 \given \psi) \, p(\psi \given f)\label{eq:full-ll}
-\end{equation}
-where
-\begin{eqnarray}
-    p(\bs{v}_1, \bs{v}_2 \given \psi) &=& \delta(|\bs{v}_1-\bs{v}_2| - \psi) \quad .
-\end{eqnarray}
-
-Because we are only interested in inferring the fraction of stars consistent
-with being comoving, we marginalize this expression over the true 3D velocities
-and velocity difference, obtaining the velocity-marginalized, per-pair
-likelihood
-\begin{equation}
-    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) =
-    \int \dd^3\bs{v}_1 \, \dd^3\bs{v}_2 \, \dd \psi \,
-    p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f)
-    \quad .
-\end{equation}
-This expression is still conditioned on the true distances, $d_1$ and $d_2$, but
-we observe the parallaxes $\varpi_1$ and $\varpi_2$.
-To remove the final per-pair latent parameters in this model, we therefore also
-have to marginalize the likelihood over these distance parameters to obtain the
-velocity-and-distance marginalized likelihood
-\begin{equation}
-    p(\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2 \given f) =
-        \int \dd d_1 \, \dd d_2 \,
-        p(\varpi_1 \given d_1) \, p(\varpi_2 \given d_2) \,
-        p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \,
-        p(d_1) \, p(d_2)
-\end{equation}
-where we assume a uniform space-density prior for the distances
-(\citealt{Bailer-Jones:2015aa}) with a maximum distance of 1~kpc.
-
-The posterior probability over the fraction of true comoving star pairs, $f$, is
-then given by
-\begin{equation}
-    p(f \given \{\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2\}_{N}) \propto
-        p(f) \, \prod_n^N
-        p(\bs{\eta}_{1,n}, \bs{\eta}_{2,n}, \varpi_{1,n}, \varpi_{2,n} \given f)
-\end{equation}
-where $N$ is the number of candidate comoving star pairs and the index $n$
-specifies the data for the $n$th pair.
-
-We approximate all of the marginalization integrals using numerical procedures.
-Given distances $d_1$ and $d_2$, we can approximate the velocity marginalization
-using Monte Carlo integration:
-\begin{equation}
-    p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \approx
-    \frac{1}{K} \, \sum^K_k \,
-    p(\psi^{(k)} \given f) \quad .
-\end{equation}
-In this expression, the $k$-th sample $\psi^{(k)}$ is defined as
-\begin{eqnarray}
-    \psi^{(k)} &=& |\bs{v}_1^{(k)} - \bs{v}_2^{(k)}| \\
-    \bs{v}_1^{(k)} &=& \mat{M}(\alpha_1, \delta_1) \cdot
-        \left(\begin{array}{c}
-            d_1\,\mu_{\alpha^*, 1}^{(k)} \\
-            d_1\,\mu_{\delta, 1}^{(k)} \\
-            v_{r, 1}^{(k)}
-        \end{array}\right)\\
-    \left(\begin{array}{c}
-        \mu_{\alpha^*, 1}^{(k)} \\
-        \mu_{\delta, 1}^{(k)} \\
-        v_{r, 1}^{(k)}
-    \end{array}\right) &\sim& \mathcal{N}(\bs{\eta}_1, \mat{C}_1)
-\end{eqnarray}
-where $\mat{C}_1$ is the covariance matrix of the \gaia\ data provided by the
-\tgas\ catalog with an additional (uncorrelated) diagonal term added for the
-uncertainty in the radial velocity measurement, and the rotation matrix
-$\mat{M}(\alpha, \delta)$ transforms from the local, tangent-space coordinates
-at $(\alpha, \delta)$ to the ICRS (see, e.g., Section~3 of \citealt{Oh:2017}).
-We perform the outer integral over the two distances, $d_1$ and $d_2$, using
-Simpson's rule on a 2D grid of distances that, in each dimension, spans the
-domain $\left[ (\varpi + 3\sigma_{\varpi})^{-1}, (\varpi -
-3\sigma_{\varpi})^{-1} \right]$, where $\sigma_{\varpi}$ is the parallax
-uncertainty reported in the \tgas\ catalog for each star.
-From convergence tests, we have found that it is sufficient to use 1024 samples
-in the Monte Carlo integral and 5 by 5 distance grid points.
-
-\clearpage
-
 \bibliographystyle{aasjournal}

### 2017-10-02 paper/comoving-followup.tex
footnote - uh oh infinite jest is rubbing off on me

@@
 \newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}
+\DeclareMathOperator{\Tr}{Tr}
 
@@
 
-% TODO: assumptions:
-% - Same velocity dispersion for all ages. Could have an age-dependent
-%   dispersion
-% -
-
 Of the 323 observed star pairs, 12 have at least one member with an unreliable
@@
     f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
-        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right] \\
+        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right]
+        \label{eq:vel-diff-dist}\\
     \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
@@
 
-Dispersion ... Nordstrom 2004
+We use an isotropic model for the field velocity distribution with dispersion
+set by taking the trace of the local velocity dispersion tensor for young
+($<2~{\rm Gyr}$) thin disk stars from \citealt{Nordstrom:2004}, $\mat{V}_{\rm
+N04}$, $\sigma_{\rm field} =
+\sqrt{\frac{1}{3}\Tr{\left(\mat{V}_{\rm N04}\right)}} \approx
+18~\kms$.\footnote{Even though the dispersion in this isotropic model for the
+field stars is only $18~\kms$, most randomly paired stars will tend to have 3D
+velocity differences $>20~\kms$. The 15th and 85th percentiles of the
+distribution of velocity differences (\eqname~\ref{eq:vel-diff-dist}) are
+$\approx 23~\kms$ and $\approx 59~\kms$, respectively.}
+This is a conservative choice: the velocity dispersion of thin disk stars
+increases steadily with age (e.g., \citealt{Aumer:2016}), so assuming that all
+of our pairs are young should lead to under-predictions of the probability that
+a given pair is truly comoving.
+A more robust model that simultaneously models the ages and age distribution of
+the comoving pairs would lead to more precise inferences of these probabilities,
+but is outside of the scope of this work.
+
+We generate samples in $f$ using \package{emcee} (\citealt{Foreman-Mackey:2013})
+with 28 walkers, keeping track of the value of the marginal likelihood under
+each mixture model component for each sample.
+We run the sampler for 256 burn-in steps from a small ball of initial positions
+centered on $f=0.5$, then restart the walkers from the final position and run
+for an additional 1024 steps.
+With the samples in $f$ and the mixture model component likelihoods, we compute
+the per-pair probabilities of belonging to the comoving component (see the final
+section of \citealt{Foreman-Mackey:2014}).
+
+% TODO: summary of assumptions:
+% - parallax-proper-motion covariances small
+% - Same velocity dispersion for all ages. Could have an age-dependent
+%   dispersion
+% - Isotropic field
 

### 2017-10-03 scripts/mixture_model.py
allow runnin on rave too

@@ -273,6 +273,8 @@ if __name__ == "__main__":
                         dest='mpi')
     parser.add_argument('--sim', action='store_true', default=False,
                         dest='simulated_data')
+    parser.add_argument('--name', required=True, dest='name',
+                        help='Name of the data - can be "apw" or "rave"')
 
     args = parser.parse_args()
 
@@ -301,12 +303,15 @@ if __name__ == "__main__":
     else:
         print("Loading real data")
 
+        if args.name not in ['apw', 'rave']:
+            raise ValueError("Invalid name '{0}'".format(args.name))
+
         # Load real data
-        _tbl1 = fits.getdata('../data/tgas_apw1.fits')
+        _tbl1 = fits.getdata('../data/tgas_{0}1.fits'.format(args.name))
         data1 = TGASData(_tbl1, rv=_tbl1['RV']*u.km/u.s,
                          rv_err=_tbl1['RV_err']*u.km/u.s)
 
-        _tbl2 = fits.getdata('../data/tgas_apw2.fits')
+        _tbl2 = fits.getdata('../data/tgas_{0}2.fits'.format(args.name))
         data2 = TGASData(_tbl2, rv=_tbl2['RV']*u.km/u.s,
                          rv_err=_tbl2['RV_err']*u.km/u.s)
 
@@ -316,8 +321,8 @@ if __name__ == "__main__":
     print("Model created")
     # plot_posterior(data1, data2)
 
-    chain_file = '../data/sampler_chain.npy'
-    blobs_file = '../data/sampler_blobs.npy'
+    chain_file = '../data/sampler_chain_{0}.npy'.format(args.name)
+    blobs_file = '../data/sampler_blobs_{0}.npy'.format(args.name)
 
     if not os.path.exists(chain_file):
         print("Couldn't find cached chain file - starting sampling")

### 2017-10-03 scripts/mixture_model.py
decrease vdisp

@@ -54,7 +54,7 @@ def ln_dv_pdf(x, sigma):
 class MixtureModel:
 
     def __init__(self, data1, data2, n_dv_samples=512, n_dist_grid=5,
-                 field_vdisp=25.*u.km/u.s):
+                 field_vdisp=18.*u.km/u.s):
         """
         Parameters
         ----------

### 2017-10-03 scripts/mixture_model.py
name probs file

@@ -239,7 +239,7 @@ def run_emcee(model, pool, chain_file, blobs_file):
         os.remove('../data/sampler_chain{0}.npy'.format(batch))
         os.remove('../data/sampler_blobs{0}.npy'.format(batch))
 
-def analyze_chain(chain, blobs):
+def analyze_chain(chain, blobs, probs_file):
     # MAGIC NUMBER: index after which walkers are converged
     ix = 256
     trim_chain = chain[:,ix:]
@@ -261,7 +261,7 @@ def analyze_chain(chain, blobs):
             norm += 1
     post_prob /= norm
 
-    np.save('../data/pair_probs.npy', post_prob)
+    np.save(probs_file, post_prob)
 
 if __name__ == "__main__":
     import schwimmbad
@@ -330,6 +330,7 @@ if __name__ == "__main__":
         pool.close()
 
     analyze_chain(np.load(chain_file),
-                  np.load(blobs_file))
+                  np.load(blobs_file),
+                  '../data/pair_probs_{0}.npy'.format(args.name))
 
     sys.exit(0)

### 2017-10-09 paper/comoving-followup.tex
Addressing @smoh comments

@@
 For the remaining 311 pairs, we construct a statistical model to determine the
-probability that a given candidate comoving pair is truly comoving.
+probability that a given candidate comoving pair is comoving in 3D velocity
+given the radial velocity and uncertainty measured in this work combined with
+\tgas\ astrometric data.
 For this inference, the data are the proper motions in the ICRS reference frame,
-$\bs{\mu}$, and parallax, $\varpi$, (from the \tgas\ catalog) and barycentric
-radial velocity, $v_r$, for each star in a candidate comoving pair; we assume
-that the sky position for each star $(\alpha, \delta)$ is known perfectly.
+$\bs{\mu} = (\mu_{\alpha^*}, \mu_\delta)$,\footnote{$\mu_{\alpha^*} = \mu_\alpha
+\, \cos\delta$} and parallax, $\varpi$, (from the
+\tgas\ catalog) and barycentric radial velocity, $v_r$, for each star in a
+candidate comoving pair; we assume that the sky position for each star $(\alpha,
+\delta)$ is known perfectly.
 For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
@@
 the observed parallax and the proper motion components (but keep the covariance
-structure between the two proper motion components).
-
-Our model is expressed in terms of the magnitude of the true (unobserved) 3D
+between the two proper motion components).
+
+Our model is expressed in terms of the magnitude of the true (latent) 3D
 velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
@@
 pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.
-Using a mixture model and inferring the mixture weight is a common trick to
-avoid explicitly having to sample over the per-pair probabilities: this
-expression is equivalent to the marginal likelihood under a model in which every
-candidate comoving pair has an additional latent parameter (a binary flag,
-$q_n$) that specifies which mixture component it belongs to
-(\citealt{Hogg:2010,Foreman-Mackey:2014}).
+Using a mixture model and inferring the mixture weight is a trick to avoid
+explicitly having to sample over the probabilities for each individual object
+(candidate comoving pair): this expression is equivalent to the marginal
+likelihood under a model in which every candidate comoving pair has an
+additional latent parameter (a binary flag, $q_n$) that specifies which mixture
+component it belongs to (\citealt{Hogg:2010,Foreman-Mackey:2014}).
 We later use this fact to compute the probability that each pair is truly
@@
 The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
-to be the probability of the magnitude of the 3D separation of two vectors under
-the assumption that the two vectors were drawn independently from an isotropic,
-3D Gaussian with dispersion $\sigma$.
+to be the probability distribution of the magnitude of the 3D difference between
+two vectors under the assumption that the two vectors were drawn independently
+from an isotropic, 3D Gaussian with dispersion $\sigma$.
+For the three-dimensional velocity vectors we consider here, the distribution is
+Maxwellian in form.
 That is,
@@
         \label{eq:vel-diff-dist}\\
-    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{0}, \sigma^2\,\mat{I}_3) \\
-    p_{\rm co}(\psi) &=& f(\psi \,;\, \sigma = 1~\kms)\\
-    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma = \sigma_{\rm field})
+    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{v} \given \bs{0}, \sigma^2\,\mat{I}_3) \\
+    p_{\rm co}(\psi) &=& \mathcal{N}(\psi \given 0, \sigma^2_{\rm co})\\
+    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma_{\rm field})
 \end{eqnarray}
-where $\sigma_{\rm field}$ is defined below, $\mathcal{N}(\bs{m},
-\mat{\Sigma})$ represents a normal distribution with mean vector $\bs{m}$ and
-covariance matrix $\mat{\Sigma}$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
+where $\sigma_{\rm co}$ and $\sigma_{\rm field}$ are defined below,
+$\mathcal{N}(\bs{v} \given \bs{m}, \mat{\Sigma})$ represents a normal
+distribution over $\bs{v}$ with mean vector $\bs{m}$ and covariance matrix
+$\mat{\Sigma}$, $\mathcal{N}(x \given m, \sigma^2)\,[a,b]$ is a truncated normal
+distribution with mean $m$ and variance $\sigma^2$ defined on the domain
+$[a,b]$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
 
@@
     p(\bs{\eta}_1 \given d_1, \bs{v}_1) \,
-    p(\bs{\eta}_2, v_{r,2} \given d_2, \bs{v}_2) \,
+    p(\bs{\eta}_2 \given d_2, \bs{v}_2) \,
     p(\bs{v}_1, \bs{v}_2 \given \psi) \, p(\psi \given f)\label{eq:full-ll}
@@
 Because we are only interested in inferring the fraction of stars consistent
-with being comoving, we marginalize this expression over the true 3D velocities
-and velocity difference, obtaining the velocity-marginalized, per-pair
-likelihood
+with being comoving, $f$, we marginalize this expression over the true 3D
+velocities and velocity difference, obtaining the velocity-marginalized,
+per-pair likelihood
 \begin{equation}
@@
     p(\bs{\eta}_1, \bs{\eta}_2, \bs{v}_1, \bs{v}_2, \psi \given d_1, d_2, f)
-    \quad .
+    \quad . \label{eq:vel-marg}
 \end{equation}
@@
         \int \dd d_1 \, \dd d_2 \,
+        p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \,
         p(\varpi_1 \given d_1) \, p(\varpi_2 \given d_2) \,
-        p(\bs{\eta}_1, \bs{\eta}_2 \given d_1, d_2, f) \,
         p(d_1) \, p(d_2)
@@
 \begin{equation}
-    p(f \given \{\bs{\eta}_1, \bs{\eta}_2, \varpi_1, \varpi_2\}_{N}) \propto
-        p(f) \, \prod_n^N
-        p(\bs{\eta}_{1,n}, \bs{\eta}_{2,n}, \varpi_{1,n}, \varpi_{2,n} \given f)
+    p(f \given \{\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2}\}_N)
+        \propto p(f) \, \prod_n^N
+        p(\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2} \given f)
 \end{equation}
-where $N$ is the number of candidate comoving star pairs and the index $n$
-specifies the data for the $n$th pair.
+where $(\bs{\eta}_{n1}, \varpi_{n1})$ and $(\bs{\eta}_{n2}, \varpi_{n2})$ refer
+to the first and second star in the $n$-th comoving pair, respectively, and  $N$
+is the total number of candidate comoving star pairs.
 
 We approximate all of the marginalization integrals using numerical procedures.
-Given distances $d_1$ and $d_2$, we can approximate the velocity marginalization
-using Monte Carlo integration:
+Given distances $d_1$ and $d_2$, we estimate the velocity marginalization
+(\eqname~\ref{eq:vel-marg}) using Monte Carlo integration:
 \begin{equation}
@@
 \end{equation}
-In this expression, the $k$-th sample $\psi^{(k)}$ is defined as
+In this expression, the $k$-th Monte Carlo sample $\psi^{(k)}$ is defined as
 \begin{eqnarray}
@@
             v_{r, 1}^{(k)}
-        \end{array}\right)\\
+        \end{array}\right)\label{eq:v1-sample} \\
     \left(\begin{array}{c}
@@
         v_{r, 1}^{(k)}
-    \end{array}\right) &\sim& \mathcal{N}(\bs{\eta}_1, \mat{C}_1)
+    \end{array}\right) &\sim& \mathcal{N}(\bs{\eta}_1, \mat{C}_1) \label{eq:mu1-sample}
 \end{eqnarray}
@@
 at $(\alpha, \delta)$ to the ICRS (see, e.g., Section~3 of \citealt{Oh:2017}).
+Expressions for star 2 are equivalent with index 1 replaced with 2 in \eqname
+s~\ref{eq:v1-sample}--\ref{eq:mu1-sample}.
 We perform the outer integral over the two distances, $d_1$ and $d_2$, using
@@
 domain $\left[ (\varpi + 3\sigma_{\varpi})^{-1}, (\varpi -
-3\sigma_{\varpi})^{-1} \right]$, where $\sigma_{\varpi}$ is the parallax
-uncertainty reported in the \tgas\ catalog for each star.
+3\sigma_{\varpi})^{-1} \right]$,\footnote{All of our stars have high enough
+parallax signal-to-noise ratios that integration over this grid is equivalent to
+integrating over much larger domains. We tested this by increasing the grid to
+$\left[ (\varpi + 10\sigma_{\varpi})^{-1}, (\varpi - 10\sigma_{\varpi})^{-1}
+\right]$ and comparing the integral estimates.} where $\sigma_{\varpi}$ is the
+parallax uncertainty reported in the \tgas\ catalog for each star.
 From convergence tests, we have found that it is sufficient to use 1024 samples
@@
 distribution of velocity differences (\eqname~\ref{eq:vel-diff-dist}) are
-$\approx 23~\kms$ and $\approx 59~\kms$, respectively.}
+$\approx 23~\kms$ and $\approx 59~\kms$, respectively.
+This is related to the fact that, for a three-dimensional Gaussian, most of the
+probability \emph{mass} is not at $\bs{0}$, but is rather concentrated in an
+annular region at finite distance from the origin.}
 This is a conservative choice: the velocity dispersion of thin disk stars
@@
 but is outside of the scope of this work.
+
+At the smallest tangential separations of our candidate pairs, ($s_{\rm tan}
+\approx 10^{-3}~\pc$), if the two stars are bound, the expected velocity
+difference is a few $\kms$ (depending on mass).
+We therefore set $\sigma_{\rm co} = 2~\kms$
 

### 2017-10-09 paper/comoving-followup.tex
overhaul results and figures

@@
   \begin{center}
-    \includegraphics[width=0.85\linewidth]{rv-comparison.pdf}
+    \includegraphics[width=\linewidth]{rv-comparison.pdf}
   \end{center}
@@
 
-\section{Methods: Comoving pair probabilities}\label{sec:methods}
+\section{Methods: Comoving pair probabilities}\label{sec:pair-probs}
 
@@
 (flagged) radial velocity measurement from the spectroscopic reduction pipeline.
-For the remaining 311 pairs, we construct a statistical model to determine the
-probability that a given candidate comoving pair is comoving in 3D velocity
-given the radial velocity and uncertainty measured in this work combined with
-\tgas\ astrometric data.
+For the remaining \npairsobs\ pairs, we construct a statistical model to
+determine the probability that a given candidate comoving pair is comoving in 3D
+velocity given the radial velocity and uncertainty measured in this work
+combined with \tgas\ astrometric data.
 For this inference, the data are the proper motions in the ICRS reference frame,
@@
 field stars is only $18~\kms$, most randomly paired stars will tend to have 3D
-velocity differences $>20~\kms$. The 15th and 85th percentiles of the
-distribution of velocity differences (\eqname~\ref{eq:vel-diff-dist}) are
-$\approx 23~\kms$ and $\approx 59~\kms$, respectively.
+velocity differences $>20~\kms$.
+The 15th and 85th percentiles of the distribution of velocity differences
+(\eqname~\ref{eq:vel-diff-dist}) are $\approx 23~\kms$ and $\approx 59~\kms$,
+respectively.
 This is related to the fact that, for a three-dimensional Gaussian, most of the
@@
 difference is a few $\kms$ (depending on mass).
-We therefore set $\sigma_{\rm co} = 2~\kms$
-
-We generate samples in $f$ using \package{emcee} (\citealt{Foreman-Mackey:2013})
-with 28 walkers, keeping track of the value of the marginal likelihood under
-each mixture model component for each sample.
+We use the phrase ``tangential separation'' (and symbol $s_{\rm tan}$) to refer
+to the chord length separation between the two pairs assuming the two stars are
+at the same distance (the smaller distance of the two, $d_{\rm min}$):
+\begin{equation}
+    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
+\end{equation}
+where $\theta$ is the angular extent of the pair.
+We set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function at zero)
+so that binary stars with very small but finite velocity differences are
+considered comoving.
+In principle, this ``tolerance'' on finite velocity difference should decrease
+with increasing separation, at least out to the radius at which most wide
+binaries are disrupted.
+At the unbinding radius, the instantaneous velocity difference between the two
+stars should be comparable to the orbital velocity of the binary at unbinding
+(typically $\approx$0.1--$0.5~\kms$; \citealt{Jiang:2010}), but the difference
+may also increase steadily through dynamical heating.
+Rather than constructing a model for the full process in this work, we use a
+constant velocity tolerance ($\sigma_{\rm co}$) and defer to future work a full
+population analysis that also infers the separation distribution of comoving
+stars, which could instead express the model as a phase-space distribution
+function so that the velocity difference tolerance effectively becomes a
+function of separation.
+We note that even with this tolerance, it is very improbable that two stars
+sampled at random from the field population will have a 3D velocity difference
+$<1~\kms$.
+
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=0.75\linewidth]{mixture-model-demo.pdf}
+  \end{center}
+  \caption{%
+    Probability distributions (normalized to have peak at 1) used for the
+    mixture model described in \sectionname~\ref{sec:pair-probs} as a function
+    of true 3D velocity difference, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$.
+    The field component, $p_{\rm field}(\psi)$, is Maxwellian in form with a
+    dispersion parameter $\sigma_{\rm field} \approx 18~\kms$.
+    The comoving component, $p_{\rm co}(\psi)$, is a normal distribution
+    centered on 0 with a dispersion $\sigma_{\rm co} = 1~\kms$, truncated such
+    that $\psi \geq 0$.
+    \label{fig:mixture-model-demo}}
+\end{figure}
+
+\figurename~\ref{fig:mixture-model-demo} shows the probability distributions for
+each mixture model component, $p_{\rm field}(\psi)$ and $p_{\rm co}(\psi)$,
+normalized so that the peak of the distribution is at 1.
+As noted above, even though the field dispersion is set to $\sigma_{\rm field}
+\approx 18~\kms$, most randomly paired stars will have 3D velocity differences
+(i.e. $\psi$) significantly larger than the value of the dispersion.
+
+To generate samples in $f$, we use Markov Chain Monte Carlo sampling as
+implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
+We use 28 walkers and keep track of the value of the marginal likelihood under
+each mixture model component for each sample at every iteration step.
 We run the sampler for 256 burn-in steps from a small ball of initial positions
@@
 
-% TODO: summary of assumptions:
-% - parallax-proper-motion covariances small
-% - Same velocity dispersion for all ages. Could have an age-dependent
-%   dispersion
-% - Isotropic field
-
-\section{Old section on FMLs}
-We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
-that the two stars are comoving with identical 3D velocity vectors (within some
-allowed small dispersion), and (2) assuming that the two stars have independent
-3D velocity vectors drawn from a prior velocity distribution.
-We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
-a mixture of three Gaussian velocity components for the prior, meant to
-represent the thin disk, thick disk, and halo velocity distributions relative to
-the sun (see \citealt{Oh:2017}).
-
-To compute the new likelihood ratio, we modify the expressions from
-\citet{Oh:2017} to include the radial velocity measurements.
-In the same notation, this requires the following changes:
-\begin{itemize}
-  \item The projection matrix, $\mat{M}$, now has to include the radial unit
-    vector
-    \begin{equation}
-      \mat{M} =
-        \left(
-          \begin{array}{c c c}
-            -\sin\alpha & \cos\alpha & 0 \\
-            -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
-            \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
-          \end{array}
-        \right)
-    \end{equation}
-    where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
-    declination) of a given star.
-  \item The data-space velocity vector, $\vec{y}$, for one star now becomes
-    \begin{equation}
-      \vec{y} =
-        \transpose{\left(
-          \begin{array}{c c c}
-            r_i\,\mu_{\alpha}^* &
-            r_i\,\mu_{\delta} &
-            v_{r}
-          \end{array}
-        \right)}
-    \end{equation}
-    where $\mu_\alpha^*$ is the proper motion in right ascension (including the
-    $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
-    $v_r$ is the radial velocity.
-  \item The covariance matrix becomes the block-diagonal matrix
-    \begin{equation}
-      \mat{\Sigma} = \left(
-        \begin{array}{c c}
-          r^2 \, \mat{C} & 0 \\
-          0 & \sigma_{v_r}^2
-        \end{array}
-      \right)
-    \end{equation}
-    where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
-    (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
-    uncertainty.
-\end{itemize}
-To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
-vector must become the 6-vector
-\begin{equation}
-  \vec{y} = \transpose{\left(
-    \begin{array}{c c}
-      \vec{y}_i &
-      \vec{y}_j
-    \end{array}
-    \right)}
-\end{equation}
-and the covariance matrix becomes the block-diagonal matrix
-\begin{equation}
-  \mat{\Sigma} = \left(
-    \begin{array}{c c}
-      \mat{\Sigma}_i & 0 \\
-      0 & \mat{\Sigma}_j
-    \end{array}
-  \right) \quad .
-\end{equation}
-
-\begin{figure}[htbp]
+\section{Results} \label{sec:results}
+
+\subsection{Probable comoving pairs}\label{sec:genuine}
+
+We use the model described in \sectionname~\ref{sec:pair-probs} to constrain the
+fraction of stars consistent with being comoving given 3D velocities for
+candidate comoving pairs (a) with radial velocities obtained and described in
+this work---\npairsobs\ pairs, and (b) with radial velocities previously
+measured and released in \DR{5} of the Radial Velocity Experiment
+(\acronym{RAVE}; \citealt{Kunder:2017})---202 pairs.
+Using the posterior samples in $f$, we compute the probability (under this
+model) that each given pair is comoving including the radial velocity
+measurements.
+\figurename~\ref{fig:f-samples} shows histograms of posterior pdf samples in $f$
+for pairs with our radial velocity measurements (blue, darker histogram) and
+with \acronym{RAVE} radial velocities (orange, lighter histogram); in either
+case, $f \approx 40\%$ implying a contamination rate $1-f \approx 60\%$.
+
+\begin{figure}[htb]
   \begin{center}
-    \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
+    \includegraphics[width=0.7\linewidth]{f_samples.pdf}
   \end{center}
   \caption{%
-    Point estimate of the 3D velocity difference between two stars in a comoving
-    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
-    log-likelihood-ratio computed when including radial velocity data.
-    Error bars are shown for a few randomly-chosen, representative pairs and
-    represent the 15th and 85th percentiles of velocity differences computed
-    from samples from the error distributions of parallax, proper motion
-    components, and radial velocity.
-    The likelihood ratio increases (i.e. the belief that the stars are comoving
-    becomes stronger) when the 3D velocity difference is small.
-    \label{fig:llr}}
+    Histograms of samples in $f$, the fraction of comoving pairs, for candidate
+    pairs with radial velocities from this work (blue, darker histogram) and
+    from \acronym{RAVE} (orange, lighter histogram).
+    For the \npairsobs\ candidate pairs observed in this work, this amounts to
+    $\approx$120 comoving pairs.
+    For the 202 candidate pairs in \acronym{RAVE}, this amounts to $\approx$90
+    comoving pairs.
+    \label{fig:f-samples}}
 \end{figure}
 
-If the new likelihood ratio is larger than that computed from astrometry alone,
-the added radial velocity information supports the belief that the two stars are
-truly comoving in 3-space.
-Conversely, if the new likelihood ratio is smaller, the radial velocity data
-indicate that the pair is likely a false-positive and are not truly comoving.
-In practice, we compute the logarithm of the likelihood ratios,
-$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
-proper-motion-only log-FML-ratio for a given pair of stars as
-\llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
-radial velocity data as \llrnew.
-We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
-\ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
-\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
-plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
-observed pairs.
-In detail, for visualization, we sample from the error distributions over proper
-motion, parallax, and radial velocity for each star individually and compute the
-magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
-samples.
-We plot the median of the resulting distribution of velocity-differences as
-dark markers (circles).
-The error bars show the 15th and 85th percentiles of the distributions for a few
-randomly chosen pairs.
-Note that pairs with large 3D velocity differences have large (negative) changes
-to their log-likelihood-ratio values.
-Some pairs with large velocity differences also have small (near zero) changes
-to the log-likelihood ratio: these are pairs where each member has a velocity
-that is an outlier with respect to the prior velocity distribution.
-
-The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
-all observed candidate comoving pairs are provided with the associated table
-(\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).
+\figurename s~\ref{fig:dx-dv} and \ref{fig:dx-dv-rave} show the magnitude of the
+3D velocity difference plotted against the tangential separation of the two
+stars in all candidate pairs with radial velocities measured in this work and
+from \acronym{RAVE}, respectively.
+The error bars show the 15--85th percentiles of the 3D velocity difference
+distributions for each candidate pair and black points mark the medians.
+The velocity difference distributions for true comoving pairs with radial
+velocity precisions of $\approx5~\kms$ often have peaks at smaller velocity
+differences, but have medians at larger values because the distributions have
+large tails.
+This is visible in the plot in that the lower error bars tend to be smaller than
+the upper error bar, indicating that the distributions tend to be more
+concentrated towards smaller velocities.
+This, combined with the fact that it is very improbable to get 3D velocity
+differences around $10~\kms$ from the field mixture component, ultimately means
+that many of the pairs with median 3D velocity differences close to $\approx
+10~\kms$ actually end up having a high probability of being comoving (as is seen
+in the blue points in these figures).
 
@@
   \begin{center}
-    \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
+    \includegraphics[width=0.9\linewidth]{dx-dv-apw.pdf}
   \end{center}
   \caption{%
-    Point estimate of the 3D velocity difference between two stars in a comoving
-    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
-    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
-    The distances used in the point estimates are computed using a
-    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
+    Three-dimensional velocity difference between two stars in a comoving pair,
+    $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the tangential
+    (sky) separation of the two stars, $s_{\rm tan}$.
+    Each point corresponds to one candidate comoving pair and is colored by the
+    probability of being comoving under the mixture model described in
+    \sectionname~\ref{sec:pair-probs}.
+    Error bars in velocity represent the 15--85th percentiles of the inferred
+    3D velocity differences for each pair.
+    The left panel shows small-separation pairs ($s_{\rm tan} < 1~\pc$) in
+    log-space; the right panel shows large-separation pairs ($1 < s_{\rm tan} <
+    10~\pc$) with a linear horizontal axis.
     \label{fig:dx-dv}}
@@
 
-\section{Results} \label{sec:results}
-
-\subsection{Confirmed comoving pairs}\label{sec:genuine}
-
-\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
-plotted against the 3D positional separation of the two stars in all observed
-pairs from this work.
-We again plot the median of the velocity difference distribution, and here use
-the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
-the stars' \tgas\ parallax measurements to compute the 3D positions of the two
-stars.
-Small, gray markers are pairs for which $\llrnew < \llrold$, i.e. where the
-radial velocity measurements weaken or change our belief that a given pair is
-comoving.
-Larger, black markers are confirmed comoving pairs ($\llrnew > \llrold$); note
-that many large-separation pairs have been ruled-out as comoving, but
-most small-separation pairs maintain their status as comoving.
-
-\figurename~\ref{fig:separation} shows histograms of the spatial separation and
-sky separation for all pairs; background, gray histogram shows all observed
-pairs, and foreground, black histogram shows all pairs for which $\llrnew >
-\llrold$.
-In this and subsequent figures, we use the phrase ``tangential separation'' (and
-symbol $s_{\rm tan}$) to refer to the chord length separation between the two
-pairs assuming the two stars are at the same distance (the smaller distance of
-the two, $d_{\rm min}$):
-\begin{equation}
-    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
-\end{equation}
-where $\theta$ is the angular extent of the pair.
-
-At sky separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
-confirmed to be comoving; interestingly, the separation distribution of the
-genuine pairs appears to be uniform or possibly growing in number with
-increasing separation.
+\begin{figure}[htb]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{dx-dv-rave.pdf}
+  \end{center}
+  \caption{%
+    Same as \figurename~\ref{fig:dx-dv} for pairs with radial velocities
+    measured by the \acronym{RAVE} survey.
+    \label{fig:dx-dv-rave}}
+\end{figure}
+
+\subsection{Separation distributions}\label{sec:separations}
+
+\figurename~\ref{fig:separation} shows histograms of the tangential separation
+for all pairs with radial velocities from this work and the \acronym{RAVE}
+survey, with each pair weighted by the probability that it is comoving.
+Interestingly, the separation distribution of the probable comoving pairs
+appears to be uniform with increasing separation.
 This strongly suggests the existence of a population of unbound wide binaries or
@@
   \caption{%
-    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
-    the stars in all observed comoving star pairs (gray), and only those
-    confirmed as comoving (black).
-    Note that the false-positive rate is $\approx 60\%$ at large separations
-    ($>1~\pc$), but there are still a significant number of comoving pairs at
-    these very wide separations.
-    \emph{Right}: Distribution of tangential separations for all comoving star
-    pairs (gray) and only those identified as comoving with 3D velocity
-    information (black).
-    Noise in the parallax measurements makes the 3D separation distribution
-    appear roughly uniform.
-    Even the distribution in tangential separation appears to be approximately
-    linear for large separations ($\gtrsim 0.1~\pc$).
+    Weighted histogram---in log-log (\emph{left panel}) and linear (\emph{right
+    panel}) scales---of the tangential separation of all comoving star pairs
+    with radial velocity measurements either from this work or from the
+    \acronym{RAVE} survey.
+    At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log) appears
+    uniform, implying a decreasing distribution in linear space that is
+    qualitatively consistent with past measurements of the wide binary
+    separation distribution.
+    At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to be
+    an underabundance of comoving pairs, but $\gtrsim 1~\pc$ the number of
+    comoving pairs appears to be uniform out to our sample separation limit of
+    $10~\pc$.
     \label{fig:separation}}
@@
 
-\begin{figure}[p]
+To see whether the separation distribution depends on heliocentric distance, we
+also plot the tangential separation of all comoving pairs with comoving
+probability $> 50\%$ as a function of heliocentric distance in
+\figurename~\ref{fig:separation-scatter}.
+The top histogram again shows the tangential separation distribution for the
+joint sample (now only for pairs with $> 50\%$ comoving probability).
+The scatter plot shows tangential separation and the pair distances, and
+the right histogram shows the distance distribution of these high-probability
+pairs.
+Reassuringly, there do not appear to be any trends of the separation
+distribution with distance.
+
+\begin{figure}[htb]
   \begin{center}
-    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
+    \includegraphics[width=0.9\linewidth]{separation-with-rave.pdf}
   \end{center}
   \caption{%
-    \emph{Top}: Distribution of tangential separation for all confirmed comoving
-      star pairs, either from the radial velocity follow-up presented in this
-      work or from RAVE \DR{5} radial velocities.
-      At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log)
-      appears uniform, consistent with past measurements of the wide binary
-      separation distribution.
-      At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to
-      be an underabundance of comoving pairs, but the number of comoving pairs
-      then begins to increase again out to our sample separation limit of
-      $10~\pc$.
+    \emph{Top}: Distribution of tangential separation for all comoving star
+      pairs with probability $>50\%$, either from the radial velocity follow-up
+      presented in this work or from \acronym{RAVE} \DR{5} radial velocities.
     \emph{Middle}: Tangential separation plotted against mean distance for each
@@
       to decline.
-    \label{fig:separation-RAVE}}
+    \label{fig:separation-scatter}}
 \end{figure}
 
-An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
-previously-measured radial velocities (for each star) reported in \DR{5} of the
-Radial Velocity Experiment (RAVE; \citealt{Kunder:2017}).
-Using the same procedure described in \sectionname~\ref{sec:catalog}, we
-identify confirmed comoving pairs using the RAVE velocities for all pairs within
-$200~\pc$ and find a similar false-positive rate at large separations ($\approx
-38\%$ are confirmed).
-Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
-that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
-We plot the confirmed comoving pairs identified through radial velocity
-follow-up in this work along with the RAVE-confirmed pairs in
-\figurename~\ref{fig:separation-RAVE}.
-Top histogram shows the tangential separation distribution for the joint
-sample, scatter plot shows log-separation and the pair distances, and right
-histogram shows the distance distribution of the confirmed pairs.
-At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
-of our method drops, as expected given the hard cut in log-FML-ratio applied to
-define the candidate sample.
-Reassuringly, there do not appear to be any trends of the separation
-distribution with distance.
-
 \figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
-diagram for all confirmed comoving pairs.
+diagram for pairs with radial velocities measured in this work with comoving
+probability $> 0.5$.
 As would be expected if all of the pairs are coeval, the majority of pairs lie
@@
 (\sectionname~\ref{sec:interesting-pairs}).
-
-\clearpage
 
@@
 \end{document}
+
+% \section{Old section on FMLs}
+% We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
+% that the two stars are comoving with identical 3D velocity vectors (within some
+% allowed small dispersion), and (2) assuming that the two stars have independent
+% 3D velocity vectors drawn from a prior velocity distribution.
+% We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
+% a mixture of three Gaussian velocity components for the prior, meant to
+% represent the thin disk, thick disk, and halo velocity distributions relative to
+% the sun (see \citealt{Oh:2017}).
+
+% To compute the new likelihood ratio, we modify the expressions from
+% \citet{Oh:2017} to include the radial velocity measurements.
+% In the same notation, this requires the following changes:
+% \begin{itemize}
+%   \item The projection matrix, $\mat{M}$, now has to include the radial unit
+%     vector
+%     \begin{equation}
+%       \mat{M} =
+%         \left(
+%           \begin{array}{c c c}
+%             -\sin\alpha & \cos\alpha & 0 \\
+%             -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
+%             \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
+%           \end{array}
+%         \right)
+%     \end{equation}
+%     where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
+%     declination) of a given star.
+%   \item The data-space velocity vector, $\vec{y}$, for one star now becomes
+%     \begin{equation}
+%       \vec{y} =
+%         \transpose{\left(
+%           \begin{array}{c c c}
+%             r_i\,\mu_{\alpha}^* &
+%             r_i\,\mu_{\delta} &
+%             v_{r}
+%           \end{array}
+%         \right)}
+%     \end{equation}
+%     where $\mu_\alpha^*$ is the proper motion in right ascension (including the
+%     $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
+%     $v_r$ is the radial velocity.
+%   \item The covariance matrix becomes the block-diagonal matrix
+%     \begin{equation}
+%       \mat{\Sigma} = \left(
+%         \begin{array}{c c}
+%           r^2 \, \mat{C} & 0 \\
+%           0 & \sigma_{v_r}^2
+%         \end{array}
+%       \right)
+%     \end{equation}
+%     where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
+%     (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
+%     uncertainty.
+% \end{itemize}
+% To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
+% vector must become the 6-vector
+% \begin{equation}
+%   \vec{y} = \transpose{\left(
+%     \begin{array}{c c}
+%       \vec{y}_i &
+%       \vec{y}_j
+%     \end{array}
+%     \right)}
+% \end{equation}
+% and the covariance matrix becomes the block-diagonal matrix
+% \begin{equation}
+%   \mat{\Sigma} = \left(
+%     \begin{array}{c c}
+%       \mat{\Sigma}_i & 0 \\
+%       0 & \mat{\Sigma}_j
+%     \end{array}
+%   \right) \quad .
+% \end{equation}
+
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
+%   \end{center}
+%   \caption{%
+%     Point estimate of the 3D velocity difference between two stars in a comoving
+%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
+%     log-likelihood-ratio computed when including radial velocity data.
+%     Error bars are shown for a few randomly-chosen, representative pairs and
+%     represent the 15th and 85th percentiles of velocity differences computed
+%     from samples from the error distributions of parallax, proper motion
+%     components, and radial velocity.
+%     The likelihood ratio increases (i.e. the belief that the stars are comoving
+%     becomes stronger) when the 3D velocity difference is small.
+%     \label{fig:llr}}
+% \end{figure}
+
+% If the new likelihood ratio is larger than that computed from astrometry alone,
+% the added radial velocity information supports the belief that the two stars are
+% truly comoving in 3-space.
+% Conversely, if the new likelihood ratio is smaller, the radial velocity data
+% indicate that the pair is likely a false-positive and are not truly comoving.
+% In practice, we compute the logarithm of the likelihood ratios,
+% $\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
+% proper-motion-only log-FML-ratio for a given pair of stars as
+% \llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
+% radial velocity data as \llrnew.
+% We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
+% \ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
+% \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
+% plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
+% observed pairs.
+% In detail, for visualization, we sample from the error distributions over proper
+% motion, parallax, and radial velocity for each star individually and compute the
+% magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
+% samples.
+% We plot the median of the resulting distribution of velocity-differences as
+% dark markers (circles).
+% The error bars show the 15th and 85th percentiles of the distributions for a few
+% randomly chosen pairs.
+% Note that pairs with large 3D velocity differences have large (negative) changes
+% to their log-likelihood-ratio values.
+% Some pairs with large velocity differences also have small (near zero) changes
+% to the log-likelihood ratio: these are pairs where each member has a velocity
+% that is an outlier with respect to the prior velocity distribution.
+
+% The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
+% all observed candidate comoving pairs are provided with the associated table
+% (\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).
+
+% \begin{figure}[htb]
+%   \begin{center}
+%     \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
+%   \end{center}
+%   \caption{%
+%     Point estimate of the 3D velocity difference between two stars in a comoving
+%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
+%     spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
+%     The distances used in the point estimates are computed using a
+%     Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
+%     \label{fig:dx-dv}}
+% \end{figure}

### 2017-10-10 paper/comoving-followup.tex
confirmed -> high-probability and moved LAMOST sec

@@
 Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
-the number of confirmed pairs per separation decreases with increasing
+the number of high-probability pairs per separation decreases with increasing
 separation (number per log separation is uniform).
@@
 appear to be comoving either in sky-projected velocities --- we refer to these
-as ``comoving pair candidates'' --- or using full-space velocity information ---
-we refer to these as ``confirmed comoving pairs.''
+as ``candidate comoving pairs'' --- or using full-space velocity information ---
+we refer to these as ``high-probability comoving pairs.''
 Radial velocity follow-up is still needed to confirm the majority of the known
@@
 larger separations have been claimed (e.g., \citealt{Yoo:2004,Quinn:2009}), but
-these claims are limited by small numbers of confirmed pairs at large
+these claims are limited by small numbers of high-probability pairs at large
 separations.
@@
 To verify the existence of a population of very wide-separation comoving pairs,
-we first need to identify a large sample of confirmed comoving stars.
+we first need to identify a large sample of high-probability comoving stars.
 This will already be possible for the brightest stars with \gaia\ \DR{2}, but
@@
 velocities for candidate comoving pairs.
-In this work, we present a catalog of confirmed comoving star pairs discovered
-through low-resolution spectroscopic follow-up of candidate comoving stars
-selected from inferences using astrometric measurements from the \tgas\ catalog
-alone (\citealt{Oh:2017}).
+In this work, we present a catalog of comoving star pairs discovered through
+low-resolution spectroscopic follow-up of candidate comoving stars selected from
+inferences using astrometric measurements from the \tgas\ catalog alone
+(\citealt{Oh:2017}).
 
@@
       overlap (squares, orange).
-    \emph{Right}: Distance distribution for all confirmed comoving pairs.
+    \emph{Right}: Distance distribution for all high-probability comoving pairs.
       Note that at distances $\gtrsim 150~\pc$, our detection efficiency appears
@@
   \caption{%
-    {\it Left}: Color-magnitude diagram of confirmed comoving star pairs plotted
-    over the density of all \tgas\ sources within 200 pc; similar to
-    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
+    {\it Left}: Color-magnitude diagram of high-probability comoving star pairs
+    plotted over the density of all \tgas\ sources within 200 pc; similar to
+    \figurename~\ref{fig:sample-cmd}, but for only the high-probability comoving pairs.
     Several interesting pairs that include main-sequence and giant component
@@
 \figurename~\ref{fig:genuine-highlighted-cmd} (right) shows a few highlighted
-pairs selected from the color-magnitude diagram of confirmed comoving stars.
+pairs selected from the color-magnitude diagram of high-probability comoving
+stars.
 We compile \project{Tycho-2} ($B$, $V$; \citealt{Hog:2000}), \tmass\
@@
         \midrule
-        HD 54318 & $4.15 \pm 0.04$ & $10030 \pm 345$ & $2.4 \pm 0.1$ &
-          \multirow{2}{*}{$8.4 \pm 1.1$}\\
-        TYC 3792-658-1 & $4.54 \pm 0.03$ & $5462 \pm 139$ & $0.89 \pm 0.06$ &\\
+        HD 78249 & $3.53 \pm 0.02$ & $4997 \pm 62$ & $1.54 \pm 0.01$ &
+          \multirow{2}{*}{$3.3 \pm 0.2$}\\
+        TYC 3813-167-1 & $4.64 \pm 0.01$ & $4503 \pm 45$ & $0.75 \pm 0.02$ &\\
       \bottomrule
@@
 calibrating and testing stellar evolution models (e.g., \citealt{Torres:2013}).
-With a larger sample of confirmed comoving pairs, identified pairs that contain
-both an evolved star and a main sequence star --- such as this pair --- will be
-of great interest.
-
-\subsubsection{HD 54318 -- TYC 3792-658-1}
-
-The components of this pair appear to be an A- or B-type dwarf (HD 54318) and a
-G-type dwarf (TYC 3792-658-1), the largest magnitude difference between all of
-the confirmed comoving pairs.
-This pair is separated by $2.5^\circ$ on the sky or $8.5~\pc$ in tangential
-separation and have a consistent distance of $\approx 190~\pc$ within their
-uncertainties.
+With a larger sample of high-probability comoving pairs, identified pairs that
+contain both an evolved star and a main sequence star --- such as this pair ---
+will be of great interest.
+
+\subsubsection{HD 78249 -- TYC 3813-167-1}
+
+The components of this pair appear to be a K-type subgiant (HD 78249) and a
+K-type dwarf (TYC 3813-167-1), the largest magnitude difference between all of
+the comoving pairs with probability $>0.5$.
+This pair is separated by $3^\circ$ on the sky or $3.3~\pc$ in tangential
+separation and distances of $55.6 \pm 1.4~\pc$ and $63.8 \pm 1.3~\pc$,
+respectively, implying a 3D separation of $\approx8$--$9~\pc$.
 The mass ratio distribution of comoving star pairs will help constrain star
@@
 
+\subsection{An additional test of comoving pairs with \acronym{LAMOST}}
+
+We use metallicities and radial velocities for an additional, independent set of
+61 candidate comoving pairs that are in
+\DR{3} of the \acronym{LAMOST} survey (\citealt{Deng:2012}) as an additional
+validation test.
+\figurename~\ref{fig:lamost} shows these pairs plotted as the difference in
+radial velocity and difference in iron abundance with points colored by the
+tangential separation of each pair.
+There is a clear over-density of pairs with abundance and radial-velocity
+differences close to zero.
+From this figure it is also clear that the false positives are predominantly the
+largest separation pairs (i.e. are the darkest markers), but a number of
+$>1~\pc$ pairs appear to have small radial-velocity differences and small
+differences in iron abundances, making them even more likely to be comoving
+pairs.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.65\linewidth]{lamost.pdf}
+  \end{center}
+  \caption{%
+    Differences in metallicity and radial velocity for candidate comoving pair
+    member stars from the \acronym{LAMOST} survey.
+    Points represent each pair, colored by the tangential separation, and
+    error bars represent the uncertainties taken from the \acronym{LAMOST}
+    \DR{3} catalog.
+    There is a clear over-density of pairs with near-zero differences in
+    metallicity and radial velocity even at large ($>1~\pc$) separations, adding
+    more confidence to the existence of comoving star pairs with these extreme
+    separations.
+    \label{fig:lamost}}
+\end{figure}
+
 \section{Discussion} \label{sec:discussion}
@@
 The results presented in this work demonstrate that even at separations larger
-than a parsec, a substantial fraction of comoving pair candidates selected from
-astrometry alone are confirmed with three-dimensional velocity information.
+than a parsec, a substantial fraction of candidate comoving pairs selected from
+astrometry alone are consistent with being comoving given three-dimensional
+velocity information.
 This is exciting because it confirms the existence of what is likely a
@@
 pair lookalikes, especially at large separations.
-
-We do not try to estimate a contamination rate in this work, but instead use
-metallicities and radial velocities from 61 candidate comoving pairs that are in
-\DR{3} of the \acronym{LAMOST} survey (\citealt{Deng:2012}) as an additional
-validation test.
-\figurename~\ref{fig:lamost} shows these pairs plotted as the difference in
-radial velocity and difference in iron abundance with points colored by the
-tangential separation of each pair.
-There is a clear over-density of pairs with abundance and radial-velocity
-differences close to zero.
-From this figure it is also clear that the false positives are predominantly the
-largest separation pairs (i.e. are the darkest markers), but a number of
-$>1~\pc$ pairs appear to have small radial-velocity differences and small
-differences in iron abundances, making them even more likely to be genuine
-pairs.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=0.9\linewidth]{lamost.pdf}
-  \end{center}
-  \caption{%
-    Differences in metallicity and radial velocity for candidate comoving pair
-    member stars from the \acronym{LAMOST} survey.
-    Points represent each pair, colored by the tangential separation, and
-    error bars represent the uncertainties taken from the \acronym{LAMOST}
-    \DR{3} catalog.
-    There is a clear over-density of pairs with near-zero differences in
-    metallicity and radial velocity even at large ($>1~\pc$) separations, adding
-    even more confidence to the existence of comoving star pairs with these
-    extreme separations.
-    \label{fig:lamost}}
-\end{figure}
+\todo{however, hard to get 2 stars with same 3d velocities}
 
@@
 With future data releases from the \gaia\ mission combined with large
-spectroscopic surveys, the anticipated catalogs of confirmed comoving star pairs
-will offer a new ``stereo'' view on the evolution and state of our Galaxy.
+spectroscopic surveys, the anticipated catalogs of high-probability comoving
+star pairs will offer a new ``stereo'' view on the evolution and state of our
+Galaxy.
 

### 2017-10-10 paper/comoving-followup.tex
sky -> tan
update table
comments to myself

@@
 \newcommand{\npairsobs}{311}
-\newcommand{\ncomoving}{127}
+\newcommand{\nhighprob}{124}
 \newcommand{\nraveoverlap}{154}
@@
 stellar associations and clusters are born with small spreads in kinematics and
-chemistry, but disrupt and dissociate on timescales comparable to their orbital
-time within the Milky Way.
-The kinematic properties of comoving star pairs---bound wide binaries or unbound
-but coeval stars---are therefore sensitive to both the smooth tidal field and
-gravitational perturbation spectrum of the Milky Way.
+chemistry, but are thought to physically dissociate over short timescales.
+The kinematic properties of comoving stars---bound, wide or unbound but coeval
+systems---are therefore sensitive to both the smooth tidal field and the
+spectrum of massive perturbers (e.g., stellar remnants, dark matter, giant
+molecular clouds) in the Milky Way.
 % Aims
@@
 % \project{Hipparcos} astrometry combined with radial velocity surveys.
-In previous work, we identified $>4,000$ candidate comoving pairs using only
-astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog and
-find a large number of pairs with separations $\gtrsim 1~\pc$.
+In previous work, we identified $>4,000$ candidate comoving star pairs using
+only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog
+and found a large number of candidate comoving pairs with separations $\gtrsim
+1~\pc$.
 Without radial velocity measurements for the vast majority of these pairs, the
-false-positive rate at these separations is significant ($\sim 50\%$).
+false-positive rate at large separations is significant ($\sim 60\%$).
 % Methods
 Here we present results from our own low-resolution radial-velocity survey of a
-random sample of the candidate pairs to identify and validate true comoving star
+random sample of the candidate pairs to identify high-probability comoving star
 pairs.
 % Results
-Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
-pairs with separations as large as $\approx 10~\pc$, our original search limit.
+Of the \npairsobs\ observed comoving pairs, \nhighprob\ comoving pairs have
+probabilities $50\%$ with separations as large as $\approx 10~\pc$, our original
+search limit.
 Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
 the number of high-probability pairs per separation decreases with increasing
-separation (number per log separation is uniform).
+separation.
 From $\approx 0.5~\pc$ to $10~\pc$, the number per separation is approximately
-uniform (number per log separation increases).
+uniform.
 % Conclusions
-We confirm the discovery of a population of comoving star pairs at very large
-separations, suggesting that disrupted wide binaries and stellar associations
-remain approximately spatially coherent.
+We thus confirm the discovery of a population of comoving star pairs at very
+large separations, suggesting that disrupted wide binaries and stellar
+associations remain approximately spatially coherent.
+\todo{something about stellar models and chemical tagging}
 \end{abstract}
-
-% The confirmed discovery of this population of coeval, comoving stars will
-% enable future dynamical tests of dark matter and XX of star formation XX.
-% as predicted from simulations of the disruption and evolution of wide binary
-% stars. Gaia DR2 -> how many do we expect? Can we say anything? Out to what
-% distance and separation for a typical MS star can we determine they are bound
-% assuming high-res spectra to get very precise RVs?
 
@@
 
-To verify the existence of a population of very wide-separation comoving pairs,
-we first need to identify a large sample of high-probability comoving stars.
-This will already be possible for the brightest stars with \gaia\ \DR{2}, but
-obtaining even larger samples will require significant efforts to obtain radial
-velocities for candidate comoving pairs.
-In this work, we present a catalog of comoving star pairs discovered through
+Our primary goal in this work is to identify a sample of high-probability, very
+wide-separation comoving pairs to be followed up with high-resolution
+spectroscopy in future work to determine detailed chemical abundances.
+This will be possible for the brightest stars with \gaia\ \DR{2}, but obtaining
+even larger samples will require significant efforts to obtain radial velocities
+for candidate comoving pairs at fainter magnitudes.
+Here we present a catalog of comoving star pairs discovered through our own
 low-resolution spectroscopic follow-up of candidate comoving stars selected from
@@
 (\citealt{Oh:2017}).
-
-We describe the data and observations in \sectionname~\ref{sec:data}.
-Details of the spectroscopic reduction and velocity measurements are described
-in the Appendix.
+For each pair, we provide a probability of being comoving as determined by
+modeling the 3D velocity information for each pair, with radial velocities
+either from this work, or from \DR{5} of the Radial Velocity Experiment
+(\acronym{RAVE}; \citealt{Kunder:2017}).
+
+We describe the data and observations in \sectionname~\ref{sec:data}, then
+present a probabilistic model to infer the comoving probability for each pair
+with radial velocity measurements in \sectionname~\ref{sec:pair-probs}.
 We present our main results in \sectionname~\ref{sec:results} and
 \sectionname~\ref{sec:discussion}.
+Details of the spectroscopic reduction and radial velocity measurements for our
+observations are described in the Appendix.
 
@@
   \caption{%
-    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
-    blue histogram) and all comoving star pair members observed in this
-    follow-up effort (black points); colored lines connect candidate comoving
-    star pairs.
+    Color-magnitude diagram for all 195,294 \tgas\ sources with \acronym{2MASS}
+    photometry within $200~\pc$ (background blue histogram) and all comoving
+    star pair members observed in this follow-up effort (black points); colored
+    lines connect candidate comoving star pairs.
     \tmass\ (\citealt{Skrutskie:2006}) $J$-band magnitudes are obtained
@@
 2).
-We marginalized over the (unknown) true velocity and distance of each star to
+We marginalized over the (unknown) true 3D velocity and distance of each star to
 obtain the ratio of fully marginalized likelihoods,
@@
 The likelihoods appropriately take into account the reported uncertainties and
-covariances of the astrometric measurements in the \tgas\ catalog
-as well as projection effect in comparing spherical
-velocities for star pairs separated by large angles on the sky.
+covariances of the astrometric measurements in the \tgas\ catalog as well as
+projection effects with comparing spherical velocities for star pairs separated
+by large angles on the sky.
 
 After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$,
-we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
-with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ for each
-star in the \tgas\ catalog.
+for each star, we compute the marginalized likelihood ratio for all nearby stars
+within $10~\pc$ and with a difference in tangential velocity $\Delta v_\perp <
+10~\kms$.
 After applying a conservative cut on the likelihood ratio motivated by computing
@@
 \figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
-observed in this observing run (markers) plotted over the density of all \tgas\
-stars within $200~\pc$.
-Candidate comoving pairs are connected by (colored) lines.
+observed in this observing run (markers) plotted over the density of 195,294
+\tgas\ stars within $200~\pc$ with \acronym{2MASS} photometry.
+Candidate comoving pairs are connected by (arbitrarily colored) lines.
 
@@
 (MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
-adopt as our global systematic error for all absolute radial velocities.
+adopt as our global systematic error for all \emph{absolute} radial velocities.
 We note, however, that for computing \emph{relative} velocities for pairs of
@@
     p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
+    \label{eq:mixture}
 \end{eqnarray}
-where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
-pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.
+where $p_{\rm co}$ is the probability distribution function (pdf) of true
+comoving pairs, and $p_{\rm field}$ is the pdf of randomly paired field stars.
 Using a mixture model and inferring the mixture weight is a trick to avoid
 explicitly having to sample over the probabilities for each individual object
-(candidate comoving pair): this expression is equivalent to the marginal
-likelihood under a model in which every candidate comoving pair has an
-additional latent parameter (a binary flag, $q_n$) that specifies which mixture
-component it belongs to (\citealt{Hogg:2010,Foreman-Mackey:2014}).
+(candidate comoving pair): this expression (\eqname~\ref{eq:mixture}) is
+equivalent to the marginal likelihood under a model in which every candidate
+comoving pair has an additional latent parameter (a binary flag, $q_n$) that
+specifies which mixture component it belongs to
+(\citealt{Hogg:2010,Foreman-Mackey:2014}).
 We later use this fact to compute the probability that each pair is truly
@@
 
-The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
-to be the probability distribution of the magnitude of the 3D difference between
-two vectors under the assumption that the two vectors were drawn independently
-from an isotropic, 3D Gaussian with dispersion $\sigma$.
+For the field mixture component, $p_{\rm field}$, we use the probability
+distribution of the magnitude of the 3D difference between two vectors under the
+assumption that the two vectors were drawn independently from an isotropic, 3D
+Gaussian with dispersion $\sigma_{\rm field}$.
 For the three-dimensional velocity vectors we consider here, the distribution is
 Maxwellian in form.
-That is,
+For the comoving mixture component, $p_{\rm co}$, we use a truncated normal
+distribution centered on 0 with dispersion $\sigma_{\rm co}$ defined over the
+domain $\psi \in [0, \infty]$.
+To summarize,
 \begin{eqnarray}
-    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
+    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{v} \given \bs{0}, \sigma^2\,\mat{I}_3) \\
+    p_{\rm co}(\psi) &=& \mathcal{N}(\psi \given 0, \sigma^2_{\rm co})\,[0, \infty]\\
+    p_{\rm field}(\psi) &=& g(\psi \,;\, \sigma_{\rm field})\\g(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
         \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right]
-        \label{eq:vel-diff-dist}\\
-    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{v} \given \bs{0}, \sigma^2\,\mat{I}_3) \\
-    p_{\rm co}(\psi) &=& \mathcal{N}(\psi \given 0, \sigma^2_{\rm co})\\
-    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma_{\rm field})
+        \label{eq:vel-diff-dist}
 \end{eqnarray}
-where $\sigma_{\rm co}$ and $\sigma_{\rm field}$ are defined below,
-$\mathcal{N}(\bs{v} \given \bs{m}, \mat{\Sigma})$ represents a normal
+where $\mathcal{N}(\bs{v} \given \bs{m}, \mat{\Sigma})$ represents a normal
 distribution over $\bs{v}$ with mean vector $\bs{m}$ and covariance matrix
@@
 distribution with mean $m$ and variance $\sigma^2$ defined on the domain
-$[a,b]$, and $\mat{I}_3$ is the 3 by 3 identity matrix.
+$[a,b]$, $\mat{I}_3$ is the 3 by 3 identity matrix, and the values of
+$\sigma_{\rm co}$ and $\sigma_{\rm field}$ are set below,
 
@@
 where $(\bs{\eta}_{n1}, \varpi_{n1})$ and $(\bs{\eta}_{n2}, \varpi_{n2})$ refer
-to the first and second star in the $n$-th comoving pair, respectively, and  $N$
-is the total number of candidate comoving star pairs.
+to the velocity and parallax data for the first and second star in the $n$-th
+comoving pair, respectively, and $N$ is the total number of candidate comoving
+star pairs.
 
@@
 \begin{equation}
-    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
+    s_{\rm tan} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
 \end{equation}
 where $\theta$ is the angular extent of the pair.
-We set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function at zero)
-so that binary stars with very small but finite velocity differences are
-considered comoving.
+We therefore set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function
+at zero) so that binary stars with very small but finite velocity differences
+are considered comoving.
 In principle, this ``tolerance'' on finite velocity difference should decrease
@@
 may also increase steadily through dynamical heating.
-Rather than constructing a model for the full process in this work, we use a
+Rather than constructing a model for this full process in this work, we use a
 constant velocity tolerance ($\sigma_{\rm co}$) and defer to future work a full
@@
 each mixture model component, $p_{\rm field}(\psi)$ and $p_{\rm co}(\psi)$,
-normalized so that the peak of the distribution is at 1.
+normalized so that the peak of each distribution is at 1.
 As noted above, even though the field dispersion is set to $\sigma_{\rm field}
@@
 
-To generate samples in $f$, we use Markov Chain Monte Carlo sampling as
-implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
+To generate posterior samples in $f$, we use Markov Chain Monte Carlo sampling
+as implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
 We use 28 walkers and keep track of the value of the marginal likelihood under
@@
 We use the model described in \sectionname~\ref{sec:pair-probs} to constrain the
-fraction of stars consistent with being comoving given 3D velocities for
+fraction of stars consistent with being comoving given 3D velocities for the
 candidate comoving pairs (a) with radial velocities obtained and described in
 this work---\npairsobs\ pairs, and (b) with radial velocities previously
-measured and released in \DR{5} of the Radial Velocity Experiment
-(\acronym{RAVE}; \citealt{Kunder:2017})---202 pairs.
+measured and released in \DR{5} of the \acronym{RAVE} survey
+(\citealt{Kunder:2017})---202 pairs.
 Using the posterior samples in $f$, we compute the probability (under this
@@
 survey, with each pair weighted by the probability that it is comoving.
-Interestingly, the separation distribution of the probable comoving pairs
-appears to be uniform with increasing separation.
-This strongly suggests the existence of a population of unbound wide binaries or
+Interestingly, the separation distribution appears to be uniform with increasing
+separation above $\approx1~\pc$.
+This suggests the existence of a population of unbound wide binaries or
 dissolved open clusters, which maintain approximate spatial coherence for long
@@
 widely-separated but comoving stars at even larger separations ($>10~\pc$).
+However, high resolution spectroscopy is needed to determine precise stellar
+parameters, ages, and high-precision radial velocities for all of our
+high-probability comoving pairs to further validate the sample.
 
@@
 
-To see whether the separation distribution depends on heliocentric distance, we
-also plot the tangential separation of all comoving pairs with comoving
-probability $> 50\%$ as a function of heliocentric distance in
+To see whether the separation distribution depends on heliocentric distance,
+which could indicate a problem with our statistical inferences, we also plot the
+tangential separation of all comoving pairs with comoving probability $> 50\%$
+as a function of heliocentric distance in
 \figurename~\ref{fig:separation-scatter}.
@@
 \figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
-diagram for pairs with radial velocities measured in this work with comoving
-probability $> 0.5$.
+diagram for high-probability ($>50\%$) pairs with radial velocities measured in
+this work.
 As would be expected if all of the pairs are coeval, the majority of pairs lie
@@
 5~\pc$.
-At first glance, it seems unlikely that the two stars could be coeval, as the
-more massive A star has not begun its post-main-sequence evolution.
+At first glance, it seems impossible that the two stars could be coeval, given
+that the more massive A star has not begun its post-main-sequence evolution.
 One possible explanation is that this pair is a false-positive in three
@@
 
-We use metallicities and radial velocities for an additional, independent set of
-61 candidate comoving pairs that are in
-\DR{3} of the \acronym{LAMOST} survey (\citealt{Deng:2012}) as an additional
-validation test.
+As an additional validation test of the candidate comoving stars, we use
+metallicities and radial velocities for an independent set of 61 candidate
+comoving pairs that are in \DR{3} of the \acronym{LAMOST} survey
+(\citealt{Deng:2012}).
 \figurename~\ref{fig:lamost} shows these pairs plotted as the difference in
@@
 pair lookalikes, especially at large separations.
-\todo{however, hard to get 2 stars with same 3d velocities}
+However, we reiterate that given the velocity dispersions in the Galactic disk,
+it is unlikely to get 3D velocity differences $<10~\kms$ (see, e.g.,
+\figurename~\ref{fig:mixture-model-demo}).
 
@@
 For example, the outcomes of planet formation may lead to large or appreciable
-abundance differences between coeval stars (e.g., \citealt{Pinsonneault:2001},
-Oh et al., submitted).
+abundance differences between coeval stars (e.g.,
+\citealt{Pinsonneault:2001,Oh:2017a}).
 We therefore caution that---to study the chemical properties of comoving pairs
@@
 pairs presented shows an abundance of comoving pairs at separations between
-$1$--$10~\pc$ (\figurename~\ref{fig:separation-RAVE}), the imposed limit in our
-initial search (\citealt{Oh:2017}).
-This result is further tentative evidence (see also \citealt{Shaya:2011}) for
-the existence of a population of unbound comoving stars that represent the
-ultimate fate of wide binaries and moving groups that slowly dissociate due to
-weak perturbations from stars (\citealt{Jiang:2010}) and other massive
-perturbers in the Galaxy.
+$1$--$10~\pc$ (\figurename~\ref{fig:separation-scatter}), the imposed limit in
+our initial search (\citealt{Oh:2017}).
+This result is further evidence (see also \citealt{Shaya:2011}) for the
+existence of a population of unbound comoving stars that represent the ultimate
+fate of wide binaries and moving groups that slowly dissociate due to weak
+perturbations from stars (\citealt{Jiang:2010}) and other massive perturbers in
+the Galaxy.
 From this work it appears that modeling the population of wide binaries or
@@
         \texttt{Oh17\_group\_id}    &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
-        \texttt{R\_mu}              &      & Log-likelihood-ratio computed with only astrometric information\\
-        \texttt{R\_rv}              &      & Log-likelihood-ratio computed with radial velocities as well\\
+        \texttt{prob}               &      & \todo{Probability ...}\\
         \texttt{sep\_tan}           & \pc  & Median of tangential separation error distribution\\

### 2017-10-10 paper/comoving-followup.tex
wrapping
referee minor concerns

@@
 associations remain approximately spatially coherent.
-\todo{something about stellar models and chemical tagging}
+% \todo{something about stellar models and chemical tagging?}
 \end{abstract}
@@
 decreasing power-law, and the power-law index, $\alpha$, is found to be $\alpha
-\approx 1.5$--$2$ out to angular separations of $\approx 0.25^\circ$ ($\approx
-0.65~\pc$ at a distance of $150~\pc$), apparently consistent with simulations
-(e.g., \citealt{Weinberg:1987}).
+\approx 1.5$--$2$ out to separations $\approx 0.65~\pc$ (angular separations of
+$\approx 0.25^\circ$ at a distance of $150~\pc$), apparently consistent with
+simulations (e.g., \citealt{Weinberg:1987}).
 Tentative reports of a steepening or cutoff of the separation distribution at
@@
 They find $\sim 200$ high probability comoving pairs with a separation
-distribution that falls of until $\sim 1~\pc$, where the number of pairs then
+distribution that falls off until $\sim 1~\pc$, where the number of pairs then
 begins rising in bins of log-separation.
@@
 To select candidate pairs for follow-up spectroscopic observations, we only
-consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
-distance.
-With the SNR cut used above and given the limitations of the \tgas\ catalog
-itself, \tgas\ is $\approx$90\% complete to $200~\pc$ (\citealt{Bovy:2017}).
-From this source target list, we randomly observed targets with airmass $\sec z
-< 1.5$ during our observing run.
+consider mutually exclusive star pairs with Heliocentric distances $d <
+200~\pc$.
+With the SNR cut used above, and within the limits (e.g., magnitude range) of
+the \tgas\ catalog itself, \tgas\ is $\approx$90\% complete to $200~\pc$
+(\citealt{Bovy:2017}).
+From this source target list, we randomly observed targets.
 \figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
@@
 90~{\rm s}$ per observation.
+All of our targets were observed at an airmass $\sec z < 1.5$.
 We are primarily interested in relative velocities between pair members and
@@
 From this work it appears that modeling the population of wide binaries or
-comoving stars with a monotonically decreasing separation distribution---as is commonly done---too simple to also represent the population of unbound systems.
+comoving stars with a monotonically decreasing separation distribution---as is
+commonly done---too simple to also represent the population of unbound systems.
 With future data releases from the \gaia\ mission combined with large
@@
         \texttt{Oh17\_group\_id}    &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
-        \texttt{prob}               &      & \todo{Probability ...}\\
+        \texttt{prob}               &      & Probability of being comoving given 3D velocity data.\\
         \texttt{sep\_tan}           & \pc  & Median of tangential separation error distribution\\

### 2017-11-02 comoving_rv/longslit/wavelength.py
if fails to fit wvln line, make plot

@@ -4,6 +4,7 @@ import numpy as np
 from numpy.polynomial.polynomial import polyval
 from celerite.modeling import Model
 from celerite import terms, GP
+from celerite.solver import LinAlgError
 
 # Project
 from .fitting.gaussian import GaussianLineFitter
@@ -51,7 +52,16 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
                                             arrs=[flux, flux_ivar])
 
         lf = GaussianLineFitter(x_, flux_, ivar_, absorp_emiss=1.)
-        lf.fit()
+        try:
+            lf.fit()
+        except LinAlgError:
+            fig,ax = plt.subplots(1,1)
+            ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
+            ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
+                        marker='', linestyle='none', zorder=-1, alpha=0.5)
+            ax.set_title("Failed to fit line!")
+            fig.tight_layout()
+            plt.show()
 
         # TODO: need to get plot path into here
         # fig = lf.plot_fit()

### 2017-11-02 comoving_rv/longslit/wavelength.py
also plot initial model

@@ -55,10 +55,14 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
         try:
             lf.fit()
         except LinAlgError:
+            lf.init_gp()
+
             fig,ax = plt.subplots(1,1)
             ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
             ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
                         marker='', linestyle='none', zorder=-1, alpha=0.5)
+            ax.plot_fit(axes=ax)
+
             ax.set_title("Failed to fit line!")
             fig.tight_layout()
             plt.show()

### 2017-11-02 comoving_rv/longslit/wavelength.py
typo

@@ -61,7 +61,7 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
             ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
             ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
                         marker='', linestyle='none', zorder=-1, alpha=0.5)
-            ax.plot_fit(axes=ax)
+            lf.plot_fit(axes=ax)
 
             ax.set_title("Failed to fit line!")
             fig.tight_layout()

### 2017-11-02 comoving_rv/longslit/wavelength.py
plot data on right axes

@@ -57,13 +57,12 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
         except LinAlgError:
             lf.init_gp()
 
-            fig,ax = plt.subplots(1,1)
+            fig = lf.plot_fit()
+            ax = fig.axes[1]
             ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
             ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
                         marker='', linestyle='none', zorder=-1, alpha=0.5)
-            lf.plot_fit(axes=ax)
 
-            ax.set_title("Failed to fit line!")
             fig.tight_layout()
             plt.show()
 

### 2017-11-02 comoving_rv/longslit/wavelength.py
allow failure

@@ -55,16 +55,20 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
         try:
             lf.fit()
         except LinAlgError:
-            lf.init_gp()
+            # lf.init_gp()
 
-            fig = lf.plot_fit()
-            ax = fig.axes[1]
-            ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
-            ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
-                        marker='', linestyle='none', zorder=-1, alpha=0.5)
+            # fig = lf.plot_fit()
+            # ax = fig.axes[1]
+            # ax.plot(x_, flux_, drawstyle='steps-mid', marker='', linestyle='-')
+            # ax.errorbar(x_, flux_, 1/np.sqrt(ivar_),
+            #             marker='', linestyle='none', zorder=-1, alpha=0.5)
 
-            fig.tight_layout()
-            plt.show()
+            # fig.tight_layout()
+            # plt.show()
+            logger.warning("Failed to fit line! Skipping...but you should be "
+                           "careful if many fits fail, you might have a bad "
+                           "comparison lamp spectrum.")
+            continue
 
         # TODO: need to get plot path into here
         # fig = lf.plot_fit()

### 2017-11-02 comoving_rv/longslit/wavelength.py
success

@@ -68,7 +68,7 @@ def fit_all_lines(pixels, flux, flux_ivar, line_waves, line_pixels,
             logger.warning("Failed to fit line! Skipping...but you should be "
                            "careful if many fits fail, you might have a bad "
                            "comparison lamp spectrum.")
-            continue
+            lf.success = False
 
         # TODO: need to get plot path into here
         # fig = lf.plot_fit()

### 2017-11-02 scripts/check_all_headers.py
script to check headers

@@ -1,0 +1,28 @@
+# Standard library
+import glob
+from os import path
+from comoving_rv.longslit import GlobImageFileCollection
+
+def main(data_path):
+
+    for night_path in glob.glob(path.join(data_path, 'n?')):
+        ic = GlobImageFileCollection(night_path)
+
+        for hdu, fname in ic.hdus(return_fname=True):
+            n_saturated = (hdu.data > 65530).sum()
+            print('{0}: IMAGETYP={1}, OBJECT={2}, {3} saturated pixels'
+                  .format(fname, hdu.header['IMAGETYP'], hdu.header['OBJECT'],
+                          n_saturated))
+
+if __name__ == "__main__":
+    from argparse import ArgumentParser
+
+    # Define parser object
+    parser = ArgumentParser(description="")
+
+    parser.add_argument('-p', '--path', dest='path', required=True,
+                        help='Path containing night directories (n1, n2, ..)')
+
+    args = parser.parse_args()
+
+    main(data_path=args.path)

### 2017-11-02 comoving_rv/longslit/fitting/core.py
nothing makes any sense

@@ -4,6 +4,7 @@ import inspect
 
 # Third-party
 from celerite import terms, GP
+from celerite.solver import LinalgError
 import matplotlib.pyplot as plt
 import numpy as np
 from scipy.optimize import minimize
@@ -254,7 +255,7 @@ class LineFitter(object):
 
         try:
             ll = self.gp.log_likelihood(self.flux)
-        except RuntimeError:
+        except (RuntimeError, LinalgError):
             return np.inf
 
         if np.isnan(ll):

### 2017-11-02 comoving_rv/longslit/fitting/core.py
capital A

@@ -4,7 +4,7 @@ import inspect
 
 # Third-party
 from celerite import terms, GP
-from celerite.solver import LinalgError
+from celerite.solver import LinAlgError
 import matplotlib.pyplot as plt
 import numpy as np
 from scipy.optimize import minimize
@@ -255,7 +255,7 @@ class LineFitter(object):
 
         try:
             ll = self.gp.log_likelihood(self.flux)
-        except (RuntimeError, LinalgError):
+        except (RuntimeError, LinAlgError):
             return np.inf
 
         if np.isnan(ll):

### 2017-11-10 paper/submit/referee-response.txt
response to ref

We thank the referee for the constructive report and suggested improvements. To
address the referee's concerns, we have redone and rewritten our population
modeling to provide a probabilistic classification of each pair as being
"comoving" or "field." Though our conclusions or results haven't changed as a
result of this change, we agree that it is more robust and more honest to
provide probabilities for each pair rather than perform a hard cut on likelihood
ratio increase. We believe that the re-analysis and re-write of our modeling
section will address many of the broader concerns the referee had with this work
and thus don't reply to every issue specifically. We've also moved away from
using the word "confirmed comoving pairs" and instead say "high-probability."
See especially sections 3 and 4 in the revised draft.

To assist the referee in interpreting the differences, we have also uploaded a
"diff"ed and rendered version of the paper that shows the changes made
(diff.pdf).

Below we address additional, specific concerns raised in the report: lines that
contain questions or comments raised by the referee begin with ">" and our
responses are listed immediately below.

> Minor concerns:
> ===============

> (1) The comparison with LAMOST metallicity and Delta-RV seems important enough
> for the case that it should probably be detailed in its own subsection, rather
> than being placed in the discussion.

We agree, and have moved it into the "Results" section - see Section 4.4

> (2) The description of the data reduction in the appendices is perhaps
> too detailed and too general, especially the 1D spectrum extraction
> and wavelength calibration. If this is standard procedure for
> MDM-MODSPEC reduction, then these details make the paper
> unnecessarily long. Publishing these details might however be useful
> if this is a novel reduction procedure - but this does not seem to be
> the case. The RV measurements on the other hand might be worth
> detailing if the procedure is truly novel, or if it is critical in
> demonstrating the level of RV precision needed for the study.

We aren't aware of a spectroscopic reduction pipeline that uses Gaussian
processes or flexible model for the spectral extraction or wavelength
calibration, so we have chosen to leave the description in this paper. We placed
it in the appendix because we agree that, while cumbersome, does describe a
novel reduction procedure. Most readers can just ignore the appendix.

> Other suggested changes:
> ========================

> Section 1, paragraph 5:
> -----------------------
> "is found to be alpha~1.5-2 out to..." Since the alpha power law is determined
> for the distribution of separations, the sentence should be reversed here: "is
> found to be alpha~1.5-2 out to projected separations ~0.65pc (~0.25degrees of
> angular separation on the sky at a distance of 150pc),"

Fixed.

> Section 1, paragraph 7:
> -----------------------
> "that falls of until" -> "that falls off until"

Fixed.

> Section 2.1, paragraph 2:
> -------------------------
> "...likelihood ratio for all stars within 10pc and with a
> difference..."
>
> Perhaps you mean:
>
> "...likelihood ratio for a star to have another star within 10 pc of
> itself, and with a difference..."

For each star in TGAS, we compute the likelihood ratio for all nearby stars
(within 10pc). We've restructured this sentence to try to clarify.

> Section 2.1, paragraph 3:
> -------------------------
> "...exclusive star pairs within 200 pc in Heliocentric distance."
> ->
> "...exclusive stars pairs with heliocentric distances d<200pc.

Fixed.

> "...used above and given the limitations of the TGAS catalog itself..."
> ->
> "...used above, and within the limits of the TGAS catalog itself
> (e.g. magnitude range), TGAS is..."

Fixed.

> "...with airmass sec z < 1.5..."
> ->
> [The airmass is an incidental parameter of the observations, not an
> intrinsic property of the targets. This statement should therefore be
> moved to section 2.2.]

Fixed.

> "...all TGAS stars within 200 pc."
> ->
> [Please cite the actual number of TGAS stars within 200pc.]

Fixed. In fact, we fixed this to read "all 195,294 TGAS stars with 2MASS
photometry within 200 pc...".

### 2017-12-05 comoving_rv/db/connect.py
don't ensure exists

@@ -11,7 +11,7 @@ from sqlalchemy.ext.declarative import declarative_base
 Session = scoped_session(sessionmaker(autoflush=True, autocommit=False))
 Base = declarative_base()
 
-def db_connect(database_path, ensure_db_exists=True):
+def db_connect(database_path, ensure_db_exists=False):
     """
     Connect to the specified database.
 

### 2017-12-05 scripts/injest_db.py

@@ -72,7 +72,7 @@ def main(db_path, run_root_path, drop_all=False, overwrite=False, **kwargs):
     # --------------------------------------------------------------------------
 
     # connect to the database
-    engine = db_connect(db_path)
+    engine = db_connect(db_path, ensure_db_exists=True)
     # engine.echo = True
     logger.debug("Connected to database at '{}'".format(db_path))
 

### 2017-12-07 paper/comoving-followup.tex
remove comments
edits for ref report 2

@@
 
+\newcommand{\edits}[1]{{\bf #1}}
+
 \definecolor{mediumpersianblue}{rgb}{0.0, 0.4, 0.65}
@@
 chemistry, but are thought to physically dissociate over short timescales.
-The kinematic properties of comoving stars---bound, wide or unbound but coeval
+The kinematic properties of comoving stars---bound, wide or unbound-but-coeval
 systems---are therefore sensitive to both the smooth tidal field and the
@@
 1~\pc$.
-Without radial velocity measurements for the vast majority of these pairs, the
+Without radial velocity measurements for the majority of these pairs, the
 false-positive rate at large separations is significant ($\sim 60\%$).
@@
 % Results
-Of the \npairsobs\ observed comoving pairs, \nhighprob\ comoving pairs have
-probabilities $50\%$ with separations as large as $\approx 10~\pc$, our original
-search limit.
+\edits{Of an initial sample of \npairsobs\ possible comoving pairs targeted
+for follow-up spectroscopy, \nhighprob\ have probabilities $50\%$ with
+separations as large as $\approx 10~\pc$, our original search limit.}
 Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
-the number of high-probability pairs per separation decreases with increasing
-separation.
-From $\approx 0.5~\pc$ to $10~\pc$, the number per separation is approximately
-uniform.
+the number of high-probability pairs per separation \edits{unit} decreases with
+increasing separation.
+\edits{On the other hand, from $\approx 0.5~\pc$ to $10~\pc$, the number per
+separation unit is approximately uniform.}
 % Conclusions
-We thus confirm the discovery of a population of comoving star pairs at very
-large separations, suggesting that disrupted wide binaries and stellar
-associations remain approximately spatially coherent.
-% \todo{something about stellar models and chemical tagging?}
+\edits{This supports the existence of a population of comoving star pairs at
+very large separations, suggestive of the expected population of disrupted wide
+binaries and stellar associations.}
+
 \end{abstract}
@@
 determine the probability that a given candidate comoving pair is comoving in 3D
-velocity given the radial velocity and uncertainty measured in this work
-combined with \tgas\ astrometric data.
+\edits{velocity space} given the radial velocity and uncertainty measured in
+this work combined with \tgas\ astrometric data.
 For this inference, the data are the proper motions in the ICRS reference frame,
@@
 and the (global) fraction of pairs that are comoving, $f$.
+\edits{We note that $f$ is \emph{not} the fraction of pairs that are coeval and
+comoving; here we only model the fraction of pairs that are consistent with
+being comoving in 3D velocity space given our measurements, some of which could
+be chance alignments of stars in full-space position and velocity.
+Therefore, $f$ is not a measure of our contamination rate due to random
+alignments of comoving stars.}
+
 For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
@@
 measurements.
+\edits{We again note that $f$ is not a measure of contamination from chance
+field star pairings, but rather just quantifies the fraction of pairs within our
+observed sample that are consistent with having a small 3D velocity difference.}
 \figurename~\ref{fig:f-samples} shows histograms of posterior pdf samples in $f$
@@
 with \acronym{RAVE} radial velocities (orange, lighter histogram); in either
-case, $f \approx 40\%$ implying a contamination rate $1-f \approx 60\%$.
+case, $f \approx 40\%$ \edits{implying that $1-f \approx 60\%$ of the observed
+pairs are not consistent with being comoving in 3D space.}
 
@@
 the Milky Way.
-With \gaia\ \DR{2}, we will be able to study the statistical properties of these
+With \gaia\ \DR{2}, we will be able to study the statistical properties of
 widely-separated but comoving stars at even larger separations ($>10~\pc$).
-However, high resolution spectroscopy is needed to determine precise stellar
-parameters, ages, and high-precision radial velocities for all of our
-high-probability comoving pairs to further validate the sample.
+\edits{However, we note that without chemical information, we don't know the
+contamination rate of chance field star pairings at large separations.
+High resolution spectroscopy is needed to determine precise stellar parameters,
+ages, and radial velocities for all of our high-probability comoving pairs to
+further validate the sample.}
 
@@
 \subsection{An additional test of comoving pairs with \acronym{LAMOST}}
+\label{sec:lamost}
 
@@
 velocity information.
-This is exciting because it confirms the existence of what is likely a
-population of nearly-coherent but unbound coeval stars.
+This is exciting because it \edits{suggests} the existence of what \edits{could
+be} the population of nearly-coherent but unbound coeval stars.
 It is also encouraging for the prospect of selecting large samples of comoving
@@
 
-Going forward, definitive confirmation of comoving star pairs may require
-precise measurements of chemical abundances combined with precise radial
-velocities, as demonstrated with the \acronym{LAMOST} data.
+Going forward, confirmation of comoving star pairs will require precise
+measurements of chemical abundances combined with precise radial velocities, as
+demonstrated above with the \acronym{LAMOST} data and \edits{in other recent
+work that found large-separation pairs with similar chemical properties}
+(\citealt{Andrews:2017b}).
 However, we would also like to \emph{test} assumptions about chemical
@@
 our initial search (\citealt{Oh:2017}).
+\edits{Though there is likely contamination even in the sample of 3D comoving
+star pairs identified in this work, the fact that other large-separation
+($>1~{\rm pc}$, comoving star pairs in \acronym{LAMOST} and the \acronym{GALAH}
+surveys (Simpson, et al., in prep.) have similar chemical properties suggests
+that some fraction are real.}
 This result is further evidence (see also \citealt{Shaya:2011}) for the
@@
 the Galaxy.
+\edits{We feel that this work justifies a concerted effort to obtain
+high-resolution spectroscopic follow-up to use chemical information to identify
+the largest-separation comoving pairs.}
 From this work it appears that modeling the population of wide binaries or
 comoving stars with a monotonically decreasing separation distribution---as is
-commonly done---too simple to also represent the population of unbound systems.
+commonly done---may be too simple to also represent the population of unbound
+systems.
 With future data releases from the \gaia\ mission combined with large
@@
 the Max-Planck-Institut f\"ur Astronomie, Heidelberg.
+
+\edits{We thank the anonymous referee for feedback that has improved the quality
+of this article.}
 
@@
 \end{document}
-
-% \section{Old section on FMLs}
-% We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
-% that the two stars are comoving with identical 3D velocity vectors (within some
-% allowed small dispersion), and (2) assuming that the two stars have independent
-% 3D velocity vectors drawn from a prior velocity distribution.
-% We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
-% a mixture of three Gaussian velocity components for the prior, meant to
-% represent the thin disk, thick disk, and halo velocity distributions relative to
-% the sun (see \citealt{Oh:2017}).
-
-% To compute the new likelihood ratio, we modify the expressions from
-% \citet{Oh:2017} to include the radial velocity measurements.
-% In the same notation, this requires the following changes:
-% \begin{itemize}
-%   \item The projection matrix, $\mat{M}$, now has to include the radial unit
-%     vector
-%     \begin{equation}
-%       \mat{M} =
-%         \left(
-%           \begin{array}{c c c}
-%             -\sin\alpha & \cos\alpha & 0 \\
-%             -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
-%             \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
-%           \end{array}
-%         \right)
-%     \end{equation}
-%     where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
-%     declination) of a given star.
-%   \item The data-space velocity vector, $\vec{y}$, for one star now becomes
-%     \begin{equation}
-%       \vec{y} =
-%         \transpose{\left(
-%           \begin{array}{c c c}
-%             r_i\,\mu_{\alpha}^* &
-%             r_i\,\mu_{\delta} &
-%             v_{r}
-%           \end{array}
-%         \right)}
-%     \end{equation}
-%     where $\mu_\alpha^*$ is the proper motion in right ascension (including the
-%     $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
-%     $v_r$ is the radial velocity.
-%   \item The covariance matrix becomes the block-diagonal matrix
-%     \begin{equation}
-%       \mat{\Sigma} = \left(
-%         \begin{array}{c c}
-%           r^2 \, \mat{C} & 0 \\
-%           0 & \sigma_{v_r}^2
-%         \end{array}
-%       \right)
-%     \end{equation}
-%     where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
-%     (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
-%     uncertainty.
-% \end{itemize}
-% To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
-% vector must become the 6-vector
-% \begin{equation}
-%   \vec{y} = \transpose{\left(
-%     \begin{array}{c c}
-%       \vec{y}_i &
-%       \vec{y}_j
-%     \end{array}
-%     \right)}
-% \end{equation}
-% and the covariance matrix becomes the block-diagonal matrix
-% \begin{equation}
-%   \mat{\Sigma} = \left(
-%     \begin{array}{c c}
-%       \mat{\Sigma}_i & 0 \\
-%       0 & \mat{\Sigma}_j
-%     \end{array}
-%   \right) \quad .
-% \end{equation}
-
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
-%   \end{center}
-%   \caption{%
-%     Point estimate of the 3D velocity difference between two stars in a comoving
-%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
-%     log-likelihood-ratio computed when including radial velocity data.
-%     Error bars are shown for a few randomly-chosen, representative pairs and
-%     represent the 15th and 85th percentiles of velocity differences computed
-%     from samples from the error distributions of parallax, proper motion
-%     components, and radial velocity.
-%     The likelihood ratio increases (i.e. the belief that the stars are comoving
-%     becomes stronger) when the 3D velocity difference is small.
-%     \label{fig:llr}}
-% \end{figure}
-
-% If the new likelihood ratio is larger than that computed from astrometry alone,
-% the added radial velocity information supports the belief that the two stars are
-% truly comoving in 3-space.
-% Conversely, if the new likelihood ratio is smaller, the radial velocity data
-% indicate that the pair is likely a false-positive and are not truly comoving.
-% In practice, we compute the logarithm of the likelihood ratios,
-% $\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
-% proper-motion-only log-FML-ratio for a given pair of stars as
-% \llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
-% radial velocity data as \llrnew.
-% We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
-% \ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
-% \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
-% plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
-% observed pairs.
-% In detail, for visualization, we sample from the error distributions over proper
-% motion, parallax, and radial velocity for each star individually and compute the
-% magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
-% samples.
-% We plot the median of the resulting distribution of velocity-differences as
-% dark markers (circles).
-% The error bars show the 15th and 85th percentiles of the distributions for a few
-% randomly chosen pairs.
-% Note that pairs with large 3D velocity differences have large (negative) changes
-% to their log-likelihood-ratio values.
-% Some pairs with large velocity differences also have small (near zero) changes
-% to the log-likelihood ratio: these are pairs where each member has a velocity
-% that is an outlier with respect to the prior velocity distribution.
-
-% The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
-% all observed candidate comoving pairs are provided with the associated table
-% (\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).
-
-% \begin{figure}[htb]
-%   \begin{center}
-%     \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
-%   \end{center}
-%   \caption{%
-%     Point estimate of the 3D velocity difference between two stars in a comoving
-%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
-%     spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
-%     The distances used in the point estimates are computed using a
-%     Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
-%     \label{fig:dx-dv}}
-% \end{figure}

### 2017-12-07 paper/comoving-followup.tex
remove galah

@@
 star pairs identified in this work, the fact that other large-separation
-($>1~{\rm pc}$, comoving star pairs in \acronym{LAMOST} and the \acronym{GALAH}
-surveys (Simpson, et al., in prep.) have similar chemical properties suggests
-that some fraction are real.}
+($>1~{\rm pc}$, comoving star pairs in \acronym{LAMOST} have similar chemical properties suggests that some fraction are real.}
 This result is further evidence (see also \citealt{Shaya:2011}) for the

### 2017-12-07 paper/comoving-followup.tex
parens

@@
 star pairs identified in this work, the fact that other large-separation
-($>1~{\rm pc}$, comoving star pairs in \acronym{LAMOST} have similar chemical properties suggests that some fraction are real.}
+($>1~{\rm pc}$), comoving star pairs in \acronym{LAMOST} have similar chemical
+properties suggests that some fraction are real.}
 This result is further evidence (see also \citealt{Shaya:2011}) for the

### 2018-04-08 scripts/mixture_model.py
mixure model

@@ -8,61 +8,19 @@ from astropy.io import fits
 import astropy.units as u
 import matplotlib.pyplot as plt
 import numpy as np
-from scipy.misc import logsumexp
-from scipy.integrate import simps
-from scipy.stats import norm
 from tqdm import tqdm
 import emcee
 from emcee.utils import MPIPool
 
 # Project
-from gwb.coords import get_tangent_basis
 from gwb.data import TGASData
 
-pc_mas_yr_per_km_s = (1*u.km/u.s).to(u.pc*u.mas/u.yr, u.dimensionless_angles()).value
-km_s_per_pc_mas_yr = 1/pc_mas_yr_per_km_s
-
-def get_icrs_samples(data, size=1, seed=None):
-    if seed is not None:
-        np.random.seed(seed)
-
-    all_samples = []
-    for i in range(len(data)):
-        star = data[i]
-        y = np.array([star.parallax.value, star.pmra.value, star.pmdec.value,
-                      star.rv.to(u.km/u.s).value])
-        Cov = star.get_cov()[2:,2:]
-        Cov[-1,-1] = star.rv_err.to(u.km/u.s).value**2
-
-        all_samples.append(np.random.multivariate_normal(y, Cov, size=size))
-    all_samples = np.array(all_samples)
-
-    ra = np.repeat(data.ra.to(u.radian).value[:,None], size, axis=1)
-    dec = np.repeat(data.dec.to(u.radian).value[:,None], size, axis=1)
-
-    # ra : radian
-    # dec : radian
-    # parallax : mas
-    # pm_ra_cosdec : mas/yr
-    # pm_dec : mas/yr
-    # rv : km/s
-    return np.dstack((ra[...,None], dec[...,None], all_samples))
-
-
-def ln_dv_pdf(x, sigma):
-    return 2*np.log(x) - x**2/(4*sigma**2) - 1.2655121234846456 - 3*np.log(sigma)
-
-
-def comoving_ln_pdf(x, sigma):
-    """ Normal distribution truncated at 0 (x >= 0) """
-    x = np.abs(x)
-    return -0.5*(x/sigma)**2 - 0.22579135264472741 - np.log(sigma)
-
+def ln_normal(x, mu, var):
+    return -0.5*((x-mu)**2 / var + np.log(2*np.pi*var))
 
 class MixtureModel:
 
-    def __init__(self, data1, data2, n_dv_samples=512, n_dist_grid=5,
[…]

### 2018-04-08 paper/submit/referee-response-2.txt
ref reports

We thank the referee for the suggestions. We hope we have addressed the major
concerns by re-doing the mixture modeling and adding text to the relevant
sections of this paper. The referee report is included below preceded by ">"
characters, with our responses below points.

> The statistical analysis presented in section 3 appears to be an
> attempt at modeling the 3D velocity differences (\psi) in a subset of
> 311 co-moving pairs, using a combination of a model that represents
> the distribution of velocity differences for real co-moving pairs,
> with one that represents the distribution of velocity differences for
> randomly selected field stars.

> While the model for the distribution of co-moving pairs (P_{co})
> appears satisfactory, the model for the distribution of randomly
> selected field stars (P_{field}) is not.

> Here's the problem:
> The subset of 311 co-moving candidates represents
> a subset of objects that have been *pre-selected* on the basis of
> having similar proper motions (from TGAS). As a result, their
> distribution of velocity differences cannot be represented by the 3D
> distribution of field stars, because these pairs represent a subset of
> objects that *a priori* have motions that are much more similar to
> each other than to normal field stars. Effectively, this subset
> represent stars that are already known to have motions that are
> similar to each other along two components of motion (motion in the
> plane of the sky), and thus the only uncertainty that remains is in
> the radial velocity direction - in other words, this is a 1D problem,
> and not a 3D problem. If you pre-select stars that have identical
> transverse motion, then the distribution of velocity differences is
> going to be a truncated Gaussian, and not a Maxwellian distribution.

> As a result, the modeled probability distribution P_{field} grossly
> underestimates the likelihood that two stars from the subset of common
> proper motion pairs will have total velocity differences near 0 km/s.

> Unfortunately, if that is what was done, the entire analysis in section 3,
> as well as the general conclusions of the manuscript, are invalid.

> Perhaps I have misunderstood the analysis presented in section
> 3. If that is the case, then I would urge the authors to present a
> revised discussion that would explain more clearly what has been done,
> in particular one that would clearly explain how the
> simulationaccounts for the fact that we are dealing with a subset of
> stars that were initially preselected based on having similar proper
> motions, and how this affects the results.

Our selection of the candidates does not involve selecting common proper motion
pairs of stars: as is described in this and in previous work, we ask whether two
stars are consistent with having the same 3D Cartesian velocity given parallax
and proper motion components. This is subtly different, however we do
acknowledge that this will skew the velocity-difference distribution of field
pairs, and the previous mixture modeling may be biased because of selection
effects.

As a more conservative and much simpler model for the comoving fraction, f, we
have re-done the analysis by modeling only the radial velocity differences
between pairs.

We think the referee has misunderstood the interpretation of "f" (the comoving
probability), so we have updated much of the language to try to clarify. f is
not the probability that a given pair is a coeval, comoving pair of stars, but
rather it is just the probability that two stars are comoving in three
dimensions. Given this distinction and the assumptions specified in the text, we
think our model is specified correctly. We have added the following text to the
section on modeling (see bold text in Section 3):

"We note that $f$ is \emph{not} the fraction of pairs that are coeval and
comoving; here we only model the fraction of pairs that are consistent with
being comoving in 3D velocity space given our measurements, some of which could
be chance alignments of stars in full-space position and velocity.
Therefore, $f$ is not a measure of our contamination rate due to random
alignments of comoving stars."

We have also added a similar note to the beginning of the results section:

"We again note that $f$ is not a measure of contamination from chance
field star pairings, but rather just quantifies the fraction of pairs within our
observed sample that are consistent with having a small 3D velocity difference."

In editing the text to clarify this point, we also have softened some of our
claims based on the resulting high-probability pairs. Without also modeling the
total distribution function of the Milky Way and all selection effects, we can't
actually know the contamination fraction of unassociated stars with the same 3D
velocity at large separations. From the LAMOST data visualized in this work, and
from recent work by Andrews et al., large separation (>1 pc) comoving in 3D
pairs with the same chemical abundances do exist, and so we now only state that
further follow-up of our sample of confidently 3D comoving pairs should allow
for a refined sample that contains likely coeval, very wide separation pairs. We
have added some text to the results section to try to clarify this - see bolded
text.

> As pointed out in the first report, it is surprising that so many of
> these common proper motion pairs (selected from Oh et al. 2017) are
> found after these follow-up spectroscopic observations to have
> velocity differences >20km/s, especially for the subset of pairs with
> separations 1pc-10pc. Remember that these are all common proper motion
> pairs, so their 3D motions must be relatively small to begin with;
> effectively their velocity differences start with being close to 0km/s
> when the radial velocity is not included, and then grows to pretty
> large values when the radial velocity data is included. This suggests
> a high level of contamination in the sample, from which one could
> argue that the majority of pairs with distances 1pc-10pc are in fact
> chance aalignment This is a serious concern that must be addressed.

We agree that there is likely significant contamination in the Oh et al. 2017
sample at large separations; this is stated in the original paper, and in this
paper text. One clarification: Oh et al. 2017 do not select pairs as being
"common proper motion," but rather construct a statistical model that accounts
for projection effects when comparing stars with the same 3D velocity that have
large (>~ 1 deg) separations.

> On the other hand, it is clear that of the ~20-25 pairs in the <1pc
> subset, the majority of them appear to be confirmed as co-moving
> objects. The manuscript should probably focus on these objects. For
> pairs with separations >1pc, the present study does not provide
> sufficient evidence to identify any of these objects as a probable
> co-moving system, and instead appear to suggest that most of them are
> chance aalignments This can only be solved with more accurate radial
> velocity information.

We believe that we have shown that some fraction of the large separation pairs
are consistent with being comoving, but agree that we can't conclude whether
these systems are coeval or chance alignments (and have stated this in the
text). We feel that it is still worth discussing these systems because of their
great interest. Precisely because contamination is an issue at large
separations, by identifying pairs that appear to be comoving with low-resolution
spectroscopy, we can improve yields for follow-up with higher resolution
spectroscopy (as is planned and being pursued by other members of our group).
We feel that we have stressed this point in the discussion, but have added some
qualifying statements to the conclusions section to help clarify.

> Other recommended corrections:
> ==============================

> Abstract:
> ---------

> "Of the 311 observed comoving pairs, we confirm 124 comoving pairs
> have probabilities 50% with separations as large as 10 pc, our
> original search limit."
> =>
> "Of an initial sample of 311 possible comoving pairs targeted for
> follow-up spectroscopy, we identify 124 pairs that have co-moving
> probabilities >50% with separations as large as 10 pc, our original
> search limit."

Changed.

> "...the number of high-probability pairs per separation decreases..."
> =>
> "...the number of high-probability pairs per separation unit
> decreases..."

Changed.

> "From ~0.5 pc to 10 pc, the number per separation is approximately
> uniform..."
> =>
> "From ~0.5 pc to 10 pc, on the other hand, the number per separation
> is approximately uniform..."

Changed.

> "We thus confirm the discovery of a population of comoving..."
> =>
> "This supports the existence of a population of comoving..."

Changed.

> Section 3:
> ----------

> "...is comoving in 3D velocity given..."
> =>
> "...is comoving in 3D velocity space given..."

Changed.
